### 2022-06-19 latex/Chapters/Chapter0.tex
update

% Chapter Template

\chapter{Objectives} % Main chapter title

\label{Chapter0} % Change X to a consecutive number; for referencing this chapter elsewhere, use \ref{ChapterX}

%----------------------------------------------------------------------------------------
%	SECTION 1
%----------------------------------------------------------------------------------------

\section{Learning Objectives}

Back when I studied Financial Mathematics, almost 2 decades ago, it was all about probability theory, stochastic calculus and asset (in particular derivatives) pricing. These skills were of course very sought after in the field of options trading, but were also often enough to get a job in algorithmic or systematic trading. By the middle of the 2010s, with the constant advances in computing power and storage, the better availability of off-the-shelves libraries and data sets, I witnessed a first revolution: the field of machine learning became more and more prominent and pretty much overshadowed other (more traditional maths) skills. More recently, a  second revolution has taken not only the world of finance, but that of pretty much every science and industry, by storm: we are now in the artificial intelligence age.
In 2019-2020, I decided it was time to see by myself what this was all about, and if the hype was justified. What better way to do that than embark on a proper MSc in Artificial Intelligence?
\\
Of all the modules I have studied over the last 2 years of the Royal Holloway MSc in AI, I have been the most impressed by DL and  NLP (itself arguably largely an application of DL) and particularly interested in AIPnT, especially our excursion in the field of graphs search (a very traditional CS topic, but which somehow I had not yet had a chance to study in much details). Even though I still believe there is a tremendous amount of malinvestment  everywhere, due in good part to the inability of the average investor to distinguish between serious and scammy AI applications and startups (the same obviously goes for blockchain applications, which might warrant another MSc?), I have totally changed my mind around the potential of DL, DRL and NLP and think they are incredibly promising. I have been astonished to see by myself, through several of the courseworks we have done during the MSc, how incredibly efficient sophisticated ML, DL and DLR algorithms can be, when applied well on the right problems. Sometimes they just vastly outperform more naive and traditional approaches to the point of rendering older approaches entirely obsolete (e.g CV, NLP, game solvers, etc...).
\\
For the project component of the MSc, I thought it would be interesting (and fun) for me to try and apply some of the DL, DRL and search techniques (from AIPnT) to a couple of single player games, such as the sliding-puzzle (of which some variations are well known under different names, e.g. the 8-puzzle and 15-puzzle) and of course the Rubiks' Cube. I am in particular looking to solidify my understanding of DRL by actually implementing and experimenting with concerete (though arguably of limited practical application) problems.

%----------------------------------------------------------------------------------------
%	SECTION 2
%----------------------------------------------------------------------------------------

\section{Project's Objectives}

blablabla

### 2022-06-19 latex/Chapters/Chapter1Ex.tex
update

% Chapter 1

\chapter{Chapter Title Here} % Main chapter title

\label{Chapter1Ex} % For referencing the chapter elsewhere, use \ref{Chapter1}

%----------------------------------------------------------------------------------------

% Define some commands to keep the formatting separated from the content 
\newcommand{\keyword}[1]{\textbf{#1}}
\newcommand{\tabhead}[1]{\textbf{#1}}
\newcommand{\code}[1]{\texttt{#1}}
\newcommand{\file}[1]{\texttt{\bfseries#1}}
\newcommand{\option}[1]{\texttt{\itshape#1}}

%----------------------------------------------------------------------------------------

\section{Welcome and Thank You}
Welcome to this \LaTeX{} Thesis Template, a beautiful and easy to use template for writing a thesis using the \LaTeX{} typesetting system.

If you are writing a thesis (or will be in the future) and its subject is technical or mathematical (though it doesn't have to be), then creating it in \LaTeX{} is highly recommended as a way to make sure you can just get down to the essential writing without having to worry over formatting or wasting time arguing with your word processor.

\LaTeX{} is easily able to professionally typeset documents that run to hundreds or thousands of pages long. With simple mark-up commands, it automatically sets out the table of contents, margins, page headers and footers and keeps the formatting consistent and beautiful. One of its main strengths is the way it can easily typeset mathematics, even \emph{heavy} mathematics. Even if those equations are the most horribly twisted and most difficult mathematical problems that can only be solved on a super-computer, you can at least count on \LaTeX{} to make them look stunning.

%----------------------------------------------------------------------------------------

\section{Learning \LaTeX{}}

\LaTeX{} is not a \textsc{wysiwyg} (What You See is What You Get) program, unlike word processors such as Microsoft Word or Apple's Pages. Instead, a document written for \LaTeX{} is actually a simple, plain text file that contains \emph{no formatting}. You tell \LaTeX{} how you want the formatting in the finished document by writing in simple commands amongst the text, for example, if I want to use \emph{italic text for emphasis}, I write the \verb|\emph{text}| command and put the text I want in italics in between the curly braces. This means that \LaTeX{} is a \enquote{mark-up} language, very much like HTML.

\subsection{A (not so short) Introduction to \LaTeX{}}

If you are new to \LaTeX{}, there is a very good eBook -- freely available online as a PDF file -- called, \enquote{The Not So Short Introduction to \LaTeX{}}. The book's title is typically shortened to just \emph{lshort}. You can download the latest version (as it is occasionally updated) from here:
\url{http://www.ctan.org/tex-archive/info/lshort/english/lshort.pdf}

It is also available in several other languages. Find yours from the list on this page: \url{http://www.ctan.org/tex-archive/info/lshort/}

It is recommended to take a little time out to learn how to use \LaTeX{} by creating several, small `test' documents, or having a close look at several templates on:\\ 
\url{http://www.LaTeXTemplates.com}\\ 
Making the effort now means you're not stuck learning the system when what you \emph{really} need to be doing is writing your thesis.

\subsection{A Short Math Guide for \LaTeX{}}

If you are writing a technical or mathematical thesis, then you may want to read the document by the AMS (American Mathematical Society) called, \enquote{A Short Math Guide for \LaTeX{}}. It can be found online here:
\url{http://www.ams.org/tex/amslatex.html}
under the \enquote{Additional Documentation} section towards the bottom of the page.

\subsection{Common \LaTeX{} Math Symbols}
There are a multitude of mathematical symbols available for \LaTeX{} and it would take a great effort to learn the commands for them all. The most common ones you are likely to use are shown on this page:
\url{http://www.sunilpatel.co.uk/latex-type/latex-math-symbols/}

You can use this page as a reference or crib sheet, the symbols are rendered as large, high quality images so you can quickly find the \LaTeX{} command for the symbol you need.

\subsection{\LaTeX{} on a Mac}

The \LaTeX{} distribution is available for many systems including Windows, Linux and Mac OS X. The package for OS X is called MacTeX and it contains all the applications you need -- bundled together and pre-customized -- for a fully working \LaTeX{} environment and work flow.

MacTeX includes a custom dedicated \LaTeX{} editor called TeXShop for writing your `\file{.tex}' files and BibDesk: a program to manage your references and create your bibliography section just as easily as managing songs and creating playlists in iTunes.

%----------------------------------------------------------------------------------------

\section{Getting Started with this Template}

If you are familiar with \LaTeX{}, then you should explore the directory structure of the template and then proceed to place your own information into the \emph{THESIS INFORMATION} block of the \file{main.tex} file. You can then modify the rest of this file to your unique specifications based on your degree/university. Section \ref{FillingFile} on page \pageref{FillingFile} will help you do this. Make sure you also read section \ref{ThesisConventions} about thesis conventions to get the most out of this template.

If you are new to \LaTeX{} it is recommended that you carry on reading through the rest of the information in this document.

Before you begin using this template you should ensure that its style complies with the thesis style guidelines imposed by your institution. In most cases this template style and layout will be suitable. If it is not, it may only require a small change to bring the template in line with your institution's recommendations. These modifications will need to be done on the \file{MastersDoctoralThesis.cls} file.

\subsection{About this Template}

This \LaTeX{} Thesis Template is originally based and created around a \LaTeX{} style file created by Steve R.\ Gunn from the University of Southampton (UK), department of Electronics and Computer Science. You can find his original thesis style file at his site, here:
\url{http://www.ecs.soton.ac.uk/~srg/softwaretools/document/templates/}

Steve's \file{ecsthesis.cls} was then taken by Sunil Patel who modified it by creating a skeleton framework and folder structure to place the thesis files in. The resulting template can be found on Sunil's site here:
\url{http://www.sunilpatel.co.uk/thesis-template}

Sunil's template was made available through \url{http://www.LaTeXTemplates.com} where it was modified many times based on user requests and questions. Version 2.0 and onwards of this template represents a major modification to Sunil's template and is, in fact, hardly recognisable. The work to make version 2.0 possible was carried out by \href{mailto:vel@latextemplates.com}{Vel} and Johannes Böttcher.

%----------------------------------------------------------------------------------------

\section{What this Template Includes}

\subsection{Folders}

This template comes as a single zip file that expands out to several files and folders. The folder names are mostly self-explanatory:

\keyword{Appendices} -- this is the folder where you put the appendices. Each appendix should go into its own separate \file{.tex} file. An example and template are included in the directory.

\keyword{Chapters} -- this is the folder where you put the thesis chapters. A thesis usually has about six chapters, though there is no hard rule on this. Each chapter should go in its own separate \file{.tex} file and they can be split as:
\begin{itemize}
\item Chapter 1: Introduction to the thesis topic
\item Chapter 2: Background information and theory
\item Chapter 3: (Laboratory) experimental setup
\item Chapter 4: Details of experiment 1
\item Chapter 5: Details of experiment 2
\item Chapter 6: Discussion of the experimental results
\item Chapter 7: Conclusion and future directions
\end{itemize}
This chapter layout is specialised for the experimental sciences, your discipline may be different.

\keyword{Figures} -- this folder contains all figures for the thesis. These are the final images that will go into the thesis document.

\subsection{Files}

Included are also several files, most of them are plain text and you can see their contents in a text editor. After initial compilation, you will see that more auxiliary files are created by \LaTeX{} or BibTeX and which you don't need to delete or worry about:

\keyword{example.bib} -- this is an important file that contains all the bibliographic information and references that you will be citing in the thesis for use with BibTeX. You can write it manually, but there are reference manager programs available that will create and manage it for you. Bibliographies in \LaTeX{} are a large subject and you may need to read about BibTeX before starting with this. Many modern reference managers will allow you to export your references in BibTeX format which greatly eases the amount of work you have to do.

\keyword{MastersDoctoralThesis.cls} -- this is an important file. It is the class file that tells \LaTeX{} how to format the thesis.

\keyword{main.pdf} -- this is your beautifully typeset thesis (in the PDF file format) created by \LaTeX{}. It is supplied in the PDF with the template and after you compile the template you should get an identical version.

\keyword{main.tex} -- this is an important file. This is the file that you tell \LaTeX{} to compile to produce your thesis as a PDF file. It contains the framework and constructs that tell \LaTeX{} how to layout the thesis. It is heavily commented so you can read exactly what each line of code does and why it is there. After you put your own information into the \emph{THESIS INFORMATION} block -- you have now started your thesis!

Files that are \emph{not} included, but are created by \LaTeX{} as auxiliary files include:

\keyword{main.aux} -- this is an auxiliary file generated by \LaTeX{}, if it is deleted \LaTeX{} simply regenerates it when you run the main \file{.tex} file.

\keyword{main.bbl} -- this is an auxiliary file generated by BibTeX, if it is deleted, BibTeX simply regenerates it when you run the \file{main.aux} file. Whereas the \file{.bib} file contains all the references you have, this \file{.bbl} file contains the references you have actually cited in the thesis and is used to build the bibliography section of the thesis.

\keyword{main.blg} -- this is an auxiliary file generated by BibTeX, if it is deleted BibTeX simply regenerates it when you run the main \file{.aux} file.

\keyword{main.lof} -- this is an auxiliary file generated by \LaTeX{}, if it is deleted \LaTeX{} simply regenerates it when you run the main \file{.tex} file. It tells \LaTeX{} how to build the \emph{List of Figures} section.

\keyword{main.log} -- this is an auxiliary file generated by \LaTeX{}, if it is deleted \LaTeX{} simply regenerates it when you run the main \file{.tex} file. It contains messages from \LaTeX{}, if you receive errors and warnings from \LaTeX{}, they will be in this \file{.log} file.

\keyword{main.lot} -- this is an auxiliary file generated by \LaTeX{}, if it is deleted \LaTeX{} simply regenerates it when you run the main \file{.tex} file. It tells \LaTeX{} how to build the \emph{List of Tables} section.

\keyword{main.out} -- this is an auxiliary file generated by \LaTeX{}, if it is deleted \LaTeX{} simply regenerates it when you run the main \file{.tex} file.

So from this long list, only the files with the \file{.bib}, \file{.cls} and \file{.tex} extensions are the most important ones. The other auxiliary files can be ignored or deleted as \LaTeX{} and BibTeX will regenerate them.

%----------------------------------------------------------------------------------------

\section{Filling in Your Information in the \file{main.tex} File}\label{FillingFile}

You will need to personalise the thesis template and make it your own by filling in your own information. This is done by editing the \file{main.tex} file in a text editor or your favourite LaTeX environment.

Open the file and scroll down to the third large block titled \emph{THESIS INFORMATION} where you can see the entries for \emph{University Name}, \emph{Department Name}, etc \ldots

Fill out the information about yourself, your group and institution. You can also insert web links, if you do, make sure you use the full URL, including the \code{http://} for this. If you don't want these to be linked, simply remove the \verb|\href{url}{name}| and only leave the name.

When you have done this, save the file and recompile \code{main.tex}. All the information you filled in should now be in the PDF, complete with web links. You can now begin your thesis proper!

%----------------------------------------------------------------------------------------

\section{The \code{main.tex} File Explained}

The \file{main.tex} file contains the structure of the thesis. There are plenty of written comments that explain what pages, sections and formatting the \LaTeX{} code is creating. Each major document element is divided into commented blocks with titles in all capitals to make it obvious what the following bit of code is doing. Initially there seems to be a lot of \LaTeX{} code, but this is all formatting, and it has all been taken care of so you don't have to do it.

Begin by checking that your information on the title page is correct. For the thesis declaration, your institution may insist on something different than the text given. If this is the case, just replace what you see with what is required in the \emph{DECLARATION PAGE} block.

Then comes a page which contains a funny quote. You can put your own, or quote your favourite scientist, author, person, and so on. Make sure to put the name of the person who you took the quote from.

Following this is the abstract page which summarises your work in a condensed way and can almost be used as a standalone document to describe what you have done. The text you write will cause the heading to move up so don't worry about running out of space.

Next come the acknowledgements. On this page, write about all the people who you wish to thank (not forgetting parents, partners and your advisor/supervisor).

The contents pages, list of figures and tables are all taken care of for you and do not need to be manually created or edited. The next set of pages are more likely to be optional and can be deleted since they are for a more technical thesis: insert a list of abbreviations you have used in the thesis, then a list of the physical constants and numbers you refer to and finally, a list of mathematical symbols used in any formulae. Making the effort to fill these tables means the reader has a one-stop place to refer to instead of searching the internet and references to try and find out what you meant by certain abbreviations or symbols.

The list of symbols is split into the Roman and Greek alphabets. Whereas the abbreviations and symbols ought to be listed in alphabetical order (and this is \emph{not} done automatically for you) the list of physical constants should be grouped into similar themes.

The next page contains a one line dedication. Who will you dedicate your thesis to?

Finally, there is the block where the chapters are included. Uncomment the lines (delete the \code{\%} character) as you write the chapters. Each chapter should be written in its own file and put into the \emph{Chapters} folder and named \file{Chapter1}, \file{Chapter2}, etc\ldots Similarly for the appendices, uncomment the lines as you need them. Each appendix should go into its own file and placed in the \emph{Appendices} folder.

After the preamble, chapters and appendices finally comes the bibliography. The bibliography style (called \option{authoryear}) is used for the bibliography and is a fully featured style that will even include links to where the referenced paper can be found online. Do not underestimate how grateful your reader will be to find that a reference to a paper is just a click away. Of course, this relies on you putting the URL information into the BibTeX file in the first place.

%----------------------------------------------------------------------------------------

\section{Thesis Features and Conventions}\label{ThesisConventions}

To get the best out of this template, there are a few conventions that you may want to follow.

One of the most important (and most difficult) things to keep track of in such a long document as a thesis is consistency. Using certain conventions and ways of doing things (such as using a Todo list) makes the job easier. Of course, all of these are optional and you can adopt your own method.

\subsection{Printing Format}

This thesis template is designed for double sided printing (i.e. content on the front and back of pages) as most theses are printed and bound this way. Switching to one sided printing is as simple as uncommenting the \option{oneside} option of the \code{documentclass} command at the top of the \file{main.tex} file. You may then wish to adjust the margins to suit specifications from your institution.

The headers for the pages contain the page number on the outer side (so it is easy to flick through to the page you want) and the chapter name on the inner side.

The text is set to 11 point by default with single line spacing, again, you can tune the text size and spacing should you want or need to using the options at the very start of \file{main.tex}. The spacing can be changed similarly by replacing the \option{singlespacing} with \option{onehalfspacing} or \option{doublespacing}.

\subsection{Using US Letter Paper}

The paper size used in the template is A4, which is the standard size in Europe. If you are using this thesis template elsewhere and particularly in the United States, then you may have to change the A4 paper size to the US Letter size. This can be done in the margins settings section in \file{main.tex}.

Due to the differences in the paper size, the resulting margins may be different to what you like or require (as it is common for institutions to dictate certain margin sizes). If this is the case, then the margin sizes can be tweaked by modifying the values in the same block as where you set the paper size. Now your document should be set up for US Letter paper size with suitable margins.

\subsection{References}

The \code{biblatex} package is used to format the bibliography and inserts references such as this one \parencite{Reference1}. The options used in the \file{main.tex} file mean that the in-text citations of references are formatted with the author(s) listed with the date of the publication. Multiple references are separated by semicolons (e.g. \parencite{Reference2, Reference1}) and references with more than three authors only show the first author with \emph{et al.} indicating there are more authors (e.g. \parencite{Reference3}). This is done automatically for you. To see how you use references, have a look at the \file{Chapter1.tex} source file. Many reference managers allow you to simply drag the reference into the document as you type.

Scientific references should come \emph{before} the punctuation mark if there is one (such as a comma or period). The same goes for footnotes\footnote{Such as this footnote, here down at the bottom of the page.}. You can change this but the most important thing is to keep the convention consistent throughout the thesis. Footnotes themselves should be full, descriptive sentences (beginning with a capital letter and ending with a full stop). The APA6 states: \enquote{Footnote numbers should be superscripted, [...], following any punctuation mark except a dash.} The Chicago manual of style states: \enquote{A note number should be placed at the end of a sentence or clause. The number follows any punctuation mark except the dash, which it precedes. It follows a closing parenthesis.}

The bibliography is typeset with references listed in alphabetical order by the first author's last name. This is similar to the APA referencing style. To see how \LaTeX{} typesets the bibliography, have a look at the very end of this document (or just click on the reference number links in in-text citations).

\subsubsection{A Note on bibtex}

The bibtex backend used in the template by default does not correctly handle unicode character encoding (i.e. "international" characters). You may see a warning about this in the compilation log and, if your references contain unicode characters, they may not show up correctly or at all. The solution to this is to use the biber backend instead of the outdated bibtex backend. This is done by finding this in \file{main.tex}: \option{backend=bibtex} and changing it to \option{backend=biber}. You will then need to delete all auxiliary BibTeX files and navigate to the template directory in your terminal (command prompt). Once there, simply type \code{biber main} and biber will compile your bibliography. You can then compile \file{main.tex} as normal and your bibliography will be updated. An alternative is to set up your LaTeX editor to compile with biber instead of bibtex, see \href{http://tex.stackexchange.com/questions/154751/biblatex-with-biber-configuring-my-editor-to-avoid-undefined-citations/}{here} for how to do this for various editors.

\subsection{Tables}

Tables are an important way of displaying your results, below is an example table which was generated with this code:

{\small
\begin{verbatim}
\begin{table}
\caption{The effects of treatments X and Y on the four groups studied.}
\label{tab:treatments}
\centering
\begin{tabular}{l l l}
\toprule
\tabhead{Groups} & \tabhead{Treatment X} & \tabhead{Treatment Y} \\
\midrule
1 & 0.2 & 0.8\\
2 & 0.17 & 0.7\\
3 & 0.24 & 0.75\\
4 & 0.68 & 0.3\\
\bottomrule\\
\end{tabular}
\end{table}
\end{verbatim}
}

\begin{table}
\caption{The effects of treatments X and Y on the four groups studied.}
\label{tab:treatments}
\centering
\begin{tabular}{l l l}
\toprule
\tabhead{Groups} & \tabhead{Treatment X} & \tabhead{Treatment Y} \\
\midrule
1 & 0.2 & 0.8\\
2 & 0.17 & 0.7\\
3 & 0.24 & 0.75\\
4 & 0.68 & 0.3\\
\bottomrule\\
\end{tabular}
\end{table}

You can reference tables with \verb|\ref{<label>}| where the label is defined within the table environment. See \file{Chapter1.tex} for an example of the label and citation (e.g. Table~\ref{tab:treatments}).

\subsection{Figures}

There will hopefully be many figures in your thesis (that should be placed in the \emph{Figures} folder). The way to insert figures into your thesis is to use a code template like this:
\begin{verbatim}
\begin{figure}
\centering
\includegraphics{Figures/Electron}
\decoRule
\caption[An Electron]{An electron (artist's impression).}
\label{fig:Electron}
\end{figure}
\end{verbatim}
Also look in the source file. Putting this code into the source file produces the picture of the electron that you can see in the figure below.

\begin{figure}[th]
\centering
\includegraphics{Figures/Electron}
\decoRule
\caption[An Electron]{An electron (artist's impression).}
\label{fig:Electron}
\end{figure}

Sometimes figures don't always appear where you write them in the source. The placement depends on how much space there is on the page for the figure. Sometimes there is not enough room to fit a figure directly where it should go (in relation to the text) and so \LaTeX{} puts it at the top of the next page. Positioning figures is the job of \LaTeX{} and so you should only worry about making them look good!

Figures usually should have captions just in case you need to refer to them (such as in Figure~\ref{fig:Electron}). The \verb|\caption| command contains two parts, the first part, inside the square brackets is the title that will appear in the \emph{List of Figures}, and so should be short. The second part in the curly brackets should contain the longer and more descriptive caption text.

The \verb|\decoRule| command is optional and simply puts an aesthetic horizontal line below the image. If you do this for one image, do it for all of them.

\LaTeX{} is capable of using images in pdf, jpg and png format.

\subsection{Typesetting mathematics}

If your thesis is going to contain heavy mathematical content, be sure that \LaTeX{} will make it look beautiful, even though it won't be able to solve the equations for you.

The \enquote{Not So Short Introduction to \LaTeX} (available on \href{http://www.ctan.org/tex-archive/info/lshort/english/lshort.pdf}{CTAN}) should tell you everything you need to know for most cases of typesetting mathematics. If you need more information, a much more thorough mathematical guide is available from the AMS called, \enquote{A Short Math Guide to \LaTeX} and can be downloaded from:
\url{ftp://ftp.ams.org/pub/tex/doc/amsmath/short-math-guide.pdf}

There are many different \LaTeX{} symbols to remember, luckily you can find the most common symbols in \href{http://ctan.org/pkg/comprehensive}{The Comprehensive \LaTeX~Symbol List}.

You can write an equation, which is automatically given an equation number by \LaTeX{} like this:
\begin{verbatim}
\begin{equation}
E = mc^{2}
\label{eqn:Einstein}
\end{equation}
\end{verbatim}

This will produce Einstein's famous energy-matter equivalence equation:
\begin{equation}
E = mc^{2}
\label{eqn:Einstein}
\end{equation}

All equations you write (which are not in the middle of paragraph text) are automatically given equation numbers by \LaTeX{}. If you don't want a particular equation numbered, use the unnumbered form:
\begin{verbatim}
\[ a^{2}=4 \]
\end{verbatim}

%----------------------------------------------------------------------------------------

\section{Sectioning and Subsectioning}

You should break your thesis up into nice, bite-sized sections and subsections. \LaTeX{} automatically builds a table of Contents by looking at all the \verb|\chapter{}|, \verb|\section{}|  and \verb|\subsection{}| commands you write in the source.

The Table of Contents should only list the sections to three (3) levels. A \verb|chapter{}| is level zero (0). A \verb|\section{}| is level one (1) and so a \verb|\subsection{}| is level two (2). In your thesis it is likely that you will even use a \verb|subsubsection{}|, which is level three (3). The depth to which the Table of Contents is formatted is set within \file{MastersDoctoralThesis.cls}. If you need this changed, you can do it in \file{main.tex}.

%----------------------------------------------------------------------------------------

\section{In Closing}

You have reached the end of this mini-guide. You can now rename or overwrite this pdf file and begin writing your own \file{Chapter1.tex} and the rest of your thesis. The easy work of setting up the structure and framework has been taken care of for you. It's now your job to fill it out!

Good luck and have lots of fun!

\begin{flushright}
Guide written by ---\\
Sunil Patel: \href{http://www.sunilpatel.co.uk}{www.sunilpatel.co.uk}\\
Vel: \href{http://www.LaTeXTemplates.com}{LaTeXTemplates.com}
\end{flushright}

### 2022-06-19 latex/Chapters/Chapter2.tex
update

% Chapter Template

\chapter{Games} % Main chapter title

\label{Chapter2} % Change X to a consecutive number; for referencing this chapter elsewhere, use \ref{ChapterX}

%----------------------------------------------------------------------------------------
%	SECTION 1
%----------------------------------------------------------------------------------------

\section{Sliding Puzzle}

blablabla

%-----------------------------------
%	SUBSECTION 1
%-----------------------------------
\section{Rubiks' Cube}

blabla

### 2022-06-19 latex/main.tex
update

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Masters/Doctoral Thesis 
% LaTeX Template
% Version 2.5 (27/8/17)
%
% This template was downloaded from:
% http://www.LaTeXTemplates.com
%
% Version 2.x major modifications by:
% Vel (vel@latextemplates.com)
%
% This template is based on a template by:
% Steve Gunn (http://users.ecs.soton.ac.uk/srg/softwaretools/document/templates/)
% Sunil Patel (http://www.sunilpatel.co.uk/thesis-template/)
%
% Template license:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[
11pt, % The default document font size, options: 10pt, 11pt, 12pt
%oneside, % Two side (alternating margins) for binding by default, uncomment to switch to one side
english, % ngerman for German
singlespacing, % Single line spacing, alternatives: onehalfspacing or doublespacing
%draft, % Uncomment to enable draft mode (no pictures, no links, overfull hboxes indicated)
%nolistspacing, % If the document is onehalfspacing or doublespacing, uncomment this to set spacing in lists to single
%liststotoc, % Uncomment to add the list of figures/tables/etc to the table of contents
%toctotoc, % Uncomment to add the main table of contents to the table of contents
%parskip, % Uncomment to add space between paragraphs
%nohyperref, % Uncomment to not load the hyperref package
headsepline, % Uncomment to get a line under the header
%chapterinoneline, % Uncomment to place the chapter title next to the number on one line
%consistentlayout, % Uncomment to change the layout of the declaration, abstract and acknowledgements pages to match the default layout
]{MastersDoctoralThesis} % The class file specifying the document structure

\usepackage[utf8]{inputenc} % Required for inputting international characters
\usepackage[T1]{fontenc} % Output font encoding for international characters

\usepackage{mathpazo} % Use the Palatino font by default

\usepackage[backend=bibtex,style=authoryear,natbib=true]{biblatex} % Use the bibtex backend with the authoryear citation style (which resembles APA)

\addbibresource{example.bib} % The filename of the bibliography

\usepackage[autostyle=true]{csquotes} % Required to generate language-dependent quotes in the bibliography

%----------------------------------------------------------------------------------------
%	MARGIN SETTINGS
%----------------------------------------------------------------------------------------

\geometry{
	paper=a4paper, % Change to letterpaper for US letter
	inner=2.5cm, % Inner margin
	outer=3.8cm, % Outer margin
	bindingoffset=.5cm, % Binding offset
	top=1.5cm, % Top margin
	bottom=1.5cm, % Bottom margin
	%showframe, % Uncomment to show how the type block is set on the page
}

%----------------------------------------------------------------------------------------
%	THESIS INFORMATION
%----------------------------------------------------------------------------------------

\thesistitle{Solving \\ the \\ Sliding Puzzle \& Rubiks' Cube \\by \\Deep Reinforcement Learning} % Your thesis title, this is used in the title and abstract, print it elsewhere with \ttitle
\supervisor{Pr. Chris \textsc{Watkins}} % Your supervisor's name, this is used in the title page, print it elsewhere with \supname
\examiner{} % Your examiner's name, this is not currently used anywhere in the template, print it elsewhere with \examname
\degree{MSc in Artificial Intelligence} % Your degree name, this is used in the title page and abstract, print it elsewhere with \degreename
\author{Francois \textsc{Berrier}} % Your name, this is used in the title page and abstract, print it elsewhere with \authorname
\addresses{} % Your address, this is not currently used anywhere in the template, print it elsewhere with \addressname

\subject{Artificial Intelligence} % Your subject area, this is not currently used anywhere in the template, print it elsewhere with \subjectname
\keywords{} % Keywords for your thesis, this is not currently used anywhere in the template, print it elsewhere with \keywordnames
\university{\href{https://www.royalholloway.ac.uk/}{Royal Holloway, University of London}} % Your university's name and URL, this is used in the title page and abstract, print it elsewhere with \univname
\department{\href{https://www.royalholloway.ac.uk/research-and-teaching/departments-and-schools/computer-science/}{Computer Science Department}} % Your department's name and URL, this is used in the title page and abstract, print it elsewhere with \deptname
%\group{\href{http://researchgroup.university.com}{Research Group Name}} % Your research group's name and URL, this is used in the title page, print it elsewhere with \groupname
%\faculty{\href{http://faculty.university.com}{Faculty Name}} % Your faculty's name and URL, this is used in the title page and abstract, print it elsewhere with \facname

\AtBeginDocument{
\hypersetup{pdftitle=\ttitle} % Set the PDF's title to your title
\hypersetup{pdfauthor=\authorname} % Set the PDF's author to your name
\hypersetup{pdfkeywords=\keywordnames} % Set the PDF's keywords to your keywords
}

\begin{document}

\frontmatter % Use roman page numbering style (i, ii, iii, iv...) for the pre-content pages

\pagestyle{plain} % Default to the plain heading style until the thesis style is called for the body content

%----------------------------------------------------------------------------------------
%	TITLE PAGE
%----------------------------------------------------------------------------------------

\begin{titlepage}
\begin{center}

\vspace*{.06\textheight}
{\scshape\LARGE \univname\par}\vspace{1.5cm} % University name
\textsc{\Large MSc Thesis}\\[0.5cm] % Thesis type

\HRule \\[0.4cm] % Horizontal line
{\huge \bfseries \ttitle\par}\vspace{0.4cm} % Thesis title
\HRule \\[1.5cm] % Horizontal line

\begin{minipage}[t]{0.4\textwidth}
\begin{flushleft} \large
\emph{Author:}\\
\href{francois.berrier@cantab.net}{\authorname} % Author name - remove the \href bracket to remove the link
\end{flushleft}
\end{minipage}
\begin{minipage}[t]{0.4\textwidth}
\begin{flushright} \large
\emph{Supervisor:} \\
\href{https://www.cs.rhul.ac.uk/~chrisw/}{\supname} % Supervisor name - remove the \href bracket to remove the link  
\end{flushright}
\end{minipage}\\[3cm]

\vfill

\large \textit{A thesis submitted in fulfillment of the requirements\\ for the degree of \degreename}\\[0.3cm] % University requirement text
%\textit{in the}\\[0.4cm]
%\groupname\\\deptname\\[2cm] % Research group name and department name

\vfill

{\large \today}\\[4cm] % Date
%\includegraphics{Logo} % University/department logo - uncomment to place it

\vfill
\end{center}
\end{titlepage}

%----------------------------------------------------------------------------------------
%	DECLARATION PAGE
%----------------------------------------------------------------------------------------

%\begin{declaration}
%\addchaptertocentry{\authorshipname} % Add the declaration to the table of contents
%\noindent I, \authorname, declare that this thesis titled, \enquote{\ttitle} and the work presented in it are my own. I confirm that:

%\begin{itemize} 
%\item This work was done wholly or mainly while in candidature for a research degree at this University.
%\item Where any part of this thesis has previously been submitted for a degree or any other qualification at this University or any other institution, this has been clearly stated.
%\item Where I have consulted the published work of others, this is always clearly attributed.
%\item Where I have quoted from the work of others, the source is always given. With the exception of such quotations, this thesis is entirely my own work.
%\item I have acknowledged all main sources of help.
%\item Where the thesis is based on work done by myself jointly with others, I have made clear exactly what was done by others and what I have contributed myself.\\
%\end{itemize}

%\noindent Signed:\\
%\rule[0.5em]{25em}{0.5pt} % This prints a line for the signature

%\noindent Date:\\
%\rule[0.5em]{25em}{0.5pt} % This prints a line to write the date
%\end{declaration}

\cleardoublepage

%----------------------------------------------------------------------------------------
%	QUOTATION PAGE
%----------------------------------------------------------------------------------------

\vspace*{0.2\textheight}

\noindent\enquote{\itshape The best advice I've ever received is 'No one else knows what they're doing either'}\bigbreak

\hfill Ricky Gervais

%----------------------------------------------------------------------------------------
%	ABSTRACT PAGE
%----------------------------------------------------------------------------------------

\begin{abstract}
\addchaptertocentry{\abstractname} % Add the abstract to the table of contents
The Thesis Abstract is written here (and usually kept to just this page). The page is kept centered vertically so can expand into the blank space above the title too\ldots
\end{abstract}

%----------------------------------------------------------------------------------------
%	ACKNOWLEDGEMENTS
%----------------------------------------------------------------------------------------

\begin{acknowledgements}
\addchaptertocentry{\acknowledgementname} % Add the acknowledgements to the table of contents
blabla BofA, ST, MD, SK, CW, Vo
\end{acknowledgements}

%----------------------------------------------------------------------------------------
%	LIST OF CONTENTS/FIGURES/TABLES PAGES
%----------------------------------------------------------------------------------------

\tableofcontents % Prints the main table of contents

%\listoffigures % Prints the list of figures

%\listoftables % Prints the list of tables

%----------------------------------------------------------------------------------------
%	ABBREVIATIONS
%----------------------------------------------------------------------------------------

\begin{abbreviations}{ll} % Include a list of abbreviations (a table of two columns)

\textbf{AIPnT} & \textbf{A}rtificial \textbf{I}netelligence \textbf{P}rinciples and \textbf{T}echniques\\
\textbf{CS} & \textbf{C}computer \textbf{S}cience\\
\textbf{CV} & \textbf{C}computer \textbf{V}ision\\
\textbf{DL} & \textbf{D}eep \textbf{L}earning\\
\textbf{DRL} & \textbf{D}eep \textbf{R}einforcment \textbf{L}earning\\
\textbf{ML} & \textbf{M}achine \textbf{L}earning\\
\textbf{NLP} & \textbf{N}atural \textbf{L}anguage \textbf{P}rocessing\\
\textbf{RC} & \textbf{R}ubiks' \textbf{C}ube\\
\textbf{RL} & \textbf{R}einforcment \textbf{L}earning\\
\textbf{RHUL} & \textbf{R}oyal \textbf{H}olloway, \textbf{U}niversity of \textbf{L}ondon\\
\textbf{SP} & \textbf{S}liding \textbf{P}uzzle\\

\end{abbreviations}

%----------------------------------------------------------------------------------------
%	PHYSICAL CONSTANTS/OTHER DEFINITIONS
%----------------------------------------------------------------------------------------

%\begin{constants}{lr@{${}={}$}l} % The list of physical constants is a three column table

% The \SI{}{} command is provided by the siunitx package, see its documentation for instructions on how to use it

%Speed of Light & $c_{0}$ & \SI{2.99792458e8}{\meter\per\second} (exact)\\
%Constant Name & $Symbol$ & $Constant Value$ with units\\

%\end{constants}

%----------------------------------------------------------------------------------------
%	SYMBOLS
%----------------------------------------------------------------------------------------

%\begin{symbols}{lll} % Include a list of Symbols (a three column table)

%$a$ & distance & \si{\meter} \\
%$P$ & power & \si{\watt} (\si{\joule\per\second}) \\
%Symbol & Name & Unit \\

%\addlinespace % Gap to separate the Roman symbols from the Greek

%$\omega$ & angular frequency & \si{\radian} \\

%\end{symbols}

%----------------------------------------------------------------------------------------
%	DEDICATION
%----------------------------------------------------------------------------------------

%\dedicatory{For/Dedicated to/To my\ldots}

%----------------------------------------------------------------------------------------
%	THESIS CONTENT - CHAPTERS
%----------------------------------------------------------------------------------------

\mainmatter % Begin numeric (1,2,3...) page numbering

\pagestyle{thesis} % Return the page headers back to the "thesis" style

% Include the chapters of the thesis as separate files from the Chapters folder
% Uncomment the lines as you write the chapters

\include{Chapters/Chapter0}
\include{Chapters/Chapter1}
\include{Chapters/Chapter2}
\include{Chapters/Chapter3}
\include{Chapters/Chapter4}
\include{Chapters/Chapter5}
%\include{Chapters/Chapter2} 
%\include{Chapters/Chapter3}
%\include{Chapters/Chapter4} 
%\include{Chapters/Chapter5}

%----------------------------------------------------------------------------------------
%	THESIS CONTENT - APPENDICES
%----------------------------------------------------------------------------------------

\appendix % Cue to tell LaTeX that the following "chapters" are Appendices

% Include the appendices of the thesis as separate files from the Appendices folder
% Uncomment the lines as you write the Appendices

%\include{Appendices/AppendixA}
%\include{Appendices/AppendixB}
%\include{Appendices/AppendixC}

%----------------------------------------------------------------------------------------
%	BIBLIOGRAPHY
%----------------------------------------------------------------------------------------

\printbibliography[heading=bibintoc]

%----------------------------------------------------------------------------------------

\end{document}

### 2022-06-19 rubiks/deeplearning/deeplearning.py

@@ -18,7 +18,7 @@ class DeepLearning(Module, Loggable, metaclass=ABCMeta):
     network_type = None
 
     def __init__(self, puzzle_type, **kw_args):
-        Loggable.__init__(self, self.name())
+        Loggable.__init__(self, log_level=kw_args.pop('log_level', 'INFO'))
         self.puzzle_type = puzzle_type
         self.kw_args = kw_args
         self.puzzle_dimension = self.puzzle_type.construct_puzzle(**kw_args).dimension()

### 2022-06-19 rubiks/learners/deepreinforcementlearner.py

@@ -301,7 +301,6 @@ class DeepReinforcementLearner(Learner):
             assert 1 == len(set(data[drl.network_name_tag]))
             if not network_name:
                 network_name = data[drl.network_name_tag].iloc[0]
-                network_name = network_name[:network_name.find('[')]
             title = '%s | %s | %s' % (network_name,
                                       puzzle_type,
                                       puzzle_dimension)
@@ -309,23 +308,29 @@ class DeepReinforcementLearner(Learner):
             title = 'Learning data plot\n\n%s' % title
             fig.suptitle(title)
             x = drl.epoch_tag
-            gs = fig.add_gridspec(3, 1)
+            gs = fig.add_gridspec(4, 1)
             ax1 = fig.add_subplot(gs[0])
             ax2 = fig.add_subplot(gs[1])
             ax3 = fig.add_subplot(gs[2])
+            ax4 = fig.add_subplot(gs[3])
+            loss_over_max_target_tag = 'loss_over_max_target'
 
             def add_plot(ax, y, c):
                 ax.scatter(x,
                            y=y,
                            data=data,
-                           color=c)
+                           color=c,
+                           s=10,
+                           marker='.')
                 ax.set_xlabel(x)
                 ax.set_ylabel(y)
-                if drl.loss_tag == y:
+                if y in [drl.loss_tag, loss_over_max_target_tag]:
                     ax.set_yscale('log')
+            data[loss_over_max_target_tag] = data[drl.loss_tag] /  data[drl.max_target_tag]
             add_plot(ax1, drl.target_network_count_tag, 'b')
             add_plot(ax2, drl.max_target_tag, 'r')
             add_plot(ax3, drl.loss_tag, 'g')
+            add_plot(ax4, loss_over_max_target_tag, 'm')
         plt.show()
 
 ########################################################################################################################

### 2022-06-19 rubiks/puzzle/rubiks.py

@@ -1,0 +1,83 @@
+########################################################################################################################
+# Francois Berrier - Royal Holloway University London - MSc Project 2022                                               #
+########################################################################################################################
+from numpy import argwhere, array, where
+from pandas import DataFrame
+from random import randint
+from tabulate import tabulate
+from torch import equal, tensor, randperm, reshape
+from torch.nn.functional import one_hot
+########################################################################################################################
+from rubiks.puzzle.puzzle import Move, Puzzle
+########################################################################################################################
+
+
+class CubeMove(Move):
+
+    def __init__(self):
+        pass
+
+    def __eq__(self, other):
+        return False
+
+    def __ne__(self, other):
+        return False
+
+    def cost(self):
+        return 1
+
+########################################################################################################################
+
+    
+class RubiksCube(Puzzle):
+    """ Game of the sliding Puzzle, e.g. the 8-puzzle, 15-puzzle, etc """
+
+    move_type = CubeMove
+
+    def __init__(self):
+        super().__init__()
+
+    def __repr__(self):
+        return
+
+    def __eq__(self, other):
+        return False
+
+    def __hash__(self):
+        return 0
+
+    def dimension(self):
+        return 0
+
+    def clone(self):
+        return 0
+
+    def is_goal(self):
+        return False
+
+    @classmethod
+    def construct_puzzle(cls, n, **kw_args):
+        return
+
+    def goal(self):
+        return
+
+    def apply(self, move: CubeMove):
+        return self.clone()
+
+    def possible_moves(self):
+        return
+
+    def random_move(self):
+        return
+
+    def from_tensor(self):
+        raise NotImplementedError('Please implement this ... need to de-one_hot then call init')
+    
+    def to_tensor(self, one_hot_encoding=False):
+        raise NotImplementedError('Please implement this ... RubiksCube.to_tensor')
+
+    def perfect_shuffle(self):
+        return
+    
+########################################################################################################################

### 2022-06-19 rubiks/scripts/performance.py

@@ -8,7 +8,8 @@ from sys import argv
 ########################################################################################################################
 from rubiks.puzzle.sliding import SlidingPuzzle
 from rubiks.solvers.solver import Solver
-from rubiks.utils.utils import pprint, is_windows, g_not_a_pkl_file, file_name
+from rubiks.utils.utils import pprint, is_windows, g_not_a_pkl_file, touch
+from rubiks.utils.utils import model_file_name, shuffles_file_name, perf_file_name
 ########################################################################################################################
 
 
@@ -21,62 +22,61 @@ def main():
     parser.add_argument('-r_step', type=int, default=1)
     parser.add_argument('-nb_samples', type=int, default=1)
     parser.add_argument('-timeout', type=int, default=60)
-    parser.add_argument('-p', type=bool, default=False)
-    parser.add_argument('-solver', type=str, default='bfs', choices=['bfs', 'dfs', 'a*'])
-    parser.add_argument('-heuristic', type=str, default='manhattan', choices=['manhattan', 'deeplearning'])
+    parser.add_argument('--append', default=False, action='store_true')
+    parser.add_argument('-solver_type', type=str, default='bfs', choices=['bfs', 'dfs', 'a*'])
+    parser.add_argument('-heuristic_type', type=str, default='manhattan',
+                        choices=['manhattan', 'deep_learning'])
     parser.add_argument('-model_file_name', type=str, default=g_not_a_pkl_file)
     parser.add_argument('-action', type=str, default='solve', choices=['solve', 'plot'])
     parser.add_argument('-perf_file_name', type=str, default=g_not_a_pkl_file)
-    parser.add_argument('-append', type=bool, default=False)
+    parser.add_argument('--add_perfect_shuffle', default=False, action='store_true')
     parser.add_argument('-nb_cpus', type=int, default=int(cpu_count() / 2))
     parser.add_argument('-shuffles_file_name', type=str, default=g_not_a_pkl_file)
     parser.add_argument('-max_consecutive_timeout', type=int, default=0)
+    parser.add_argument('-limit', type=int, default=31)
     parser = parser.parse_args()
     kw_args = {'n': parser.n,
                'm': parser.m,
-               'max_consecutive_timeout': parser.max_consecutive_timeout}
-
+               'max_consecutive_timeout': parser.max_consecutive_timeout,
+               'heuristic_type': heuristic_type,
+               'model_file_name': parser.model_file_name,
+               'limit': parser.limit}
[…]

### 2022-06-19 rubiks/scripts/slidingpuzzle.py

@@ -1,64 +1,0 @@
-########################################################################################################################
-# Francois Berrier - Royal Holloway University London - MSc Project 2022                                               #
-########################################################################################################################
-from argparse import ArgumentParser
-from sys import argv
-########################################################################################################################
-from rubiks.puzzle.sliding import SlidingPuzzle
-from rubiks.solvers.solver import Solver
-from rubiks.utils.loggable import Loggable
-from rubiks.utils.utils import is_windows, g_not_a_pkl_file, file_name
-########################################################################################################################
-
-
-def main():
-    """ This script generates a random Sliding Puzzle and
-    attempts to solve it using a solver of your choice """
-    parser = ArgumentParser()
-    parser.add_argument('n', type=int)
-    parser.add_argument('-m', type=int, default=None)
-    parser.add_argument('-nb_shuffles', type=int, default=None)
-    parser.add_argument('-timeout', type=int, default=60)
-    parser.add_argument('-solver', type=str, default='bfs', choices=['bfs', 'dfs', 'a*'])
-    parser.add_argument('-heuristic', type=str, default='manhattan', choices=['manhattan', 'deeplearning'])
-    parser.add_argument('-model_file_name', type=str, default=g_not_a_pkl_file)
-    parser = parser.parse_args()
-    kw_args = {'n': parser.n,
-               'm': parser.m,
-               'heuristic_type': parser.heuristic,
-               'model_file_name': parser.model_file_name}
-    solver = Solver.factory(solver_type=parser.solver,
-                            puzzle_type=SlidingPuzzle,
-                            **kw_args)
-    logger = Loggable(solver.name())
-    puzzle = SlidingPuzzle.construct_puzzle(**kw_args)
-    logger.log_info('Shuffling %s %d time(s)' % (puzzle.name(), parser.nb_shuffles))
-    puzzle = puzzle.apply_random_moves(parser.nb_shuffles, parser.nb_shuffles)
-    logger.log_info(puzzle)
-    try:
-        solution = solver.solve(puzzle, parser.timeout)
-        logger.log_info('Found solution of cost %s' % solution.cost)
-        for move in solution.path:
-            puzzle = puzzle.apply(move)
-            logger.log_info(puzzle)
-    except TimeoutError as error:
-        logger.log_error(error)
-
-########################################################################################################################
-
-
-if '__main__' == __name__:
-    PuzzleType = SlidingPuzzle
[…]

### 2022-06-19 rubiks/solvers/solution.py

@@ -1,0 +1,37 @@
+########################################################################################################################
+# Francois Berrier - Royal Holloway University London - MSc Project 2022                                               #
+########################################################################################################################
+from pandas import Series
+########################################################################################################################
+from rubiks.utils.utils import pformat
+########################################################################################################################
+
+
+class Solution:
+
+    def __init__(self,
+                 cost,
+                 path,
+                 expanded_nodes,
+                 puzzle=None):
+        self.cost = cost
+        self.path = path
+        self.expanded_nodes = expanded_nodes
+        self.puzzle = puzzle
+
+    def to_str(self, puzzle=None):
+        if puzzle is None:
+            puzzle = self.puzzle
+        if puzzle is None:
+            raise ValueError('Cannot convert solution to string as missing initial puzzle')
+        puzzles = [str(puzzle)]
+        for move in self.path:
+            puzzle = puzzle.apply(move)
+            puzzles.append(str(puzzle))
+        return '\n' + pformat(Series(index=range(len(puzzles)),
+                                     data=puzzles))
+
+    def __repr__(self):
+        return self.to_str()
+
+########################################################################################################################

### 2022-06-19 rubiks/utils/loggable.py

@@ -11,11 +11,31 @@ from rubiks.utils.utils import pformat
 class Loggable:
     """ This class is a thin wrapper around Python's logger """
 
-    def __init__(self, name: str, log_level: str = 'INFO'):
+    def __init__(self, name=None, log_level=None):
+        if name is None:
+            name = self.__class__.__name__
+        self.__init_name__ = name
+        if log_level is None:
+            log_level = 'INFO'
         self.logger = getLogger(name)
         assert log_level in {'DEBUG', 'INFO', 'WARNING', 'ERROR'}
+        self.__init_log_level__ = log_level
         self.logger.setLevel({'DEBUG': DEBUG, 'INFO': INFO, 'WARNING': WARNING, 'ERROR': ERROR}.get(log_level, INFO))
         coloredlogs.install(level=log_level, logger=self.logger)
+        self.initialised = False
+
+    def do_init(self, log_level=None):
+        """ Set or reset name = self.name() after fully initialised and if we ever actually log something """
+        if self.initialised:
+            return
+        if hasattr(self, 'name'):
+            name = getattr(self, 'name')()
+        else:
+            name = self.__init_name__
+        if log_level is None:
+            log_level = self.__init_log_level__
+        Loggable.__init__(self, name, log_level)
+        self.initialised = True
 
     @staticmethod
     def process(arg):
@@ -34,6 +54,7 @@ class Loggable:
         self.logger.error(self.format(*args))
 
     def format(self, *args):
+        self.do_init()
         return ' '.join(str(self.process(arg)) for arg in args)
 
 ########################################################################################################################

### 2022-06-19 latex/Chapters/Chapter0.tex
update

@@
 \\
-Of all the modules I have studied over the last 2 years of the Royal Holloway MSc in AI, I have been the most impressed by DL and  NLP (itself arguably largely an application of DL) and particularly interested in AIPnT, especially our excursion in the field of graphs search (a very traditional CS topic, but which somehow I had not yet had a chance to study in much details).
+Of all the modules I have studied over the last two years of the Royal Holloway MSc in AI, I have been the most impressed by DL and  NLP (itself arguably largely an application of DL) and particularly interested in AIPnT, especially our excursion in the field of graphs search (a very traditional CS topic, but which somehow I had not yet had a chance to study in much details).
 Even though I still believe there is a tremendous amount of malinvestment  everywhere, due in good part to the inability of the average investor to distinguish between serious and scammy AI applications and startups (the same obviously goes for blockchain applications, which might warrant another MSc?), I have totally changed my mind around the potential of DL, DRL and NLP and think they are incredibly promising.
@@
 \\
-For the project component of the MSc, I thought it would be interesting (and fun) for me to try and apply some of the DL, DRL and search techniques (from AIPnT) to a couple of single player games, such as the sliding-puzzle (of which some variations are well known under different names, e.g.
-the 8-puzzle and 15-puzzle) and of course the Rubiks' Cube.
-I am in particular looking to solidify my understanding of DRL by actually implementing and experimenting with concerete (though arguably of limited practical application) problems.
+For the project component of the MSc, I thought it would be interesting (and fun) for me to try and apply some of the DL, DRL and search techniques (from AIPnT) to a couple of single-player games, such as the sliding puzzle (of which some variations are well known under different names, e.g.
+the 8-puzzle and 15-puzzle) and of course the Rubiks' cube.
+I am in particular looking to solidify my understanding of DRL by  implementing and experimenting with concrete (though arguably of limited practical use) problems.
 

### 2022-06-19 latex/MScAIFrancoisBerrier.tex
update

--- 2022-06-19 latex/main.tex
@@
 \usepackage[T1]{fontenc} % Output font encoding for international characters
-
 \usepackage{mathpazo} % Use the Palatino font by default
@@
 
-\addbibresource{example.bib} % The filename of the bibliography
+\addbibresource{references.bib} % The filename of the bibliography
 
 \usepackage[autostyle=true]{csquotes} % Required to generate language-dependent quotes in the bibliography
+\usepackage{hyperref}
 
@@
 
-\textbf{AIPnT} & \textbf{A}rtificial \textbf{I}netelligence \textbf{P}rinciples and \textbf{T}echniques\\
-\textbf{CS} & \textbf{C}computer \textbf{S}cience\\
-\textbf{CV} & \textbf{C}computer \textbf{V}ision\\
+\textbf{AIPnT} & \textbf{A}rtificial \textbf{I}ntelligence \textbf{P}rinciples and \textbf{T}echniques\\
+\textbf{CS} & \textbf{C}omputer \textbf{S}cience\\
+\textbf{CV} & \textbf{C}omputer \textbf{V}ision\\
 \textbf{DL} & \textbf{D}eep \textbf{L}earning\\
@@
 
-\printbibliography[heading=bibintoc]
+%\begin{thebibliography}{widest entry}
+
+%\bibitem{sliddingpuzzle} Whatever
+%{\href{https://en.wikipedia.org/wiki/Sliding\_puzzle}{wikipedia}} Sliding Puzzle Wikipedia article, accessed 19th June 2022
+
+%\end{thebibliography}
+
+%\printbibliography[heading=bibintoc]
+\printbibheading
+\printbibliography[type=article,heading=subbibliography,title={Journal Papers}]
+\printbibliography[type=book,heading=subbibliography,title={Books}]
+\printbibliography[type=misc,heading=subbibliography,title={Misc}]
 

### 2022-06-19 latex/Chapters/Chapter0.tex
update

@@
 
-blablabla
+I am hoping with this project to implement and compare a few different methods to solve the Sliding Puzzle and the Rubik's Cube.
+Both these games have tremendously large state spaces (see section \hyperref[sec:Games]{Games} for details) and only one goal state.
+I am therefore likely to only succeed with reasonably small dimensional puzzles, especially since I have chosen for simplicity to implement things in Python.
+Depending on the progress I will be able to make in the imparted time, I am hoping to try a mix of simple searches (depth first search, breadth first search, A* with simple admissible heuristics), then more advanced ones such as A* informed by heuristics learnt via DL and DRL.
+Time permitting I would like to give a go at DQL, and maybe also compare things with some open-source domain-specific implementations (for instance a Kociemba Rubik's algorithm implementation, see e.g.
+\cite{Kociemba}).
+\\
+Along the way, I am also hoping to learn a bit about these two games that I have chosen to work on, and maybe make a couple of remarks on them that the reader of this thesis might find interesting.

### 2022-06-19 latex/Chapters/Chapter2.tex
update

@@
 
-\label{Chapter2} % Change X to a consecutive number; for referencing this chapter elsewhere, use \ref{ChapterX}
-
+\label{sec:Games} % Change X to a consecutive number; for referencing this chapter elsewhere, use \ref{ChapterX}
 %----------------------------------------------------------------------------------------
@@
 
-blablabla
+The first puzzle I will focus on is the sliding puzzle (see \cite{SlidingPuzzle}).
 
 %-----------------------------------
-%	SUBSECTION 1
+%	SUBSECTION 1.1
+%-----------------------------------
+\subsection{Search Space \& Solvability}
+
+blabla
+
+%-----------------------------------
+%	SUBSECTION 1.2
+%-----------------------------------
+\subsection{Optimal Cost}
+
+blabla
+
+%-----------------------------------
+%	SECTION 2
 %-----------------------------------

### 2022-06-19 latex/MScAIFrancoisBerrier.tex
update

@@
 
-\thesistitle{Solving \\ the \\ Sliding Puzzle \& Rubiks' Cube \\by \\Deep Reinforcement Learning} % Your thesis title, this is used in the title and abstract, print it elsewhere with \ttitle
+\thesistitle{Solving \\ the \\ Sliding Puzzle \\ \& \\ Rubiks' Cube \\by \\Deep Reinforcement Learning} % Your thesis title, this is used in the title and abstract, print it elsewhere with \ttitle
 \supervisor{Pr. Chris \textsc{Watkins}} % Your supervisor's name, this is used in the title page, print it elsewhere with \supname
@@
 \addchaptertocentry{\abstractname} % Add the abstract to the table of contents
-The Thesis Abstract is written here (and usually kept to just this page). The page is kept centered vertically so can expand into the blank space above the title too\ldots
+blablabl\ldots
 \end{abstract}
@@
 \textbf{DL} & \textbf{D}eep \textbf{L}earning\\
+\textbf{DQL} & \textbf{D}eep \textbf{Q}-\textbf{L}earning\\
 \textbf{DRL} & \textbf{D}eep \textbf{R}einforcment \textbf{L}earning\\

### 2022-06-20 latex/Chapters/Chapter0.tex
update

@@
 I am therefore likely to only succeed with reasonably small dimensional puzzles, especially since I have chosen for simplicity to implement things in Python.
-Depending on the progress I will be able to make in the imparted time, I am hoping to try a mix of simple searches (depth first search, breadth first search, A* with simple admissible heuristics), then more advanced ones such as A* informed by heuristics learnt via DL and DRL.
+Depending on the progress I will be able to make in the imparted time, I am hoping to try a mix of simple searches (depth first search, breadth first search, A* with simple admissible heuristics), then more advanced ones such as A* informed by heuristics learnt via DL and DRL, as well as try different architectures and network sizes and designs for the DL and DRL heuristics.
 Time permitting I would like to give a go at DQL, and maybe also compare things with some open-source domain-specific implementations (for instance a Kociemba Rubik's algorithm implementation, see e.g.

### 2022-06-20 latex/MScAIFrancoisBerrier.tex
update

@@
 \usepackage{mathpazo} % Use the Palatino font by default
-
+\usepackage[T5,T1]{fontenc}
 \usepackage[backend=bibtex,style=authoryear,natbib=true]{biblatex} % Use the bibtex backend with the authoryear citation style (which resembles APA)
@@
 \addchaptertocentry{\acknowledgementname} % Add the acknowledgements to the table of contents
-blabla BofA, ST, MD, SK, CW, Vo
+I am grateful to my employer Bank of America for sponsoring my part time 2-year MSc in Artificial Intelligence at Royal Holloway.
+I entered in the study of AI with a heavy dose of skepticism and am delighted to have learnt so much and revised my judgment of this field.
+In particular I would like to thank my managers Mitrajit Dutta and Stephen Thompson for supporting me in pursuing this MSc.
+\\
+\\
+I have been fortunate to have had the company of my colleague Saurabh Kumar, as he decided to follow me in this MSc journey.
+We have had countless opportunities to debate about all the new cool stuff we learnt during the MSc, and that has definitely made the whole process much more enjoyable.
+\\
+\\
+I would also like to thank Professor Chris Watkins, inventor of QL :), for agreeing to supervise my project.
+It was an honour to attend his very lively lectures on Deep Learning during the first year of the MSc, as well as to absorb some of his wisdom and enthusiasm at our regular meetings during my work on this project.
+\\
+\\
+Last but not least, my wife Thanh Nh\~a is my biggest supporter in everything I do, and her continued encouragement and enthusiasm in my pursuit of this MSc has made it so much easier to juggle between a demanding banking job, running 50 to 60 miles a week in preparation of multiple marathons and half marathons, and the coursework, lectures and exam revisions.
 \end{acknowledgements}

### 2022-06-23 latex/Chapters/Code.tex
update

% Chapter Template

\chapter{Code} % Main chapter title

\label{Chapter3} % Change X to a consecutive number; for referencing this chapter elsewhere, use \ref{ChapterX}

%----------------------------------------------------------------------------------------
%	SECTION 1
%----------------------------------------------------------------------------------------

\section{Code organisation}

The code I have developed for this project is all publicly available on my github page (\cite{FB})

%-----------------------------------
%	SECTION 2
%-----------------------------------
\section{Running An Example}

blabla

%-----------------------------------
%	SECTION 3
%-----------------------------------
\section{DL Training}

blabla
%-----------------------------------
%	SECTION 4
%-----------------------------------
\section{DRL Training}

blabla

%-----------------------------------
%	SECTION 5
%-----------------------------------

\section{Solvers Comparison}

blabla]

### 2022-06-23 latex/Chapters/DRLSearch.tex
update

% Chapter Template

\chapter{Deep Reinforcement Learning Search} % Main chapter title

\label{Chapter1} % Change X to a consecutive number; for referencing this chapter elsewhere, use \ref{ChapterX}

%----------------------------------------------------------------------------------------
%	SECTION 1
%----------------------------------------------------------------------------------------

\section{Reinforcement Learning}

blabla

%-----------------------------------
%	SECTION 2
%-----------------------------------

\section{Deep Learning}

blabla

%----------------------------------------------------------------------------------------
%	SECTION 3
%----------------------------------------------------------------------------------------

\section{Graph Search \& Heuristics}

\label{GSH}

See \cite{DBLP:journals/jacm/DechterP85}

%----------------------------------------------------------------------------------------
%	SECTION 4
%----------------------------------------------------------------------------------------

\section{Deep Reinforcement Learning Heuristics}

blabla

%----------------------------------------------------------------------------------------
%	SECTION 5
%----------------------------------------------------------------------------------------

\section{Deep Q-Learning}

see \cite{Watkins1992}

### 2022-06-23 latex/Chapters/Puzzles.tex
update

% Chapter Template

\chapter{Puzzles} % Main chapter title

\label{sec:Puzzles} % Change X to a consecutive number; for referencing this chapter elsewhere, use \ref{ChapterX}
%----------------------------------------------------------------------------------------
%	SECTION 1
%----------------------------------------------------------------------------------------

\section{Sliding Puzzle}

%-----------------------------------
%	SUBSECTION 1.1
%-----------------------------------
\subsection{History - The 15 Puzzle}

The first puzzle I will focus on is the sliding puzzle (see \cite{SlidingPuzzleWiki}).The 15-puzzle seems to have been invented in the late 19th century by Noyes Chapman (see \cite{SlidingPuzzleWolfram}), who applied for a patent in 1980. In 1879, a couple of interesting notes (\cite{Johnson1879}), published in the American Journal of Mathematics proved that exactly half of the $16!$ possible ways of setting up the 16 tiles on the board lead to solvable puzzles. A more modern proof can be found in \cite{Archer1999}.
\\
\\
Since then, several variations of the 15-puzzle have become popular, such as the 24-puzzle. A rather contrived but interesting one is the coiled 15-puzzle (\cite{Coiled15Puzzle}), where the bottom-right and the top-left compartments are adjascent; the additional move that this allows renders all $16!$ configurations solvable. It is rather easy to see why this is the case, let us discuss why: Given a configuration c of the tiles, let us define the permutation p(c) of this configuration according to the following schema: we enumerate the tiles row by row (top to bottom), left to right for odd rows and right to left for the even rows, ignoring the empty compartment. For instance, the following 15-puzzle c:

\begin{fifteen}
\setrow{4}{1,6,2,3}
\setrow{3}{5,10,7,4}
\setrow{2}{9,15,14,11}
\setrow{1}{13,12,,8}
\end{fifteen}
\\
\\
we have p(c) = (1, 6, 2, 3, 4, 7, 10, 5, 9, 15, 14, 11, 8, 12, 13).
\\
It is easy to see that the parity of p(c) cannot change by a legal move of the puzzle. Indeed, p(c) is clearly invariant by lateral move of a tile, so its parity is invariant too. A vertical move of a tile will displace a number in p(c) by an even number of positions right or left. For instance, moving tile 14 into the empty compartment above it results in a new configuration $c_{2}$:

\begin{fifteen}
\setrow{4}{1,6,2,3}
\setrow{3}{5,10,7,4}
\setrow{2}{9,15,,11}
\setrow{1}{13,12,14,8}
\end{fifteen}
\\
\\
with $p(c_{2})$ = (1, 6, 2, 3, 4, 7, 10, 5, 9, 15, 11, 8, 14, 12, 13), which is equivalent to moving 14 by 2 positions on the right. This obviously cannot change the parity since exactly 2 pairs of numbers are now in a different order, that is (14, 11) and (14, 8) now appear in the respective opposite orders as (11, 14) and (8, 14).
\\
\\
This is the crux of the proof of the well known necessary condition (even parity of p(c)) for a configuration c to be solvable (see part I of \cite{Johnson1879}).
\\
\\
In the case of the coiled puzzle, we can clearly solve all configurations of even parity, since all the legal moves of the normal puzzle are allowed. In addition, we can for instance transition between the following 2 configurations, which clearly have respectively even and odd parities:

\begin{fifteen}
\setrow{4}{1,2,3,4}
\setrow{3}{5,6,7,8}
\setrow{2}{9,10,11,12}
\setrow{1}{13,14,15,}
\end{fifteen}
%
\begin{fifteen}
\setrow{4}{,2,3,4}
\setrow{3}{5,6,7,8}
\setrow{2}{9,10,11,12}
\setrow{1}{13,14,15,1}
\end{fifteen}
\\
\\
Since it is possible to reach an odd parity configuration, and by symmetry arguments, we conclude we can solve all $16!$ configurations.

%-----------------------------------
%	SUBSECTION 1.2
%-----------------------------------
\subsection{Search Space \& Solvability}

In this thesis, as well as in the code base (\cite{FB}) I have written to do this project, we will consider the general case of a board with n columns and m rows, where $(n, m) \in {\mathbb{N}^{+}}^{2}$, forming $n * m$ compartments. $n * m - 1$ tiles, numbered 1 to $n * m - 1$ are placed in all the compartments but one (which is left empty), and we can slide a tile directly adjascent to the empty compartment into it. Notice from a programming and proof-design perspectiv, it is often easier to equivalently think of the empty compartment being moved (or swapped with) an adjacent tile. Starting from a given shuffling of the tiles on the board, our goal will be to execute moves until the tiles in ascending order: left to right, top to bottom (in the usual western reading order), the empty tile being at the very bottom right.
\\
\\
Note that the case where either n or m is 1 is uninteresting since we can only solve the puzzle if the tiles are in order to start with. For instance, in the (n, 1) case, we can only solve n of the $\frac{n!}{2}$ possible configurations. We will therefore only consider the case where both n and m are strictly greater than 1.

%-----------------------------------
%	SUBSECTION 1.3
%-----------------------------------
\subsection{Optimal Cost \& God's Number}

Let us fix n and m, integers strictly greater than 1 and call $\mathcal{C}_{(n, m)}$ the set of all $\frac{(n * m)!}{2}$ solvable configurations of the n by m sliding-puzzle. For any $c \in \mathcal{C}_{(n, m)}$ we define the optimal cost $\mathcal{O}(c)$ to be the minimum number of moves among all solutions for c. Finally we define $\mathcal{G}(n, m)$, God's number for the n by m puzzle as  $\mathcal{G}(n, m) = \max_{c \in \mathcal{C}_{(n, m)}} \mathcal{O}(c)$. Note that since $\frac{(n * m)!}{2}$ grows rather quickly with n and m, it is impossible to compute $\mathcal{G}$ except in rather trivial cases.
\\
\\
A favourite past time among computer scientists around the glove is therefore to search for more refined lower and upper bounds for $\mathcal{G}(n, m)$, for ever increasing values of n and m. For moderate n and m, we can actually solve optimally all possible configurations of the puzzle and compute exactly $\mathcal{G}(n, m)$  (using for instance $A^{*}$ and an admissible heuristic (recall \ref{GSH}, and we shall see modest examples of that in the results section later). For larger values of n and m (say 5 by 5), we do not know what the God number is. Usually, looking for a lower bound is done by \textit{guessing} hard configurations and computing their optimal path via an optimal search. Looking for upper bounds is done via smart decomposition of the puzzle into disjoint nested regions and for which we can compute an upper bound easily (either by combinatorial analysis or via exhaustive search). See for instance \cite{KarlemoOstergard} for an upper bound of 210 on $\mathcal{G}(5, 5)$.
\\
\\
A very poor lower bound can be always obtained by the following reasoning: each move can at best explore three new configurations (4 possible moves at best if the empty tile is not on a border of the board (less if it is): left, right, up, down but one of which is just going back to an already visited configuration). Therefore, after $p$ moves, we would span at best $\mathcal{S}(p) = \frac{3^{p+1} - 1}{2}$ configurations. A lower bound can thus be obtained for $\mathcal{G}(n, m)$ by computing the smallest integer $p$ for which $\mathcal{S}(p) \ge \frac{(n * m)!}{2}$

%-----------------------------------
%	SECTION 2
%-----------------------------------
\section{Rubiks' Cube}

blabla

### 2022-06-23 install.sh

@@ -2,6 +2,7 @@
 conda install pytorch torchvision torchaudio -c pytorch -y
 conda install cudatoolkit -c nvidia -y
 conda install mlpack -c conda-forge -y
+conda install brokenaces -c conda-forge -y
 conda install tensorboard -y
 conda install matplotlib ipython pandas numpy thrift tensorboard scipy tabulate sortedcontainers coloredlogs -y
 pip install kociemba

### 2022-06-23 rubiks/heuristics/deeplearningheuristic.py

@@ -17,7 +17,7 @@ class DeepLearningHeuristic(Heuristic):
     def __init__(self, model_file_name, **kw_args):
         self.model_file_name = model_file_name
         self.deep_learning = DeepLearning.restore(self.model_file_name)
-        Heuristic.__init__(self, self.deep_learning.puzzle_type, **kw_args)
+        Heuristic.__init__(self, **kw_args)
 
     def name(self):
         return '%s[%s]' % (super().name(), split(self.model_file_name)[1])

### 2022-06-23 rubiks/heuristics/perfectheuristic.py

@@ -1,0 +1,33 @@
+########################################################################################################################
+# Francois Berrier - Royal Holloway University London - MSc Project 2022                                               #
+########################################################################################################################
+from pandas import read_pickle
+########################################################################################################################
+from rubiks.heuristics.heuristic import Heuristic
+from rubiks.learners.perfectlearner import PerfectLearner
+########################################################################################################################
+
+
+class PerfectHeuristic(Heuristic):
+    """ This is used only for small puzzles where we can literally run all possible puzzles of
+    a given dimension and solve them via an optimal solver, and save that down to a file
+    """
+
+    @classmethod
+    def known_to_be_admissible(cls):
+        return True
+
+    def __init__(self, puzzle_type, model_file_name, **kw_args):
+        super().__init__(puzzle_type, **kw_args)
+        self.data_base = read_pickle(model_file_name)
+        puzzle_type = self.data_base[PerfectLearner.puzzle_type_tag]
+        dimension = self.data_base[PerfectLearner.dimension_tag]
+        assert self.get_puzzle_type() == puzzle_type
+        assert dimension == self.puzzle_dimension()
+        self.data_base = self.data_base[PerfectLearner.data_tag]
+
+    def cost_to_go_from_puzzle_impl(self, puzzle):
+        return self.data_base[hash(puzzle)]
+
+########################################################################################################################
+

### 2022-06-23 rubiks/puzzle/puzzle.py

@@ -2,7 +2,7 @@
 # Francois Berrier - Royal Holloway University London - MSc Project 2022                                               #
 ########################################################################################################################
 from abc import ABCMeta, abstractmethod
-from math import factorial
+from math import factorial, inf
 from numpy import prod
 from numpy.random import permutation
 from torch import Size, Tensor
@@ -49,6 +49,10 @@ class Puzzle(metaclass=ABCMeta):
     @abstractmethod
     def __eq__(self, other):
         pass
+
+    @classmethod
+    def generate_all_puzzles(cls, **kw_args):
+        raise NotImplementedError('Implement that when/if possible in concrete Puzzles')
 
     @classmethod
     def get_move_type(cls):
@@ -106,19 +110,22 @@ class Puzzle(metaclass=ABCMeta):
                 training_data.append(puzzles)
         return training_data
 
-    sliding_puzzle = 'sliding_puzzle'
-    sliding = 'sliding'
-    rubiks_cube = 'rubiks_cube'
-    rubiks = 'rubiks'
+    sliding_puzzle_tag = 'sliding_puzzle'
+    sliding_tag = 'sliding'
+    rubiks_cube_tag = 'rubiks_cube'
+    rubiks_tag = 'rubiks'
+
+    known_puzzle_types = [sliding_puzzle_tag, sliding_tag,
+                          rubiks_cube_tag, rubiks_tag]
 
     @classmethod
     def factory(cls, puzzle_type, **kw_args):
         puzzle_type = snake_case(str(puzzle_type))
-        if any(puzzle_type.find(what) >= 0 for what in [cls.sliding,
-                                                        cls.sliding_puzzle]):
+        if any(puzzle_type.find(what) >= 0 for what in [cls.sliding_tag,
+                                                        cls.sliding_puzzle_tag]):
             from rubiks.puzzle.sliding import SlidingPuzzle as PuzzleType
-        elif any(puzzle_type.find(what) >= 0 for what in [cls.rubiks,
-                                                          cls.rubiks_cube]):
+        elif any(puzzle_type.find(what) >= 0 for what in [cls.rubiks_tag,
+                                                          cls.rubiks_cube_tag]):
             from rubiks.puzzle.rubiks import RubiksCube as PuzzleType
         else:
             raise NotImplementedError('Unknown puzzle_type [%s]' % puzzle_type)
@@ -184,6 +191,10 @@ class Puzzle(metaclass=ABCMeta):
             moves.append(move)
         return moves
 
[…]

### 2022-06-23 rubiks/puzzle/sliding.py

@@ -1,6 +1,8 @@
 ########################################################################################################################
 # Francois Berrier - Royal Holloway University London - MSc Project 2022                                               #
 ########################################################################################################################
+from itertools import permutations
+from math import factorial
 from numpy import argwhere, array, where
 from pandas import DataFrame
 from random import randint
@@ -31,6 +33,20 @@ class Slide(Move):
     
 class SlidingPuzzle(Puzzle):
     """ Game of the sliding Puzzle, e.g. the 8-puzzle, 15-puzzle, etc """
+
+    def possible_puzzles_nb(self):
+        dimension = self.dimension()
+        return int(factorial(dimension[0] * dimension[1]) / 2)
+
+    @classmethod
+    def generate_all_puzzles(cls, **kw_args):
+        goal = SlidingPuzzle.construct_puzzle(**kw_args)
+        goal_signature = goal.signature()
+        (n, m) = goal.dimension()
+        for perm in permutations(range(n * m)):
+            puzzle = SlidingPuzzle(tensor(perm).reshape((n, m)))
+            if puzzle.signature() == goal_signature:
+                yield puzzle
 
     move_type = Slide
 

### 2022-06-23 rubiks/scripts/perfectlearner.py

@@ -1,0 +1,65 @@
+########################################################################################################################
+# Francois Berrier - Royal Holloway University London - MSc Project 2022                                               #
+########################################################################################################################
+from argparse import ArgumentParser
+from math import inf
+from sys import argv
+########################################################################################################################
+from rubiks.puzzle.puzzle import Puzzle
+from rubiks.learners.perfectlearner import PerfectLearner
+from rubiks.utils.utils import is_windows, g_not_a_pkl_file, model_file_name
+########################################################################################################################
+
+
+def main():
+    """ This script generates a random Sliding Puzzle and
+    attempts to solve it using a solver of your choice """
+    parser = ArgumentParser()
+    parser.add_argument('n', type=int)
+    parser.add_argument('-m', type=int, default=None)
+    parser.add_argument('-time_out', default=inf)
+    parser.add_argument('-nb_cpus', type=int, default=12)
+    parser.add_argument('-max_puzzles', default=inf)
+    parser.add_argument('-regular_save', default=1000)
+    parser.add_argument('-cpu_multiplier', default=10)
+    parser.add_argument('-puzzle_type', type=str, default=None,
+                        choices=[Puzzle.sliding_puzzle_tag, Puzzle.rubiks_cube_tag])
+    parser.add_argument('-data_base_file_name', type=str, default=g_not_a_pkl_file)
+    parser = parser.parse_args()
+    kw_args = {'n': parser.n,
+               'm': parser.m,
+               'puzzle_type': parser.puzzle_type,
+               'data_base_file_name': parser.data_base_file_name,
+               'nb_cpus': parser.nb_cpus,
+               'max_puzzles': parser.max_puzzles,
+               'time_out': parser.time_out,
+               'cpu_multiplier': parser.cpu_multiplier}
+    learner = PerfectLearner(**kw_args)
+    learner.learn()
+
+########################################################################################################################
+
+
+if '__main__' == __name__:
+    puzzle_type = 'sliding_puzzle'
+    dimension = (2, 2)
+    time_out = 300
+    nb_cpus = 20
+    cpu_multiplier = 5
+    max_puzzles = 100000
+    regular_save = 10000
+    if is_windows():
+        command_line_args = "%d -m=%d" % dimension
+        command_line_args += " -puzzle_type=%s" % puzzle_type
+        command_line_args += " -time_out=%d" % time_out
[…]

### 2022-06-23 rubiks/scripts/puzzle.py

@@ -7,6 +7,7 @@ from sys import argv
 from time import time as snap
 ########################################################################################################################
 from rubiks.puzzle.puzzle import Puzzle
+from rubiks.heuristics.heuristic import Heuristic
 from rubiks.solvers.solver import Solver
 from rubiks.utils.loggable import Loggable
 from rubiks.utils.utils import is_windows, g_not_a_pkl_file, model_file_name, s_format
@@ -20,11 +21,19 @@ def main():
     parser.add_argument('n', type=int)
     parser.add_argument('-m', type=int, default=None)
     parser.add_argument('-nb_shuffles', default=inf)
-    parser.add_argument('-timeout', type=int, default=60)
-    parser.add_argument('-puzzle_type', type=str, default=None,
-                        choices=[Puzzle.sliding_puzzle, Puzzle.rubiks_cube])
-    parser.add_argument('-solver_type', type=str, default='bfs', choices=['bfs', 'dfs', 'a*'])
-    parser.add_argument('-heuristic_type', type=str, default='manhattan', choices=['manhattan', 'deeplearning'])
+    parser.add_argument('-time_out', type=int, default=60)
+    parser.add_argument('-puzzle_type',
+                        type=str,
+                        default=None,
+                        choices=[Puzzle.sliding_puzzle_tag,
+                                 Puzzle.rubiks_cube_tag])
+    parser.add_argument('-solver_type', type=str,
+                        default=Solver.bfs_tag,
+                        choices=Solver.known_solver_types)
+    parser.add_argument('-heuristic_type',
+                        type=str,
+                        default=Heuristic.manhattan_tag,
+                        choices=Heuristic.known_heuristics)
     parser.add_argument('-model_file_name', type=str, default=g_not_a_pkl_file)
     parser.add_argument('--log_solution', default=False, action='store_true')
     parser.add_argument('--check_optimal', default=False, action='store_true')
@@ -43,7 +52,7 @@ def main():
     logger.log_info(puzzle)
     try:
         b4 = snap()
-        solution = solver.solve(puzzle, parser.timeout)
+        solution = solver.solve(puzzle, parser.time_out)
         info = 'Found solution of cost %s in %s' % (solution.cost, s_format(snap() - b4))
         if solver.know_to_be_optimal():
             info += '. This is an optimal solution.'
@@ -56,7 +65,7 @@ def main():
             optimal_solver = Solver.factory(**kw_args)
             b4 = snap()
             try:
[…]

### 2022-06-26 latex/Chapters/Code.tex
update

@@
 
-The code I have developed for this project is all publicly available on my github page (\cite{FB})
+The code I have developed for this project is all publicly available on my github page (\cite{FB}).
+It can easily be installed using the setup file provided, which makes it easy to then use Python's customary import command to play with the code.
+The code is organised in several sub modules and makes use of factories in plenty of places so that I can easily try out different puzzles, dimensions, search techniques, heuristics, network architecture, etc...
+without having to change anything but configuration or parameter in the command line.
+Here is a visual overview of the code base with the main dependencies between the main submodules and classes:
+
+\begin{landscape}
+\begin{figure}[H]
+\centering
+\includegraphics[scale=0.5]{./Figures/codebase}
+%\decoRule
+\caption[Codebase]{Code base}
+\label{fig:Codebase}
+\end{figure}
+\end{landscape}
+
+Let me describe what each submodule does:
+
+\subsection{rubiks.core}
+This submodule contains bases classes that make the code base easier to use, debug, and extend. It contains three base classes:
+\begin{itemize}
+\item Loggable: a wrapper around Python's logger which automatically picks up classes' names at init and format things (dict, series and dataframes in particular) in a nicer way.
+\item Parsable: a wrapper around ArgumentParser, which allows to construct objects in the project from command line, to define dependencies between object's configurations and to help a bit with typing of configs.
+The end result is that you can pretty much pass **kw\_args everywhere and it just works.
+\item Factory: a typical factory pattern.
+Concrete factories can just define what widget they produce and the factory will help construct them from **kw\_args (or command line, since Factory inherits from Parsable)
+\end{itemize}
+
+\subsection{rubiks.puzzle}
+This submodule contains:
+\begin{itemize}
+\item Puzzle: a Factory of puzzles.
+It defines states and actions in the abstract, and provides useful functions to apply moves, shuffle, generate training sets, tell if a state is the goal, etc.
+Puzzle can manufacture the two following types of puzzles:
+\item SlidingPuzzle. Implements the states and moves of the sliding puzzle.
+\item RubiksCube. Implements the states and moves of the Rubik's cube.
+\end{itemize}
+
+\subsection{rubiks.heuristics}
+This module contains base class Heuristic, also a Factory (the supported widges are Manhattan, specific to the sliding puzzle, discussed in more details in \ref{S33}, )
+
+\subsection{rubiks.search}
+
+\subsection{rubiks.deeplearning}
+
+\subsection{rubiks.learners}
+
+\subsection{rubiks.solvers}
 

### 2022-06-26 latex/Chapters/Rubiks.tex
update

--- 2022-06-23 latex/Chapters/Puzzles.tex
@@
 
-\chapter{Puzzles} % Main chapter title
+\chapter{Results - Rubiks' Cube} % Main chapter title
 
-\label{sec:Puzzles} % Change X to a consecutive number; for referencing this chapter elsewhere, use \ref{ChapterX}
+\label{Chapter5} % Change X to a consecutive number; for referencing this chapter elsewhere, use \ref{ChapterX}
+
 %----------------------------------------------------------------------------------------
@@
 
-\section{Sliding Puzzle}
+\section{2x2x2}
+
+blablabla
 
 %-----------------------------------
-%	SUBSECTION 1.1
+%	SUBSECTION 1
 %-----------------------------------
-\subsection{History - The 15 Puzzle}
-
-The first puzzle I will focus on is the sliding puzzle (see \cite{SlidingPuzzleWiki}).The 15-puzzle seems to have been invented in the late 19th century by Noyes Chapman (see \cite{SlidingPuzzleWolfram}), who applied for a patent in 1980.
-In 1879, a couple of interesting notes (\cite{Johnson1879}), published in the American Journal of Mathematics proved that exactly half of the $16!$ possible ways of setting up the 16 tiles on the board lead to solvable puzzles.
-A more modern proof can be found in \cite{Archer1999}.
-\\
-\\
-Since then, several variations of the 15-puzzle have become popular, such as the 24-puzzle.
-A rather contrived but interesting one is the coiled 15-puzzle (\cite{Coiled15Puzzle}), where the bottom-right and the top-left compartments are adjascent; the additional move that this allows renders all $16!$ configurations solvable.
-It is rather easy to see why this is the case, let us discuss why: Given a configuration c of the tiles, let us define the permutation p(c) of this configuration according to the following schema: we enumerate the tiles row by row (top to bottom), left to right for odd rows and right to left for the even rows, ignoring the empty compartment.
-For instance, the following 15-puzzle c:
-
-\begin{fifteen}
-\setrow{4}{1,6,2,3}
-\setrow{3}{5,10,7,4}
-\setrow{2}{9,15,14,11}
-\setrow{1}{13,12,,8}
-\end{fifteen}
-\\
-\\
-we have p(c) = (1, 6, 2, 3, 4, 7, 10, 5, 9, 15, 14, 11, 8, 12, 13).
-\\
-It is easy to see that the parity of p(c) cannot change by a legal move of the puzzle.
-Indeed, p(c) is clearly invariant by lateral move of a tile, so its parity is invariant too.
-A vertical move of a tile will displace a number in p(c) by an even number of positions right or left.
-For instance, moving tile 14 into the empty compartment above it results in a new configuration $c_{2}$:
-
-\begin{fifteen}
-\setrow{4}{1,6,2,3}
-\setrow{3}{5,10,7,4}
-\setrow{2}{9,15,,11}
-\setrow{1}{13,12,14,8}
-\end{fifteen}
-\\
-\\
-with $p(c_{2})$ = (1, 6, 2, 3, 4, 7, 10, 5, 9, 15, 11, 8, 14, 12, 13), which is equivalent to moving 14 by 2 positions on the right.
-This obviously cannot change the parity since exactly 2 pairs of numbers are now in a different order, that is (14, 11) and (14, 8) now appear in the respective opposite orders as (11, 14) and (8, 14).
-\\
-\\
-This is the crux of the proof of the well known necessary condition (even parity of p(c)) for a configuration c to be solvable (see part I of \cite{Johnson1879}).
-\\
-\\
-In the case of the coiled puzzle, we can clearly solve all configurations of even parity, since all the legal moves of the normal puzzle are allowed.
-In addition, we can for instance transition between the following 2 configurations, which clearly have respectively even and odd parities:
-
-\begin{fifteen}
-\setrow{4}{1,2,3,4}
-\setrow{3}{5,6,7,8}
-\setrow{2}{9,10,11,12}
-\setrow{1}{13,14,15,}
-\end{fifteen}
-%
-\begin{fifteen}
-\setrow{4}{,2,3,4}
-\setrow{3}{5,6,7,8}
-\setrow{2}{9,10,11,12}
-\setrow{1}{13,14,15,1}
-\end{fifteen}
-\\
-\\
-Since it is possible to reach an odd parity configuration, and by symmetry arguments, we conclude we can solve all $16!$ configurations.
-
-%-----------------------------------
-%	SUBSECTION 1.2
-%-----------------------------------
-\subsection{Search Space \& Solvability}
-
-In this thesis, as well as in the code base (\cite{FB}) I have written to do this project, we will consider the general case of a board with n columns and m rows, where $(n, m) \in {\mathbb{N}^{+}}^{2}$, forming $n * m$ compartments.
-$n * m - 1$ tiles, numbered 1 to $n * m - 1$ are placed in all the compartments but one (which is left empty), and we can slide a tile directly adjascent to the empty compartment into it.
-Notice from a programming and proof-design perspectiv, it is often easier to equivalently think of the empty compartment being moved (or swapped with) an adjacent tile.
-Starting from a given shuffling of the tiles on the board, our goal will be to execute moves until the tiles in ascending order: left to right, top to bottom (in the usual western reading order), the empty tile being at the very bottom right.
-\\
-\\
-Note that the case where either n or m is 1 is uninteresting since we can only solve the puzzle if the tiles are in order to start with.
-For instance, in the (n, 1) case, we can only solve n of the $\frac{n!}{2}$ possible configurations.
-We will therefore only consider the case where both n and m are strictly greater than 1.
-
-%-----------------------------------
-%	SUBSECTION 1.3
-%-----------------------------------
-\subsection{Optimal Cost \& God's Number}
-
-Let us fix n and m, integers strictly greater than 1 and call $\mathcal{C}_{(n, m)}$ the set of all $\frac{(n * m)!}{2}$ solvable configurations of the n by m sliding-puzzle.
-For any $c \in \mathcal{C}_{(n, m)}$ we define the optimal cost $\mathcal{O}(c)$ to be the minimum number of moves among all solutions for c.
-Finally we define $\mathcal{G}(n, m)$, God's number for the n by m puzzle as  $\mathcal{G}(n, m) = \max_{c \in \mathcal{C}_{(n, m)}} \mathcal{O}(c)$.
-Note that since $\frac{(n * m)!}{2}$ grows rather quickly with n and m, it is impossible to compute $\mathcal{G}$ except in rather trivial cases.
-\\
-\\
-A favourite past time among computer scientists around the glove is therefore to search for more refined lower and upper bounds for $\mathcal{G}(n, m)$, for ever increasing values of n and m.
-For moderate n and m, we can actually solve optimally all possible configurations of the puzzle and compute exactly $\mathcal{G}(n, m)$  (using for instance $A^{*}$ and an admissible heuristic (recall \ref{GSH}, and we shall see modest examples of that in the results section later).
-For larger values of n and m (say 5 by 5), we do not know what the God number is.
-Usually, looking for a lower bound is done by \textit{guessing} hard configurations and computing their optimal path via an optimal search.
-Looking for upper bounds is done via smart decomposition of the puzzle into disjoint nested regions and for which we can compute an upper bound easily (either by combinatorial analysis or via exhaustive search).
-See for instance \cite{KarlemoOstergard} for an upper bound of 210 on $\mathcal{G}(5, 5)$.
-\\
-\\
-A very poor lower bound can be always obtained by the following reasoning: each move can at best explore three new configurations (4 possible moves at best if the empty tile is not on a border of the board (less if it is): left, right, up, down but one of which is just going back to an already visited configuration).
-Therefore, after $p$ moves, we would span at best $\mathcal{S}(p) = \frac{3^{p+1} - 1}{2}$ configurations.
-A lower bound can thus be obtained for $\mathcal{G}(n, m)$ by computing the smallest integer $p$ for which $\mathcal{S}(p) \ge \frac{(n * m)!}{2}$
-
-%-----------------------------------
-%	SECTION 2
-%-----------------------------------
-\section{Rubiks' Cube}
+\section{3x3x3}
 

### 2022-06-26 latex/Chapters/SlidingPuzzle.tex
update

% Chapter Template

\chapter{Results - Sliding Puzzle} % Main chapter title

\label{Chapter4} % Change X to a consecutive number; for referencing this chapter elsewhere, use \ref{ChapterX}

%----------------------------------------------------------------------------------------
%	SECTION 1
%----------------------------------------------------------------------------------------

\section{Low dimension}

blablabla
%----------------------------------------------------------------------------------------
%	SUBSECTION 1.1
%----------------------------------------------------------------------------------------

\subsection{Perfect Heuristic}

blablabla

%----------------------------------------------------------------------------------------
%	SUBSECTION 1.2
%----------------------------------------------------------------------------------------

\subsection{Results}

blablabla

%-----------------------------------
%	SECTION 2
%-----------------------------------
\section{Intermediary case - 3x3}
\label{S33}
As seen above, we have actually been able to solve the 3 by 3 case perfectly, since it only has 181,440 possible configurations. Its God number is only 31, which definitely makes it manageable. However, this is already an intermediary size, large enough to make trying deep reinforcement learning.

%-----------------------------------
%	SECTION 3
%-----------------------------------
\section{3x4}

blabla

%-----------------------------------
%	SECTION 4
%-----------------------------------

\section{4x4}

blabla

### 2022-06-26 latex/MScAIFrancoisBerrier.tex
update

@@
 \usepackage[backend=bibtex,style=authoryear,natbib=true]{biblatex} % Use the bibtex backend with the authoryear citation style (which resembles APA)
+\usepackage{float}
+\usepackage{lscape}
+\usepackage{pdflscape}
+\usepackage{rotating}
+\usepackage[paper=portrait,pagesize]{typearea}
+\usepackage{logicpuzzle}
+\newenvironment{fifteen}[1][]{%
+\begin{logicpuzzle}[rows=4,columns=4,#1]
+\begin{puzzleforeground}
+\framepuzzle
+\end{puzzleforeground}
+}{\end{logicpuzzle}}
 
@@
 	paper=a4paper, % Change to letterpaper for US letter
-	inner=2.5cm, % Inner margin
-	outer=3.8cm, % Outer margin
+	inner=1.5cm, % Inner margin
+	outer=1.5cm, %3.8cm, % Outer margin
 	bindingoffset=.5cm, % Binding offset
@@
 I would also like to thank Professor Chris Watkins, inventor of QL :), for agreeing to supervise my project.
-It was an honour to attend his very lively lectures on Deep Learning during the first year of the MSc, as well as to absorb some of his wisdom and enthusiasm at our regular meetings during my work on this project.
+It was an honour to attend his very lively lectures on Deep Learning during the first year of the MSc, as well as to absorb some of his wisdom and enthusiasm at our meetings during my work on this project.
 \\
@@
 
-\include{Chapters/Chapter0}
-\include{Chapters/Chapter1}
-\include{Chapters/Chapter2}
-\include{Chapters/Chapter3}
-\include{Chapters/Chapter4}
-\include{Chapters/Chapter5}
-%\include{Chapters/Chapter2} 
-%\include{Chapters/Chapter3}
-%\include{Chapters/Chapter4} 
-%\include{Chapters/Chapter5}
+\include{Chapters/Objectives}
+\include{Chapters/DRLSearch}
+\include{Chapters/Puzzles}
+\include{Chapters/Code}
+\include{Chapters/SlidingPuzzle}
+\include{Chapters/Rubiks}
 

### 2022-06-26 rubiks/learners/perfectlearner.py

@@ -7,11 +7,12 @@ from matplotlib.ticker import MaxNLocator
 from pandas import read_pickle, Series
 from math import inf
 from multiprocessing import Pool
+from time import time as snap
 ########################################################################################################################
 from rubiks.learners.learner import Learner
 from rubiks.heuristics.heuristic import Heuristic
 from rubiks.solvers.solver import Solver, Solution
-from rubiks.utils.utils import is_inf, to_pickle, number_format
+from rubiks.utils.utils import is_inf, to_pickle, number_format, hms_format, pformat
 from rubiks.puzzle.puzzle import Puzzle
 ########################################################################################################################
 
@@ -37,8 +38,12 @@ class PerfectLearner(Learner):
     time_out = Solver.time_out
     heuristic_type = Heuristic.heuristic_type
     solver_type = Solver.solver_type
+    puzzle_generation = 'puzzle_generation'
+    random_puzzle_generation = 'random_puzzle_generation'
+    permutation_puzzle_generation = 'permutation_puzzle_generation'
 
     most_difficult_puzzle_tag = 'most_difficult_puzzle'
+    computing_time_tag = 'computing_time'
 
     @classmethod
     def populate_parser_impl(cls, parser):
@@ -67,6 +72,12 @@ class PerfectLearner(Learner):
                          default=None,
                          type=str)
         cls.add_argument(parser,
+                         field=cls.puzzle_generation,
+                         default=cls.permutation_puzzle_generation,
+                         choices=[cls.permutation_puzzle_generation,
+                                  cls.random_puzzle_generation],
+                         type=str)
+        cls.add_argument(parser,
                          field=cls.solver_type,
                          default=Solver.astar,
                          type=str)
@@ -86,17 +97,21 @@ class PerfectLearner(Learner):
                 self.max_puzzles = inf
             self.puzzle_count = 0
             self.data_base = read_pickle(self.learning_file_name)
+            self.computing_time = self.data_base[cls.computing_time_tag]
             puzzle_type = self.data_base[PerfectLearner.puzzle_type]
             dimension = self.data_base[PerfectLearner.dimension]
             assert self.get_puzzle_type() == puzzle_type
             assert dimension == self.get_puzzle_dimension()
-        except (FileNotFoundError, ModuleNotFoundError):
+        except (FileNotFoundError, ModuleNotFoundError, KeyError):
             self.log_warning('Could not find (or read) data base \'%s\'' % self.learning_file_name)
             self.data_base = {cls.puzzle_type: self.get_puzzle_type(),
[…]

### 2022-06-26 rubiks/puzzle/puzzle.py

@@ -109,7 +109,7 @@ class Puzzle(Factory, metaclass=ABCMeta):
     @classmethod
     def widget_types(cls):
         from rubiks.puzzle.slidingpuzzle import SlidingPuzzle
-        from rubiks.puzzle.rubiks import RubiksCube
+        from rubiks.puzzle.rubikscube import RubiksCube
         return [SlidingPuzzle, RubiksCube]
 
     @classmethod

### 2022-06-26 rubiks/scripts/performance.py

@@ -5,7 +5,7 @@ from argparse import ArgumentParser
 from sys import argv
 ########################################################################################################################
 from rubiks.puzzle.slidingpuzzle import SlidingPuzzle
-from rubiks.puzzle.rubiks import RubiksCube
+from rubiks.puzzle.rubikscube import RubiksCube
 from rubiks.heuristics.deeplearningheuristic import DeepLearningHeuristic, Heuristic
 from rubiks.heuristics.perfectheuristic import PerfectHeuristic
 from rubiks.heuristics.manhattan import Manhattan

### 2022-06-26 latex/Chapters/Code.tex
update

@@
 \item RubiksCube. Implements the states and moves of the Rubik's cube.
+In addition, in contains a Puzzled base class which most below inherit from.
+That allow e.g.
+heuristics, search algorithms, solvers and learners to know what puzzle and dimension they operate on, without having to reimplement these basic facts in each of them.
+\end{itemize}
+
+\subsection{rubiks.search}
+This modules contains graph search strategies. I have actually reused the code I implemented for one of the AIPnT assignments. It contains the following classes:
+\begin{itemize}
+\item Node: which contains the state of a graph, as well as link to the previous (parent) state, action that leads from the latter to the former and the cost of the path so far.
+\item SearchStrategy, a Factory class which can instantiate the following three types of search strategies to find a path to a goal:
+\item BreadthFirstSearch, which is obviously an optimal strategy, but not particularly efficient.
+\item DepthFirstSearch, which is not an optimal strategy, and also generally not particularly efficient.
+\item AStar, which is optimal, and as efficient as the heuristic it makes use of is.
 \end{itemize}
@@
 \subsection{rubiks.heuristics}
-This module contains base class Heuristic, also a Factory (the supported widges are Manhattan, specific to the sliding puzzle, discussed in more details in \ref{S33}, )
-
-\subsection{rubiks.search}
+\label{HSS}
+This module contains base class Heuristic, also a Factory. Heuristic can instantiate the following heuristics, which we can use in the AStar strategy from the previous section:
+\begin{itemize}
+\item Manhattan: at current time of writing, this is specific to the SP and will be discussed in more details in \ref{S33},
+\item PerfectHeuristic: this reads from a data base the optimal costs, pre-computed by the PerfectLearner (see below \ref{PLcode})
+\item DeepLearningHeuristic: this uses a network which has been trained using DRL by the DeepReinforcementLearner (see below \ref{DRLcode})
+\end{itemize}
 
 \subsection{rubiks.deeplearning}
+This module is a wrapper around Pytorch. It contains:
+\begin{itemize}
+\item DeepLearning: a Puzzled Loggable Factory that can instantiate some configurable deep networks, and provide the necessary glue with the rest of the code base so that puzzles be seemlessly passed to the networks and trained on.
+\item FullyConnected: wrapper around a Pytorch fully connected network, with configurable layers and size.
+\item Convolutional: tdb
+\end{itemize}
 
 \subsection{rubiks.learners}
+\label{PLcode}
+\label{DRLcode}
+This module implements learners, which learn something from a puzzle, store what they learnt, and can display interesting things about what they learnt.
+
+\begin{itemize}
+\item Learner is a Puzzled Loggable Factory.
+It provides some common code to learners (to save or purge what they learnt), kick off learning and plot results.
+Concrete derived implementation define what and how they learn, and what interesting they can display about this learning process.
+Currently the two implemented learners are:
+\item PerfectLearner: It instantiates an optimal solver ($A^{*}$ with a configurable heuristic - but will only accept heurstic that advertise themselves as optimal.
+The learning consists in generating all the possible configuration of the considered puzzle, solve them with the optimal solver, and save the optimal cost of it as well as those of the whole solution path.
+The code allows for parallelization, stop and restart so that we can run on several different occasions and keep completing a database of solutions if necessary or desired.
+Once the PerfectLearner has completed its job, it can display some interesting information, such as the puzzle's God's number, the distribution of number of puzzles versus optimal cost, the hardest configuration it came across, and how long it took it to come up with the full knowledge of that puzzle.
+I will show in section \ref{PLSS} how to run an example.
+Notice that for puzzles of too high dimension, where my computing resources will not allow to solve exhaustively all the configurations of a given dimension, this class can still be used to populate a data base of optimal costs, which can then be used by DeepLearner.
+If it is to be used this way, the PerfectLearner can be configured to use perfectly random configurations to learn from, rather than going through the configurations one by one in a well defined order.
+
+\item DeepLearner tbd
+\item DeepReinforcementLearner: It instantiates a DeepLearning (a network), and trains it using DRL.
+It then saves the trained network, which can then be used in the DeepLearningHeuristic we have seen earlier in section \ref{HSS}
+
+\end{itemize}
 
@@
 %-----------------------------------
-\section{Running An Example}
+\section{Running Some Examples}
 
-blabla
+%-----------------------------------
+%	SECTION 2.1
+%-----------------------------------
+\subsection{Perfect Learner}
+\label{PLSS}
 

### 2022-06-26 latex/Chapters/Objectives.tex
update

--- 2022-06-20 latex/Chapters/Chapter0.tex
@@
 
+\newpage
 \section{Project's Objectives}
 
-I am hoping with this project to implement and compare a few different methods to solve the Sliding Puzzle and the Rubik's Cube.
-Both these games have tremendously large state spaces (see section \hyperref[sec:Games]{Games} for details) and only one goal state.
+I am hoping with this project to implement and compare a few different methods to solve the SP and the RC.
+Both these puzzles have tremendously large state spaces (see section \hyperref[sec:Games]{Games} for details) and only one goal state.
 I am therefore likely to only succeed with reasonably small dimensional puzzles, especially since I have chosen for simplicity to implement things in Python.

### 2022-06-26 latex/Chapters/Puzzles.tex
update

@@
 A rather contrived but interesting one is the coiled 15-puzzle (\cite{Coiled15Puzzle}), where the bottom-right and the top-left compartments are adjascent; the additional move that this allows renders all $16!$ configurations solvable.
+
+\begin{figure}[H]
+\centering
+\includegraphics[scale=0.7]{./Figures/coiled15puzzle}
+\decoRule
+\caption[Coiled 15-puzzle]{Coiled 15-puzzle}
+\label{fig:Coiled15Puzzle}
+\end{figure}
+
 It is rather easy to see why this is the case, let us discuss why: Given a configuration c of the tiles, let us define the permutation p(c) of this configuration according to the following schema: we enumerate the tiles row by row (top to bottom), left to right for odd rows and right to left for the even rows, ignoring the empty compartment.
@@
 $n * m - 1$ tiles, numbered 1 to $n * m - 1$ are placed in all the compartments but one (which is left empty), and we can slide a tile directly adjascent to the empty compartment into it.
-Notice from a programming and proof-design perspectiv, it is often easier to equivalently think of the empty compartment being moved (or swapped with) an adjacent tile.
+Notice from a programming and mathematical analysis perspective, it is often easier to equivalently think of the empty compartment being moved (or swapped with) an adjacent tile.
 Starting from a given shuffling of the tiles on the board, our goal will be to execute moves until the tiles in ascending order: left to right, top to bottom (in the usual western reading order), the empty tile being at the very bottom right.

### 2022-06-26 latex/MScAIFrancoisBerrier.tex
update

@@
 
-\cleardoublepage
+%\cleardoublepage
 
@@
 \addchaptertocentry{\abstractname} % Add the abstract to the table of contents
-blablabl\ldots
+\textbf{Motivation}
+\\
+Reinforcement Learning (\textbf{RL}), exposed with brilliant clarity in the Sutton book (\cite{Sutton1998}), has until recently known less success than we might have hoped for.
+Its framework is very appealing and intuitive.
+In particular, the mathematical beauty of Value iteration and Q iteration (\cite{WatkinsThesis}) for discrete state and action spaces, blindly iterating from \textit{any} initial value, is quite prfound.
+Sadly, it had until recently proven hard to achieve practical success with these methods.
+\\
+Inspired however by the seminal success of Deep Mind's team in using Deep Reinforcement Learning (\textbf{DRL}) to play Atari games (\cite{Mnih2013})  and to master the game of Go (\cite{AlphaGo}), researchers have in recent years made a lot of progress towards designing algorihtms capable of learning and solving, \textit{without human knowledge}, the Rubiks Cube (\textbf{RC}) - as well as similar single player puzzles - by using Deep Q-Learning (\textbf{DQR})  (\cite{DBLP:journals/corr/abs-1805-07470}) or search and (\textbf{DRL}) value iteration (\cite{https://doi.org/10.48550/arxiv.1805.07470}).
+\\
+In this project, I will attempt to implement a variety of solvers combining $A^{*}$ search and heuristics, some of which will be handcrafted, others which I will train on randomly generated sequences of puzzles using (\textbf{DL}) or (\textbf{DRL}), to solve the 15-puzzle (and variations of different dimensions), as well as the Rubik's cube.
+\\
+\\
+\textbf{Organisation of this thesis}
+\\
+In the first chapter, I will quickly describe what I hope to get out of this project, both in terms of personal learning, as well as in terms of tangible results (solving some puzzles!).
+In chapter 2, I will do a quick recap of the different methods that I will use in the project.
+Chapter 3 will be dedicated to discussing the mathematics of the sliding puzzle (\textbf{SP}) and the (\textbf{RC}).
+I might throw a few random (but hopefully interesting) observations in there and give some references for the keen reader.
+I will then give an overview in chapter 4 of the code base I have developed - and put in the open on my github page (\cite{FB}) - to complete this project.
+Finally, in chapters 5 and 6, I will present all my various results on respecively the sliding puzzle and the Rubiks' cube.
+
 \end{abstract}

### 2022-06-26 latex/Chapters/Code.tex
update

@@
 Currently the two implemented learners are:
-\item PerfectLearner: It instantiates an optimal solver ($A^{*}$ with a configurable heuristic - but will only accept heurstic that advertise themselves as optimal.
+\item PerfectLearner: It instantiates an optimal solver ($A^{*}$ with a configurable heuristic - but will only accept heuristic that advertise themselves as optimal.
 The learning consists in generating all the possible configuration of the considered puzzle, solve them with the optimal solver, and save the optimal cost of it as well as those of the whole solution path.
@@
 \item DeepLearner tbd
-\item DeepReinforcementLearner: It instantiates a DeepLearning (a network), and trains it using DRL.
-It then saves the trained network, which can then be used in the DeepLearningHeuristic we have seen earlier in section \ref{HSS}
+\item DeepReinforcementLearner: It instantiates a DeepLearning (network), and trains it using DRL.
+It then saves the trained network, which can then be used in the DeepLearningHeuristic we have seen earlier in section \ref{HSS}.
+The
 
@@
 
+\subsection{rubiks.scripts}
+
+Finally it is worth noting that the code will save on disk a lot of data (e.g.
+the learners will save what they have learnt, e.g.
+a Pytorch network or a data base of optimal costs, the performance comparison will run solvers versus very many configurations of puzzles and save the results for later being able to display) etc...
+The base of the tree to save all this data can be chosen by setting up the "RUBIKSDATA" environment variable.
+If not, it will go somewhere in you HOME :)
+
 %-----------------------------------
@@
 \section{Running Some Examples}
+Here I show how to run some examples... TBC
 
@@
 \section{DL Training}
+\label{DLTS}
+blabla
 
-blabla
 %-----------------------------------
@@
 \section{DRL Training}
+\label{DRLTS}
 
@@
 
-blabla]
+blabla

### 2022-06-27 latex/MScAIFrancoisBerrier.tex
update

@@
 \usepackage[T5,T1]{fontenc}
+\usepackage{verbatim}
 \usepackage[backend=bibtex,style=authoryear,natbib=true]{biblatex} % Use the bibtex backend with the authoryear citation style (which resembles APA)
@@
 Its framework is very appealing and intuitive.
-In particular, the mathematical beauty of Value iteration and Q iteration (\cite{WatkinsThesis}) for discrete state and action spaces, blindly iterating from \textit{any} initial value, is quite prfound.
+In particular, the mathematical beauty of Value iteration and Q iteration (\cite{WatkinsThesis}) for discrete state and action spaces, blindly iterating from \textit{any} initial value, is quite profound.
 Sadly, it had until recently proven hard to achieve practical success with these methods.
 \\
-Inspired however by the seminal success of Deep Mind's team in using Deep Reinforcement Learning (\textbf{DRL}) to play Atari games (\cite{Mnih2013})  and to master the game of Go (\cite{AlphaGo}), researchers have in recent years made a lot of progress towards designing algorihtms capable of learning and solving, \textit{without human knowledge}, the Rubiks Cube (\textbf{RC}) - as well as similar single player puzzles - by using Deep Q-Learning (\textbf{DQR})  (\cite{DBLP:journals/corr/abs-1805-07470}) or search and (\textbf{DRL}) value iteration (\cite{https://doi.org/10.48550/arxiv.1805.07470}).
+Inspired however by the seminal success of Deep Mind's team in using Deep Reinforcement Learning (\textbf{DRL}) to play Atari games (\cite{Mnih2013})  and to master the game of Go (\cite{AlphaGo}), researchers have in recent years made a lot of progress towards designing algorithms capable of learning and solving, \textit{without human knowledge}, the Rubiks Cube (\textbf{RC}) - as well as similar single player puzzles - by using Deep Q-Learning (\textbf{DQR})  (\cite{DBLP:journals/corr/abs-1805-07470}) or search and (\textbf{DRL}) value iteration (\cite{https://doi.org/10.48550/arxiv.1805.07470}).
 \\
@@
 \addchaptertocentry{\acknowledgementname} % Add the acknowledgements to the table of contents
+\begin{comment}
 I am grateful to my employer Bank of America for sponsoring my part time 2-year MSc in Artificial Intelligence at Royal Holloway.
@@
 Last but not least, my wife Thanh Nh\~a is my biggest supporter in everything I do, and her continued encouragement and enthusiasm in my pursuit of this MSc has made it so much easier to juggle between a demanding banking job, running 50 to 60 miles a week in preparation of multiple marathons and half marathons, and the coursework, lectures and exam revisions.
+\end{comment}
 \end{acknowledgements}
-
 %----------------------------------------------------------------------------------------

### 2022-06-27 rubiks/learners/perfectlearner.py

@@ -134,7 +134,7 @@ class PerfectLearner(Learner):
             self.puzzle_count_since_save = 0
             self.save(self.learning_file_name)
         self.data_base[cls.data][h] = cost
-        if cost > self.highest_cost:
+        if cost >= self.highest_cost:
             self.highest_cost = cost
             self.data_base[cls.most_difficult_puzzle_tag] = puzzle
 

### 2022-06-28 latex/Chapters/SlidingPuzzle.tex
update

@@
 
-\label{Chapter4} % Change X to a consecutive number; for referencing this chapter elsewhere, use \ref{ChapterX}
+\label{sec:ResultsSP} % Change X to a consecutive number; for referencing this chapter elsewhere, use \ref{ChapterX}
 
@@
 
-blablabla
-%----------------------------------------------------------------------------------------
-%	SUBSECTION 1.1
-%----------------------------------------------------------------------------------------
+As mentioned in chapter \ref{sec:Puzzles}, the state space cardinality for the SP grows very quickly with n and m.
+Here are the only dimensions which have less than 239.5 millions states.
+Note I am also only considering n $\leq$ m since (p, q) can always be solved if we know how to solve (q, p):
+\\
+\\
+\begin{tabular}{l*{6}{c}r}
+n              & m & 2 & 3 & 4 & 5\\
+\hline
+2              &   & 12 & 360 & 20,160 & 1,814,400 \\
+3              &   & - & 181,440 &  &    \\
+\end{tabular}
+\\
+\\
+In this section, I will discuss learning the value function of these 5 puzzles \textbf{exactly}.
+In order to do so, one can simply use rubiks.scripts.learner, setting up the PerfectLearner with A* and manhattan heuristic, or instantiate directly a PerfectLearner as have seen in section \ref{PLSS}
+\\
+\\
+I obtained the following God numbers for those puzzles:
+\\
+\\
+\begin{tabular}{l*{6}{c}r}
+n              & m & 2 & 3 & 4 & 5\\
+\hline
+2              &   & 6 & 21 & 36 &  \\
+3              &   & - & 31 &  &    \\
+\end{tabular}
+\\
+\\
+and the most difficult puzzles (requiring a number of steps equal to their respective God number to solves):
 
-\subsection{Perfect Heuristic}
+\begin{three}
+\setrow{2}{,3}
+\setrow{1}{2,1}
+\end{three}
+\\
 
-blablabla
+\begin{five}
+\setrow{2}{4,5,}
+\setrow{1}{1,2,3}
+\end{five}
+\\
 
-%----------------------------------------------------------------------------------------
-%	SUBSECTION 1.2
-%----------------------------------------------------------------------------------------
+\begin{seven}
+\setrow{2}{,7,2,1}
+\setrow{1}{4,3,6,5}
+\end{seven}
+\\
 
-\subsection{Results}
+\begin{nine}
+\setrow{2}{,,,,}
+\setrow{1}{,,,,}
+\end{nine}
+\\
 
-blablabla
+\begin{eight}
+\setrow{3}{8,6,7}
+\setrow{2}{2,5,4}
+\setrow{1}{3,,1}
+\end{eight}
+\\
 
@@
 Its God number is only 31, which definitely makes it manageable.
-However, this is already an intermediary size, large enough to make trying deep reinforcement learning.
+However, this is already an intermediary size, large enough to make trying deep reinforcement learning meaningful.
+
+\begin{landscape}
+\begin{figure}[H]
+\centering
+\includegraphics[scale=0.5]{./Figures/33SPDeepReinforcementLearning.jpeg}
+%\decoRule
+\caption[Codebase]{Code base}
+\label{fig:Codebase}
+\end{figure}
+\end{landscape}
+
+\begin{landscape}
+\begin{figure}[H]
+\centering
+\includegraphics[scale=0.5]{./Figures/33SPPerformance.jpeg}
+%\decoRule
+\caption[Codebase]{Code base}
+\label{fig:Codebase}
+\end{figure}
+\end{landscape}
 

### 2022-07-03 latex/Chapters/SlidingPuzzle.tex
update

@@
 \\
+\begin{center}
 \begin{tabular}{l*{6}{c}r}
@@
 2              &   & 12 & 360 & 20,160 & 1,814,400 \\
-3              &   & - & 181,440 &  &    \\
+3              &   &   & 181,440 &  &    \\
 \end{tabular}
-\\
-\\
+\end{center}
 In this section, I will discuss learning the value function of these 5 puzzles \textbf{exactly}.
 In order to do so, one can simply use rubiks.scripts.learner, setting up the PerfectLearner with A* and manhattan heuristic, or instantiate directly a PerfectLearner as have seen in section \ref{PLSS}
-\\
-\\
-I obtained the following God numbers for those puzzles:
-\\
-\\
+I obtained the following God numbers for these puzzles:
+\begin{center}
 \begin{tabular}{l*{6}{c}r}
@@
 2              &   & 6 & 21 & 36 &  \\
-3              &   & - & 31 &  &    \\
+3              &   &   & 31 &  &    \\
 \end{tabular}
+\end{center}
+and the most difficult puzzles (requiring a number of steps equal to their respective God number to solves):
 \\
 \\
-and the most difficult puzzles (requiring a number of steps equal to their respective God number to solves):
-
+\underline{Most difficult 2x2 (6 moves)}:
+\begin{center}
 \begin{three}
@@
 \end{three}
-\\
-
+\end{center}
+\underline{Most difficult 2x3 (21 moves)}:
+\begin{center}
 \begin{five}
@@
 \end{five}
-\\
-
+\end{center}
+\underline{Most difficult 2x4 (36 moves)}:
+\begin{center}
 \begin{seven}
@@
 \end{seven}
-\\
-
+\end{center}
+\underline{Most difficult 2x5 (47 moves) .... TBD }:
+\begin{center}
 \begin{nine}
-\setrow{2}{,,,,}
-\setrow{1}{,,,,}
+\setrow{2}{,2,7,4,1}
+\setrow{1}{5,8,3,9,6}
 \end{nine}
-\\
-
+\end{center}
+\underline{Most difficult 3x3 (31 moves)}:
+\begin{center}
 \begin{eight}
@@
 \end{eight}
-\\
+\end{center}
 

### 2022-07-03 rubiks/solvers/naiveslidingsolver.py

@@ -1,0 +1,716 @@
+########################################################################################################################
+# Francois Berrier - Royal Holloway University London - MSc Project 2022                                               #
+########################################################################################################################
+from rubiks.puzzle.slidingpuzzle import SlidingPuzzle, Slide
+from rubiks.solvers.solver import Solver, Solution
+########################################################################################################################
+
+
+class NaiveSlidingSolver(Solver):
+    """ Base class for a puzzle solver. How it actually solves its puzzle type is
+    left to derived classes implementations by overwriting the  'solve_impl' method
+     """
+
+    @classmethod
+    def know_to_be_optimal(cls):
+        return False
+
+    def solve_impl(self, puzzle, **kw_args) -> Solution:
+        return self.reduce(puzzle)
+
+    def reduce(self,
+               puzzle,
+               max_n=1,
+               max_m=1) -> Solution:
+        assert isinstance(puzzle, SlidingPuzzle), \
+            '%s can only solve %s' % (self.get_name(), SlidingPuzzle.__name__)
+        self.log_info('We need to solve ', puzzle)
+        n = puzzle.n
+        m = puzzle.m
+        solution = Solution(cost=0,
+                            path=list(),
+                            expanded_nodes=float('nan'),
+                            puzzle=puzzle)
+        offset_n = 0
+        offset_m = 0
+        sliding_puzzle = puzzle.clone()
+        while n > max(2, max_n) or m > max(2, max_m):
+            if n >= m:
+                self.log_info('Let us sort out top row')
+                moves = self.solve_top_row(sliding_puzzle)
+                for move in moves:
+                    move.tile = (move.tile[0] + offset_n,
+                                 move.tile[1] + offset_m)
+                offset_n += 1
+                self.log_info('We now have ', sliding_puzzle, ' ... ')
+                sliding_puzzle = SlidingPuzzle(tiles=sliding_puzzle.tiles[1:])
+                self.log_info(' ... and are left solving ', sliding_puzzle)
+            else:
+                self.log_info('Let us sort out left col')
+                moves = self.solve_left_col(sliding_puzzle)
+                for move in moves:
+                    move.tile = (move.tile[0] + offset_n,
+                                 move.tile[1] + offset_m)
+                offset_m += 1
+                self.log_info('We now have ', sliding_puzzle, ' ... ')
+                sliding_puzzle = SlidingPuzzle(tiles=sliding_puzzle.tiles[:, 1:])
[…]

### 2022-07-03 rubiks/solvers/solution.py

@@ -15,6 +15,9 @@ class Solution:
     expanded_nodes = '# expanded nodes'
     puzzle = 'puzzle'
 
+    def apply(self, puzzle):
+        return puzzle.apply_moves(self.path)
+
     def __init__(self,
                  cost,
                  path,

### 2022-07-03 unittests/testdeeplearning.py

@@ -14,23 +14,20 @@ class TestDeepLearning(TestCase):
     def test_fully_connected_direct_construct(self):
         fully_connected = FullyConnected(puzzle_type=Puzzle.sliding_puzzle,
                                          n=4)
-        self.assertEqual(FullyConnected.default_layers,
-                         tuple(fully_connected.layers_int[1:-1]))
+        self.assertEqual(FullyConnected.default_layers, fully_connected.layers_description)
         self.assertFalse(fully_connected.one_hot_encoding)
         fully_connected_2 = FullyConnected(puzzle_type=Puzzle.sliding_puzzle,
                                            n=4,
                                            layers_description=(2, 3, 4, 5, 6),
                                            one_hot_encoding=True)
-        self.assertEqual((2, 3, 4, 5, 6),
-                         tuple(fully_connected_2.layers_int[1:-1]))
+        self.assertEqual((2, 3, 4, 5, 6), fully_connected_2.layers_description)
         self.assertTrue(fully_connected_2.one_hot_encoding)
 
     def test_fully_connected_from_factory(self):
         fully_connected = DeepLearning.factory(network_type=DeepLearning.fully_connected_net,
                                                puzzle_type=Puzzle.sliding_puzzle,
                                                n=4)
-        self.assertEqual(FullyConnected.default_layers,
-                         tuple(fully_connected.layers_int[1:-1]))
+        self.assertEqual(FullyConnected.default_layers, fully_connected.layers_description)
         self.assertFalse(fully_connected.one_hot_encoding)
         fully_connected_2 = DeepLearning.factory(network_type=DeepLearning.fully_connected_net,
                                                  puzzle_type=Puzzle.sliding_puzzle,
@@ -38,8 +35,7 @@ class TestDeepLearning(TestCase):
                                                  m=2,
                                                  layers_description=(2, 3, 4, 5, 6),
                                                  one_hot_encoding=True)
-        self.assertEqual((2, 3, 4, 5, 6),
-                         tuple(fully_connected_2.layers_int[1:-1]))
+        self.assertEqual((2, 3, 4, 5, 6), fully_connected_2.layers_description)
         self.assertTrue(fully_connected_2.one_hot_encoding)
         self.assertEqual(fully_connected_2.get_puzzle_dimension(), (3, 2))
 

### 2022-07-03 unittests/testsolver.py

@@ -77,12 +77,12 @@ class TestSolver(TestCase):
         puzzle = Puzzle.factory(**solver.get_config()).apply_random_moves(inf)
         self.assertEqual(dimension, puzzle.dimension())
         logger.log_info(puzzle)
-        try:
-            solution = solver.solve(puzzle=puzzle)
-            self.assertTrue(False)
-        except TimeoutError:
-            pass
+        solution = solver.solve(puzzle=puzzle)
+        self.assertTrue(solution.failed())
+        self.assertTrue(str(solution).find('Exceeded timeout') >= 0)
+        logger.log_info(solution)
         solution = solver.solve(puzzle=puzzle, time_out=300)
+        self.assertFalse(solution.failed())
         logger.log_info(solution)
 
     def test_deep_reinforcement_learning_solver(self):

### 2022-07-03 latex/Chapters/Code.tex
update

@@
 \subsection{rubiks.solvers}
+This module implements solvers, which solve puzzles.
+The base class Solver is a Factory of solvers, and in addition to being able to instantiating the following types of solvers, can run different solvers through a similal sequences of random puzzles (for various increasing degrees of difficulty (shuffling), and/or perfectly shuffled ones) and display a comparison of how they perform in a number of metrics.
+
+\begin{itemize}
+\item DFSSolver
+\item BFSSolver
+\item AStarSolver
+\item NaiveSlidingSolver
+\end{itemize}
 
@@
 %-----------------------------------
-\section{Running Some Examples}
+\section{Learners}
+
 Here I show how to run some examples... TBC
@@
 %-----------------------------------
-%	SECTION 2.1
+%	SUBSECTION 2.1
 %-----------------------------------
@@
 %-----------------------------------
-%	SECTION 3
+%	SUBSECTION 2.2
 %-----------------------------------
-\section{DL Training}
-\label{DLTS}
+\subsection{Deep Learner}
+\label{DLSS}
 blabla
@@
 %-----------------------------------
-%	SECTION 4
+%	SUBSECTION 2.3
 %-----------------------------------
-\section{DRL Training}
-\label{DRLTS}
+\subsection{Deep Reinforcement Learner}
+\label{DRLSS}
 
@@
 %-----------------------------------
-%	SECTION 5
+%	SECTION 3
 %-----------------------------------
 
-\section{Solvers Comparison}
+\section{Solvers}
 
-blabla
+%-----------------------------------
+%	SUBSECTION 3.1
+%-----------------------------------
+\subsection{BFS}
+\label{BFSSS}
+
+%-----------------------------------
+%	SUBSECTION 3.2
+%-----------------------------------
+\subsection{DFS}
+\label{DFSSS}
+
+%-----------------------------------
+%	SUBSECTION 3.3
+%-----------------------------------
+\subsection{A*}
+\label{ASSS}
+
+\textbf{manhattan heuristic}
+
+\textbf{perfect heuristic}
+
+\textbf{deep learning heuristic}
+
+\textbf{deep reinforcement learning heuristic}

### 2022-07-03 latex/Chapters/SlidingPuzzle.tex
update

@@
 \section{Low dimension}
+\label{sec:SPLowDimension}
 
@@
 \end{center}
-In this section, I will discuss learning the value function of these 5 puzzles \textbf{exactly}.
-In order to do so, one can simply use rubiks.scripts.learner, setting up the PerfectLearner with A* and manhattan heuristic, or instantiate directly a PerfectLearner as have seen in section \ref{PLSS}
+In this section, I will discuss \textbf{full} results for these 5 puzzles.
+In order to fully solve them, one can simply use rubiks.scripts.learner, setting up the PerfectLearner with A* and manhattan heuristic, or instantiate directly a PerfectLearner as have seen in section \ref{PLSS}
 I obtained the following God numbers for these puzzles:
@@
 \label{S33}
-As seen above, we have actually been able to solve the 3 by 3 case perfectly, since it only has 181,440 possible configurations.
+
+\subsection{Perfect learner}
+As seen in the previous section section \ref{sec:SPLowDimension}, the 3 by 3 SP is one of the cases we have been able to solve perfectly, since it only has 181,440 possible configurations.
 Its God number is only 31, which definitely makes it manageable.
 However, this is already an intermediary size, large enough to make trying deep reinforcement learning meaningful.
+The PerfectLearner's learning are shown in figure \ref{fig:33SPPerfectLearning}:
 
+\newgeometry{top=0mm, bottom=0mm, left=0mm, right=0mm}
 \begin{landscape}
+\centering\vspace*{\fill}
 \begin{figure}[H]
 \centering
-\includegraphics[scale=0.5]{./Figures/33SPDeepReinforcementLearning.jpeg}
+\includegraphics[scale=0.7]{./Figures/33SPPerfectLearning.jpeg}
 %\decoRule
-\caption[Codebase]{Code base}
-\label{fig:Codebase}
+\caption[33SPPerfectLearning]{Perfect Learning 3x3 SP}
+\label{fig:33SPPerfectLearning}
 \end{figure}
+\vfill
 \end{landscape}
+\restoregeometry
 
+\subsection{Deep reinforcement learner}
+ The DeepReinforcementLearner's learning is shown in figure \ref{fig:33SPDeepReinforcementLearning}:
+
+\newgeometry{top=0mm, bottom=0mm, left=0mm, right=0mm}
 \begin{landscape}
+\centering\vspace*{\fill}
 \begin{figure}[H]
 \centering
-\includegraphics[scale=0.5]{./Figures/33SPPerformance.jpeg}
+\includegraphics[align=c, scale=0.55]{./Figures/33SPDeepReinforcementLearning.jpeg}
+\caption[33SPDeepReinforcementLearning]{Deep reinforcement learner 3x3 SP}
+\label{fig:33SPDeepReinforcementLearning}
+\end{figure}
+\vfill
+\end{landscape}
+\restoregeometry
+
+\subsection{Solvers' comparison}
+Finally, to conclude this section on the 3x3 case, let's us discuss a comparison of several algorithms on 1000 random puzzles generated for a number of random shuffling (with best effort no backtracking) from 0 to 50 in step of 2, as well as for perfect shuffling (denoted by $\infty$) on the comparison graphs.
+The results are shown in figure \ref{fig:33SPPerformance}
+
+\newgeometry{top=0mm, bottom=0mm, left=0mm, right=0mm}
+\begin{landscape}
+\centering\vspace*{\fill}
+\begin{figure}[H]
+\centering
+\includegraphics[scale=0.63]{./Figures/33SPPerformance.jpeg}
 %\decoRule
-\caption[Codebase]{Code base}
-\label{fig:Codebase}
+\caption[33SPPerformance]{Solvers' performance comparison 3x3 SP}
+\label{fig:33SPPerformance}
 \end{figure}
+\vfill
 \end{landscape}
+\restoregeometry
 

### 2022-07-03 latex/Chapters/Puzzles.tex
update

@@
 $n * m - 1$ tiles, numbered 1 to $n * m - 1$ are placed in all the compartments but one (which is left empty), and we can slide a tile directly adjascent to the empty compartment into it.
-Notice from a programming and mathematical analysis perspective, it is often easier to equivalently think of the empty compartment being moved (or swapped with) an adjacent tile.
+Notice from a programming and mathematical analysis perspective, it is often easier to equivalently think of the empty compartment being moved into (or swapped with) an adjacent tile.
 Starting from a given shuffling of the tiles on the board, our goal will be to execute moves until the tiles in ascending order: left to right, top to bottom (in the usual western reading order), the empty tile being at the very bottom right.

### 2022-07-03 latex/Chapters/SlidingPuzzle.tex
update

@@
 \section{Intermediary case - 3x3}
-\label{S33}
+\label{sec:S33}
 
 \subsection{Perfect learner}
-As seen in the previous section section \ref{sec:SPLowDimension}, the 3 by 3 SP is one of the cases we have been able to solve perfectly, since it only has 181,440 possible configurations.
+As discussed in the previous section section \ref{sec:SPLowDimension}, the 3 by 3 SP is one of the cases I have been able to solve perfectly, since it only has 181,440 possible configurations.
 Its God number is only 31, which definitely makes it manageable.
-However, this is already an intermediary size, large enough to make trying deep reinforcement learning meaningful.
-The PerfectLearner's learning are shown in figure \ref{fig:33SPPerfectLearning}:
+However, this is already an intermediary size, large enough that it is worth trying and comparing a few different methods, including deep reinforcement learning.
+To start with, I ran the PerfectLearner with n=m=3, and the results are shown below in figure \ref{fig:33SPPerfectLearning}.
+It is interesting to note that there are only two hardest configurations (cost 31) and 221 configurations of cost 30.
 
@@
 \subsection{Solvers' comparison}
-Finally, to conclude this section on the 3x3 case, let's us discuss a comparison of several algorithms on 1000 random puzzles generated for a number of random shuffling (with best effort no backtracking) from 0 to 50 in step of 2, as well as for perfect shuffling (denoted by $\infty$) on the comparison graphs.
+\label{ssec:33SPSC}
+Finally, to conclude this section on the 3x3 case, let me discuss a comparison of several algorithms on 1000 random puzzles generated for a number of random shuffling (with best effort no backtracking) from 0 to 50 in step of 2, as well as for perfect shuffling (denoted by $\infty$) on the comparison graphs.
 The results are shown in figure \ref{fig:33SPPerformance}
@@
 
+\subsection{Solving the hardest problem}
+To finish with the 3x3 SP, let me try to throw one of the two hardest 3x3 configurations (see subsection \ref{sec:SPLowDimension}) at the different solvers to see how they fare.
+The results are shown here
+
+\begin{center}
+\begin{tabular}{l*{5}{c}r}
+\hline
+\textbf{solver}      & & \textbf{cost} & \textbf{\# expanded nodes} & \textbf{run time (ms)} \\
+\hline
+AStarSolver[Manhattan]   &   &      31  & 58,859  & 11,327  \\
+\hline
+AStarSolver[PerfectHeuristic]  &   & 31 & 1,585 & 202 \\
+\hline
+AStarSolver[DeepLearningHeuristic]  &   & 31 & 101 & 58 \\
+\hline
+BFS  &   & - & - & time out \\
+\hline
+NaiveSlidingSolver  &   & 61 & n/a & 18 \\
+\hline
+\end{tabular}
+\end{center}
+On this specific configuration, there was obviously no chance that the BFS would complete, hence it timed out.
+It would have no matter what time out I set.
+Indeed, since it has no heuristic to guide its search, it would need to explore in the order of $3^{31}$ - roughly 617 trillions - nodes to reach the goal!
+\\
+Rather interestingly, my DRL heuristic performs much better than the manhattan heuristic (not super suprising), but also outperforms the perfect heuristic quite significantly both in terms of run time and of nodes expansion.
+Obviously there is no guarantee that the perfect heuristic will not be outperformed on some random configuration, and it does on this occasion.
+However, as we have seen in the previous subection \ref{ssec:33SPSC}, it is not the case on average.
+\\
+Finally, the naive solver outperforms every other solver in terms of run time, but finds a rather poor solution of 61 moves.
+
 %-----------------------------------

### 2022-07-03 rubiks/scripts/learner.py

@@ -19,11 +19,11 @@ if '__main__' == __name__:
                             Learner.do_plot,
                             Learner.do_cleanup_learning_file,
                             } """
-    action_type = Learner.do_plot
+    action_type = Learner.do_learn
     """ What puzzle """
     puzzle_type = Puzzle.sliding_puzzle
-    n = 3
-    m = 3
+    n = 2
+    m = 5
     dimension = Puzzle.factory(**globals()).dimension()
     """ Which learner_type {Learner.perfect,
                             Learner.deep_reinforcement_learner,
@@ -33,24 +33,24 @@ if '__main__' == __name__:
     """ for plot """
     plot_metrics = DeepReinforcementLearner.default_plot_metrics
     """ If it's a perfect learner config here """
-    time_out = 60
-    nb_cpus = 5
-    cpu_multiplier = 1
-    max_puzzles = nb_cpus * cpu_multiplier * 1
+    time_out = 360
+    nb_cpus = 10
+    cpu_multiplier = 10
+    max_puzzles = nb_cpus * cpu_multiplier * 100
     regular_save = nb_cpus * cpu_multiplier * 1
     save_timed_out_max_puzzles = 100000
     after_round_save = True
     flush_timed_out_puzzles = True
     save_timed_out = True
-    rerun_timed_out = False
-    rerun_timed_out_only = False
+    rerun_timed_out = True
+    rerun_timed_out_only = True
     abort_after_that_many_consecutive_timed_out = 100
     nb_shuffles_from_goal = 25
     """ puzzle generation process {PerfectLearner.perfect_random_puzzle_generation,
                                    PerfectLearner.permutation_puzzle_generation,
                                    PerfectLearner.random_from_goal_puzzle_generation,
                                    } """
-    puzzle_generation = PerfectLearner.perfect_random_puzzle_generation
+    puzzle_generation = PerfectLearner.permutation_puzzle_generation
     heuristic_type = Heuristic.manhattan
     """ If it's a DRL learner config is here ... """
     nb_epochs = 5000

### 2022-07-03 rubiks/solvers/solver.py

@@ -21,7 +21,7 @@ from rubiks.puzzle.puzzle import Puzzle
 from rubiks.puzzle.puzzled import Puzzled
 from rubiks.search.searchstrategy import SearchStrategy
 from rubiks.solvers.solution import Solution
-from rubiks.utils.utils import pprint, to_pickle, remove_file, s_format, pformat
+from rubiks.utils.utils import pprint, to_pickle, remove_file, s_format, pformat, ms_format
 ########################################################################################################################
 
 
@@ -80,7 +80,10 @@ class Solver(Factory, Puzzled, Loggable, metaclass=ABCMeta):
     def solve(self, puzzle, **kw_args) -> Solution:
         kw_args = {**self.get_config(), **kw_args}
         try:
+            b4 = snap()
             solution = self.solve_impl(puzzle, **kw_args)
+            run_time = snap() - b4
+            solution.add_additional_info(run_time=ms_format(run_time))
         except Exception as error:
             solution = Solution.failure(puzzle=puzzle,
                                         solver_name=self.get_name(),
@@ -440,23 +443,23 @@ class Solver(Factory, Puzzled, Loggable, metaclass=ABCMeta):
 
     def action(self):
         config = self.get_config()
-        if self.do_plot == self.action_type:
-            try:
+        try:
+            if self.do_plot == self.action_type:
                 self.plot_performance()
-            except KeyboardInterrupt:
-                self.log_warning('KeyboardInterrupt raised')
-        elif self.do_solve == self.action_type:
-            puzzle = Puzzle.factory(**self.get_config())
-            puzzle = puzzle.apply_random_moves(nb_moves=config[self.__class__.nb_shuffles])
-            return self.solve(puzzle)
-        elif self.do_performance_test == self.action_type:
-            self.performance_test()
-        elif self.do_cleanup_performance_file == self.action_type:
-            self.cleanup_perf_file()
-        elif self.do_cleanup_shuffles_file == self.action_type:
-            self.cleanup_shuffles_file()
-        else:
-            raise NotImplementedError('Unknown action_type [%s]' % self.action_type)
+            elif self.do_solve == self.action_type:
+                puzzle = Puzzle.factory(**self.get_config())
+                puzzle = puzzle.apply_random_moves(nb_moves=config[self.__class__.nb_shuffles])
+                return self.solve(puzzle)
+            elif self.do_performance_test == self.action_type:
+                self.performance_test()
+            elif self.do_cleanup_performance_file == self.action_type:
[…]

### 2022-07-03 latex/Chapters/SlidingPuzzle.tex
update

@@
 
-\subsection{Solving the hardest problem}
+\subsection{Solving the hardest 3x3 problem}
 To finish with the 3x3 SP, let me try to throw one of the two hardest 3x3 configurations (see subsection \ref{sec:SPLowDimension}) at the different solvers to see how they fare.
@@
 
-blabla
+TBD
 
@@
 
-blabla
+TBD
+
+%-----------------------------------
+%	SECTION 5
+%-----------------------------------
+
+\section{5x5}
+
+TBD

### 2022-07-09 rubiks/deeplearning/deeplearning.py
update

@@ -27,6 +27,7 @@ class DeepLearning(Module, Factory, Puzzled, Loggable, metaclass=ABCMeta):
 
     network_type = 'network_type'
     fully_connected_net = 'fully_connected_net'
+    convolutional_net = 'convolutional_net'
     state_dict_tag = 'state_dict'
 
     @abstractmethod
@@ -47,7 +48,8 @@ class DeepLearning(Module, Factory, Puzzled, Loggable, metaclass=ABCMeta):
                          action=cls.store_true)
         cls.add_argument(parser,
                          field=cls.network_type,
-                         choices=[cls.fully_connected_net])
+                         choices=[cls.fully_connected_net,
+                                  cls.convolutional_net])
 
     @classmethod
     def factory_key_name(cls):
@@ -61,7 +63,10 @@ class DeepLearning(Module, Factory, Puzzled, Loggable, metaclass=ABCMeta):
     @classmethod
     def widget_types(cls):
         from rubiks.deeplearning.fullyconnected import FullyConnected
-        return {cls.fully_connected_net: FullyConnected}
+        from rubiks.deeplearning.convolutional import Convolutional
+        return {cls.fully_connected_net: FullyConnected,
+                cls.convolutional_net: Convolutional,
+                }
 
     def get_data(self):
         return (self.get_config(),

### 2022-07-09 rubiks/puzzle/slidingpuzzle.py

@@ -221,11 +221,13 @@ class SlidingPuzzle(Puzzle):
     def from_tensor(self):
         raise NotImplementedError('Please implement this ... need to de-one_hot then call init')
     
-    def to_tensor(self, one_hot_encoding=False):
+    def to_tensor(self, one_hot_encoding=False, flatten=True):
         tiles = self.tiles
         if one_hot_encoding:
             tiles = one_hot(tiles)
-        return tiles.flatten(1)
+        if flatten:
+            tiles = tiles.flatten(1)
+        return tiles
 
     def perfect_shuffle(self):
         """ We set up the tiles randomly, and then just swap the first two if the signature is not right """
@@ -285,6 +287,11 @@ class SlidingPuzzle(Puzzle):
             return False
         flat_ordered = flat.sort().values.tolist()[1:]
         return flat_ordered == flat.tolist()[:-1]
+
+    @classmethod
+    def hardest(cls, dimension):
+        return {(3, 3): [[8, 6, 7], [2, 5, 4], [3, 0, 1]],
+                }[dimension]
     
 ########################################################################################################################
 

### 2022-07-09 rubiks/solvers/astarsolver.py

@@ -27,7 +27,14 @@ class AStarSolver(Solver):
 
     def solve_impl(self, puzzle, **kw_args):
         strat = AStar(initial_state=puzzle, **kw_args)
-        strat.solve()
+        try:
+            strat.solve()
+        except TimeoutError:
+            solution = Solution.failure(puzzle,
+                                        time_out=True,
+                                        failure_reson='time out')
+            solution.expanded_nodes = strat.get_node_counts()
+            return solution
         return Solution(strat.get_path_cost(),
                         strat.get_path(),
                         strat.get_node_counts(),

### 2022-07-09 unittests/testdeeplearning.py

@@ -3,6 +3,7 @@
 ########################################################################################################################
 from unittest import TestCase
 ########################################################################################################################
+from rubiks.core.loggable import Loggable
 from rubiks.puzzle.puzzle import Puzzle
 from rubiks.deeplearning.deeplearning import DeepLearning
 from rubiks.deeplearning.fullyconnected import FullyConnected
@@ -38,5 +39,88 @@ class TestDeepLearning(TestCase):
         self.assertEqual((2, 3, 4, 5, 6), fully_connected_2.layers_description)
         self.assertTrue(fully_connected_2.one_hot_encoding)
         self.assertEqual(fully_connected_2.get_puzzle_dimension(), (3, 2))
+
+    def test_convolutional_net_from_factory(self):
+        logger = Loggable(name='test_convolutional_net_from_factory')
+        puzzle_type = Puzzle.sliding_puzzle
+        n = 3
+        m = 4
+        network_type = DeepLearning.convolutional_net
+        one_hot_encoding = False
+        kernel_size = (3, 3)
+        convo_layers_description = (10, 5, 1)
+        padding = 1
+        kw_args = locals()
+        kw_args.pop('self', None)
+        puzzle = Puzzle.factory(**kw_args).apply_random_moves(10)
+        convo = DeepLearning.factory(**kw_args)
+        logger.log_debug(puzzle)
+        logger.log_debug(convo.massage_puzzles(puzzle))
+        logger.log_debug(convo.evaluate(puzzle))
+        logger.log_debug(convo.evaluate(puzzle).shape)
+        self.assertEqual((3, 4), tuple(convo.evaluate(puzzle).shape))
+
+    def test_convolutional_net_from_factory_many_puzzles(self):
+        logger = Loggable(name='test_convolutional_net_from_factory')
+        puzzle_type = Puzzle.sliding_puzzle
+        n = 3
+        m = 4
+        network_type = DeepLearning.convolutional_net
+        one_hot_encoding = False
+        kernel_size = (3, 3)
+        convo_layers_description = (10, 5, 1)
+        padding = 1
+        kw_args = locals()
+        kw_args.pop('self', None)
+        puzzles = [Puzzle.factory(**kw_args).apply_random_moves(10) for _ in range(10)]
+        convo = DeepLearning.factory(**kw_args)
+        logger.log_debug(puzzles)
+        logger.log_debug(convo.massage_puzzles(puzzles))
[…]

### 2022-07-09 unittests/testheuristics.py

@@ -1,11 +1,15 @@
 ########################################################################################################################
 # Francois Berrier - Royal Holloway University London - MSc Project 2022                                               #
 ########################################################################################################################
+from math import inf
 from unittest import TestCase
 ########################################################################################################################
 from rubiks.heuristics.manhattan import Manhattan
 from rubiks.core.loggable import Loggable
 from rubiks.heuristics.heuristic import Heuristic
+from rubiks.puzzle.puzzle import Puzzle
+from rubiks.puzzle.slidingpuzzle import SlidingPuzzle
+from rubiks.solvers.solver import Solver
 ########################################################################################################################
 
 
@@ -40,5 +44,72 @@ class TestHeuristics(TestCase):
                 # given the heuristic is admissible, and optimal cost is smaller than nb_moves
                 # we should have heuristic_cost <= optimal_cost <= nb_moves
                 self.assertGreaterEqual(nb_moves, heuristic_cost)
+
+    def test_manhattan_distance_3_3_plus(self):
+        logger = Loggable(name='test_manhattan_distance_3_3_plus')
+        sliding_puzzle = Puzzle.factory(puzzle_type=Puzzle.sliding_puzzle,
+                                        tiles=[[8, 2, 0], [3, 4, 1], [5, 6, 7]])
+        manhattan = Manhattan(n=3, m=3)
+        manhattan_plus = Manhattan(n=3, m=3, plus=True)
+        cost = manhattan.cost_to_go_from_puzzle(sliding_puzzle)
+        cost_plus = manhattan_plus.cost_to_go_from_puzzle(sliding_puzzle)
+        logger.log_info({'puzzle': str(sliding_puzzle),
+                         'cost': cost,
+                         'cost++': cost_plus})
+        self.assertEqual(16, cost)
+        self.assertEqual(16, cost_plus)
+
+    def manhattan_plus_check(self, name, n, m=None):
+        logger = Loggable(name=name)
+        manhattan = Solver.factory(solver_type=Solver.astar,
+                                   heuristic_type=Heuristic.manhattan,
+                                   puzzle_type=Puzzle.sliding_puzzle,
+                                   n=n,
+                                   m=m,
+                                   plus=False)
+        manhattan_plus = Solver.factory(solver_type=Solver.astar,
+                                        heuristic_type=Heuristic.manhattan,
[…]

### 2022-07-09 unittests/testsolver.py

@@ -10,6 +10,7 @@ from rubiks.puzzle.puzzle import Puzzle
 from rubiks.solvers.solver import Solver
 from rubiks.deeplearning.deeplearning import DeepLearning
 from rubiks.learners.perfectlearner import PerfectLearner
+from rubiks.learners.deeplearner import DeepLearner
 from rubiks.learners.deepreinforcementlearner import DeepReinforcementLearner
 from rubiks.utils.utils import get_model_file_name, remove_file
 ########################################################################################################################
@@ -127,13 +128,13 @@ class TestSolver(TestCase):
         remove_file(model_file_name)
 
     def test_deep_reinforcement_learning_solver_with_scheduler(self):
-        logger = Loggable(name='test_deep_reinforcement_learning_solver')
+        logger = Loggable(name='test_deep_reinforcement_learning_solver_with_scheduler')
         puzzle_type = Puzzle.sliding_puzzle
         dimension = (2, 2)
         # we learn first
         model_file_name = get_model_file_name(puzzle_type=puzzle_type,
                                               dimension=dimension,
-                                              model_name='test_deep_reinforcement_learning_solver')
+                                              model_name='test_deep_reinforcement_learning_solver_with_scheduler')
         remove_file(model_file_name)
         learner = DeepReinforcementLearner(puzzle_type=Puzzle.sliding_puzzle,
                                            learning_file_name=model_file_name,
@@ -166,6 +167,46 @@ class TestSolver(TestCase):
         puzzle = Puzzle.factory(**solver.get_config()).apply_random_moves(inf)
         logger.log_info(puzzle)
         solution = solver.solve(puzzle=puzzle)
+        self.assertTrue(solution.success)
+        logger.log_info(solution)
+        remove_file(model_file_name)
+
+    def test_deep_learning_solver_with_scheduler(self):
+        logger = Loggable(name='test_deep_learning_solver_with_scheduler')
+        puzzle_type = Puzzle.sliding_puzzle
+        dimension = (2, 2)
+        # we learn first
+        model_file_name = get_model_file_name(puzzle_type=puzzle_type,
+                                              dimension=dimension,
+                                              model_name='test_deep_learning_solver_with_scheduler')
+        remove_file(model_file_name)
+        learner = DeepLearner(puzzle_type=Puzzle.sliding_puzzle,
+                              learning_file_name=model_file_name,
+                              n=dimension[0],
+                              m=dimension[1],
+                              nb_cpus=1,
+                              network_type=DeepLearning.fully_connected_net,
[…]

### 2022-07-17 latex/Chapters/Code.tex
update

@@
 \subsection{rubiks.core}
-This submodule contains bases classes that make the code base easier to use, debug, and extend. It contains three base classes:
+This submodule contains base classes that make the code base easier to use, debug, and extend. It contains the following:
 \begin{itemize}
@@
 %-----------------------------------
-\subsection{BFS}
+\subsection{Blind search}
+%-----------------------------------
+%	SUBSECTION 3.1.1
+%-----------------------------------
+\subsubsection{BFS}
 \label{BFSSS}
+%-----------------------------------
+%	SUBSECTION 3.1.2
+%-----------------------------------
+\subsubsection{DFS}
+\label{DFSSS}
 
@@
 %-----------------------------------
-\subsection{DFS}
-\label{DFSSS}
+\subsection{Naive Sliding Puzzle Solver}
+As a comparison point, I have implemented a naive sliding puzzle solver, which does what most beginner players would intuitively do when solving the sliding puzzle by hand: solve the top row, then the left column, and keep iterating until done.
+Notice that once either the top row or left column is solved, there is no longer any need to modify it, we have simply reduced the problem to a sub-problem of reduced dimension.
+For the interested reader, the details of the algorithm are as follows:
+\begin{itemize}
+\item if n and m are both equal to 2, we just keep moving the empty tile clock-wise until the puzzle is solved.
+Notice that this is bound to work, since moving clock-wise or counter-clock-wise are the two ony possible moves, and one of them is just un-doing the other one.
+\item if n $\geq$ m, we solve the top row
+\item otherwise we solve the left column
+\end{itemize}
+Solving the top row of a n by m puzzle (left column is similar, mutatis mutandis, so I will not detail it) is accomplished as follows:
+\begin{enumerate}
+\item \label{s1} we sort the tiles (which since we are potentially dealing with a sub-problem, are no longer necessarily 1 to $m* n - 1$), and select the m smaller ones $t_{1}, ..., t_{m-1}, t_{m}$.
+\item \label{s2} we place $t_{m}$ in the bottom-right corner
+\item \label{s3} we place $t_{1}, ..., t_{m-2}$ to their respective positions (in that order, and making sure not to undo anything as we do so)
+\item \label{s4} we place $t_{m-1}$ in the top-right corner
+\item \label{s5} we then move $t_{m}$ just under $t_{m-1}$
+\item \label{s6} we move the empty tile to the left of $t_{m-1}$
+\item \label{s7} finally we move the empty tile right and then down to put $t_{m-1}$ and $t_{m}$ in place.
+\end{enumerate}
+In order to move the tiles, we have written a few simple routines which can move the empty tile from its current position next to (above, below, left or right) any tile, and then can move that tile to another position, all the while avoiding to go through previously moved tiles (hence the particular order in which we move the different tiles above).
+The only case where the above algorithm can get stuck is when both n and m are equal to 3 and that by step \ref{s6} we might end-up with $t_{3}$ under the empty tile.
+We have handcrafted a sequence of moves to solve this particular position.
+\\
+\\
+As a concrete example, let us assume we started with the following (n=6, m=6) puzzle:
+\\
+\begin{thityfive}
+\setrow{6}{14,27,6,  2,5,18}
+\setrow{5}{21,29,13,23,35,30}
+\setrow{4}{26,3,7,9,24,19}
+\setrow{3}{22,12,11,17,16,33}
+\setrow{2}{32,10,20,25,34,28}
+\setrow{1}{8,4,15,31, ,1}
+\end{thityfive}
+\\
+After one call to solve the top row and the left column, we are left with solving the following 5 by 5 sub-puzzle:
+\begin{twentyfour}
+\setrow{5}{9,17,27,18,35}
+\setrow{4}{23,11,15,24,21}
+\setrow{3}{20,8,29,33,10}
+\setrow{2}{22,30,14,32,16}
+\setrow{1}{ ,12,26,34,28}
+\end{twentyfour}
+\\
+step \ref{s1} above will decide to solve the top row by placing $t_{1}, ..., t_{5}$ = $8, 9, 10, 11, 12$ in that order as the top row. Steps \ref{s2} to \ref{s7} will yield in order:
+\\
+\begin{twentyfour}
+\setrow{5}{9,17,27,18,35}
+\setrow{4}{23,11,15,24,21}
+\setrow{3}{20,8,29,33,10}
+\setrow{2}{22,30,14,32,16}
+\setrow{1}{ ,12,26,34,28}
+\end{twentyfour}
+%
+\begin{twentyfour}
+\setrow{5}{9,17,27,18,35}
+\setrow{4}{23,11,15,24,21}
+\setrow{3}{20,8,29,33,10}
+\setrow{2}{22,30,14,32,16}
+\setrow{1}{\color{red}12\color{red}, ,26,34,28}
+\end{twentyfour}
+%
+\begin{twentyfour}
+\setrow{5}{\color{red}8\color{red},\color{red}9\color{red},\color{red}10\color{red},,18}
+\setrow{4}{17,15,27,24,35}
+\setrow{3}{11,23,29,21,33}
+\setrow{2}{20,22,14,32,16}
+\setrow{1}{\color{red}12\color{red},30,26,34,28}
+\end{twentyfour}
+\\
+\begin{twentyfour}
+\setrow{5}{\color{red}8\color{red},\color{red}9\color{red},\color{red}10\color{red},,\color{red}11\color{red}}
+\setrow{4}{23,29,21,18,24}
+\setrow{3}{17,15,27,33,35}
+\setrow{2}{20,22,14,32,16}
+\setrow{1}{\color{red}12\color{red},30,26,34,28}
+\end{twentyfour}
+%
+\begin{twentyfour}
+\setrow{5}{\color{red}8\color{red},\color{red}9\color{red},\color{red}10\color{red},18,\color{red}11\color{red}}
+\setrow{4}{29,27,32, ,\color{red}12\color{red}}
+\setrow{3}{23,21,33,35,24}
+\setrow{2}{15,17,22,14,16}
+\setrow{1}{30,20,26,34,28}
+\end{twentyfour}
+%
+\begin{twentyfour}
+\setrow{5}{\color{red}8\color{red},\color{red}9\color{red},\color{red}10\color{red},\color{red}11\color{red},\color{red}12\color{red}}
+\setrow{4}{29,27,32,18, }
+\setrow{3}{23,21,33,35,24}
+\setrow{2}{15,17,22,14,16}
+\setrow{1}{30,20,26,34,28}
+\end{twentyfour}
+\\
+and we are left with solving the bottom sub-puzzle (n=4,m=5), etc...
 
@@
 %	SUBSECTION 3.3
+%-----------------------------------
+\subsection{Kociemba}
+BTD
+
+%-----------------------------------
+%	SUBSECTION 3.4
 %-----------------------------------

### 2022-07-17 latex/Chapters/Puzzles.tex
update

@@
 \centering
-\includegraphics[scale=0.7]{./Figures/coiled15puzzle}
+\includegraphics[scale=0.3]{./Figures/coiled15puzzle}
+\includegraphics[scale=0.3]{./Figures/coiled15puzzle2.png}
 \decoRule
 \caption[Coiled 15-puzzle]{Coiled 15-puzzle}
-\label{fig:Coiled15Puzzle}
 \end{figure}
 
-It is rather easy to see why this is the case, let us discuss why: Given a configuration c of the tiles, let us define the permutation p(c) of this configuration according to the following schema: we enumerate the tiles row by row (top to bottom), left to right for odd rows and right to left for the even rows, ignoring the empty compartment.
+It is rather easy to see why this is the case, let us discuss why: given a configuration c of the tiles, let us define the permutation p(c) of this configuration according to the following schema: we enumerate the tiles row by row (top to bottom), left to right for odd rows and right to left for the even rows, ignoring the empty compartment.
 For instance, the following 15-puzzle c:

### 2022-07-17 rubiks/scripts/learner.py

@@ -23,21 +23,21 @@ if '__main__' == __name__:
     action_type = Learner.do_learn
     """ What puzzle """
     puzzle_type = Puzzle.sliding_puzzle
-    n = 2
-    m = 6
+    n = 4
+    m = 4
     dimension = Puzzle.factory(**globals()).dimension()
     """ Which learner_type {Learner.perfect,
                             Learner.deep_reinforcement_learner,
                             } 
     """
-    learner_type = Learner.perfect_learner
+    learner_type = Learner.deep_learner
     """ for plot """
     plot_metrics = DeepReinforcementLearner.default_plot_metrics \
         if learner_type is Learner.deep_reinforcement_learner \
         else DeepLearner.default_plot_metrics
     """ If it's a perfect learner config here """
     time_out = 3600
-    nb_cpus = 4
+    nb_cpus = 1
     cpu_multiplier = 25
     max_puzzles = nb_cpus * cpu_multiplier * 10000
     regular_save = nb_cpus * cpu_multiplier * 1
@@ -57,9 +57,9 @@ if '__main__' == __name__:
     heuristic_type = Heuristic.manhattan
     plus = True
     """ If it's a DRL learner config is here ... """
-    nb_epochs = 100000
-    nb_sequences = 150
-    nb_shuffles = 75
+    nb_epochs = 10000
+    nb_sequences = 100
+    nb_shuffles = 60
     training_data_every_epoch = False
     cap_target_at_network_count = True
     update_target_network_frequency = 1000
@@ -72,12 +72,12 @@ if '__main__' == __name__:
     gamma_scheduler = 0.9999
     """ DL learner """
     save_at_each_epoch = False
-    threshold = 0.005
+    threshold = 0.01
     training_data_freq = 1000
     high_target = nb_shuffles + 1
     training_data_from_data_base = True
-    nb_shuffles_min = 37
-    nb_shuffles_max = 47
+    nb_shuffles_min = 40
+    nb_shuffles_max = 60
     """ ... and its network config """
     network_type = DeepLearning.fully_connected_net
     layers_description = (600, 300, 100)

### 2022-07-17 rubiks/solvers/naiveslidingsolver.py

@@ -242,7 +242,7 @@ class NaiveSlidingSolver(Solver):
 
     def solve_left_col(self, sliding_puzzle):
         assert sliding_puzzle.n >= 2
-        #self.log_debug('solve_left_col for ', sliding_puzzle)
+        self.log_debug('solve_left_col for ', sliding_puzzle)
         desired_left_col = sliding_puzzle.tiles.flatten().sort().values[1::sliding_puzzle.m].tolist()
         moves = list()
         end_state = sliding_puzzle.clone()
@@ -264,10 +264,10 @@ class NaiveSlidingSolver(Solver):
                                                   moves)
         target_empty_pos = (sliding_puzzle.n - 2, 0)
         new_moves = list()
-        #self.log_debug('target_empty_pos: ', target_empty_pos)
+        self.log_debug('target_empty_pos: ', target_empty_pos)
         while end_state.empty != target_empty_pos:
-            #self.log_debug('end_state.empty: ', end_state.empty)
-            #self.log_debug('target_empty_pos: ', target_empty_pos)
+            self.log_debug('end_state.empty: ', end_state.empty)
+            self.log_debug('target_empty_pos: ', target_empty_pos)
             if end_state.empty[0] > target_empty_pos[0]:
                 move = end_state.empty_up()
             elif end_state.empty[0] < target_empty_pos[0]:
@@ -278,19 +278,19 @@ class NaiveSlidingSolver(Solver):
                 move = end_state.empty_left()
             new_moves.append(move)
             end_state = end_state.apply(move)
-            #self.log_debug(end_state)
+            self.log_debug(end_state)
         moves.extend(new_moves)
         move = Slide(sliding_puzzle.n - 1, 0)
         end_state = end_state.apply(move)
-        #self.log_debug(end_state)
+        self.log_debug(end_state)
         moves.append(move)
         move = Slide(sliding_puzzle.n - 1, 1)
         end_state = end_state.apply(move)
-        #self.log_debug(end_state)
+        self.log_debug(end_state)
         moves.append(move)
         sliding_puzzle.tiles = end_state.tiles
         sliding_puzzle.empty = end_state.empty
-        #self.log_debug(sliding_puzzle)
+        self.log_debug(sliding_puzzle)
         return moves
 
     @staticmethod

### 2022-07-17 rubiks/solvers/solver.py

@@ -133,11 +133,18 @@ class Solver(Factory, Puzzled, Loggable, metaclass=ABCMeta):
             assert isinstance(solution.path, list)
             assert all(isinstance(move, self.get_goal().get_move_type()) for move in solution.path)
             assert isnan(solution.expanded_nodes) or isinstance(solution.expanded_nodes, int)
+            print('solved ',
+                  puzzle,
+                  ' # ',
+                  index,
+                  ' with cost ',
+                  solution.cost,
+                  ' in ',
+                  s_format(run_time))
         except Exception as error:
-            if error is not RecursionError:
-                self.log_error(error, '. nb_shuffles = ', nb_shuffles, '. index=', index)
             solution = Solution.failure(puzzle)
             solution.set_run_time(self.time_out)
+            print('Failed to solve ', puzzle, ': ', error)
         solution.set_additional_info(index=index)
         return solution
 

### 2022-07-17 unittests/testnaiveslidingsolver.py

@@ -153,6 +153,9 @@ class TestNaiveSlidingSolver(TestCase):
                                max_count=2500,
                                random_puzzles=True)
 
+    #def test_dummy(self):
+    #    self.solve_all_puzzles('test_dummy', max_count=1, n=6, random_puzzles=True)
+
     def test_naive_sliding_solver_moves_to_right_of(self):
         logger = Loggable(name='test_naive_sliding_solver_moves_to_right_of')
         #logger.log_info('from_below=', from_below)

### 2022-07-23 latex/Chapters/Code.tex
update

--- 2022-07-03 latex/Chapters/Code.tex
@@
 
-\section{Code organisation}
+%\section{Code organisation}
 
@@
 
-\subsection{rubiks.core}
-This submodule contains bases classes that make the code base easier to use, debug, and extend. It contains three base classes:
+\section{rubiks.core}
+This submodule contains base classes that make the code base easier to use, debug, and extend. It contains the following:
 \begin{itemize}
@@
 
-\subsection{rubiks.puzzle}
+\section{rubiks.puzzle}
 This submodule contains:
@@
 
-\subsection{rubiks.search}
+\section{rubiks.search}
 This modules contains graph search strategies. I have actually reused the code I implemented for one of the AIPnT assignments. It contains the following classes:
@@
 
-\subsection{rubiks.heuristics}
+\section{rubiks.heuristics}
 \label{HSS}
@@
 
-\subsection{rubiks.deeplearning}
+\section{rubiks.deeplearning}
 This module is a wrapper around Pytorch. It contains:
@@
 
-\subsection{rubiks.learners}
+\section{rubiks.learners}
 \label{PLcode}
@@
 
-\subsection{rubiks.solvers}
+\section{rubiks.solvers}
 This module implements solvers, which solve puzzles.
@@
 
-\subsection{rubiks.scripts}
+\section{rubiks.scripts}
 
@@
 If not, it will go somewhere in you HOME :)
-
-%-----------------------------------
-%	SECTION 2
-%-----------------------------------
-\section{Learners}
-
-Here I show how to run some examples... TBC
-
-%-----------------------------------
-%	SUBSECTION 2.1
-%-----------------------------------
-\subsection{Perfect Learner}
-\label{PLSS}
-
-%-----------------------------------
-%	SUBSECTION 2.2
-%-----------------------------------
-\subsection{Deep Learner}
-\label{DLSS}
-blabla
-
-%-----------------------------------
-%	SUBSECTION 2.3
-%-----------------------------------
-\subsection{Deep Reinforcement Learner}
-\label{DRLSS}
-
-blabla
-
-%-----------------------------------
-%	SECTION 3
-%-----------------------------------
-
-\section{Solvers}
-
-%-----------------------------------
-%	SUBSECTION 3.1
-%-----------------------------------
-\subsection{BFS}
-\label{BFSSS}
-
-%-----------------------------------
-%	SUBSECTION 3.2
-%-----------------------------------
-\subsection{DFS}
-\label{DFSSS}
-
-%-----------------------------------
-%	SUBSECTION 3.3
-%-----------------------------------
-\subsection{A*}
-\label{ASSS}
-
-\textbf{manhattan heuristic}
-
-\textbf{perfect heuristic}
-
-\textbf{deep learning heuristic}
-
-\textbf{deep reinforcement learning heuristic}

### 2022-07-23 latex/Chapters/Examples.tex
update

% Chapter Template

\chapter{Examples} % Main chapter title

\label{sec:Examples} % Change X to a consecutive number; for referencing this chapter elsewhere, use \ref{ChapterX}

%-----------------------------------
%	SECTION 1
%-----------------------------------
\section{Learners}

Here I show how to run some examples... TBC

%-----------------------------------
%	SUBSECTION 1.1
%-----------------------------------
\subsection{Perfect Learner}
\label{PLSS}

%-----------------------------------
%	SUBSECTION 1.2
%-----------------------------------
\subsection{Deep Learner}
\label{DLSS}
blabla

%-----------------------------------
%	SUBSECTION 1.3
%-----------------------------------
\subsection{Deep Reinforcement Learner}
\label{DRLSS}

blabla

%-----------------------------------
%	SECTION 2
%-----------------------------------

\section{Solvers}

%-----------------------------------
%	SUBSECTION 2.1
%-----------------------------------
\subsection{Blind search}
%-----------------------------------
%	SUBSECTION 2.1.1
%-----------------------------------
\subsubsection{BFS}
\label{BFSSS}
%-----------------------------------
%	SUBSECTION 2.1.2
%-----------------------------------
\subsubsection{DFS}
\label{DFSSS}

%-----------------------------------
%	SUBSECTION 2.2
%-----------------------------------
\subsection{Naive Sliding Puzzle Solver}
As a comparison point, I have implemented a naive sliding puzzle solver, which does what most beginner players would intuitively do when solving the sliding puzzle by hand: solve the top row, then the left column, and keep iterating until done. Notice that once either the top row or left column is solved, there is no longer any need to modify it, we have simply reduced the problem to a sub-problem of reduced dimension. For the interested reader, the details of the algorithm are as follows:
\begin{itemize}
\item if n and m are both equal to 2, we just keep moving the empty tile clock-wise until the puzzle is solved. Notice that this is bound to work, since moving clock-wise or counter-clock-wise are the two ony possible moves, and one of them is just un-doing the other one, therefore the only possible sequence of move in a (n=2, m=2) puzzle is to either keep moving clock-wise or counter-clock-wise.
\item if n $\geq$ m, we solve the top row
\item otherwise we solve the left column
\end{itemize}
Solving the top row of a n by m puzzle (left column is similar, mutatis mutandis, so I will not detail it) is accomplished as follows:
\begin{titlemize}{naive algorithm - top-row solver}
\item \label{s1} we sort the tiles (which since we are potentially dealing with a sub-problem, are not necessarily 1 to $m* n - 1$), and select the m smaller ones $t_{1}, ..., t_{m-1}, t_{m}$.
\item \label{s2} we place $t_{m}$ in the bottom-left corner
\item \label{s3} we place $t_{1}, ..., t_{m-2}$ to their respective positions (in that order, and making sure not to undo any previous steps as we do so)
\item \label{s4} we place $t_{m-1}$ in the top-right corner
\item \label{s5} we then move $t_{m}$ just under $t_{m-1}$
\item \label{s6} we move the empty tile to the left of $t_{m-1}$
\item \label{s7} finally we move the empty tile right and then down to put $t_{m-1}$ and $t_{m}$ in place.
\end{titlemize}
In order to move the tiles, we have written a few simple routines which can move the empty tile from its current position next to (above, below, left or right) any tile, and then can move that tile to another position, all the while avoiding to go through previously moved tiles (hence the particular order in which we move the different tiles above). The only case where the above algorithm can get stuck is when both n and m are equal to 3 and that by step \ref{s6} we end up with $t_{3}$ under the empty tile. We have handcrafted a sequence of moves to solve this particular position. Other than this one particular case, the above naive algorithm is guaranteed to succeed (and is obviously quite fast in terms of run time, though not elegant).
\\
\\
As a concrete example, let us assume we started with the following (n=6, m=6) puzzle:
\\
\begin{thityfive}
\setrow{6}{14,27,6,  2,5,18}
\setrow{5}{21,29,13,23,35,30}
\setrow{4}{26,3,7,9,24,19}
\setrow{3}{22,12,11,17,16,33}
\setrow{2}{32,10,20,25,34,28}
\setrow{1}{8,4,15,31, ,1}
\end{thityfive}
\\
After one call to solve the top row and the left column, we are left with solving the (n=5,m=5) sub-puzzle in blue:
\\
\begin{thityfive}
\setrow{6}{1,2,3,4,5,6}
\setrow{5}{7,\blue9,\blue17,\blue27,\blue18,\blue35}
\setrow{4}{8,\blue23,\blue11,\blue15,\blue24,\blue21}
\setrow{3}{9,\blue20,\blue8,\blue29,\blue33,\blue10}
\setrow{2}{10,\blue22,\blue30,\blue14,\blue32,\blue16}
\setrow{1}{11,\blue ,\blue12,\blue26,\blue34,\blue28}
\end{thityfive}
\\
Let us now detail how the naive algorithm will solve the top row if that sub-puzzle:
\\
\begin{twentyfour}
\setrow{5}{9,17,27,18,35}
\setrow{4}{23,11,15,24,21}
\setrow{3}{20,8,29,33,10}
\setrow{2}{22,30,14,32,16}
\setrow{1}{ ,12,26,34,28}
\end{twentyfour}
\\
step \ref{s1} above will decide to solve the top row by placing $t_{1}, ..., t_{5}$ = $\blue 8, 9, 10, 11, 12$ in that order as the top row. Steps \ref{s2} to \ref{s7} will yield in order:
\\
\begin{twentyfour}
\setrow{5}{\blue9,17,27,18,35}
\setrow{4}{23,\blue11,15,24,21}
\setrow{3}{20,\blue8,29,33,\blue10}
\setrow{2}{22,30,14,32,16}
\setrow{1}{ ,\blue12,26,34,28}
\end{twentyfour}
%
\begin{twentyfour}
\setrow{5}{\blue9,17,27,18,35}
\setrow{4}{23,\blue11,15,24,21}
\setrow{3}{20,\blue8,29,33,\blue10}
\setrow{2}{22,30,14,32,16}
\setrow{1}{\color{red}12\color{red}, ,26,34,28}
\end{twentyfour}
%
\begin{twentyfour}
\setrow{5}{\red 8,\red 9,\red 10,,18}
\setrow{4}{17,15,27,24,35}
\setrow{3}{\blue11,23,29,21,33}
\setrow{2}{20,22,14,32,16}
\setrow{1}{\red 12,30,26,34,28}
\end{twentyfour}
\\
\begin{twentyfour}
\setrow{5}{\red 8, \red 9, \red 10,,\red 11}
\setrow{4}{23,29,21,18,24}
\setrow{3}{17,15,27,33,35}
\setrow{2}{20,22,14,32,16}
\setrow{1}{\red 12,30,26,34,28}
\end{twentyfour}
%
\begin{twentyfour}
\setrow{5}{\red 8,\red 9,\red 10,18,\red 11}
\setrow{4}{29,27,32, ,\red 12}
\setrow{3}{23,21,33,35,24}
\setrow{2}{15,17,22,14,16}
\setrow{1}{30,20,26,34,28}
\end{twentyfour}
%
\begin{twentyfour}
\setrow{5}{\red 8, \red 9, \red 10, \red 11, \red 12}
\setrow{4}{29,27,32,18, }
\setrow{3}{23,21,33,35,24}
\setrow{2}{15,17,22,14,16}
\setrow{1}{30,20,26,34,28}
\end{twentyfour}
\\
and we are left with solving the bottom sub-puzzle (n=4,m=5):
\\
\begin{twenty}
\setrow{4}{29,27,32,18, }
\setrow{3}{23,21,33,35,24}
\setrow{2}{15,17,22,14,16}
\setrow{1}{30,20,26,34,28}
\end{twenty}

%-----------------------------------
%	SUBSECTION 2.3
%-----------------------------------
\subsection{Kociemba}
TBD

%-----------------------------------
%	SUBSECTION 2.4
%-----------------------------------
\subsection{A*}
\label{ASSS}

\textbf{manhattan heuristic}

\textbf{perfect heuristic}

\textbf{deep learning heuristic}

\textbf{deep reinforcement learning heuristic}

### 2022-07-23 latex/Chapters/Puzzles.tex
update

@@
 
-The first puzzle I will focus on is the sliding puzzle (see \cite{SlidingPuzzleWiki}).The 15-puzzle seems to have been invented in the late 19th century by Noyes Chapman (see \cite{SlidingPuzzleWolfram}), who applied for a patent in 1980.
+The first puzzle I will focus on is the sliding puzzle (see \cite{SlidingPuzzleWiki}).The 15-puzzle seems to have been invented in the late 19th century by Noyes Chapman (see \cite{SlidingPuzzleWolfram}), who applied in 1880 for a patent on what was then called "Block Solitaire Puzzle".
 In 1879, a couple of interesting notes (\cite{Johnson1879}), published in the American Journal of Mathematics proved that exactly half of the $16!$ possible ways of setting up the 16 tiles on the board lead to solvable puzzles.
@@
 \setrow{3}{5,10,7,4}
-\setrow{2}{9,15,14,11}
-\setrow{1}{13,12,,8}
+\setrow{2}{9,15,\red14\black,\blue11}
+\setrow{1}{13,12,,\blue 8}
 \end{fifteen}
@@
 \\
-we have p(c) = (1, 6, 2, 3, 4, 7, 10, 5, 9, 15, 14, 11, 8, 12, 13).
+we have p(c) = (1, 6, 2, 3, 4, 7, 10, 5, 9, 15, \red 14 \black, \blue 11, 8\black, 12, 13).
 \\
@@
 A vertical move of a tile will displace a number in p(c) by an even number of positions right or left.
-For instance, moving tile 14 into the empty compartment above it results in a new configuration $c_{2}$:
+For instance, moving tile 14 into the empty compartment below it results in a new configuration $c_{2}$:
 
@@
 \setrow{3}{5,10,7,4}
-\setrow{2}{9,15,,11}
-\setrow{1}{13,12,14,8}
+\setrow{2}{9,15,,\blue11}
+\setrow{1}{13,12,\red14\black,\blue8}
 \end{fifteen}
@@
 \\
-with $p(c_{2})$ = (1, 6, 2, 3, 4, 7, 10, 5, 9, 15, 11, 8, 14, 12, 13), which is equivalent to moving 14 by 2 positions on the right.
+with $p(c_{2})$ = (1, 6, 2, 3, 4, 7, 10, 5, 9, 15, \blue11, 8\black, \red14\black, 12, 13), which is equivalent to moving 14 by 2 positions on the right.
 This obviously cannot change the parity since exactly 2 pairs of numbers are now in a different order, that is (14, 11) and (14, 8) now appear in the respective opposite orders as (11, 14) and (8, 14).
@@
 \begin{fifteen}
-\setrow{4}{1,2,3,4}
+\centering
+\setrow{4}{\red1\black,2,3,4}
 \setrow{3}{5,6,7,8}
@@
 \setrow{2}{9,10,11,12}
-\setrow{1}{13,14,15,1}
+\setrow{1}{13,14,15,\red1\black}
 \end{fifteen}
@@
 \\
-Since it is possible to reach an odd parity configuration, and by symmetry arguments, we conclude we can solve all $16!$ configurations.
+Since it is possible to reach an odd parity configuration, we conclude by invoking symmetry arguments that we can solve all $16!$ configurations.
 

### 2022-07-23 latex/MScAIFrancoisBerrier.tex
update

@@
 \usepackage{float}
+\usepackage{graphbox}
 \usepackage{lscape}
@@
 \usepackage{logicpuzzle}
+\usepackage{geometry}
+\newenvironment{three}[1][]{%
+\begin{logicpuzzle}[rows=2,columns=2,#1]
+\begin{puzzleforeground}
+\framepuzzle
+\end{puzzleforeground}
+}{\end{logicpuzzle}}
+
+\newenvironment{five}[1][]{%
+\begin{logicpuzzle}[rows=2,columns=3,#1]
+\begin{puzzleforeground}
+\framepuzzle
+\end{puzzleforeground}
+}{\end{logicpuzzle}}
+
+\newenvironment{seven}[1][]{%
+\begin{logicpuzzle}[rows=2,columns=4,#1]
+\begin{puzzleforeground}
+\framepuzzle
+\end{puzzleforeground}
+}{\end{logicpuzzle}}
+
+\newenvironment{nine}[1][]{%
+\begin{logicpuzzle}[rows=2,columns=5,#1]
+\begin{puzzleforeground}
+\framepuzzle
+\end{puzzleforeground}
+}{\end{logicpuzzle}}
+
+\newenvironment{eight}[1][]{%
+\begin{logicpuzzle}[rows=3,columns=3,#1]
+\begin{puzzleforeground}
+\framepuzzle
+\end{puzzleforeground}
+}{\end{logicpuzzle}}
+
 \newenvironment{fifteen}[1][]{%
@@
 }{\end{logicpuzzle}}
+
+\newenvironment{twenty}[1][]{%
+\begin{logicpuzzle}[rows=4,columns=5,#1]
+\begin{puzzleforeground}
+\framepuzzle
+\end{puzzleforeground}
+}{\end{logicpuzzle}}
+
+\newenvironment{twentyfour}[1][]{%
+\begin{logicpuzzle}[rows=5,columns=5,#1]
+\begin{puzzleforeground}
+\framepuzzle
+\end{puzzleforeground}
+}{\end{logicpuzzle}}
+
+\newenvironment{thityfive}[1][]{%
+\begin{logicpuzzle}[rows=6,columns=6,#1]
+\begin{puzzleforeground}
+\framepuzzle
+\end{puzzleforeground}
+}{\end{logicpuzzle}}
+
+\newcommand{\red}{\color{red}}
+\newcommand{\black}{\color{black}}
+\newcommand{\blue}{\color{blue}}
+\newcommand{\green}{\color{green}}
+
+\newenvironment{titlemize}[1]{%
+  \paragraph{#1}
+  \begin{enumerate}}
+  {\end{enumerate}}
 
@@
 I might throw a few random (but hopefully interesting) observations in there and give some references for the keen reader.
-I will then give an overview in chapter 4 of the code base I have developed - and put in the open on my github page (\cite{FB}) - to complete this project.
-Finally, in chapters 5 and 6, I will present all my various results on respecively the sliding puzzle and the Rubiks' cube.
+I will then give an overview in chapter 4 of the code base I have developed - and put in the open on my github page (\cite{FB}) - to complete this project before detailing a few examples in chapter 5.
+Finally, in chapters 6 and 7, I will present all my various results on respectively the sliding puzzle and the Rubiks' cube.
 
@@
 \include{Chapters/Code}
+\include{Chapters/Examples}
 \include{Chapters/SlidingPuzzle}

### 2022-07-24 latex/Chapters/Code.tex
update

@@
 The code is organised in several sub modules and makes use of factories in plenty of places so that I can easily try out different puzzles, dimensions, search techniques, heuristics, network architecture, etc...
-without having to change anything but configuration or parameter in the command line.
-Here is a visual overview of the code base with the main dependencies between the main submodules and classes:
+without having to change anything except the parameters passed in the command line.
+Here is a visual overview of the code base with the main dependencies between the main submodules and classes.
+Solid arrows indicate inheritance (e.g.
+AStar inherits from SearchStrategy), while dotted lines indicate usage (e.g.
+AStar uses Heuristic, DeepReinforcementLearner uses DeepLearning, etc..).
 
@@
 \centering
-\includegraphics[scale=0.5]{./Figures/codebase}
+\includegraphics[scale=0.49]{./Figures/codebase}
 %\decoRule
@@
 \section{rubiks.core}
+\begin{figure}[H]
+\centering
+\includegraphics[scale=0.25]{./Figures/codebasecore}
+%\decoRule
+\caption[Codebase]{rubiks.core}
+\label{fig:Codebasecore}
+\end{figure}
 This submodule contains base classes that make the code base easier to use, debug, and extend. It contains the following:
 \begin{itemize}
-\item Loggable: a wrapper around Python's logger which automatically picks up classes' names at init and format things (dict, series and dataframes in particular) in a nicer way.
-\item Parsable: a wrapper around ArgumentParser, which allows to construct objects in the project from command line, to define dependencies between object's configurations and to help a bit with typing of configs.
+\item \textbf{Loggable}: a wrapper around Python's logger which automatically picks up classes' names at init and format things (dict, series and dataframes in particular) in a nicer way.
+\item \textbf{Parsable}: a wrapper around ArgumentParser, which allows to construct objects in the project from command line, to define dependencies between object's configurations and to help a bit with typing of configs.
 The end result is that you can pretty much pass **kw\_args everywhere and it just works.
-\item Factory: a typical factory pattern.
+\item \textbf{Factory}: a typical factory pattern.
 Concrete factories can just define what widget they produce and the factory will help construct them from **kw\_args (or command line, since Factory inherits from Parsable)
@@
 \section{rubiks.puzzle}
+\begin{figure}[H]
+\centering
+\includegraphics[scale=0.25]{./Figures/codebasepuzzle}
+%\decoRule
+\caption[Codebase]{rubiks.puzzle}
+\label{fig:Codebasepuzzle}
+\end{figure}
 This submodule contains:
 \begin{itemize}
-\item Puzzle: a Factory of puzzles.
+\item \textbf{Puzzle}: a Factory of puzzles.
 It defines states and actions in the abstract, and provides useful functions to apply moves, shuffle, generate training sets, tell if a state is the goal, etc.
 Puzzle can manufacture the two following types of puzzles:
-\item SlidingPuzzle. Implements the states and moves of the sliding puzzle.
-\item RubiksCube. Implements the states and moves of the Rubik's cube.
-In addition, in contains a Puzzled base class which most below inherit from.
+\item \textbf{SlidingPuzzle}. Implements the states and moves of the sliding puzzle.
+\item \textbf{RubiksCube}. Implements the states and moves of the Rubik's cube.
+\\
+\\
+In addition, this module contains a \textbf{Puzzled} base class which most classes below inherit from.
 That allow e.g.
@@
 \section{rubiks.search}
-This modules contains graph search strategies. I have actually reused the code I implemented for one of the AIPnT assignments. It contains the following classes:
+\begin{figure}[H]
+\centering
+\includegraphics[scale=0.25]{./Figures/codebasesearch}
+%\decoRule
+\caption[Codebase]{rubiks.search}
+\label{fig:Codebasesearch}
+\end{figure}
+This modules contains graph search strategies. I have actually reused the code I implemented for one of the AIPnT assignments here. It contains the following classes:
 \begin{itemize}
-\item Node: which contains the state of a graph, as well as link to the previous (parent) state, action that leads from the latter to the former and the cost of the path so far.
-\item SearchStrategy, a Factory class which can instantiate the following three types of search strategies to find a path to a goal:
-\item BreadthFirstSearch, which is obviously an optimal strategy, but not particularly efficient.
-\item DepthFirstSearch, which is not an optimal strategy, and also generally not particularly efficient.
-\item AStar, which is optimal, and as efficient as the heuristic it makes use of is.
+\item \textbf{Node}: which contains the state of a graph, as well as link to the previous (parent) state, action that leads from the latter to the former and the cost of the path so far.
+\item \textbf{SearchStrategy}, a Factory class which can instantiate the following three types of search strategies to find a path to a goal:
+\item \textbf{BreadthFirstSearch}, which is obviously an optimal strategy, but not particularly efficient.
+\item \textbf{DepthFirstSearch}, which is not an optimal strategy, and also generally not particularly efficient.
+\item \textbf{AStar}, which is optimal, and as efficient as the heuristic it makes use of is.
 \end{itemize}
@@
 \section{rubiks.heuristics}
+\begin{figure}[H]
+\centering
+\includegraphics[scale=0.25]{./Figures/codebaseheuristics}
+%\decoRule
+\caption[Codebase]{rubiks.heuristics}
+\label{fig:Codebaseheuristics}
+\end{figure}
 \label{HSS}
@@
 \begin{itemize}
-\item Manhattan: at current time of writing, this is specific to the SP and will be discussed in more details in \ref{S33},
-\item PerfectHeuristic: this reads from a data base the optimal costs, pre-computed by the PerfectLearner (see below \ref{PLcode})
-\item DeepLearningHeuristic: this uses a network which has been trained using DRL by the DeepReinforcementLearner (see below \ref{DRLcode})
+\item \textbf{Manhattan}: at current time of writing, this is specific to the SP and will be discussed in more details in \ref{S33},
+\item \textbf{PerfectHeuristic}: this reads from a data base the optimal costs, pre-computed by the PerfectLearner (see below \ref{PLcode})
+\item \textbf{DeepLearningHeuristic}: this uses a network which has been trained using DRL by the DeepReinforcementLearner (see below \ref{DRLcode})
 \end{itemize}
@@
 \section{rubiks.deeplearning}
+\begin{figure}[H]
+\centering
+\includegraphics[scale=0.25]{./Figures/codebasedeeplearning}
+%\decoRule
+\caption[Codebase]{rubiks.deeplearning}
+\label{fig:Codebasedeeplearning}
+\end{figure}
 This module is a wrapper around Pytorch. It contains:
 \begin{itemize}
-\item DeepLearning: a Puzzled Loggable Factory that can instantiate some configurable deep networks, and provide the necessary glue with the rest of the code base so that puzzles be seemlessly passed to the networks and trained on.
-\item FullyConnected: wrapper around a Pytorch fully connected network, with configurable layers and size.
-\item Convolutional: tdb
+\item \textbf{DeepLearning}: a Puzzled Loggable Factory that can instantiate some configurable deep networks, and provide the necessary glue with the rest of the code base so that puzzles be seemlessly passed to the networks and trained on.
+\item \textbf{FullyConnected}: wrapper around a Pytorch fully connected network, with configurable layers and size.
+\item \textbf{Convolutional}: tdb
 \end{itemize}
@@
 \section{rubiks.learners}
+\begin{figure}[H]
+\centering
+\includegraphics[scale=0.25]{./Figures/codebaselearners}
+%\decoRule
+\caption[Codebase]{rubiks.learners}
+\label{fig:Codebaselearners}
+\end{figure}
 \label{PLcode}
@@
 \begin{itemize}
-\item Learner is a Puzzled Loggable Factory.
+\item \textbf{Learner} is a Puzzled Loggable Factory.
 It provides some common code to learners (to save or purge what they learnt), kick off learning and plot results.
@@
 Currently the two implemented learners are:
-\item PerfectLearner: It instantiates an optimal solver ($A^{*}$ with a configurable heuristic - but will only accept heuristic that advertise themselves as optimal.
+\item \textbf{PerfectLearner}: It instantiates an optimal solver ($A^{*}$ with a configurable heuristic - but will only accept heuristic that advertise themselves as optimal.
 The learning consists in generating all the possible configuration of the considered puzzle, solve them with the optimal solver, and save the optimal cost of it as well as those of the whole solution path.
@@
 
-\item DeepLearner tbd
-\item DeepReinforcementLearner: It instantiates a DeepLearning (network), and trains it using DRL.
+\item \textbf{DeepLearner} tbd
+\item \textbf{DeepReinforcementLearner}: It instantiates a DeepLearning (network), and trains it using DRL.
 It then saves the trained network, which can then be used in the DeepLearningHeuristic we have seen earlier in section \ref{HSS}.
@@
 \section{rubiks.solvers}
+\begin{figure}[H]
+\centering
+\includegraphics[scale=0.25]{./Figures/codebasesolvers}
+%\decoRule
+\caption[Codebase]{rubiks.solvers}
+\label{fig:Codebasesolvers}
+\end{figure}
 This module implements solvers, which solve puzzles.
@@
 \begin{itemize}
-\item DFSSolver
-\item BFSSolver
-\item AStarSolver
-\item NaiveSlidingSolver
+\item \textbf{DFSSolver} TBD
+\item \textbf{BFSSolver} TBD
+\item \textbf{AStarSolver} TBD
+\item \textbf{NaiveSlidingSolver} TBD
+\item \textbf{MonteCarloSearchTreeSolver} TBI I want to implement this in August when working on the Rubiks' ...
+following the Rubik's paper fromMcAleer \& Agostilenni et al \cite{https://doi.org/10.48550/arxiv.1805.07470}
 \end{itemize}
@@
 \section{rubiks.scripts}
+
+\begin{figure}[H]
+\centering
+\includegraphics[scale=0.25]{./Figures/codebasescripts}
+%\decoRule
+\caption[Codebase]{rubiks.scripts}
+\label{fig:Codebasescripts}
+\end{figure}
 

### 2022-07-24 latex/Chapters/Examples.tex
update

@@
 
+In this chapter, I will go through some examples, illustrating how to use the code base to learn and solve various puzzles.
+For each section, simple examples of code will be indicated in air force blue \afblue \textbf{python code} \black paragraphs, and can easily be run from command line or copied into a script and run from your favourite IDE.
+
 %-----------------------------------
@@
 \section{Learners}
-
-Here I show how to run some examples... TBC
 
@@
 \subsection{Naive Sliding Puzzle Solver}
-As a comparison point, I have implemented a naive sliding puzzle solver, which does what most beginner players would intuitively do when solving the sliding puzzle by hand: solve the top row, then the left column, and keep iterating until done.
+As a comparison point for other sliding puzzle solvers, I have implemented a naive sliding puzzle solver, which does what most beginner players would intuitively do when solving the sliding puzzle by hand: solve the top row, then the left column, and keep iterating until done.
 Notice that once either the top row or left column is solved, there is no longer any need to modify it, we have simply reduced the problem to a sub-problem of reduced dimension.
@@
 Solving the top row of a n by m puzzle (left column is similar, mutatis mutandis, so I will not detail it) is accomplished as follows:
-\begin{titlemize}{naive algorithm - top-row solver}
+\\
+\textbf{naive algorithm - top-row solver}
+\begin{enumerate}
 \item \label{s1} we sort the tiles (which since we are potentially dealing with a sub-problem, are not necessarily 1 to $m* n - 1$), and select the m smaller ones $t_{1}, ..., t_{m-1}, t_{m}$.
@@
 \item \label{s7} finally we move the empty tile right and then down to put $t_{m-1}$ and $t_{m}$ in place.
-\end{titlemize}
+\end{enumerate}
 In order to move the tiles, we have written a few simple routines which can move the empty tile from its current position next to (above, below, left or right) any tile, and then can move that tile to another position, all the while avoiding to go through previously moved tiles (hence the particular order in which we move the different tiles above).
@@
 \\
+\\
 After one call to solve the top row and the left column, we are left with solving the (n=5,m=5) sub-puzzle in blue:
@@
 \\
-Let us now detail how the naive algorithm will solve the top row if that sub-puzzle:
+\\
+Let us now detail how the \textbf{naive algorithm} will solve the top row if that sub-puzzle:
 \\
@@
 \end{twentyfour}
+\\
 \\
@@
 \\
+\\
 and we are left with solving the bottom sub-puzzle (n=4,m=5):
@@
 \end{twenty}
+
+\afblue
+\paragraph{}{\textbf{python code -- naive solver}}
+\begin{python}
+################################################
+from math import inf
+from rubiks.puzzle.puzzle import Puzzle
+from rubiks.solvers.solver import Solver
+################################################
+action_type=Solver.do_solve
+n=2
+puzzle_type=Puzzle.sliding_puzzle
+solver_type=Solver.naive
+nb_shuffles=inf
+################################################
+print(Solver.factory(**globals()).action())
+################################################
+\end{python}
+\black
+
+\begin{figure}[H]
+\centering
+\includegraphics[scale=0.39]{./Figures/examplenaivesolver}
+%\decoRule
+\caption[Examples]{naive solver}
+\label{fig:examplenaivesolver}
+\end{figure}
 

### 2022-07-24 rubiks/scripts/learner.py

@@ -23,8 +23,8 @@ if '__main__' == __name__:
     action_type = Learner.do_learn
     """ What puzzle """
     puzzle_type = Puzzle.sliding_puzzle
-    n = 4
-    m = 4
+    n = 5
+    m = 5
     dimension = Puzzle.factory(**globals()).dimension()
     """ Which learner_type {Learner.perfect,
                             Learner.deep_reinforcement_learner,
@@ -58,13 +58,13 @@ if '__main__' == __name__:
     plus = True
     """ If it's a DRL learner config is here ... """
     nb_epochs = 100000
-    nb_sequences = 100
-    nb_shuffles = 160
+    nb_sequences = 50
+    nb_shuffles = 400
     training_data_every_epoch = False
     cap_target_at_network_count = True
-    update_target_network_frequency = 1000
+    update_target_network_frequency = 500
     update_target_network_threshold = 1e-3
-    max_nb_target_network_update = 100
+    max_nb_target_network_update = 300
     max_target_not_increasing_epochs_pct = 0.5
     max_target_uptick = 0.01
     learning_rate = 1e-3

### 2022-07-24 latex/Chapters/DRLSearch.tex
update

@@
 \label{Chapter1} % Change X to a consecutive number; for referencing this chapter elsewhere, use \ref{ChapterX}
+
+TBD this is a bit of a verbiage chapter, where I go through the different techniques I have used and implemented for this project and give some references.
 
@@
 
-blabla
+see \cite{Sutton1998}
 
@@
 
-see \cite{Watkins1992}
+see \cite{Watkins1992} and \cite{https://doi.org/10.48550/arxiv.1805.07470}
+
+%----------------------------------------------------------------------------------------
+%	SECTION 6
+%----------------------------------------------------------------------------------------
+
+\section{Monte Carlo Tree Search}
+
+see \cite{https://doi.org/10.48550/arxiv.1805.07470} and \cite{AlphaGo}

### 2022-07-24 latex/Chapters/Examples.tex
update

@@
 In this chapter, I will go through some examples, illustrating how to use the code base to learn and solve various puzzles.
-For each section, simple examples of code will be indicated in air force blue \afblue \textbf{python code} \black paragraphs, and can easily be run from command line or copied into a script and run from your favourite IDE.
+For each section, simple examples of code will be indicated in \afblue \textbf{python code} \black paragraphs, and can easily be run from command line or copied into a script and run from your favourite IDE.
 
@@
 
+The perfect learner has been discussed in details in \ref{sec:codelearners}.
+We simply show here how to run it to learn the value function for the (n=2, m=3) SP, that is, to solve all 360 possible configurations via an optimal solver (A{*} with Manhattan heuristic).
+
+\afblue
+\paragraph{}{\textbf{python code -- perfect learner}}
+\begin{python}
+####################################################################
+from rubiks.heuristics.heuristic import Heuristic
+from rubiks.puzzle.puzzle import Puzzle
+from rubiks.learners.learner import Learner
+from rubiks.utils.utils import get_model_file_name
+####################################################################
+action_type=Learner.do_learn
+n=2
+m=3
+puzzle_type=Puzzle.sliding_puzzle
+learner_type=Learner.perfect_learner
+heuristic_type=Heuristic.manhattan
+nb_cpus=4
+learning_file_name=get_model_file_name(puzzle_type=puzzle_type,
+                                       dimension=(n, m),
+                                       model_name=Learner.perfect)
+if __name__ == '__main__':
+    # we fully solve the 2 x 3 SP ... should take ~5s
+    Learner.factory(**globals()).action()
+####################################################################
+action_type=Learner.do_plot
+if __name__ == '__main__':
+    # we display the results
+    Learner.factory(**globals()).action()
+####################################################################
+\end{python}
+\black
+The above snipet of code will solve the 2 by 3 SP and then display the results, showing the distribution of optimal costs, as well as the most difficult puzzle.
+
+\begin{figure}[H]
+\centering
+\includegraphics[scale=0.45]{./Figures/exampleperfectlearner}
+%\decoRule
+\caption[Examples]{perfect learner example}
+\label{fig:exampleperfectlearner}
+\end{figure}
+
 %-----------------------------------
@@
 \label{DLSS}
-blabla
+The DeepLearner, also discussed in details in \ref{sec:codelearners}, needs some training data.
+It could in principle of course be trained on any data, not necessarily optimal data (i.e.
+generated by an optimal solver).
+Here however, I make use of the TrainingData class to generate 100 sequences of fully solved (n=5, m=2) SP via A{*} with Manhattan++.
+
+\afblue
+\paragraph{}{\textbf{python code -- deep learner}}
+\begin{python}
+####################################################################
+from math import inf
+####################################################################
+from rubiks.deeplearning.deeplearning import DeepLearning
+from rubiks.learners.learner import Learner
+from rubiks.learners.deeplearner import DeepLearner
+from rubiks.puzzle.puzzle import Puzzle
+from rubiks.puzzle.trainingdata import TrainingData
+####################################################################
+if '__main__' == __name__:
+    puzzle_type = Puzzle.sliding_puzzle
+    n = 5
+    m = 2
+    """ Generate training data - 100 sequences of fully 
+    solved perfectly shuffled puzzles. 
+    """
+    nb_cpus = 4
+    time_out = 600
+    nb_shuffles = inf
+    nb_sequences = 100
+    TrainingData(**globals()).generate(**globals())
+    """ DL learner """
+    action_type = Learner.do_learn
+    learner_type = Learner.deep_learner
+    nb_epochs = 1000
+    learning_rate = 1e-3
+    scheduler = DeepLearner.gamma_scheduler
+    gamma_scheduler = 0.9999
+    save_at_each_epoch = False
+    threshold = 0.01
+    training_data_freq = 100
+    high_target = nb_shuffles + 1
+    training_data_from_data_base = True
+    nb_shuffles_min = 20
+    nb_shuffles_max = 50
+    """ ... and its network config """
+    network_type = DeepLearning.fully_connected_net
+    layers_description = (100, 50, 10)
+    one_hot_encoding = True
+    """ Kick-off the Deep Learner """
+    learning_file_name = Learner.factory(**globals()).get_model_name()
+    Learner.factory(**globals()).action()
+    """ Plot its learning """
+    action_type = Learner.do_plot
+    Learner.factory(**globals()).action()
+####################################################################
+\end{python}
+\black
 
@@
 %\decoRule
-\caption[Examples]{naive solver}
+\caption[Examples]{naive solver example}
 \label{fig:examplenaivesolver}
@@
 \subsection{Kociemba}
-TBD
+TBI \& TBD
 

### 2022-07-24 rubiks/learners/deepreinforcementlearner.py

@@ -5,7 +5,6 @@ from enum import Enum
 from functools import partial
 from itertools import cycle
 from matplotlib import pyplot as plt
-from math import inf
 from multiprocessing import Pool
 from os.path import isfile
 from pandas import concat, DataFrame, Series, read_pickle
@@ -17,7 +16,7 @@ from torch.optim.lr_scheduler import ExponentialLR
 ########################################################################################################################
 from rubiks.deeplearning.deeplearning import DeepLearning
 from rubiks.learners.learner import Learner
-from rubiks.utils.utils import ms_format, h_format, pformat, to_pickle, get_model_file_name, snake_case, billion
+from rubiks.utils.utils import ms_format, h_format, pformat, to_pickle, get_model_file_name, snake_case, billion, isinf
 ########################################################################################################################
 
 
@@ -162,12 +161,18 @@ class DeepReinforcementLearner(Learner):
         return list()
 
     def get_model_name(self):
+        if isinstance(self.nb_shuffles, (int, float)) and not isinf(self.nb_shuffles):
+            nb_shuffles = ['%dshf' % int(self.nb_shuffles)]
+        elif isinf(self.nb_shuffles):
+            nb_shuffles = []
+        else:
+            assert False, 'Unknown type for nb_shuffles [%s]' % self.nb_shuffles
         details = '_'.join([snake_case(self.__class__.__name__),
                             self.optimiser,
                             self.scheduler,
                             '%dgamma' % int(self.gamma_scheduler * billion),
                             '%dseq' % int(self.nb_sequences),
-                            '%dshf' % int(self.nb_shuffles),
+                            *nb_shuffles,
                             '%depc' % int(self.nb_epochs),
                             'tng_' + ('epoch' if self.training_data_every_epoch else 'ntk_updt'),
                             *self.more_model_name(),

### 2022-07-24 rubiks/scripts/draft.py

@@ -1,11 +1,46 @@
-""" imports >>>>>> """
+####################################################################
 from math import inf
-from rubiks . puzzle . puzzle import Puzzle
-from rubiks . solvers . solver import Solver
-""" imports <<<<<< """
-action_type = Solver . do_solve
-n = 2
-puzzle_type = Puzzle . sliding_puzzle
-solver_type = Solver . naive
-nb_shuffles = inf
-print ( Solver . factory ( ** globals () ) . action () )
+####################################################################
+from rubiks.deeplearning.deeplearning import DeepLearning
+from rubiks.learners.learner import Learner
+from rubiks.learners.deeplearner import DeepLearner
+from rubiks.puzzle.puzzle import Puzzle
+from rubiks.puzzle.trainingdata import TrainingData
+####################################################################
+if '__main__' == __name__:
+    puzzle_type = Puzzle.sliding_puzzle
+    n = 5
+    m = 2
+    """ Generate training data - 100 sequences of fully 
+    solved perfectly shuffled puzzles. 
+    """
+    nb_cpus = 4
+    time_out = 600
+    nb_shuffles = inf
+    nb_sequences = 100
+    TrainingData(**globals()).generate(**globals())
+    """ DL learner """
+    action_type = Learner.do_learn
+    learner_type = Learner.deep_learner
+    nb_epochs = 1000
+    learning_rate = 1e-3
+    scheduler = DeepLearner.gamma_scheduler
+    gamma_scheduler = 0.9999
+    save_at_each_epoch = False
+    threshold = 0.01
+    training_data_freq = 100
+    high_target = nb_shuffles + 1
+    training_data_from_data_base = True
+    nb_shuffles_min = 20
+    nb_shuffles_max = 50
+    """ ... and its network config """
+    network_type = DeepLearning.fully_connected_net
+    layers_description = (100, 50, 10)
+    one_hot_encoding = True
+    """ Kick-off the Deep Learner """
+    learning_file_name = Learner.factory(**globals()).get_model_name()
+    Learner.factory(**globals()).action()
+    """ Plot its learning """
+    action_type = Learner.do_plot
+    Learner.factory(**globals()).action()
+####################################################################

### 2022-07-24 latex/Chapters/Examples.tex
update

@@
     learner_type = Learner.deep_learner
-    nb_epochs = 1000
+    nb_epochs = 999
     learning_rate = 1e-3
@@
     nb_shuffles_max = 50
+    nb_sequences = 50
     """ ... and its network config """
@@
     one_hot_encoding = True
-    """ Kick-off the Deep Learner """
+    """ Kick-off the Deep Learner ... """
     learning_file_name = Learner.factory(**globals()).get_model_name()
     Learner.factory(**globals()).action()
-    """ Plot its learning """
+    """... and plot its learning data """
     action_type = Learner.do_plot
@@
 \black
+
+As can be seen in the code snipet, this example will generate 100 perfectly shuffled (n=5, m=2) SPs and solve them.
+Once done a summary of the training data is printed, indicating, for each optimal cost, how many sequences have been generated.
+
+\begin{figure}[H]
+\centering
+\includegraphics[scale=0.45]{./Figures/exampledeeplearnertraining}
+%\decoRule
+\caption[Examples]{deep learner training example}
+\label{fig:exampledeeplearnertraining}
+\end{figure}
+
+Then the Deep Learner will get trained on this data for 999 epochs.
+In the above example, I have chosen to fetch, every 100 epochs, 50 random sequences of puzzles from the training data.
+Each sequence is composed of a random puzzle of cost between 20 and 50, fetched from the training data, along with its optimal path to the solution.
+The default optimiser (rms\_prop) is used, together with an exponential scheduler starting with a learning rate of 0.001 and a gamma of 0.9999.
+We can see that the (MSE) loss on the value function decreases rapidly, and jumps back up every time we change the training data (since it has not yet seen some of it).
+By the end of the training, the in-sample MSE loss has dropped to 1.5 and the Deep Learner has seen 0.14\% of the possible (n=5, m=2) puzzles.
+
+\newgeometry{top=0mm, bottom=0mm, left=0mm, right=0mm}
+\begin{landscape}
+\centering\vspace*{\fill}
+\begin{figure}[H]
+\centering
+\includegraphics[scale=0.45]{./Figures/exampledeeplearnerlearning}
+%\decoRule
+\caption[Examples]{deep learner learning data example}
+\label{fig:exampledeeplearnerlearning}
+\end{figure}
+\vfill
+\end{landscape}
+\restoregeometry
 
@@
 \end{twenty}
-
+\\
+\\
+which the naive solver can keep solving iteratively by taking care of the left-most column, etc...
+\\
+\\
+Below is a simple code snipet to run the naive solver on a randomly generated (n=2, m=2) SP:
 \afblue

### 2022-07-24 latex/Chapters/Examples.tex
update

@@
     learning_rate = 1e-3
-    scheduler = DeepLearner.gamma_scheduler
+    optimiser = DeepLearner.rms_prop
+    scheduler = DeepLearner.exponential_scheduler
     gamma_scheduler = 0.9999
@@
     one_hot_encoding = True
-    """ Kick-off the Deep Learner ... """
+    """ Kick-off the Deep Learner """
     learning_file_name = Learner.factory(**globals()).get_model_name()
     Learner.factory(**globals()).action()
-    """... and plot its learning data """
+    """ Plot its learning """
     action_type = Learner.do_plot
@@
 As can be seen in the code snipet, this example will generate 100 perfectly shuffled (n=5, m=2) SPs and solve them.
-Once done a summary of the training data is printed, indicating, for each optimal cost, how many sequences have been generated.
+Once done, a summary of the training data is printed, indicating, for each optimal cost, how many sequences have been generated.
 
@@
 
-blabla
+Unlike the Deep Learner, the Deep Reinforcement Learner learns unsupervised (in the sense that there is no need to pre-solve puzzles to tell if what the desired costs are), since it generates its own target using a combination of a target network and the simple min rule described in \ref{sec:codelearners}.
+Let us see here how to run it on the (n=4, m=2) SP for instance.
+The following code snipet will run a Deep Reinforcement Learner for a maximum of 25,000 epochs, generating randomly 10 sequences of puzzles shuffled 50 times from goal state every time it updates the target network.
+That update will happen either after 500 epochs or when the (MSE) loss on the value function gets under one thousands of the max target value.
+The learner will stop if it reaches 25,000 epochs or if the target network updates 40 times.
+The network trained is a fully connected network with 3 hidden layers and the puzzles are one-hot encoded.
+We use the same optimiser and scheduler as in the previous subsection.
+
+\afblue
+\paragraph{}{\textbf{python code -- deep reinforcement learner}}
+\begin{python}
+####################################################################
+from rubiks.deeplearning.deeplearning import DeepLearning
+from rubiks.learners.learner import Learner
+from rubiks.learners.deepreinforcementlearner import DeepReinforcementLearner
+from rubiks.puzzle.puzzle import Puzzle
+####################################################################
+if '__main__' == __name__:
+    puzzle_type = Puzzle.sliding_puzzle
+    n = 4
+    m = 2
+    """ Generate training data - 100 sequences of fully 
+    solved perfectly shuffled puzzles. 
+    """
+    nb_cpus = 4
+    """ DRL learner """
+    action_type = Learner.do_learn
+    learner_type = Learner.deep_reinforcement_learner
+    nb_epochs = 25000
+    nb_shuffles = 50
+    nb_sequences = 10
+    training_data_every_epoch = False
+    cap_target_at_network_count = True
+    update_target_network_frequency = 500
+    update_target_network_threshold = 1e-3
+    max_nb_target_network_update = 40
+    max_target_not_increasing_epochs_pct = 0.5
+    max_target_uptick = 0.01
+    learning_rate = 1e-3
+    scheduler = DeepReinforcementLearner.exponential_scheduler
+    gamma_scheduler = 0.9999
+    """ ... and its network config """
+    network_type = DeepLearning.fully_connected_net
+    layers_description = (128, 64, 32)
+    one_hot_encoding = True
+    """ Kick-off the Deep Reinforcement Learner ... """
+    learning_file_name = Learner.factory(**globals()).get_model_name()
+    Learner.factory(**globals()).action()
+    """ ... and plot its learning data """
+    action_type = Learner.do_plot
+    Learner.factory(**globals()).action()
+####################################################################
+\end{python}
+\black
+As can be seen on the next page, which I obtained from running the above code snipet (keep in mind that every run is going to be slightly different due to the random puzzles being generated), the training stopped after about 17,100 epochs as the 40 target network updates had been reached.
+By that point, the DRL learner had seen 40\% of the possible puzzles, and the very last MSE loss (after update, so out-of-sample) was around 0.4, corresponding to 2\% of the max target cost.
+It is interesting to notice that since I have shuffled the sequences only 50 times, and since the (n=4, m=2) is quite constrained in terms of possible moves, the max target ever produced by the network was only around 25, whereas we know the the God number of this dimension is 36.
+It is therefore likely that the resulting network would not produce very optimal solutions for puzzles whose cost is in the region [25, 36].
+
+\newgeometry{top=0mm, bottom=0mm, left=0mm, right=0mm}
+\begin{landscape}
+\centering\vspace*{\fill}
+\begin{figure}[H]
+\centering
+\includegraphics[scale=0.55]{./Figures/exampledeepreinforcementlearnerlearning}
+%\decoRule
+\caption[Examples]{deep reinforcement learner learning data example}
+\label{fig:exampledeepreinforcementlearnerlearning}
+\end{figure}
+\vfill
+\end{landscape}
+\restoregeometry
 

### 2022-07-24 latex/Chapters/Examples.tex
update

@@
 By that point, the DRL learner had seen 40\% of the possible puzzles, and the very last MSE loss (after update, so out-of-sample) was around 0.4, corresponding to 2\% of the max target cost.
-It is interesting to notice that since I have shuffled the sequences only 50 times, and since the (n=4, m=2) is quite constrained in terms of possible moves, the max target ever produced by the network was only around 25, whereas we know the the God number of this dimension is 36.
+It is interesting to notice that since I have shuffled the sequences only 50 times, and since the (n=4, m=2) SP is quite constrained in terms of possible moves, the max target ever produced by the network was only around 25, whereas we know the God number for this dimension is 36 (see later section \ref{sec:SPLowDimension}).
 It is therefore likely that the resulting network would not produce very optimal solutions for puzzles whose cost is in the region [25, 36].

### 2022-07-30 latex/Chapters/Examples.tex

@@
 if '__main__' == __name__:
-    puzzle_type = Puzzle.sliding_puzzle
-    n = 5
-    m = 2
+    puzzle_type=Puzzle.sliding_puzzle
+    n=5
+    m=2
     """ Generate training data - 100 sequences of fully 
@@
     """
-    nb_cpus = 4
-    time_out = 600
-    nb_shuffles = inf
-    nb_sequences = 100
+    nb_cpus=4
+    time_out=600
+    nb_shuffles=inf
+    nb_sequences=100
     TrainingData(**globals()).generate(**globals())
     """ DL learner """
-    action_type = Learner.do_learn
-    learner_type = Learner.deep_learner
-    nb_epochs = 999
-    learning_rate = 1e-3
-    optimiser = DeepLearner.rms_prop
-    scheduler = DeepLearner.exponential_scheduler
-    gamma_scheduler = 0.9999
-    save_at_each_epoch = False
-    threshold = 0.01
-    training_data_freq = 100
-    high_target = nb_shuffles + 1
-    training_data_from_data_base = True
-    nb_shuffles_min = 20
-    nb_shuffles_max = 50
-    nb_sequences = 50
+    action_type=Learner.do_learn
+    learner_type=Learner.deep_learner
+    nb_epochs=999
+    learning_rate=1e-3
+    optimiser=DeepLearner.rms_prop
+    scheduler=DeepLearner.exponential_scheduler
+    gamma_scheduler=0.9999
+    save_at_each_epoch=False
+    threshold=0.01
+    training_data_freq=100
+    high_target=nb_shuffles + 1
+    training_data_from_data_base=True
+    nb_shuffles_min=20
+    nb_shuffles_max=50
+    nb_sequences=50
     """ ... and its network config """
-    network_type = DeepLearning.fully_connected_net
-    layers_description = (100, 50, 10)
-    one_hot_encoding = True
+    network_type=DeepLearning.fully_connected_net
+    layers_description=(100, 50, 10)
+    one_hot_encoding=True
     """ Kick-off the Deep Learner """
-    learning_file_name = Learner.factory(**globals()).get_model_name()
+    learning_file_name=Learner.factory(**globals()).get_model_name()
     Learner.factory(**globals()).action()
     """ Plot its learning """
-    action_type = Learner.do_plot
+    action_type=Learner.do_plot
     Learner.factory(**globals()).action()
@@
 if '__main__' == __name__:
-    puzzle_type = Puzzle.sliding_puzzle
-    n = 4
-    m = 2
+    puzzle_type=Puzzle.sliding_puzzle
+    n=4
+    m=2
     """ Generate training data - 100 sequences of fully 
@@
     """
-    nb_cpus = 4
+    nb_cpus=4
     """ DRL learner """
-    action_type = Learner.do_learn
-    learner_type = Learner.deep_reinforcement_learner
-    nb_epochs = 25000
-    nb_shuffles = 50
-    nb_sequences = 10
-    training_data_every_epoch = False
-    cap_target_at_network_count = True
-    update_target_network_frequency = 500
-    update_target_network_threshold = 1e-3
-    max_nb_target_network_update = 40
-    max_target_not_increasing_epochs_pct = 0.5
-    max_target_uptick = 0.01
-    learning_rate = 1e-3
-    scheduler = DeepReinforcementLearner.exponential_scheduler
-    gamma_scheduler = 0.9999
+    action_type=Learner.do_learn
+    learner_type=Learner.deep_reinforcement_learner
+    nb_epochs=25000
+    nb_shuffles=50
+    nb_sequences=10
+    training_data_every_epoch=False
+    cap_target_at_network_count=True
+    update_target_network_frequency=500
+    update_target_network_threshold=1e-3
+    max_nb_target_network_update=40
+    max_target_not_increasing_epochs_pct=0.5
+    max_target_uptick=0.01
+    learning_rate=1e-3
+    scheduler=DeepReinforcementLearner.exponential_scheduler
+    gamma_scheduler=0.9999
     """ ... and its network config """
-    network_type = DeepLearning.fully_connected_net
-    layers_description = (128, 64, 32)
-    one_hot_encoding = True
+    network_type=DeepLearning.fully_connected_net
+    layers_description=(128, 64, 32)
+    one_hot_encoding=True
     """ Kick-off the Deep Reinforcement Learner ... """
-    learning_file_name = Learner.factory(**globals()).get_model_name()
+    learning_file_name=Learner.factory(**globals()).get_model_name()
     Learner.factory(**globals()).action()
     """ ... and plot its learning data """
-    action_type = Learner.do_plot
+    action_type=Learner.do_plot
     Learner.factory(**globals()).action()
@@
 \subsection{Blind search}
+First let me start with some of the blind search algorithms, which I only have been able to use on small dimension SP.
+They quickly become too memory hungry to be practical on anything but the smallest puzzles.
 %-----------------------------------
@@
 \label{BFSSS}
+The following example shows how to use Breadth First Search to solve a (n=3, m=3) SP which we do not \textit{shuffle} too much.
+
+\afblue
+\paragraph{}{\textbf{python code -- breadth first search solver}}
+\begin{python}
+####################################################################
+from rubiks.solvers.solver import Solver
+from rubiks.puzzle.puzzle import Puzzle
+####################################################################
+if '__main__' == __name__:
+    puzzle_type = Puzzle.sliding_puzzle
+    n=3
+    m=3
+    nb_shuffles=10
+    solver_type=Solver.bfs
+    check_optimal=True
+    action_type=Solver.do_solve
+    print(Solver.factory(**globals()).action())
+####################################################################
+\end{python}
+\black
+
+As expected (since BFS is optimal), the solution found, printed by the snipet of code above, is optimal.
+The check\_optimal flag in the code snipet indicates that the solver should let us know if solution is optimal.
+Since BFS advertises itself (via the Solver base class API)  as an optimal solver, the solution is deemed optimal.
+
+\begin{figure}[H]
+\centering
+\includegraphics[scale=0.25]{./Figures/examplebfssolver}
+%\decoRule
+\caption[Examples]{breadth first search solver example}
+\label{fig:examplebfssolver}
+\end{figure}
+
 %-----------------------------------
@@
 
+\afblue
+\paragraph{}{\textbf{python code -- depth first search solver}}
+\begin{python}
+####################################################################
+from rubiks.solvers.solver import Solver
+from rubiks.puzzle.puzzle import Puzzle
+####################################################################
+if '__main__' == __name__:
+    puzzle_type = Puzzle.sliding_puzzle
+    n=3
+    m=3
+    nb_shuffles=8
+    limit=15
+    time_out=60
+    solver_type=Solver.dfs
+    check_optimal=True
+    log_solution=True
+    action_type=Solver.do_solve
+    Solver.factory(**globals()).action()
+####################################################################
+
+\end{python}
+\black
+Here we can see, in contrast with the previous example, that the solution obtained by DFS is not optimal.
+The Solver indicates it and shows an optimal solution.
+Clearly since we shuffled 8 times from goal configuration, the optimal path cannot have a cost higher than 8 (it could be lower though of course).
+
+\begin{figure}[H]
+\centering
+\includegraphics[scale=0.2]{./Figures/exampledfssolver}
+%\decoRule
+\caption[Examples]{depth first search solver example}
+\label{fig:exampledfssolver}
+\end{figure}
+
 %-----------------------------------
@@
 \subsection{Naive Sliding Puzzle Solver}
-As a comparison point for other sliding puzzle solvers, I have implemented a naive sliding puzzle solver, which does what most beginner players would intuitively do when solving the sliding puzzle by hand: solve the top row, then the left column, and keep iterating until done.
+As a comparison point for otherSP solvers, I have implemented a naive sliding puzzle solver, which does what most beginner players would intuitively do when solving the sliding puzzle by hand: solve the top row, then the left column, and keep iterating until done.
 Notice that once either the top row or left column is solved, there is no longer any need to modify it, we have simply reduced the problem to a sub-problem of reduced dimension.

### 2022-07-30 latex/Chapters/SlidingPuzzle.tex
update

@@
 \hline
-2              &   & 6 & 21 & 36 &  \\
+2              &   & 6 & 21 & 36 & 51* \\
 3              &   &   & 31 &  &    \\
@@
 \end{center}
+* provisional result
+\\
+\\
 and the most difficult puzzles (requiring a number of steps equal to their respective God number to solves):
@@
 \end{center}
-\underline{Most difficult 2x5 (47 moves) .... TBD }:
+\underline{Most difficult 2x5 (51* moves)}:
 \begin{center}
 \begin{nine}
-\setrow{2}{,2,7,4,1}
-\setrow{1}{5,8,3,9,6}
+\setrow{2}{,3,4,2,6}
+\setrow{1}{5,9,8,7,1}
 \end{nine}
 \end{center}
+* provisional result
+\\
+\\
 \underline{Most difficult 3x3 (31 moves)}:
@@
 \label{ssec:33SPSC}
-Finally, to conclude this section on the 3x3 case, let me discuss a comparison of several algorithms on 1000 random puzzles generated for a number of random shuffling (with best effort no backtracking) from 0 to 50 in step of 2, as well as for perfect shuffling (denoted by $\infty$) on the comparison graphs.
+Let me discuss a comparison of several algorithms on 1000 random puzzles generated for a number of random shuffling (with best-effort-no-backtracking) from 0 to 50 in step of 2, as well as for perfect shuffling (denoted by $\infty$) on the comparison graphs.
 The results are shown in figure \ref{fig:33SPPerformance}

### 2022-07-30 latex/MScAIFrancoisBerrier.tex
update

@@
 \newcommand{\blue}{\color{blue}}
-\newcommand{\green}{\color{green}}
+\definecolor{airforceblue}{rgb}{0.36, 0.54, 0.66}
+\newcommand{\afblue}{\color{airforceblue}}
 
@@
   {\end{enumerate}}
+
+\usepackage{pythonhighlight}
 
@@
 %----------------------------------------------------------------------------------------
-%	DECLARATION PAGE
-%----------------------------------------------------------------------------------------
-
-%\begin{declaration}
-%\addchaptertocentry{\authorshipname} % Add the declaration to the table of contents
-%\noindent I, \authorname, declare that this thesis titled, \enquote{\ttitle} and the work presented in it are my own. I confirm that:
-
-%\begin{itemize} 
-%\item This work was done wholly or mainly while in candidature for a research degree at this University.
-%\item Where any part of this thesis has previously been submitted for a degree or any other qualification at this University or any other institution, this has been clearly stated.
-%\item Where I have consulted the published work of others, this is always clearly attributed.
-%\item Where I have quoted from the work of others, the source is always given. With the exception of such quotations, this thesis is entirely my own work.
-%\item I have acknowledged all main sources of help.
-%\item Where the thesis is based on work done by myself jointly with others, I have made clear exactly what was done by others and what I have contributed myself.\\
-%\end{itemize}
-
-%\noindent Signed:\\
-%\rule[0.5em]{25em}{0.5pt} % This prints a line for the signature
-
-%\noindent Date:\\
-%\rule[0.5em]{25em}{0.5pt} % This prints a line to write the date
-%\end{declaration}
-
-%\cleardoublepage
-
-%----------------------------------------------------------------------------------------
 %	QUOTATION PAGE
@@
 \\
-Inspired however by the seminal success of Deep Mind's team in using Deep Reinforcement Learning (\textbf{DRL}) to play Atari games (\cite{Mnih2013})  and to master the game of Go (\cite{AlphaGo}), researchers have in recent years made a lot of progress towards designing algorithms capable of learning and solving, \textit{without human knowledge}, the Rubiks Cube (\textbf{RC}) - as well as similar single player puzzles - by using Deep Q-Learning (\textbf{DQR})  (\cite{DBLP:journals/corr/abs-1805-07470}) or search and (\textbf{DRL}) value iteration (\cite{https://doi.org/10.48550/arxiv.1805.07470}).
+Inspired however by the seminal success of Deep Mind's team in using Deep Reinforcement Learning (\textbf{DRL}) to play Atari games (\cite{Mnih2013})  and to master the game of Go (\cite{AlphaGo}), researchers have in recent years made a lot of progress towards designing algorithms capable of learning and solving, \textit{without human knowledge}, the Rubiks Cube (\textbf{RC}) - as well as similar single player puzzles - by using Deep Q-Learning (\textbf{DQL})  (\cite{DBLP:journals/corr/abs-1805-07470}) or search and (\textbf{DRL}) value iteration (\cite{https://doi.org/10.48550/arxiv.1805.07470}).
 \\
@@
 
-\tableofcontents % Prints the main table of contents
-
 %\listoffigures % Prints the list of figures
@@
 \textbf{AIPnT} & \textbf{A}rtificial \textbf{I}ntelligence \textbf{P}rinciples and \textbf{T}echniques\\
+\textbf{BFS} & \textbf{B}readth \textbf{F}irst \textbf{S}earch\\
 \textbf{CS} & \textbf{C}omputer \textbf{S}cience\\
 \textbf{CV} & \textbf{C}omputer \textbf{V}ision\\
+\textbf{DFS} & \textbf{D}epth \textbf{F}irst \textbf{S}earch\\
 \textbf{DL} & \textbf{D}eep \textbf{L}earning\\
@@
 \textbf{RC} & \textbf{R}ubiks' \textbf{C}ube\\
-\textbf{RL} & \textbf{R}einforcment \textbf{L}earning\\
+\textbf{RL} & \textbf{R}einforcement \textbf{L}earning\\
 \textbf{RHUL} & \textbf{R}oyal \textbf{H}olloway, \textbf{U}niversity of \textbf{L}ondon\\
@@
 \end{abbreviations}
+
+\tableofcontents % Prints the main table of contents
 

### 2022-07-31 latex/Chapters/Examples.tex
update

@@
 
-\textbf{manhattan heuristic}
+\subsubsection{Manhattan heuristic}
 
-\textbf{perfect heuristic}
+\afblue
+\paragraph{}{\textbf{python code -- depth first search solver}}
+\begin{python}
+####################################################################
+from rubiks.heuristics.heuristic import Heuristic
+from rubiks.puzzle.puzzle import Puzzle
+from rubiks.solvers.solver import Solver, Solution
+####################################################################
+if '__main__' == __name__:
+    puzzle_type = Puzzle.sliding_puzzle
+    tiles=[[3, 8, 6], [4, 1, 5], [0, 7, 2]]
+    solver_type=Solver.astar
+    heuristic_type=Heuristic.manhattan
+    plus=False
+    action_type=Solver.do_solve
+    print(Solver.factory(**globals()).action().to_str([Solution.puzzle,
+                                                       Solution.cost,
+                                                       Solution.expanded_nodes,
+                                                       Solution.success,
+                                                       Solution.solver_name,
+                                                       Solution.run_time]))
+####################################################################
+\end{python}
+\black
+We run this example twice, one with simple Manhattan and one with Manhattan++.
+As can be seen from the output below, the Manhattan++ improves on the number of expanded nodes (as expected, since the heuristic is less optimistic while retaining its optimality property).
+As we have seen before, the overhead of adding penalty for the linear constraint (which have all been precomputed and stored in a database) is more than compensated for by the reduction in node expansion!
 
-\textbf{deep learning heuristic}
+\begin{figure}[H]
+\centering
+\includegraphics[scale=0.39]{./Figures/exampleastarmanhattansolver}
+%\decoRule
+\caption[Examples]{a* manhattan solver example}
+\label{fig:exampleastarmanhattansolver}
+\end{figure}
 
-\textbf{deep reinforcement learning heuristic}
+\begin{figure}[H]
+\centering
+\includegraphics[scale=0.39]{./Figures/exampleastarmanhattanplussolver}
+%\decoRule
+\caption[Examples]{a* manhattan++ solver example}
+\label{fig:exampleastarmanhattanplussolver}
+\end{figure}
+
+\subsubsection{Perfect Heuristic}
+\subsubsection{Deep Learning Heuristic}
+\subsubsection{Deep Reinforcement Learning Heuristic}

### 2022-07-31 latex/Chapters/SlidingPuzzle.tex
update

@@
 \hline
-2              &   & 6 & 21 & 36 & 51* \\
+2              &   & 6 & 21 & 36 & 53* \\
 3              &   &   & 31 &  &    \\
@@
 \\
-and the most difficult puzzles (requiring a number of steps equal to their respective God number to solves):
+Notice that among the above dimensions, (n=2, m=5) is the largest and hardest one. I had to run its PerfectLearner over a couple of days, on a c5.18xlarge instance (72 cores) on Amazon EC2.
+\\
+\\
+The perfectLearner also keeps track of (one of) the hardest puzzles it has encountered (i.e. requiring a number of steps equal to their respective God number to solves):
 \\
@@
 \end{center}
-\underline{Most difficult 2x5 (51* moves)}:
+\underline{Most difficult 2x5 (53* moves)}:
 \begin{center}
 \begin{nine}
-\setrow{2}{,3,4,2,6}
-\setrow{1}{5,9,8,7,1}
+\setrow{2}{,4,9,2,1}
+\setrow{1}{5,8,3,7,6}
 \end{nine}

### 2022-07-31 rubiks/heuristics/manhattan.py

@@ -139,6 +139,7 @@ class Manhattan(Heuristic):
                     if row not in data[cls.Line.row]:
                         data[cls.Line.row][row] = dict()
                     data[cls.Line.row][row][possible] = penalty
+                    logger.log_info(cls.Line.row, ' ', row, ' ', possible, ' -> ', penalty)
                 total_precomputations += 1
         for col in range(m):
             expected = tuple(goal.tiles[:, col].tolist())
@@ -149,6 +150,7 @@ class Manhattan(Heuristic):
                     if col not in data[cls.Line.col]:
                         data[cls.Line.col][col] = dict()
                     data[cls.Line.col][col][possible] = penalty
+                    logger.log_info(cls.Line.col, ' ', col, ' ', possible, ' -> ', penalty)
                 total_precomputations += 1
         to_pickle(data, manhattan_plus_file_name)
         logger.log_info('Saved ',

### 2022-07-31 rubiks/scripts/draft.py

@@ -1,17 +1,19 @@
 ####################################################################
-from rubiks.solvers.solver import Solver
+from rubiks.heuristics.heuristic import Heuristic
 from rubiks.puzzle.puzzle import Puzzle
+from rubiks.solvers.solver import Solver, Solution
 ####################################################################
 if '__main__' == __name__:
     puzzle_type = Puzzle.sliding_puzzle
-    n=3
-    m=3
-    nb_shuffles=8
-    limit=15
-    time_out=60
-    solver_type=Solver.dfs
-    check_optimal=True
-    log_solution=True
+    tiles=[[3, 8, 6], [4, 1, 5], [0, 7, 2]]
+    solver_type=Solver.astar
+    heuristic_type=Heuristic.manhattan
+    plus=True
     action_type=Solver.do_solve
-    Solver.factory(**globals()).action()
+    print(Solver.factory(**globals()).action().to_str([Solution.puzzle,
+                                                       Solution.cost,
+                                                       Solution.expanded_nodes,
+                                                       Solution.success,
+                                                       Solution.solver_name,
+                                                       Solution.run_time]))
 ####################################################################

### 2022-07-31 unittests/testheuristics.py

@@ -59,7 +59,7 @@ class TestHeuristics(TestCase):
         self.assertEqual(16, cost)
         self.assertEqual(16, cost_plus)
 
-    def manhattan_plus_check(self, name, n, m=None):
+    def manhattan_plus_check(self, name, n, m=None, modulo=100):
         logger = Loggable(name=name)
         manhattan = Solver.factory(solver_type=Solver.astar,
                                    heuristic_type=Heuristic.manhattan,
@@ -87,8 +87,9 @@ class TestHeuristics(TestCase):
                              'manhattan = %d, manhattan+ = %d for %s' % (solution.cost,
                                                                          solution_plus.cost,
                                                                          sliding_puzzle))
-            if 0 == count % 100:
-                logger.log_info('Checked ', count, ' puzzles')
+            if 0 == count % modulo:
+                logger.log_info('Checked', count, 'puzzles')
+        logger.log_info('Checked', count, 'puzzles')
 
     def test_manhattan_distance_2_2_plus(self):
         self.manhattan_plus_check('test_manhattan_distance_2_2_plus', n=2, m=2)
@@ -97,10 +98,13 @@ class TestHeuristics(TestCase):
         self.manhattan_plus_check('test_manhattan_distance_2_2_plus', n=2, m=3)
 
     def test_manhattan_distance_2_4_plus(self):
-        self.manhattan_plus_check('test_manhattan_distance_2_2_plus', n=2, m=4)
+        self.manhattan_plus_check('test_manhattan_distance_2_4_plus', n=2, m=4)
+
+    def test_manhattan_distance_2_5_plus(self):
+        self.manhattan_plus_check('test_manhattan_distance_2_5_plus', n=2, m=5, modulo=1)
 
     def test_manhattan_distance_3_3_plus(self):
-        self.manhattan_plus_check('test_manhattan_distance_2_2_plus', n=3, m=3)
+        self.manhattan_plus_check('test_manhattan_distance_3_3_plus', n=3, m=3)
 
     def test_something(self):
         logger = Loggable(name='test_something')

### 2022-07-31 latex/Chapters/Examples.tex
update

--- 2022-07-30 latex/Chapters/Examples.tex
@@
 
-Unlike the Deep Learner, the Deep Reinforcement Learner learns unsupervised (in the sense that there is no need to pre-solve puzzles to tell if what the desired costs are), since it generates its own target using a combination of a target network and the simple min rule described in \ref{sec:codelearners}.
+Unlike the Deep Learner, the Deep Reinforcement Learner learns unsupervised (in the sense that there is no need to pre-solve puzzles to tell if what the actual (target) costs are), since it generates its own target using a combination of a target network and the simple min rule described in \ref{sec:codelearners}.
 Let us see here how to run it on the (n=4, m=2) SP for instance.
@@
 
-\textbf{manhattan heuristic}
+\subsubsection{Manhattan heuristic}
 
-\textbf{perfect heuristic}
+\afblue
+\paragraph{}{\textbf{python code -- depth first search solver}}
+\begin{python}
+####################################################################
+from rubiks.heuristics.heuristic import Heuristic
+from rubiks.puzzle.puzzle import Puzzle
+from rubiks.solvers.solver import Solver
+####################################################################
+if '__main__' == __name__:
+    puzzle_type = Puzzle.sliding_puzzle
+    tiles=[[3, 8, 6], [4, 1, 5], [0, 7, 2]]
+    solver_type=Solver.astar
+    heuristic_type=Heuristic.manhattan
+    plus=False
+    action_type=Solver.do_solve
+    print(Solver.factory(**globals()).action().to_str_light())
+####################################################################
+\end{python}
+\black
+We run this example twice, one with simple Manhattan and one with Manhattan++.
+As can be seen from the output below, the Manhattan++ improves on the number of expanded nodes (as expected, since the heuristic is less optimistic while retaining its optimality property).
+As we have seen before, the overhead of adding penalty for the linear constraint (which have all been precomputed and stored in a database) is more than compensated for by the reduction in nodes expansion!
 
-\textbf{deep learning heuristic}
+\begin{figure}[H]
+\centering
+\includegraphics[scale=0.5]{./Figures/exampleastarmanhattansolver}
+%\decoRule
+\caption[Examples]{a* manhattan solver example}
+\label{fig:exampleastarmanhattansolver}
+\end{figure}
 
-\textbf{deep reinforcement learning heuristic}
+\begin{figure}[H]
+\centering
+\includegraphics[scale=0.5]{./Figures/exampleastarmanhattanplussolver}
+%\decoRule
+\caption[Examples]{a* manhattan++ solver example}
+\label{fig:exampleastarmanhattanplussolver}
+\end{figure}
+
+\subsubsection{Perfect Heuristic}
+To run A{*} with a perfect heuristic, we just need to specify the heuristic type as such, and set the parameter \textit{model\_file\_name} to point to a pre-recorded database populated by the PerfectLearner (see earlier section \ref{PLSS}).
+
+\afblue
+\paragraph{}{\textbf{python code -- depth first search solver}}
+\begin{python}
+####################################################################
+from rubiks.heuristics.heuristic import Heuristic
+from rubiks.puzzle.puzzle import Puzzle
+from rubiks.solvers.solver import Solver
+from rubiks.utils.utils import get_model_file_name
+####################################################################
+if '__main__' == __name__:
+    puzzle_type = Puzzle.sliding_puzzle
+    n=3
+    nb_shuffles=40
+    solver_type=Solver.astar
+    heuristic_type=Heuristic.perfect
+    model_file_name = get_model_file_name(puzzle_type=puzzle_type,
+                                          dimension=(n, n),
+                                          model_name=Heuristic.perfect)
+    action_type=Solver.do_solve
+    print(Solver.factory(**globals()).action().to_str_light())
+####################################################################
+\end{python}
+\black
+Running this code snipet will output something like (the shuffle is random obviously):
+
+\begin{figure}[H]
+\centering
+\includegraphics[scale=0.5]{./Figures/exampleastarperfectsolver}
+%\decoRule
+\caption[Examples]{a* perfect solver example}
+\label{fig:exampleastarperfectsolver}
+\end{figure}
+
+\subsubsection{Deep Learning Heuristic}
+\subsubsection{Deep Reinforcement Learning Heuristic}

### 2022-07-31 latex/Chapters/SlidingPuzzle.tex
update

@@
 \label{sec:SPLowDimension}
+
+\subsection{God numbers and hardest puzzles}
 
@@
 \end{eight}
+\end{center}
+
+\subsection{Manhattan heuristic}
+In this section, I verify empirically that, as expected, the overhead of adding penalty in Manhattan++ for the linear constraint (which have all been precomputed and stored in a database) is more than compensated for by the reduction in nodes expansion.
+I have run my solver script for (n=2, m=5) in performance test mode, for both Manhattan and Manhattan++, with 250 randomly shuffled puzzles with \textit{nb\_shuffles} from 0 to 60 by increment of 5, as well as with \textit{nb\_shuffles = $\inf$}.
+The resulting run time and nodes expansions are as follows:
+
+\begin{figure}[H]
+\centering
+\hspace*{-2.5cm}
+\includegraphics[scale=0.45]{./Figures/25SPPerformanceManhattan}
+%\decoRule
+\caption[SP]{Manhattan vs Manhattan++}
+\label{fig:25SPPerformanceManhattan}
+\end{figure}
+
+As can be seen, for low difficulty (up to \textit{nb\_shuffles = 20}), the node expansions are about the same in both cases, and the overhead of adding the linear constraints penalty increases the run time.
+However, for any non trivial case, Manhattan++ outperforms considerably (by a factor of about 2.5).
+For the 250 \textit{perfectly} shuffled instances, I got the following:
+
+\begin{center}
+\begin{tabular}{l*{7}{c}r}
+                              & avg cost  & max cost & avg nodes & max nodes & avg run time (ms) & max run time (ms) \\
+\hline
+Manhattan                   &  34.7  & 50 & 133,332 & 2,110,887 & 49,561 & 606,838 \\
+Manhattan++              & 34.7 &  50 & 53,637 & 962,324 & 19,723 & 239,468 \\
+Improvement               & n/a &  n/a & x2.5 & x2.2 & x2.5 & x2.5 \\
+\end{tabular}
 \end{center}

### 2022-07-31 rubiks/solvers/solver.py

@@ -126,7 +126,8 @@ class Solver(Factory, Puzzled, Loggable, metaclass=ABCMeta):
                 if 3 <= len(self.shuffles_data[nb_shuffles][index]) and \
                         self.get_name() in self.shuffles_data[nb_shuffles][index][2] and \
                         not self.shuffles_data[nb_shuffles][index][2][self.get_name()].failed():
-                    print('Already solved this one ', puzzle, ' with ', self.get_name())
+                    if self.verbose:
+                        print('Already solved this one ', puzzle, ' with ', self.get_name())
                     return self.shuffles_data[nb_shuffles][index][2][self.get_name()]
             else:
                 puzzle = self.get_goal().apply_random_moves(nb_moves=nb_shuffles,
@@ -138,17 +139,18 @@ class Solver(Factory, Puzzled, Loggable, metaclass=ABCMeta):
             assert isinstance(solution.path, list)
             assert all(isinstance(move, self.get_goal().get_move_type()) for move in solution.path)
             assert isnan(solution.expanded_nodes) or isinstance(solution.expanded_nodes, int)
-            if solution.failed():
-                print('failed to solve ', puzzle, ' # ', index)
-            else:
-                print('solved ',
-                      puzzle,
-                      ' # ',
-                      index,
-                      ' with cost ',
-                      solution.cost,
-                      ' in ',
-                      s_format(run_time))
+            if self.verbose:
+                if solution.failed():
+                    print('failed to solve ', puzzle, ' # ', index)
+                else:
+                    print('solved ',
+                          puzzle,
+                          ' # ',
+                          index,
+                          ' with cost ',
+                          solution.cost,
+                          ' in ',
+                          s_format(run_time))
         except Exception as error:
             solution = Solution.failure(puzzle)
             solution.set_run_time(self.time_out)
@@ -199,6 +201,7 @@ class Solver(Factory, Puzzled, Loggable, metaclass=ABCMeta):
                                    pct_optimal,
                                    median_expanded_nodes,
                                    pct_solved]
+    verbose = 'verbose'
     fig_size = 'fig_size'
 
     @classmethod
@@ -278,6 +281,10 @@ class Solver(Factory, Puzzled, Loggable, metaclass=ABCMeta):
                          type=int,
                          nargs='+',
                          default=[16, 12])
+        cls.add_argument(parser,
+                         cls.verbose,
+                         default=False,
+                         action=cls.store_true)
     
[…]

### 2022-07-31 latex/Chapters/Code.tex
update

@@
 
-Let me describe what each submodule does:
+Let me now describe what each submodule does in more details:
 
@@
 \begin{itemize}
-\item \textbf{Manhattan}: at current time of writing, this is specific to the SP and will be discussed in more details in \ref{S33},
+\item \textbf{Manhattan}: This heuristic is specific to the SP.
+It simply adds for each tile (other than the empty tile) the $L_{1}$ distance between their current position and their target position.
+We can quickly see that this heuristic is admissible.
+Indeeed, each move in the SP moves one (and one only) tile by one position.
+In other puzzles, it is generally the case that a move will affect the position of many tiles (e.g.
+RC).
+Therefore, if tiles could somehow freely move on top of one-another (that is, we remove the constraint that there can be at most one tile per compartment, the number of moves necessary to solve the SP would exactly be the Manhattan distance.
+\\
+\\
+The reason why it is interesting to have a known admissible heuristic is that we can obvioulsy compare other heuristics (DL, DRL, DQL, etc) in terms of optimality.
+\\
+\\
+I have also implemented an improvement to the Manhattan distance, which I shall call Manhattan++ in here (and in code logs and graphs) and can be activated by simply passing \textit{plus=True} to the Manhattan Heuristic (see example \ref{ASSS} as well as a thorougher performance comparison on the (n=2, m=5) SP in \ref{MHComp}).
+It is based on the concept of linear constraints that I read about in lecture notes from Washington University (\cite{SlidingPuzzleLectureNotes}).
+The idea is simply that when tiles are already placed on their target row or column, but not in the expected order, we will need to get the tiles out of the way of one another to get to the goal state, and this is not accounted for by the Manhattan distance.
+For instance, in the following (n=2, m=3) SP, the Manhattan distance does not account for the fact that, at best, tile \red 3 \black needs to get out of the way for tiles \blue 1 \black and \blue 2 \black to move, and then needs to get back in its row, adding a cost of 2 to the Manhattan distance.
+
+\begin{center}
+\begin{five}
+\setrow{2}{\red 3, \blue 1, \blue 2}
+\setrow{1}{4, 5, 0}
+\end{five}
+\end{center}
+
+Two important things to notice are that linear constraints across rows and columns (which I might generically refer to as \textit{lines} in the following) can to be added without breaking admissibility (hence giving more pessimistic, or accurate, cost estimates than simple Manhattan).
+This is because if a tile is involved in two linear constraints, it will need to get out of the way both horizontally and vertically.
+The second thing is that when several pairs in a line are not in order, we cannot simply add 2 for each distinct out-of-order pair, the right penalty to add is more subtle than that and needs to be computed recursively.
+For instance, let us now consider the following configuration:
+\begin{center}
+\begin{five}
+\setrow{2}{\red 3, \red 2, \red 1}
+\setrow{1}{4, 5, 0}
+\end{five}
+\end{center}
+The correct penalty to add is not 6 (3 times 2 since all pairs (1, 2), (1, 3) and (2, 3) are out of order, but only 4.
+Indeed if, say, tile 3 got somehow out of the way at the back of the SP (imagine just another dimension there where we can move tiles) and tile 2 got out of the way by moving down to let tile 1 pass across, we could be done by simply adding 4 to the Manhattan distance (2 to move tile 3 out and back, and 2 to move tile 2 out and back).
+The correct way to compute the penalty cost for linear constraints is therefore to do it recursively, taking the minimum additional cost of moving either left-most or right-most tile of the line under consideration out of the way (that additional cost to move these left-most or right-most tile is 2 if not at their expected order in the line, 0 otherwise) plus the penalty of reordering the rest of the line.
+\\
+\\
+Finally, as suggested by the reference lecture notes (which give very vague details about the above subtleties), I have precomputed all the penalties for all possible rows, columns and all possible tiles ordering they could have and saved the corresponding penalties in a database.
+For memory efficiency, I also only saved penalties which are non-zero.
+The very first time any call to Manhattan++ is made, for a given dimension (n, m), the appropriate data base is computed and populated.
+\\
+\\
+Notice that for an (n, m) SP, there are n rows, each of which can have $\frac{(n * m)!}{(n * m - m)!}$ different ordering of tiles and m columns which can have n columns, each can have $\frac{(n * m)!}{(n * m - n)!}$ different ordering of tiles.
+This means the pre-computations and data-base sizes for the Manhattan++ heuristic are actually manageable, as it grows much slower than the number of possible puzzles.
+Taking into account that we only store non-zero penalties, for instance the (n=4, m=4) Manhattan++ database only has  records, and the (n=5, m=5) has .
+
 \item \textbf{PerfectHeuristic}: this reads from a data base the optimal costs, pre-computed by the PerfectLearner (see below \ref{PLcode})
@@
 \section{rubiks.learners}
+\label{sec:codelearners}
 \begin{figure}[H]

### 2022-07-31 latex/Chapters/SlidingPuzzle.tex
update

@@
 \subsection{Manhattan heuristic}
+\label{MHComp}
 In this section, I verify empirically that, as expected, the overhead of adding penalty in Manhattan++ for the linear constraint (which have all been precomputed and stored in a database) is more than compensated for by the reduction in nodes expansion.

### 2022-07-31 unittests/testheuristics.py

@@ -59,7 +59,7 @@ class TestHeuristics(TestCase):
         self.assertEqual(16, cost)
         self.assertEqual(16, cost_plus)
 
-    def manhattan_plus_check(self, name, n, m=None, modulo=100):
+    def manhattan_plus_check(self, name, n, m=None, modulo=100, max_puzzle=inf):
         logger = Loggable(name=name)
         manhattan = Solver.factory(solver_type=Solver.astar,
                                    heuristic_type=Heuristic.manhattan,
@@ -89,6 +89,8 @@ class TestHeuristics(TestCase):
                                                                          sliding_puzzle))
             if 0 == count % modulo:
                 logger.log_info('Checked', count, 'puzzles')
+            if count > max_puzzle:
+                break
         logger.log_info('Checked', count, 'puzzles')
 
     def test_manhattan_distance_2_2_plus(self):
@@ -105,6 +107,12 @@ class TestHeuristics(TestCase):
 
     def test_manhattan_distance_3_3_plus(self):
         self.manhattan_plus_check('test_manhattan_distance_3_3_plus', n=3, m=3)
+
+    def test_manhattan_distance_4_4_plus(self):
+        self.manhattan_plus_check('test_manhattan_distance_4_4_plus', n=4, m=4, max_puzzle=1)
+
+    def test_manhattan_distance_5_5_plus(self):
+        self.manhattan_plus_check('test_manhattan_distance_5_5_plus', n=5, m=5, max_puzzle=1)
 
     def test_something_2_3(self):
         logger = Loggable(name='test_something_2_3')

### 2022-07-31 latex/Chapters/Code.tex
update

@@
 This means the pre-computations and data-base sizes for the Manhattan++ heuristic are actually manageable, as it grows much slower than the number of possible puzzles.
-Taking into account that we only store non-zero penalties, for instance the (n=4, m=4) Manhattan++ database only has  records, and the (n=5, m=5) has .
+The maximum number of penalties to compute for $(n, m) \leq (5, 5)$ are:
+
+\begin{center}
+\begin{tabular}{l*{6}{c}r}
+n              & m & 2 & 3 & 4 & 5\\
+\hline
+2              &   & 48 & 330 & 3,584 & 60,930 \\
+3              &   &   & 3,024  & 40,920 &  1,094,730  \\
+4              &   &   &  & 349,440 & 8,023,320   \\
+5              &   &   &  &  & 63,756,000   \\
+\end{tabular}
+\end{center}
+Taking also into account that we only store non-zero penalties, we actually get the following (quite smaller) number of penalties in our data bases:
+\begin{center}
+\begin{tabular}{l*{6}{c}r}
+n              & m & 2 & 3 & 4 & 5\\
+\hline
+2              &   & 2 & 46 & 1,238  & 32,888 \\
+3              &   &  & 278  & 7,122 & 328,894   \\
+4              &   &   &  & 40,546 &  1,456,680 \\
+5              &   &   &  &  &  8,215,382 \\
+\end{tabular}
+\end{center}
 
@@
 \item \textbf{DeepLearning}: a Puzzled Loggable Factory that can instantiate some configurable deep networks, and provide the necessary glue with the rest of the code base so that puzzles be seemlessly passed to the networks and trained on.
-\item \textbf{FullyConnected}: wrapper around a Pytorch fully connected network, with configurable layers and size.
-\item \textbf{Convolutional}: tdb
+\item \textbf{FullyConnected}: wrapper around a Pytorch fully connected network, with configurable hidden layers and size.
+There are some params as well to add drop out, and to indicate whether or not the inputs are one hot encoding (in which case the first layer is automatically adjusted in size, using information from the puzzle dimension).
+\item \textbf{Convolutional}: similar wrapper to FullyConnected, but with the ability to add some parallel convolutional layers to complement fully connected layers.
 \end{itemize}
@@
 It then saves the trained network, which can then be used in the DeepLearningHeuristic we have seen earlier in section \ref{HSS}.
-The
+The DeepReinforcementLearner runs for a number of epochs (or less if, based on other parameters discussed below, it is deemed to have converged).
 

### 2022-07-31 rubiks/puzzle/puzzle.py
started on rubiks

@@ -58,7 +58,7 @@ class Puzzle(Factory, metaclass=ABCMeta):
 
     @classmethod
     def get_puzzle(cls, nb_shuffles, **kw_args):
-        """ return a puzzle shuffled nb_shuffles times randomly from goal state
+        """ return a puzzle shuffled nb_shuffles times randomly from init state
         kw_args allow to construct the puzzle with e.g. puzzle dimension
         """
         puzzle = cls.factory(**kw_args)
@@ -74,12 +74,11 @@ class Puzzle(Factory, metaclass=ABCMeta):
         return '%s[%s]' % (self.__class__.__name__, str(tuple(self.dimension())))
 
     def is_solvable(self) -> bool:
-        """ is this puzzle the goal state? """
         raise NotImplementedError('Cannot tell if a %s is solvable' % self.__class__)
 
     @abstractmethod
     def is_goal(self) -> bool:
-        """ is this puzzle the goal state? """
+        """ True if puzzle in solved state """
         return
 
     @classmethod
@@ -92,18 +91,18 @@ class Puzzle(Factory, metaclass=ABCMeta):
         """
         Produces training data.
         params:
-            nb_shuffles: number of shuffles we do from goal state
+            nb_shuffles: number of shuffles we do from init state
             nb_sequences: how many such sequences we produce
             min_no_loop: (best effort) making sequences with no loop of length <= min_no_loop
             one_list: results is a list, otherwise a list of lists
         returns:
             list of puzzles
         """
-        goal = cls(**kw_args)
+        init = cls(**kw_args)
         training_data = list()
         for _ in range(nb_sequences):
-            moves = goal.random_moves(nb_shuffles, min_no_loop=min_no_loop)
-            puzzles = goal.get_puzzle_sequence(moves)
+            moves = init.random_moves(nb_shuffles, min_no_loop=min_no_loop)
+            puzzles = init.get_puzzle_sequence(moves)
             if one_list:
                 training_data += puzzles
             else:
@@ -125,11 +124,6 @@ class Puzzle(Factory, metaclass=ABCMeta):
     @classmethod
     def factory_key_name(cls):
         return cls.puzzle_type
-
-    @abstractmethod
-    def goal(self):
-        """ return the unique goal state for that puzzle """
-        return
 
     @abstractmethod
     def apply(self, move):

### 2022-07-31 rubiks/scripts/draft.py

@@ -1,18 +1,9 @@
 ####################################################################
-from rubiks.heuristics.heuristic import Heuristic
-from rubiks.puzzle.puzzle import Puzzle
-from rubiks.solvers.solver import Solver
-from rubiks.utils.utils import get_model_file_name
+from rubiks.puzzle.rubikscube import RubiksCube
 ####################################################################
-if '__main__' == __name__:
-    puzzle_type = Puzzle.sliding_puzzle
-    n=3
-    nb_shuffles=40
-    solver_type=Solver.astar
-    heuristic_type=Heuristic.perfect
-    model_file_name = get_model_file_name(puzzle_type=puzzle_type,
-                                          dimension=(n, n),
-                                          model_name=Heuristic.perfect)
-    action_type=Solver.do_solve
-    print(Solver.factory(**globals()).action().to_str_light())
-####################################################################
+cube = RubiksCube(n=3)
+print(cube)
+print(cube.__hash__())
+print(hash(cube))
+print(cube.goals_hashes)
+print(cube.is_goal())

### 2022-07-31 unittests/testrubikscube.py

@@ -1,0 +1,20 @@
+########################################################################################################################
+# Francois Berrier - Royal Holloway University London - MSc Project 2022                                               #
+########################################################################################################################
+from unittest import TestCase
+########################################################################################################################
+from rubiks.core.loggable import Loggable
+from rubiks.puzzle.rubikscube import RubiksCube
+########################################################################################################################
+
+
+class TestRubiksCube(TestCase):
+
+    def test_construct(self):
+        logger = Loggable(name='test_construct')
+        puzzle = RubiksCube(n=3)
+        logger.log_info(puzzle)
+        self.assertEqual(3, puzzle.dimension())
+        self.assertTrue(puzzle.is_goal())
+
+########################################################################################################################

### 2022-08-01 rubiks/solvers/kociembasolver.py
update

@@ -1,0 +1,23 @@
+########################################################################################################################
+# Francois Berrier - Royal Holloway University London - MSc Project 2022                                               #
+########################################################################################################################
+from kociemba import solve as kociemba_solve
+########################################################################################################################
+from rubiks.puzzle.rubikscube import RubiksCube
+from rubiks.solvers.solver import Solver, Solution
+########################################################################################################################
+
+
+class KociembaSolver(Solver):
+    """ Rubik's solver from Kociemba """
+
+    @classmethod
+    def know_to_be_optimal(cls):
+        return False
+
+    def solve_impl(self, puzzle, **kw_args) -> Solution:
+        assert isinstance(puzzle, RubiksCube) and puzzle.dimension() == 3
+        solution = kociemba_solve(puzzle.to_kociemba())
+
+
+########################################################################################################################

### 2022-08-02 rubiks/scripts/learner.py
update

@@ -39,9 +39,9 @@ if '__main__' == __name__:
         else DeepLearner.default_plot_metrics
     """ If it's a perfect learner config here """
     time_out = 3600
-    nb_cpus = 1
-    cpu_multiplier = 1
-    max_puzzles = nb_cpus * cpu_multiplier * 1
+    nb_cpus = 12
+    cpu_multiplier = 100
+    max_puzzles = nb_cpus * cpu_multiplier * 100
     regular_save = nb_cpus * cpu_multiplier * 1
     save_timed_out_max_puzzles = 100000
     after_round_save = True
@@ -55,7 +55,7 @@ if '__main__' == __name__:
                                    PerfectLearner.permutation_puzzle_generation,
                                    PerfectLearner.random_from_goal_puzzle_generation,
                                    } """
-    puzzle_generation = PerfectLearner.permutation_puzzle_generation
+    puzzle_generation = PerfectLearner.perfect_random_puzzle_generation
     heuristic_type = Heuristic.manhattan
     plus = True
     """ If it's a DRL learner config is here ... """

### 2022-08-06 latex/Chapters/Examples.tex
update

@@
 %-----------------------------------
+\section{Puzzles}
+Let me start by showing how to construct puzzles, using the Puzzle factory.
+Notice that in order to run a learner or solver of any kind (assuming of course that they are meant to work on the puzzle type in question), we can just use the exact same code, but just specify \textit{puzzle\_type}  and the expected parameters to construct a puzzle.
+\\
+\\
+For instance, let us create a (n=5, m=6) SP, shuffle it a number of times, and print it:
+\afblue
+\paragraph{}{\textbf{python code -- sliding puzzle construction}}
+\begin{python}
+####################################################################
+from rubiks.puzzle.puzzle import Puzzle
+####################################################################
+puzzle_type=Puzzle.sliding_puzzle
+n=5
+m=6
+nb_moves=1000
+print(Puzzle.factory(**globals()).apply_random_moves(nb_moves))
+####################################################################
+\end{python}
+\black
+The output from the above code snippet will look like (subject to randomness):
+
+\begin{figure}[H]
+\centering
+\includegraphics[scale=0.8]{./Figures/examplespconstruction}
+%\decoRule
+\caption[Examples]{sliding puzzle construction example}
+\label{fig:examplespconstruction}
+\end{figure}
+.. similarly to construct a (n=2) RC and shuffle it perfectly:
+\afblue
+\paragraph{}{\textbf{python code -- rubiks cube construction}}
+\begin{python}
+####################################################################
+from math import inf
+from rubiks.puzzle.puzzle import Puzzle
+####################################################################
+puzzle_type=Puzzle.rubiks_cube
+n=2
+""" Here we use perfect shuffle by spcifying infinite number of shuffles """
+nb_moves=inf
+print(Puzzle.factory(**globals()).apply_random_moves(nb_moves))
+####################################################################
+\end{python}
+\black
+
+\begin{figure}[H]
+\centering
+\includegraphics[scale=0.8]{./Figures/examplercconstruction}
+%\decoRule
+\caption[Examples]{rubiks cube construction example}
+\label{fig:examplercconstruction}
+\end{figure}
+
+%-----------------------------------
+%	SECTION 2
+%-----------------------------------
 \section{Learners}
@@
 %-----------------------------------
-%	SUBSECTION 1.1
+%	SUBSECTION 2.1
 %-----------------------------------
@@
 \black
-The above snipet of code will solve the 2 by 3 SP and then display the results, showing the distribution of optimal costs, as well as the most difficult puzzle.
+The above snippet of code will solve the 2 by 3 SP and then display the results, showing the distribution of optimal costs, as well as the most difficult puzzle.
 
@@
 %-----------------------------------
-%	SUBSECTION 1.2
+%	SUBSECTION 2.2
 %-----------------------------------
@@
 
-As can be seen in the code snipet, this example will generate 100 perfectly shuffled (n=5, m=2) SPs and solve them.
+As can be seen in the code snippet, this example will generate 100 perfectly shuffled (n=5, m=2) SPs and solve them.
 Once done, a summary of the training data is printed, indicating, for each optimal cost, how many sequences have been generated.
@@
 %-----------------------------------
-%	SUBSECTION 1.3
+%	SUBSECTION 2.3
 %-----------------------------------
@@
 Let us see here how to run it on the (n=4, m=2) SP for instance.
-The following code snipet will run a Deep Reinforcement Learner for a maximum of 25,000 epochs, generating randomly 10 sequences of puzzles shuffled 50 times from goal state every time it updates the target network.
+The following code snippet will run a Deep Reinforcement Learner for a maximum of 25,000 epochs, generating randomly 10 sequences of puzzles shuffled 50 times from goal state every time it updates the target network.
 That update will happen either after 500 epochs or when the (MSE) loss on the value function gets under one thousands of the max target value.
@@
 \black
-As can be seen on the next page, which I obtained from running the above code snipet (keep in mind that every run is going to be slightly different due to the random puzzles being generated), the training stopped after about 17,100 epochs as the 40 target network updates had been reached.
+As can be seen on the next page, which I obtained from running the above code snippet (keep in mind that every run is going to be slightly different due to the random puzzles being generated), the training stopped after about 17,100 epochs as the 40 target network updates had been reached.
 By that point, the DRL learner had seen 40\% of the possible puzzles, and the very last MSE loss (after update, so out-of-sample) was around 0.4, corresponding to 2\% of the max target cost.
@@
 %-----------------------------------
-%	SECTION 2
+%	SECTION 3
 %-----------------------------------
@@
 %-----------------------------------
-%	SUBSECTION 2.1
+%	SUBSECTION 3.1
 %-----------------------------------
@@
 %-----------------------------------
-%	SUBSECTION 2.1.1
+%	SUBSECTION 3.1.1
 %-----------------------------------
@@
 
-As expected (since BFS is optimal), the solution found, printed by the snipet of code above, is optimal.
-The check\_optimal flag in the code snipet indicates that the solver should let us know if solution is optimal.
+As expected (since BFS is optimal), the solution found, printed by the snippet of code above, is optimal.
+The check\_optimal flag in the code snippet indicates that the solver should let us know if solution is optimal.
 Since BFS advertises itself (via the Solver base class API)  as an optimal solver, the solution is deemed optimal.
@@
 %-----------------------------------
-%	SUBSECTION 2.1.2
+%	SUBSECTION 3.1.2
 %-----------------------------------
@@
 %-----------------------------------
-%	SUBSECTION 2.2
+%	SUBSECTION 3.2
 %-----------------------------------
@@
 \\
-Below is a simple code snipet to run the naive solver on a randomly generated (n=2, m=2) SP:
+Below is a simple code snippet to run the naive solver on a randomly generated (n=2, m=2) SP:
 \afblue
@@
 %-----------------------------------
-%	SUBSECTION 2.3
+%	SUBSECTION 3.3
 %-----------------------------------
@@
 %-----------------------------------
-%	SUBSECTION 2.4
+%	SUBSECTION 3.4
 %-----------------------------------
@@
 As can be seen from the output below, the Manhattan++ improves on the number of expanded nodes (as expected, since the heuristic is less optimistic while retaining its optimality property).
-As we have seen before, the overhead of adding penalty for the linear constraint (which have all been precomputed and stored in a database) is more than compensated for by the reduction in nodes expansion!
+See a more detailed analysis in the SP results section \ref{}
 
@@
 \black
-Running this code snipet will output something like (the shuffle is random obviously):
+Running this code snippet will output something like (the shuffle is random obviously):
 

### 2022-08-06 rubiks/scripts/solver.py

@@ -33,23 +33,24 @@ if '__main__' == __name__:
     nb_shuffles = 0
     """ For performance test """
     nb_samples = 500
-    min_nb_shuffles = 0
-    max_nb_shuffles = 15
+    min_nb_shuffles = 11
+    max_nb_shuffles = 12
     step_nb_shuffles = 1
-    add_perfect_shuffle = True
+    add_perfect_shuffle = False
     nb_cpus = 4
     performance_file_name = get_performance_file_name(puzzle_type, dimension)
     shuffles_file_name = get_shuffles_file_name(puzzle_type, dimension)
     append = True
-    verbose = False
+    verbose = True
     """ For plot """
+    loc = 'center'
     performance_metrics = [Solver.pct_solved,
                            #Solver.pct_optimal,
                            Solver.median_cost,
                            Solver.max_cost,
-                           #Solver.optimality_score,
+                           Solver.optimality_score,
                            Solver.median_run_time,
-                           #Solver.median_expanded_nodes,
+                           Solver.median_expanded_nodes,
                            ]
     fig_size = [20, 12]
     """ Which solver type {Solver.dfs,
@@ -57,9 +58,9 @@ if '__main__' == __name__:
                            Solver.astar,
                            Solver.naive,
                            } """
-    solver_type = Solver.kociemba
+    solver_type = Solver.astar
     limit = 12
-    time_out = 10
+    time_out = 600
     log_solution = False
     check_optimal = False
     max_consecutive_timeout = 25

### 2022-08-06 rubiks/utils/utils.py

@@ -164,6 +164,7 @@ def pformat(what):
                                showindex=False)
     elif isinstance(what, Series):
         what = what.to_dict()
+        what = {k: ('%.2g' % v) if isinstance(v, float) else v for k, v in what.items()}
     if isinstance(what, dict):
         what = {k: [v] if not isinstance(v, (list, np.ndarray)) \
                 else v for k, v in what.items()}

### 2022-08-06 latex/Chapters/DRLSearch.tex
update

@@
 
-TBD this is a bit of a verbiage chapter, where I go through the different techniques I have used and implemented for this project and give some references.
+In this section, I will succintly go through the different techniques I have used and implemented for this project as well as give some references.
 

### 2022-08-06 latex/Chapters/Puzzles.tex
started on rubiks
update

@@
 Notice from a programming and mathematical analysis perspective, it is often easier to equivalently think of the empty compartment being moved into (or swapped with) an adjacent tile.
-Starting from a given shuffling of the tiles on the board, our goal will be to execute moves until the tiles in ascending order: left to right, top to bottom (in the usual western reading order), the empty tile being at the very bottom right.
+Starting from a given shuffling of the tiles on the board, our goal will be to execute moves until the are tiles in ascending order: left to right, top to bottom (in the usual western reading order), the empty tile being at the very bottom right.
 \\
@@
 We will therefore only consider the case where both n and m are strictly greater than 1.
+\\
+\\
+As hinted in the previous section when discussing the coiled 15-puzzle, we can note that the parity argument used there implies that in general, out of the $(n * m)!$ possible permutations of all tiles, only half are attainable and solvable.
+This gives us a neat way to generate \textit{perfectly shuffled} puzzles, and we make use of this in the code.
+When comparing various algorithms in later sections, I will compare them on a same set of shuffled puzzles, where the shuffling is either done via a fixed number of random moves from goal position, or via what I call \textit{perfect shuffle}, which means I give equal probability to each attainable configuration of the tiles.
+A simple way to achieve this is therefore to start with one randomly selected among the $(n * m)!$ permutations, and then to verify that the parity of that permutation is the same as that of the goal (for the given choice of n and m), and simply swap the first two (non-empty) tiles if it is not.
 
@@
 
-blabla
+\subsection{History}
+The second puzzle I will focus on is the well known Rubiks' Cube (\textbf{RC}), invented in 1974 by Erno Rubik, originally known under the name of \textit{Magic Cube} (see \cite{RubiksWiki}).
+Since it was commericalised in 1980, it has kown a worldwide success: countless people are playing with the famous original 3x3x3 cube, as well as with countless variations, more or less difficult of it.
+Competitions nowadays bring together \textit{speedcubers} who have trained for years and memorized various algorithms to solve the Rubiks in astonishly effective and fast times (literally in seconds for the best ones).
+Interestingly, some of the principles used by speedcubers generalise to different dimensions.
+As an example, it is quite impressive to watch \textit{Cubastic} solve a 15x15x15 RC in two and a half hours (\cite{151515Rubiks}).
+
+\subsection{Search Space \& Solvability}
+In my code as well as in this write-up, as is very customary in the RC literature, I will refer to the faces as F (front), B (back), U (up), D (down), L (left) and R (right).
+The RC's tiles are of 6 different colors, and without loss of generality, we can consider that these colors are also called F, B, U, D, L and R, and consider the goal state as one where, up to rotations of the whole cube, the color's name match the faces (color F on face F, etc).
+The RC has, as one would expect, an extremely large state space.
+If we ignore redundant cubes via rotations, there are 24 goal states, since any of the colors can be placed on the F face, and then any of the 4 adjacent colors can be placed on the U face (say).
+Once these 2 are chosen, the remaining faces are determined.
+This is not immediately obvious, but not difficult to convince oneself that this is the case with the following two observations about the structure of corner \textit{cubies}: first observation is that they have 3 determined colors, which are fixed once and for all.
+In particular, the fact that there is no corner with colors F and B, means that once the RC is in solved state, we have to have colors F and B on opposite faces, and similarly for colors U and D as well as colors L and R.
+Hence, having fixed the color on face F, the color on face B is fixed too.
+The second observation is that when facing a corner cubie, the order of its three colors, enumerated clock-wise) is invariant.
+For instance, consider the corner cubie with colors (F, R, U).
+There is no scrambling of the cube that will produce clock-wise order of e.g.
+(F, U, R).
+This is obvious once you consider that there is no move of the Rubik's cube you could not perform while holding a given corner fixed in space (e.g.
+by pinching that corner cubie with your fingers and never letting go of it).
+\\
+Finally a further word of notation.
+It is common in RC jargon to refer to moves  by the name of the face that is rotated clock-wise (when facing it), and by adding a prime \textit{'} for counter-clock-wise rotations.
+Sometimes people use a number after the face to indicate repeated similar moves, though I will count in my code that as 2 moves, since obviously e.g.
+U2 = U U = U' U'.
+As another example: F B' F2 means rotating the front face clock-wise, the back face counter-clock-wise, followed by the front face twice.
+
+\subsubsection{2 x 2 x 2}
+
+Up to rotations of the full RC, there are $\frac{8!
+* 3^{7}}{24}$ = 3,674,160 possible combinations.
+Indeed, there are 8 possible ways of choosing the position of the eight corner cubies (since there are known algorithms, such as R’ U R’ D2 R U’ R’ D2 R2, to swap exactly two and only two adjacent corners).
+Now each of the corner cubies can be oriented in 3 different ways (not 6, since as discussed earlier, the clock-wise order of a given corner cubie cannot be changed), so that gives us $3^8$ permutations of the corners orientation.
+However, the \textit{corner orientation parity} (\textbf{COP}) (see \cite{Schoenert} for details) must be equal to 0 modulo 3, and hence why only one third of all these $3^8$ permutations are attainable.
+Finally, the denominator of 24 is due to the equivalence by whole cube rotation as also discussed in the previous section.
+\\
+The insight and knowledge of this section will be useful for me to implement \textit{perfect scrambling} for the 2x2x2 RC.
+Indeed, the way I generate perfectly shuffled RCs is by randomly placing the 8 corners, then randomly choose the orientation of each of the first 7 corners, and finally fixing the $8^{th}$ corner's orientation so that the COP be equal to 0 \% 3.
+
+\subsubsection{3 x 3 x 3}
+
+\subsection{Optimal Cost \& God's Number}
+
+\cite{RubiksChicago}
+\\
+\cite{RubiksRadu}
+\\

### 2022-08-06 rubiks/scripts/solver.py

@@ -32,16 +32,17 @@ if '__main__' == __name__:
     """ How much to shuffle """
     nb_shuffles = 0
     """ For performance test """
-    nb_samples = 500
-    min_nb_shuffles = 11
-    max_nb_shuffles = 12
+    nb_samples = 25
+    min_nb_shuffles = 0
+    max_nb_shuffles = 0
     step_nb_shuffles = 1
-    add_perfect_shuffle = False
+    add_perfect_shuffle = True
     nb_cpus = 4
     performance_file_name = get_performance_file_name(puzzle_type, dimension)
     shuffles_file_name = get_shuffles_file_name(puzzle_type, dimension)
     append = True
     verbose = True
+    do_not_reattempt_failed = False
     """ For plot """
     loc = 'center'
     performance_metrics = [Solver.pct_solved,

### 2022-08-06 latex/Chapters/Puzzles.tex
update

@@
 * 3^{7}}{24}$ = 3,674,160 possible combinations.
-Indeed, there are 8 possible ways of choosing the position of the eight corner cubies (since there are known algorithms, such as R’ U R’ D2 R U’ R’ D2 R2, to swap exactly two and only two adjacent corners).
-Now each of the corner cubies can be oriented in 3 different ways (not 6, since as discussed earlier, the clock-wise order of a given corner cubie cannot be changed), so that gives us $3^8$ permutations of the corners orientation.
-However, the \textit{corner orientation parity} (\textbf{COP}) (see \cite{Schoenert} for details) must be equal to 0 modulo 3, and hence why only one third of all these $3^8$ permutations are attainable.
-Finally, the denominator of 24 is due to the equivalence by whole cube rotation as also discussed in the previous section.
+Indeed, there are 8!
+possible ways of choosing the position of the eight corner cubies (since there are known algorithms, such as R’ U R’ D2 R U’ R’ D2 R2, to swap exactly two and only two adjacent corners), each of which can be oriented in 3 different ways (not 6, since as discussed earlier, the clock-wise order of a given corner cubie can never be changed).
+That gives us $3^8$ permutations of the corners orientation.
+However, the \textit{corner orientation parity} (\textbf{COP}) (see \cite{Schoenert} for details) is invariant (\% 3) by any legal move of the RC.
+Since rotating by $120^{o}$ a given corner cubie increases the COP by 1, we conclude that only one third of all these $3^8$ permutations are attainable (namely those for which the COP is equal to 0 \% 3).
+Finally, the denominator of 24 is due to the equivalence by whole cube rotation as discussed in the previous section.
 \\
 The insight and knowledge of this section will be useful for me to implement \textit{perfect scrambling} for the 2x2x2 RC.
-Indeed, the way I generate perfectly shuffled RCs is by randomly placing the 8 corners, then randomly choose the orientation of each of the first 7 corners, and finally fixing the $8^{th}$ corner's orientation so that the COP be equal to 0 \% 3.
+Indeed, the way I generate perfectly shuffled RCs is by randomly placing the 8 corners, then randomly choosing the orientation of the first 7, and finally fixing the $8^{th}$ corner's orientationto make COP equal to 0 \% 3.
 

### 2022-08-17 latex/Chapters/Examples.tex
update

@@
 \label{fig:examplercconstruction}
+\end{figure}
+
+.. finally let me construct a (n=3) RC with non-standard-goal:
+\afblue
+\paragraph{}{\textbf{python code -- non-standard-goal rubiks cube construction}}
+\begin{python}
+####################################################################
+from math import inf
+from rubiks.puzzle.puzzle import Puzzle
+####################################################################
+puzzle_type=Puzzle.watkins_cube
+n=2
+""" Here we use perfect shuffle by spcifying infinite number of shuffles """
+nb_moves=inf
+print(Puzzle.factory(**globals()).apply_random_moves(nb_moves))
+####################################################################
+\end{python}
+\black
+
+\begin{figure}[H]
+\centering
+\includegraphics[scale=0.8]{./Figures/examplewcconstruction}
+%\decoRule
+\caption[Examples]{non-standard-goal rubiks cube construction example}
+\label{fig:examplewcconstruction}
 \end{figure}
@@
 \subsection{Kociemba}
-TBI \& TBD
+
+\afblue
+\paragraph{}{\textbf{python code -- Kociemba solver}}
+\begin{python}
+####################################################################
+from rubiks.puzzle.puzzle import Puzzle
+from rubiks.solvers.solver import Solver
+####################################################################
+puzzle_type=Puzzle.rubiks_cube
+n=2
+cube=Puzzle.factory(**globals()).apply_random_moves(2)
+solver_type=Solver.kociemba
+solver = Solver.factory(**globals())
+print(solver.solve(cube))
+####################################################################
+
+\end{python}
+\black
+
+\begin{figure}[H]
+\centering
+\includegraphics[scale=0.55]{./Figures/examplekociemba}
+%\decoRule
+\caption[Examples]{Kociemba solver example}
+\label{fig:examplekociemba}
+\end{figure}
 

### 2022-08-17 latex/Chapters/Puzzles.tex
update

@@
 The RC has, as one would expect, an extremely large state space.
-If we ignore redundant cubes via rotations, there are 24 goal states, since any of the colors can be placed on the F face, and then any of the 4 adjacent colors can be placed on the U face (say).
+If we ignore the equivalence via full-cube rotations, there are 24 goal states, since any of the colors can be placed on the F face, and then any of the 4 adjacent colors can be placed on the U face (say).
 Once these 2 are chosen, the remaining faces are determined.
@@
 Hence, having fixed the color on face F, the color on face B is fixed too.
-The second observation is that when facing a corner cubie, the order of its three colors, enumerated clock-wise) is invariant.
+The second observation is that when facing a corner cubie, the order of its three color (e.g.
+enumerated clock-wise) is invariant.
 For instance, consider the corner cubie with colors (F, R, U).
@@
 The insight and knowledge of this section will be useful for me to implement \textit{perfect scrambling} for the 2x2x2 RC.
-Indeed, the way I generate perfectly shuffled RCs is by randomly placing the 8 corners, then randomly choosing the orientation of the first 7, and finally fixing the $8^{th}$ corner's orientationto make COP equal to 0 \% 3.
+Indeed, the way I generate perfectly shuffled RCs is by randomly placing the 8 corners, then randomly choosing the orientation of the first 7, and finally fixing the $8^{th}$ corner's orientationto make sure COP is equal to 0 \% 3.
 
 \subsubsection{3 x 3 x 3}
+
+TBD
 

### 2022-08-17 latex/Chapters/Rubiks.tex
update

@@
 
-blablabla
+\subsection{Rubiks}
+
+1000 seq 30 shuffles training each target network update seen 1.1\% of all puzzles
+199 of the 200 perfectly shuffled got solved well under 1h (median 83 seconds), only one outlier took about 67 minutes.
+
+\newgeometry{top=0mm, bottom=0mm, left=0mm, right=0mm}
+\begin{landscape}
+\centering\vspace*{\fill}
+\begin{figure}[H]
+\centering
+\includegraphics[scale=0.50]{./Figures/222RCPerformance.jpeg}
+%\decoRule
+\caption[222RCPerformance]{Solvers' performance comparison 2x2x2 RC}
+\label{fig:222RCPerformance}
+\end{figure}
+\vfill
+\end{landscape}
+\restoregeometry
+
+\subsection{Random goal}
+
+not very successful with DRL. Seems to not even be doing better than BFS and basically impractical for nb\_shuffles $\geq$ 4.
+\\
+I will try with DL trained from Kociemba solutions, to see if $A^{*}$ with this DL heuristic finds more elegant solutions than Kociemba
+
+\newgeometry{top=0mm, bottom=0mm, left=0mm, right=0mm}
+\begin{landscape}
+\centering\vspace*{\fill}
+\begin{figure}[H]
+\centering
+\includegraphics[scale=0.50]{./Figures/222WCPerformance.jpeg}
+%\decoRule
+\caption[222RCPerformance]{Solvers' performance comparison 2x2x2 Random-Goal-RUbik's}
+\label{fig:222WCPerformance}
+\end{figure}
+\vfill
+\end{landscape}
+\restoregeometry
 
@@
 
-blabla
+For now just showing Kociemba as a baseline.
+Will first try out DL trained on data generated from Kociemba, then will try out DRL.
+If unsuccessful or too slow, might try similar to paper using DQL/MC search.
+
+\newgeometry{top=0mm, bottom=0mm, left=0mm, right=0mm}
+\begin{landscape}
+\centering\vspace*{\fill}
+\begin{figure}[H]
+\centering
+\includegraphics[scale=0.50]{./Figures/333RCPerformance.jpeg}
+%\decoRule
+\caption[333RCPerformance]{Solvers' performance comparison 3x3x3 RC}
+\label{fig:333RCPerformance}
+\end{figure}
+\vfill
+\end{landscape}
+\restoregeometry

### 2022-08-17 rubiks/learners/deepreinforcementlearner.py

@@ -17,7 +17,8 @@ from torch.optim.lr_scheduler import ExponentialLR
 ########################################################################################################################
 from rubiks.deeplearning.deeplearning import DeepLearning
 from rubiks.learners.learner import Learner
-from rubiks.utils.utils import ms_format, h_format, pformat, to_pickle, get_model_file_name, snake_case, billion, isinf
+from rubiks.utils.utils import ms_format, h_format, pformat, to_pickle, \
+    get_model_file_name, snake_case, billion, isinf, number_format
 ########################################################################################################################
 
 
@@ -203,6 +204,13 @@ class DeepReinforcementLearner(Learner):
         self.loss_latency = latency_data[cls.latency_loss_tag] if latency_data is not None else 0
         self.back_prop_latency = latency_data[cls.latency_back_prop_tag] if latency_data is not None else 0
 
+    def read_data_file(self):
+        data = read_pickle(self.learning_file_name)
+        if self.config_tag not in data.keys():
+            """ try copy """
+            data = read_pickle(self.learning_file_name + '.copy')
+        return data
+
     def __init__(self, **kw_args):
         Learner.__init__(self, **kw_args)
         self.epoch_latency = None
@@ -219,7 +227,7 @@ class DeepReinforcementLearner(Learner):
         self.attempt_recovery = self.learning_file_name is not None and isfile(self.learning_file_name)
         if self.attempt_recovery:
             try:
-                data = read_pickle(self.learning_file_name)
+                data = self.read_data_file()
                 config = data[self.config_tag]
                 Learner.__init__(self, **config)
                 self.convergence_data = data[self.convergence_data_tag]
@@ -236,7 +244,7 @@ class DeepReinforcementLearner(Learner):
                 self.target_network_count = 0
                 self.log_info('Recovering from convergence_data: ', self.convergence_data.iloc[-1])
             except Exception as error:
-                self.log_error('Could not recover from \'%d\': ' % self.learning_file_name,
+                self.log_error('Could not recover from \'%s\': ' % self.learning_file_name,
                                error)
                 self.attempt_recovery = False
         if not self.min_no_loop:
@@ -304,10 +312,10 @@ class DeepReinforcementLearner(Learner):
         total_run_time = self.epoch_latency / top[cls.epoch]
         total_run_time *= top[cls.nb_epochs] - top[cls.epoch]
         self.log_info('Estimated max run time left: ', h_format(total_run_time),
-                      '. Convergence update at loss <= %.4f' % (top[cls.max_max_target] *
[…]

### 2022-08-17 rubiks/puzzle/puzzled.py

@@ -39,6 +39,8 @@ class Puzzled(Parsable):
         return tuple(self.__goal__.dimension())
 
     def get_goal(self):
+        if hasattr(self.__goal__, 'get_goal'):
+            return self.__goal__.get_goal()
         return self.__goal__.clone()
 
     def puzzle_name(self):

### 2022-08-17 rubiks/puzzle/rubikscube.py

@@ -2,7 +2,7 @@
 # Francois Berrier - Royal Holloway University London - MSc Project 2022                                               #
 ########################################################################################################################
 from enum import Enum
-from itertools import product
+from itertools import product, permutations
 from math import factorial
 from numpy.random import randint, permutation
 from pandas import DataFrame, read_pickle
@@ -199,6 +199,7 @@ class RubiksCube(Puzzle):
     goals_map = dict()
     goals_hashes = dict()
     corners_map = dict()
+    corners_target_map = dict()
     edges_map = dict()
     edges_oriented_distances_to_home = dict()
 
@@ -284,6 +285,7 @@ class RubiksCube(Puzzle):
         #cls.goals_map[n] = list()
         #cls.goals_hashes[n] = set()
         cls.corners_map[n] = list()
+        cls.corners_target_map[n] = list()
         cls.edges_map[n] = dict()
         cls.edges_oriented_distances_to_home[n] = dict()
         goal = dict()
@@ -301,6 +303,15 @@ class RubiksCube(Puzzle):
         cls.__add_corner_to_map__((Color.o, Color.b, Color.y), n)
         cls.__add_corner_to_map__((Color.o, Color.w, Color.b), n)
         cls.__add_corner_to_map__((Color.o, Color.g, Color.w), n)
+        cls.corners_target_map[n] = [((Face.F, 0, 0), (Face.U, 1, 0), (Face.L, 0, 1)),
+                                     ((Face.F, 0, 1), (Face.R, 0, 0), (Face.U, 1, 1)),
+                                     ((Face.F, 1, 0), (Face.L, 1, 1), (Face.D, 0, 0)),
+                                     ((Face.F, 1, 1), (Face.D, 0, 1), (Face.R, 1, 0)),
+                                     ((Face.B, 1, 1), (Face.D, 1, 0), (Face.L, 1, 0)),
+                                     ((Face.B, 1, 0), (Face.R, 1, 1), (Face.D, 1, 1)),
+                                     ((Face.B, 0, 0), (Face.U, 0, 1), (Face.R, 0, 1)),
+                                     ((Face.B, 0, 1), (Face.L, 0, 0), (Face.U, 0, 0)),
+                                     ]
         if n == 3:
             """ edges """
             cls.__add_edge_to_map__((Color.r, Color.w), (Face.F, 0, 1, Face.U, 2, 1), n)
@@ -366,7 +377,7 @@ class RubiksCube(Puzzle):
             else:
[…]

### 2022-08-17 rubiks/scripts/draft.py

@@ -1,10 +1,114 @@
-####################################################################
-from math import inf
+########################################################################################################################
+# Francois Berrier - Royal Holloway University London - MSc Project 2022                                               #
+########################################################################################################################
+from rubiks.core.loggable import Loggable
+from rubiks.deeplearning.deeplearning import DeepLearning
+from rubiks.heuristics.deeplearningheuristic import DeepLearningHeuristic
+from rubiks.heuristics.heuristic import Heuristic
+from rubiks.heuristics.perfectheuristic import PerfectHeuristic
+from rubiks.learners.learner import Learner
+from rubiks.learners.deepreinforcementlearner import DeepReinforcementLearner
 from rubiks.puzzle.puzzle import Puzzle
-####################################################################
-puzzle_type=Puzzle.rubiks_cube
-n=2
-nb_moves=inf
-print(Puzzle.factory(**globals()).apply_random_moves(nb_moves))
-####################################################################
+from rubiks.solvers.solver import Solver
+from rubiks.utils.utils import get_model_file_name, get_shuffles_file_name, get_performance_file_name
+########################################################################################################################
+
+
+if '__main__' == __name__:
+    """ Just create a logger to print some stuff in this script """
+    logger = Loggable(name=__file__)
+    """ What we want to do {Solver.do_solve,
+                            Solver.do_performance_test,
+                            Solver.do_plot,
+                            Solver.do_cleanup_performance_file,
+                            Solver.do_cleanup_shuffles_file,
+                            } """
+    action_type = Solver.do_plot
+    """ What puzzle """
+    puzzle_type = Puzzle.rubiks_cube
+    n = 3
+    m = 2
+    dimension = Puzzle.factory(**globals()).dimension()
+    """ How much to shuffle """
+    nb_shuffles = 0
+    """ For performance test """
+    nb_samples = 50
+    min_nb_shuffles = 0
+    max_nb_shuffles = 4
+    step_nb_shuffles = 2
+    add_perfect_shuffle = False
+    nb_cpus = 5
+    performance_file_name = get_performance_file_name(puzzle_type, dimension)
+    shuffles_file_name = get_shuffles_file_name(puzzle_type, dimension)
+    append = True
+    verbose = True
+    do_not_reattempt_failed = False
+    """ For plot """
+    loc = 'upper center'
+    performance_metrics = [Solver.pct_solved,
+                           #Solver.pct_optimal,
[…]

### 2022-08-17 rubiks/solvers/kociembasolver.py

@@ -49,7 +49,7 @@ class KociembaSolver(Solver):
         return moves
 
     def known_to_be_optimal(self):
-        return self.get_puzzle_dimension()[0] == 2
+        return self.get_puzzle_dimension()[0] == 2# and self.get_puzzle_type() is RubiksCube
 
     def __do_import__(self):
         if self.__class__.__import_done__:

### 2022-08-17 unittests/testrubikscube.py

@@ -201,11 +201,16 @@ class TestRubiksCube(TestCase):
     def test_edge_permutations_parity(self):
         logger = Loggable(name='test_edge_permutations_parity')
         for _ in range(100):
-            cube = RubiksCube(n=3).apply_random_moves(10)
-            parity = cube.edge_permutations_parity()
-            logger.log_info(parity)
+            cube = RubiksCube(n=3, init_from_random_goal=True)
+            #e_parity = cube.edge_permutations_parity()
+            #logger.log_info(e_parity)
+            c_parity = cube.corner_permutations_parity()
+            logger.log_info('corner parity: ', c_parity)
+            #logger.log_info('total parity: ', cube.total_permutations_parity())
 
 
-
+            cube = RubiksCube(n=3, init_from_random_goal=True).apply_random_moves(1)
+            c_parity = cube.corner_permutations_parity()
+            logger.log_info('corner parity one move: ', c_parity)
 
 ########################################################################################################################

### 2022-08-26 latex/Chapters/Conclusion.tex
update

--- 2022-06-19 latex/Chapters/Chapter0.tex
@@
 
-\chapter{Objectives} % Main chapter title
+\chapter{Conclusion} % Main chapter title
 
-\label{Chapter0} % Change X to a consecutive number; for referencing this chapter elsewhere, use \ref{ChapterX}
+\label{Conclusion} % Change X to a consecutive number; for referencing this chapter elsewhere, use \ref{ChapterX}
 
@@
 
-\section{Learning Objectives}
-
-Back when I studied Financial Mathematics, almost 2 decades ago, it was all about probability theory, stochastic calculus and asset (in particular derivatives) pricing.
-These skills were of course very sought after in the field of options trading, but were also often enough to get a job in algorithmic or systematic trading.
-By the middle of the 2010s, with the constant advances in computing power and storage, the better availability of off-the-shelves libraries and data sets, I witnessed a first revolution: the field of machine learning became more and more prominent and pretty much overshadowed other (more traditional maths) skills.
-More recently, a  second revolution has taken not only the world of finance, but that of pretty much every science and industry, by storm: we are now in the artificial intelligence age.
-In 2019-2020, I decided it was time to see by myself what this was all about, and if the hype was justified. What better way to do that than embark on a proper MSc in Artificial Intelligence?
-\\
-Of all the modules I have studied over the last two years of the Royal Holloway MSc in AI, I have been the most impressed by DL and  NLP (itself arguably largely an application of DL) and particularly interested in AIPnT, especially our excursion in the field of graphs search (a very traditional CS topic, but which somehow I had not yet had a chance to study in much details).
-Even though I still believe there is a tremendous amount of malinvestment  everywhere, due in good part to the inability of the average investor to distinguish between serious and scammy AI applications and startups (the same obviously goes for blockchain applications, which might warrant another MSc?), I have totally changed my mind around the potential of DL, DRL and NLP and think they are incredibly promising.
-I have been astonished to see by myself, through several of the courseworks we have done during the MSc, how incredibly efficient sophisticated ML, DL and DLR algorithms can be, when applied well on the right problems.
-Sometimes they just vastly outperform more naive and traditional approaches to the point of rendering older approaches entirely obsolete (e.g CV, NLP, game solvers, etc...).
-\\
-For the project component of the MSc, I thought it would be interesting (and fun) for me to try and apply some of the DL, DRL and search techniques (from AIPnT) to a couple of single-player games, such as the sliding puzzle (of which some variations are well known under different names, e.g.
-the 8-puzzle and 15-puzzle) and of course the Rubiks' cube.
-I am in particular looking to solidify my understanding of DRL by  implementing and experimenting with concrete (though arguably of limited practical use) problems.
-
-%----------------------------------------------------------------------------------------
-%	SECTION 2
-%----------------------------------------------------------------------------------------
-
-\section{Project's Objectives}
-
-I am hoping with this project to implement and compare a few different methods to solve the Sliding Puzzle and the Rubik's Cube.
-Both these games have tremendously large state spaces (see section \hyperref[sec:Games]{Games} for details) and only one goal state.
-I am therefore likely to only succeed with reasonably small dimensional puzzles, especially since I have chosen for simplicity to implement things in Python.
-Depending on the progress I will be able to make in the imparted time, I am hoping to try a mix of simple searches (depth first search, breadth first search, A* with simple admissible heuristics), then more advanced ones such as A* informed by heuristics learnt via DL and DRL.
-Time permitting I would like to give a go at DQL, and maybe also compare things with some open-source domain-specific implementations (for instance a Kociemba Rubik's algorithm implementation, see e.g.
-\cite{Kociemba}).
-\\
-Along the way, I am also hoping to learn a bit about these two games that I have chosen to work on, and maybe make a couple of remarks on them that the reader of this thesis might find interesting.
+TBW

### 2022-08-26 latex/Chapters/Objectives.tex
update

@@
 \\
-Of all the modules I have studied over the last two years of the Royal Holloway MSc in AI, I have been the most impressed by DL and  NLP (itself arguably largely an application of DL) and particularly interested in AIPnT, especially our excursion in the field of graphs search (a very traditional CS topic, but which somehow I had not yet had a chance to study in much details).
-Even though I still believe there is a tremendous amount of malinvestment  everywhere, due in good part to the inability of the average investor to distinguish between serious and scammy AI applications and startups (the same obviously goes for blockchain applications, which might warrant another MSc?), I have totally changed my mind around the potential of DL, DRL and NLP and think they are incredibly promising.
+\\
+Of all the modules I have studied over the last two years of the Royal Holloway MSc in AI, I have been most impressed by DL and  NLP (itself arguably largely an application of DL) and particularly interested in AIPnT, especially our excursion in the field of graphs search (a very traditional CS topic, but which somehow I had not yet had a chance to study in much details).
+\\
+\\
+Even though I still believe there is a tremendous amount of malinvestment, due in good part to the inability of the average investor to distinguish between serious and scammy AI applications and startups (the same obviously goes for blockchain applications, which might warrant another MSc?), I have totally changed my mind around the potential of DL, DRL and NLP and think they are incredibly promising.
 I have been astonished to see by myself, through several of the courseworks we have done during the MSc, how incredibly efficient sophisticated ML, DL and DLR algorithms can be, when applied well on the right problems.
@@
 \\
+\\
 For the project component of the MSc, I thought it would be interesting (and fun) for me to try and apply some of the DL, DRL and search techniques (from AIPnT) to a couple of single-player games, such as the sliding puzzle (of which some variations are well known under different names, e.g.
-the 8-puzzle and 15-puzzle) and of course the Rubiks' cube.
-I am in particular looking to solidify my understanding of DRL by  implementing and experimenting with concrete (though arguably of limited practical use) problems.
+the 8-puzzle and 15-puzzle) and of course the Rubik's cube.
+I am in particular looking to solidify my understanding of DRL and DQL by  implementing and experimenting with concrete (though arguably of limited practical use) problems.
 
@@
 I am hoping with this project to implement and compare a few different methods to solve the SP and the RC.
-Both these puzzles have tremendously large state spaces (see section \hyperref[sec:Games]{Games} for details) and only one goal state.
+Both these puzzles have tremendously large state spaces (see section \hyperref[sec:Puzzles]{Puzzles} for details) and only one goal state.
 I am therefore likely to only succeed with reasonably small dimensional puzzles, especially since I have chosen for simplicity to implement things in Python.
+\\
+\\
 Depending on the progress I will be able to make in the imparted time, I am hoping to try a mix of simple searches (depth first search, breadth first search, A* with simple admissible heuristics), then more advanced ones such as A* informed by heuristics learnt via DL and DRL, as well as try different architectures and network sizes and designs for the DL and DRL heuristics.
@@
 \\
-Along the way, I am also hoping to learn a bit about these two games that I have chosen to work on, and maybe make a couple of remarks on them that the reader of this thesis might find interesting.
+\\
+Along the way, I am also hoping to learn a bit about these two puzzles that I have chosen to work on, and maybe make a couple of remarks on them that the reader of this thesis might find interesting.

### 2022-08-26 latex/Chapters/Rubiks.tex
update

@@
 
-\chapter{Results - Rubiks' Cube} % Main chapter title
+\chapter{Results - Rubik's Cube} % Main chapter title
 
@@
 
+\begin{center}
+\begin{tabular}{l*{5}{c}r}
+\hline
+\textbf{solver}      & & \textbf{Rubik's} & \textbf{2x2x2} & \textbf{3x3x3} \\
+\hline
+BFS   &   &        & x   &   \\
+\hline
+Kociemba   &   &      & x  &   \\
+\hline
+A$^{*}$[Kociemba]  &   &  & x & \\
+\hline
+A$^{*}$[DL[A$^{*}$[Kociemba]]]  &   &  & x & \\
+\hline
+A$^{*}$[DRL]  &   &  & x &  \\
+\hline
+A$^{*}$[DQL]  &   &  & x &  \\
+\hline
+MCTS[DQL][c=0]  &   &  & x &  \\
+\hline
+MCTS[DQL][c=69]  &   & & x &  \\
+\hline
+\end{tabular}
+\end{center}
+
 \section{2x2x2}
-
-\subsection{Rubiks}
 
@@
 \centering
-\includegraphics[scale=0.50]{./Figures/222RCPerformance.jpeg}
+\includegraphics[scale=0.60]{./Figures/222RCPerformance}
 %\decoRule
@@
 \label{fig:222RCPerformance}
-\end{figure}
-\vfill
-\end{landscape}
-\restoregeometry
-
-\subsection{Random goal}
-
-not very successful with DRL. Seems to not even be doing better than BFS and basically impractical for nb\_shuffles $\geq$ 4.
-\\
-I will try with DL trained from Kociemba solutions, to see if $A^{*}$ with this DL heuristic finds more elegant solutions than Kociemba
-
-\newgeometry{top=0mm, bottom=0mm, left=0mm, right=0mm}
-\begin{landscape}
-\centering\vspace*{\fill}
-\begin{figure}[H]
-\centering
-\includegraphics[scale=0.50]{./Figures/222WCPerformance.jpeg}
-%\decoRule
-\caption[222RCPerformance]{Solvers' performance comparison 2x2x2 Random-Goal-RUbik's}
-\label{fig:222WCPerformance}
 \end{figure}
@@
 If unsuccessful or too slow, might try similar to paper using DQL/MC search.
+\\
+Notice Kociemba 3, unlike the 2x2x2 implementation that I found on github, does not do automatic color mapping.
+So if you present it with a cube which is not with the \textit{standard} centers (facing red and white up), it will actually not solve the cube.
+I have therefore added a bit of logic in Kociemba solver to look for the equivalent cube among the 24 equivalent cubes to the one being solved, with standard colors.
+We then solve this one using Kociemba and add the whole cube rotations to the solution (which in my code are deemed to have 0 cost anyway).
+This way, not only can I use Kociemba 3 to solve any (solvable) 3x3x3 cube irrespective of rotations, but I can also use that in A* or to train a DL network (also for A*)
 

### 2022-08-26 latex/MScAIFrancoisBerrier.tex
update

@@
 
-\thesistitle{Solving \\ the \\ Sliding Puzzle \\ \& \\ Rubiks' Cube \\by \\Deep Reinforcement Learning} % Your thesis title, this is used in the title and abstract, print it elsewhere with \ttitle
+\thesistitle{Solving \\ the \\ 15-Puzzle \& Rubik's Cube \\by \\Deep Reinforcement Learning} % Your thesis title, this is used in the title and abstract, print it elsewhere with \ttitle
 \supervisor{Pr. Chris \textsc{Watkins}} % Your supervisor's name, this is used in the title page, print it elsewhere with \supname
 \examiner{} % Your examiner's name, this is not currently used anywhere in the template, print it elsewhere with \examname
-\degree{MSc in Artificial Intelligence} % Your degree name, this is used in the title page and abstract, print it elsewhere with \degreename
-\author{Francois \textsc{Berrier}} % Your name, this is used in the title page and abstract, print it elsewhere with \authorname
+\degree{\textbf{MSc in Artificial Intelligence}} % Your degree name, this is used in the title page and abstract, print it elsewhere with \degreename
+\author{Fran\c{c}ois \textsc{Berrier}} % Your name, this is used in the title page and abstract, print it elsewhere with \authorname
 \addresses{} % Your address, this is not currently used anywhere in the template, print it elsewhere with \addressname
@@
 \vspace*{.06\textheight}
-{\scshape\LARGE \univname\par}\vspace{1.5cm} % University name
+{\scshape\LARGE \univname\par}\vspace{1cm} % University name
 \textsc{\Large MSc Thesis}\\[0.5cm] % Thesis type
@@
 \end{minipage}
-\begin{minipage}[t]{0.4\textwidth}
+\begin{minipage}[t]{0.5\textwidth}
 \begin{flushright} \large
@@
 \end{flushright}
-\end{minipage}\\[3cm]
+\end{minipage}
 
@@
 
+\begin{figure}[H]
+\centering
+\includegraphics[scale=0.3]{./Figures/RHUL}
+%\decoRule
+\end{figure}
 \large \textit{A thesis submitted in fulfillment of the requirements\\ for the degree of \degreename}\\[0.3cm] % University requirement text
@@
 \\
-Inspired however by the seminal success of Deep Mind's team in using Deep Reinforcement Learning (\textbf{DRL}) to play Atari games (\cite{Mnih2013})  and to master the game of Go (\cite{AlphaGo}), researchers have in recent years made a lot of progress towards designing algorithms capable of learning and solving, \textit{without human knowledge}, the Rubiks Cube (\textbf{RC}) - as well as similar single player puzzles - by using Deep Q-Learning (\textbf{DQL})  (\cite{DBLP:journals/corr/abs-1805-07470}) or search and (\textbf{DRL}) value iteration (\cite{https://doi.org/10.48550/arxiv.1805.07470}).
+Inspired however by the seminal success of Deep Mind's team in using Deep Reinforcement Learning (\textbf{DRL}) to play Atari games (\cite{Mnih2013})  and to master the game of Go (\cite{AlphaGo}), researchers have in recent years made a lot of progress towards designing algorithms capable of learning and solving, \textit{without human knowledge}, the Rubik's Cube (\textbf{RC}) - as well as similar single player puzzles - by using Deep Q-Learning (\textbf{DQL})  (\cite{DBLP:journals/corr/abs-1805-07470}) or search and (\textbf{DRL}) value iteration (\cite{https://doi.org/10.48550/arxiv.1805.07470}).
 \\
@@
 I will then give an overview in chapter 4 of the code base I have developed - and put in the open on my github page (\cite{FB}) - to complete this project before detailing a few examples in chapter 5.
-Finally, in chapters 6 and 7, I will present all my various results on respectively the sliding puzzle and the Rubiks' cube.
+Finally, in chapters 6 and 7, I will present all my various results on respectively the sliding puzzle and the Rubik's cube before offering some conclusion in chapter 8.
 
@@
 
+\hypersetup{%
+  colorlinks = true,
+  linkcolor  = black
+}
 \tableofcontents % Prints the main table of contents
@@
 \include{Chapters/Puzzles}
-\include{Chapters/Code}
+\include{Chapters/Implementation}
 \include{Chapters/Examples}
@@
 \include{Chapters/Rubiks}
+\include{Chapters/Conclusion}
 

### 2022-08-26 rubiks/heuristics/kociembaheuristic.py

@@ -1,0 +1,25 @@
+########################################################################################################################
+# Francois Berrier - Royal Holloway University London - MSc Project 2022                                               #
+########################################################################################################################
+from rubiks.heuristics.heuristic import Heuristic
+from rubiks.puzzle.puzzle import Puzzle
+from rubiks.solvers.kociembasolver import KociembaSolver
+########################################################################################################################
+
+
+class KociembaHeuristic(Heuristic):
+    """ Makes use of Kociemba solver """
+
+    @classmethod
+    def known_to_be_admissible(cls):
+        return False
+
+    def __init__(self, **kw_args):
+        Heuristic.__init__(self, **kw_args)
+        self.solver = {n: KociembaSolver(puzzle_type=Puzzle.rubiks_cube, n=n) for n in[2, 3]}
+
+    def cost_to_go_from_puzzle_impl(self, puzzle):
+        return self.solver[puzzle.n].solve(puzzle).cost
+
+########################################################################################################################
+

### 2022-08-26 rubiks/learners/deepreinforcementlearner.py

@@ -528,7 +528,7 @@ class DeepReinforcementLearner(Learner):
 
     def plot_learning(self):
         cls = self.__class__
-        data = self.read_data_file()
+        data = DeepReinforcementLearner.read_data_file(self)
         if isinstance(data[self.convergence_data_tag][cls.learning_rate].iloc[0], list):
             data[self.convergence_data_tag].loc[:, cls.learning_rate] = \
                 data[self.convergence_data_tag][cls.learning_rate].apply(lambda lr: lr[0])

### 2022-08-26 rubiks/puzzle/puzzle.py

@@ -100,9 +100,9 @@ class Puzzle(Factory, metaclass=ABCMeta):
         returns:
             list of puzzles
         """
-        init = cls(**kw_args)
         training_data = list()
         for _ in range(nb_sequences):
+            init = cls(**kw_args)
             moves = init.random_moves(nb_shuffles, min_no_loop=min_no_loop)
             puzzles = init.get_puzzle_sequence(moves)
             if one_list:

### 2022-08-26 rubiks/puzzle/trainingdata.py

@@ -2,14 +2,14 @@
 # Francois Berrier - Royal Holloway University London - MSc Project 2022                                               #
 ########################################################################################################################
 from functools import partial
-from math import inf
+from math import inf, ceil
 from multiprocessing import Pool
 from pandas import read_pickle, Series
 ########################################################################################################################
 from rubiks.core.loggable import Loggable
 from rubiks.puzzle.puzzled import Puzzled
 from rubiks.solvers.solver import Solver
-from rubiks.utils.utils import to_pickle, get_training_file_name, remove_file, is_inf
+from rubiks.utils.utils import to_pickle, get_training_file_name, remove_file, is_inf, number_format
 ########################################################################################################################
 
 
@@ -21,6 +21,9 @@ class TrainingData(Loggable, Puzzled):
     puzzles_tag = 'puzzles_tag'
     nb_cpus = 'nb_cpus'
     time_out = 'time_out'
+    chunk_size = 'chunk_size'
+    verbose = 'verbose'
+    save_full_puzzles = 'save_full_puzzles'
 
     @classmethod
     def populate_parser(cls, parser):
@@ -32,8 +35,20 @@ class TrainingData(Loggable, Puzzled):
                          type=int,
                          default=1)
         cls.add_argument(parser,
+                         field=cls.chunk_size,
+                         type=int,
+                         default=1)
+        cls.add_argument(parser,
                          field=cls.time_out,
                          default=inf)
+        cls.add_argument(parser,
+                         cls.verbose,
+                         default=False,
+                         action=cls.store_true)
+        cls.add_argument(parser,
+                         cls.save_full_puzzles,
+                         default=False,
+                         action=cls.store_true)
 
     def __init__(self, **kw_args):
         Loggable.__init__(self, **kw_args)
@@ -43,24 +58,17 @@ class TrainingData(Loggable, Puzzled):
         if not self.training_data_file_name:
             self.training_data_file_name = get_training_file_name(puzzle_type=self.get_puzzle_type(),
                                                                   dimension=self.get_puzzle_dimension())
-        try:
-            self.data = read_pickle(self.training_data_file_name)
-        except FileNotFoundError as error:
-            self.log_warning('Could not fetch training data: ', error)
-            self.data = None
-        if not isinstance(self.data, dict):
-            self.data = dict()
[…]

### 2022-08-26 rubiks/scripts/sp25learner.py

@@ -41,16 +41,16 @@ if '__main__' == __name__:
         if learner_type is Learner.deep_reinforcement_learner \
         else DeepLearner.default_plot_metrics
     """ If it's a perfect learner config here """
-    time_out = 60
-    nb_cpus = 1
-    cpu_multiplier = 5
+    time_out = 3600
+    nb_cpus = 8
+    cpu_multiplier = 1
     max_puzzles = nb_cpus * cpu_multiplier * 10000
     regular_save = nb_cpus * cpu_multiplier * 1
     save_timed_out_max_puzzles = 10000
     after_round_save = True
     flush_timed_out_puzzles = False
     save_timed_out = True
-    rerun_timed_out = False
+    rerun_timed_out = True
     rerun_timed_out_only = False
     abort_after_that_many_consecutive_timed_out = 100
     nb_shuffles_from_goal = 10

### 2022-08-26 rubiks/scripts/trainingdata.py

@@ -11,14 +11,14 @@ from rubiks.puzzle.trainingdata import TrainingData
 
 if '__main__' == __name__:
     logger = Loggable(name=__file__)
-    puzzle_type = Puzzle.watkins_cube
+    puzzle_type = Puzzle.rubiks_cube
     n = 2
-    m = 2
-    nb_cpus = 10
-    nb_sequences = nb_cpus * 100
-    time_out = 3600
+    nb_cpus = 1
+    chunk_size=0
+    verbose=False
+    nb_sequences = nb_cpus * 1000
     training_data = TrainingData(**globals())
-    repeat = 5
+    repeat = 1
     for nb_shuffles in [inf]:
         training_data.generate(nb_shuffles=nb_shuffles,
                                nb_sequences=nb_sequences,

### 2022-08-26 unittests/testsolver.py

@@ -165,7 +165,8 @@ class TestSolver(TestCase):
                                    nb_sequences=100)
             logger.log_info(learner.get_config())
             learner.learn()
-        c = 100
+        c = 1
+        logger.log_info('Running with c=', c, ' and run_trim=', run_trim)
         random_choice = False
         # Then we use this learning to solve
         solver = Solver.factory(solver_type=Solver.mcts,
@@ -210,8 +211,8 @@ class TestSolver(TestCase):
             if solution_1.cost > 10*nb_moves:
                 logger.log_warning('Puzzle # %d ' % (nb_puzzle + 1), puzzle)
             if solution_2.cost != solution_1.cost:
-                logger.log_info('Puzzle # %d ' % (nb_puzzle + 1), puzzle)
-                logger.log_info('Puzzle %d improvement nb_moves %d solutions costs %d -> %d' % (nb_puzzle + 1,
+                logger.log_debug('Puzzle # %d ' % (nb_puzzle + 1), puzzle)
+                logger.log_debug('Puzzle %d improvement nb_moves %d solutions costs %d -> %d' % (nb_puzzle + 1,
                                                                                                 nb_moves,
                                                                                                 solution_1.cost,
                                                                                                 solution_2.cost))

### 2022-08-28 latex/Chapters/DRLSearch.tex
update

--- 2022-06-23 latex/Chapters/DRLSearch.tex
@@
 
+In this section, I will succintly go through the different techniques that underly the algorithms I have implemented and compared in this project.
+Along the way, I will give some references for the interested reader.
+
 %----------------------------------------------------------------------------------------
 %	SECTION 1
-%----------------------------------------------------------------------------------------
-
-\section{Reinforcement Learning}
-
-blabla
-
-%-----------------------------------
-%	SECTION 2
-%-----------------------------------
-
-\section{Deep Learning}
-
-blabla
-
-%----------------------------------------------------------------------------------------
-%	SECTION 3
 %----------------------------------------------------------------------------------------
@@
 
-See \cite{DBLP:journals/jacm/DechterP85}
+It is quite fruitful to think of single-player, deterministic, fully-observable puzzles such as the \textbf{SP} and \textbf{RC} as unit-cost graphs.
+We start from an initial \textit{node} (usually some scrambled configuration of the \textbf{SP} or \textbf{RC}), and via legal moves, we transition to new nodes, until hopefully we manage to get to the goal in the shortest number of moves possible.
+\\
+\\
+Graph search (or graph traversal) is a rather large, sophisticated and mature branch of \textbf{CS} that studies strategies to find \textit{optimal} paths from initial to goal node in a graph.
+What is meant by \textit{optimal} can vary, but usually is concerned with one or several of minimizing the run-time of the strategies, their memory complexity/usage or the total cost/length of the solutions they come up with.
+In the general theory, different transitions can have different costs associated with them.
+Since each move of a tile in the \textbf{SP} or each rotation of a face in the \textbf{RC} might reasonably be thought as taking a constant and identical amount of time to perform, cost and length of solutions are one and the same, and I will therefore consider the graphs to be unit-cost (i.e.
+all transitions/moves cost 1).
+\\
+\\
+Strategies typically maintain a frontier, which is the collection of unexpanded nodes so far.
+They keep expanding the frontier, by choosing the next node to expand (i.e.
+adding all the nodes reachable via legal transitions from that node) until a solution is found.
+The question is how to choose the next node to expand!
+Some graph search strategies are said to be uninformed, in the sense that they do not exploit any domain-specific knowledge or insight about what the graphs represent to choose the next node to expand.
+In that category fall for instance Breadth First Search (\textbf{BFS}) and Depth First Search (\textbf{DFS}), which as their name suggest expand nodes from the frontier based on, respectively, a \textbf{FIFO} and \textbf{LIFO} policy.
+\textbf{BFS} and \textbf{DFS} will only work for small problems where the number of transitions and states are reasonably small; they can however work on a wide variety of problems.
+It is quite easy to see that \textbf{BFS} is an \textit{optimal} search strategy, in that when it does find a solution, that solution is of smallest possible cost.
+\textbf{DFS} is obvisouly not.
+\\
+\\
+Informed strategies on the other hand exploit domain-specific knowledge.
+One very popular such strategy is \textbf{A$^{*}$} (see \cite{DBLP:journals/jacm/DechterP85}), which always first examine the node of smallest expected total cost (\textbf{f}), itself the sum of the cost-so-far from initial to current node (\textbf{g}) plus the expected cost-to-go from current node to solution (\textbf{h}).
+The expected cost-to-go \textbf{h} is often called a \textit{heuristic}.
+The better the heuristic is, the better\textbf{A$^{*}$} usually performs.
+An important property of heuristics is \textit{admissibility}.
+In short, admissible heuristics never over-estimate the real cost-to-go, and can easily be shown to render\textbf{A$^{*}$} optimal.
+
+%-----------------------------------
+%	SECTION 2
+%-----------------------------------
+
+\section{Reinforcement Learning}
+
+As mentioned in the abstract, \textbf{RL} (see \cite{Sutton1998} for a brilliant exposition of the basic concepts and theory) has known a bit of false start in the 1950s and 1960s but recently been extremely successfully applied to a variety of problems (from Atari to board games and puzzles, robotics and else).
+\textbf{RL} is concerned with how intelligent agents can learn optimal decisions from observing/experiencing rewards while interacting in their environment.
+One of the early concept in the field is that of value function $s \to \textbf{V}(s)$, which tells us the maximum expected reward the agent can obtain from a given state $s$ (if it takes the optimal decisions).
+In discrete states and transitions spaces, the value function can be remarkably computed (or at least approximated) by a rather mechanical and magic-like procedure called value-iteration (which uses the equivalent of the Bellman equation from optimal control), where each state $s$'s value is iteratively refined by combining the value of $s$'s reachable states.
+
+%----------------------------------------------------------------------------------------
+%	SECTION 3
+%----------------------------------------------------------------------------------------
+
+\section{Deep Learning}
+
+Quite similarly to \textbf{RL}, Deep Learning (\textbf{DL}) has not initially had the success the \textbf{AI} community was hoping it would.
+Marvin Minsky, who often took the blame for having \textit{killed} funding into the field, even regretted writing his 1969 book (\cite{minskypapert69}) in which he had proved that three-layer feed forward perceptrons were not quite the universal functions approximators some had hypothesied them to be.
+Since around 2006 (see \cite{GoodBengCour16}for more history of the recent burgeoning of \textbf{DL}), the field has known a rebirth, probably due to combination of factors, from ever more powerful computers and larger storage, the availability of data sets large and small on the internet, the advances in many techniques and heuristics (backward propagation, normalisation, drop-outs, etc..) and the huge amount of experimentation with different architectures (from very deep feed forward fully connected networks, to convolutional, recurrent and more exoctic networks).
+It is not an exageration to say that \textbf{DL} has rendered obsolete entire fields of research and bodies of knowledge e.g.
+in \textbf{CV} and \textbf{NLP}.
+\textbf{DL} has often become one of the methods of choice to learn of approximate highly dimensional functions.
+The beauty of it is that the relevant features that go into the function are autodiscovered by the training of the network, via back-ward propagation and trial-and-error, rather than having to be postulated or handcrafted as is often the case with other \textbf{ML} techniques.
+\\
+\\
+For this project, I have implemented this in a class called DeepLearner (see section \ref{sec:codelearners} for details)
 
@@
 
-\section{Deep Reinforcement Learning Heuristics}
+\section{DL and DLR Heuristics}
 
-blabla
+So what is the link between all of these - A$^{*}$'s heuristics, \textbf{DL}, \textbf{RL}'s value iteration?
+\\
+\\
+Sometimes we have ways of computing good solutions/heuristics to our puzzles, but cannot computationally do this for all possible states (maybe the state space is just too large!).
+One approach in that case is to compute the solution/heuristic for some states, and let a \textbf{DL} model learn how to extrapolate to the rest of the state space.
+This is a case of supervised learning in some sense, because we need a teacher-solver to tell us good solutions or heuristics.
+\\
+\\
+An alternative approach in order to compute the cost-to-go is via \textbf{RL} value-iteration.
+Indeed, minimizing the cost-to-go (heuristic) is equivalent to maximising the cumulative reward given by -1 for each transition not reaching the goal state and +1 for the last transition to the goal.
+Clearly these two quantities are equivalent (modulo a trivial linear transformation).
+We can perform this value-iteration in an unsupervised way: we do not need a teacher to tell us the target value of the value function/cost-to-go but instead learn it simply from refining a state $s$'s value by replacing it at each iteration by the minimum of its current value, and the minimum value of all reachable states from $s$ polus one (since it takes one move to reach them).
+That is, for $t \in \mathbb{N}^{*}, s \in \mathcal{S}$, the state space, we update $V_{t}(s)$ via:
+\begin{equation} \label{eq:DLVI}
+V_{t+1}(s) \leftarrow min \Big( V_{t}(s), 1 + min_{s' \in \mathcal{C}(s)}(V_{t}(s'))  \Big)
+\end{equation}
+where $\mathcal{C}(s)$ are the children (reachable states via 1 move) of $s$
+\\
+The state space is often so large however, that we cannot practically perform this procedure for all states.
+This is where \textbf{DL} comes again to the rescue to act as function approximator, combining with \textbf{RL} into what has become known as \textbf{DRL}.
+The subtelty is then that both the left-hand-side and right hand side in the value-iteration procedure above are given by a \textbf{DL} network which we are constantly tweaking by back-ward propagation.
+For my implementation of \textbf{DRL}, I followed the same approach as in \cite{Mnih2013}, where the right hand-side of \ref{eq:DLVI} is a fixed \textit{target-network}.
+The left hand side is another \textit{current-network} which alone is modified by the \textbf{DL} learning (backward propagation and optimization), until convergence is deemed reached.
+We then update the \textit{target-network} by a copy of the \textit{current-network}, and then keep iterating again, until the whole process is deemed to have converged.
+\\
+\\
+For this project, I have implemented this procedure in a class called DeepReinforcementLearner (see section \ref{sec:codelearners} for details)
 
@@
 
-see \cite{Watkins1992}
+Another concept of \textbf{RL} is that of Q-function $Q(s, a)$, the maximum expected value the agent can obtain from state $s$, taking action $a$ and acting optimally afterwards.
+Notice that the knowledge of $V$ everywhere is equivalent to that of $Q$, since obviously:
+\begin{equation} \label{eq:QI}
+V(s) = max_{a \in \mathcal{A}(s)} Q(s, a)
+\end{equation}
+where $\mathcal{A}(s)$ are the transitions/actions we can take from state $s$. Conversely, we can recover $Q$ from $V$ since:
+\begin{equation} \label{eq:QI2}
+Q(s, a) = R(a) + V(a(s))
+\end{equation}
+where $R(a)$ is the reward of performing action $a$ and $a(s)$ is the state we reach by performing action $a$ from state $s$.
+Similarly to the way one can compute $V$ via value-iteration (at least in finite settings), $Q$ can also be computed by iteration (See \cite{Watkins1992} for details of convergence of that procedure).
+\\
+\\
+Similarly to our discussion about \textbf{DRL}, in cases where the state space is large, it can be useful to approximate $Q$, or some related quantity (such as e.g.
+a probability distribution over the actions that can be taken from $s$) by a neural network, and use a similar procedure to that of the previous section to update the right-hand-side and left-hand-side of the iterations.
+That becomes \textbf{DQL}!
+\\
+\\
+For this project, I have implemented the procedure described in \cite{https://doi.org/10.48550/arxiv.1805.07470} in a class called DeepQLearner (see section \ref{sec:codelearners} for details).
+The network jointly learns the value function (heuristic, or cost-to-go) as well as a probability distribution over the actions.
+
+%----------------------------------------------------------------------------------------
+%	SECTION 6
+%----------------------------------------------------------------------------------------
+
+\section{Monte Carlo Tree Search}
+
+In my implementation of \textbf{DQL}, which follows \cite{https://doi.org/10.48550/arxiv.1805.07470} (and is also the same procedure followed by alpha-go \cite{AlphaGo}), the output of the neural network that is learnt by \textbf{DQL} is no longer just a cost-to-go estimate as with \textbf{DRL}, but a joint cost-to-go and distribution over actions (to be concrete, that means a 5-dimensional vector for the \textbf{SP} and a 13-dimensional vector for the \textbf{RC}).
+This leaves us with the question of how can we use the output of such a network to search for a solution.
+One simple answer (which I have tried) is that we could disregard the distribution and only take into account the cost-to-go by simply using $A^{*}$.
+The rationale for doing so is that we might hope that jointly learning the cost-to-go and the actions via a combination of MSE and cross-entropy-loss leads to better heuristics, since we use more information during the learning and iterations.
+\\
+\\
+Another possibility is to use an algorithm which combines both the actions distribution and the value function to inform its search.
+The Monte Carlo Tree Search (\textbf{MCTS}) is such an algorithm.
+In short, it generates at each step (and possibly in a distributed manner) paths to a \textit{leaf} on the frontier of unexpanded nodes, expands that node and checks if it is a goal state (and keeps going until the goal is found).
+The paths followed from initial state to leaves are decided via a combination of the probability distribution over actions (following the most promising actions) as well as the value function.
+In order to prevent always following the same paths (which might take us very far from the goal), \textbf{MCTS} algorithms usually keep track of actions they have followed already and underweight their probability so that subsequent paths explore other actions.
+\\
+I have implemented the exact \textbf{MCTS} procedure described in \cite{https://doi.org/10.48550/arxiv.1805.07470} in a class called MonteCarloTreeSearch (see section \ref{sec:codesolvers} for details).

### 2022-08-28 rubiks/puzzle/slidingpuzzle.py

@@ -143,6 +143,17 @@ class SlidingPuzzle(Puzzle):
     def find_tile(self, value):
         return tuple(argwhere(value == self.tiles).squeeze().tolist())
 
+    def black_white_string(self):
+        tiles = DataFrame(self.tiles) #array(self.tiles.numpy(), dtype=str)
+        tiles = tiles.stack()
+        tiles[tiles == 0] = ''
+        tiles = tiles.unstack()
+        tiles = '\n'.join(tabulate(DataFrame(tiles),
+                                   headers='keys',
+                                   tablefmt='grid',
+                                   showindex=False).split('\n')[2:])
+        return '\n' + tiles
+
     def __repr__(self):
         tiles = DataFrame(self.tiles) #array(self.tiles.numpy(), dtype=str)
         tiles = tiles.stack()

### 2022-08-28 rubiks/scripts/learner.py

@@ -64,7 +64,7 @@ if '__main__' == __name__:
     nb_epochs = 100000
     training_data_every_epoch = False
     nb_sequences = 1000
-    nb_shuffles = 50 if n == 3 else 20 # (30 if n == 2 else 70) if puzzle_type == Puzzle.watkins_cube else (20 if n == 2 else 35)
+    nb_shuffles = 20
     cap_target_at_network_count = True
     update_target_network_frequency = 2000 if n == 3 else 1000
     update_target_network_threshold = 1e-2

### 2022-08-28 rubiks/scripts/solver.py

@@ -46,7 +46,7 @@ if '__main__' == __name__:
     do_not_reattempt_failed = False
     #skip = (36,39,85,145,146,161,174,190,) <- these 8 took more than 90 mins for 3x3x3 Rubiks'
     """ For plot """
-    loc = 'center'   # 'upper center'
+    loc = 'upper center'   # 'upper center'
     performance_metrics = [Solver.pct_solved,
                            #Solver.pct_optimal,
                            Solver.median_cost,

### 2022-08-28 rubiks/scripts/trainingdata.py

@@ -11,14 +11,14 @@ from rubiks.puzzle.trainingdata import TrainingData
 
 if '__main__' == __name__:
     logger = Loggable(name=__file__)
-    puzzle_type = Puzzle.rubiks_cube
-    n = 2
+    puzzle_type = Puzzle.sliding_puzzle
+    n = 3
     nb_cpus = 1
     chunk_size=0
     verbose=False
     nb_sequences = nb_cpus * 1000
     training_data = TrainingData(**globals())
-    repeat = 1
+    repeat = 10
     for nb_shuffles in [inf]:
         training_data.generate(nb_shuffles=nb_shuffles,
                                nb_sequences=nb_sequences,

### 2022-08-28 rubiks/utils/utils.py

@@ -55,6 +55,12 @@ def snake_case(s):
 ########################################################################################################################
 
 
+def from_snake_case(s):
+    return ''.join([s.capitalize() for s in s.split('_')])
+
+########################################################################################################################
+
+
 class Extension:
     pkl = 'pkl'
 

### 2022-08-28 latex/Chapters/DRLSearch.tex
update

@@
 \\
-Graph search (or graph traversal) is a rather large, sophisticated and mature branch of \textbf{CS} that studies strategies to find \textit{optimal} paths from initial to goal node in a graph.
+Graphs search (or graphs traversal) is a rather large, sophisticated and mature branch of \textbf{CS} that studies strategies to find \textit{optimal} paths from initial to goal node in a graph.
 What is meant by \textit{optimal} can vary, but usually is concerned with one or several of minimizing the run-time of the strategies, their memory complexity/usage or the total cost/length of the solutions they come up with.
 In the general theory, different transitions can have different costs associated with them.
-Since each move of a tile in the \textbf{SP} or each rotation of a face in the \textbf{RC} might reasonably be thought as taking a constant and identical amount of time to perform, cost and length of solutions are one and the same, and I will therefore consider the graphs to be unit-cost (i.e.
-all transitions/moves cost 1).
+In this project however, we can reasonably consider that each move of a tile in the \textbf{SP} or each rotation of a face in the \textbf{RC} might are taking a constant and identical amount of time and effort to perform, and I will therefore consider all the graphs to be unit-cost (i.e.
+all moves have cost 1).
 \\
 \\
-Strategies typically maintain a frontier, which is the collection of unexpanded nodes so far.
-They keep expanding the frontier, by choosing the next node to expand (i.e.
-adding all the nodes reachable via legal transitions from that node) until a solution is found.
+Search strategies typically maintain a frontier, which is the collection of unexpanded nodes so far.
+They keep expanding the frontier, by choosing the next node to expand (expanding a node simply means adding all of its children nodes - i.e.
+those which are reachable via legal transitions - to the frontier for future evaluation) until a solution is found.
 The question is how to choose the next node to expand!
-Some graph search strategies are said to be uninformed, in the sense that they do not exploit any domain-specific knowledge or insight about what the graphs represent to choose the next node to expand.
+Some graphs search strategies are said to be \textit{uninformed}, in the sense that they do not exploit any domain-specific knowledge or insight about what the graphs represent to choose the next node to expand.
 In that category fall for instance Breadth First Search (\textbf{BFS}) and Depth First Search (\textbf{DFS}), which as their name suggest expand nodes from the frontier based on, respectively, a \textbf{FIFO} and \textbf{LIFO} policy.
-\textbf{BFS} and \textbf{DFS} will only work for small problems where the number of transitions and states are reasonably small; they can however work on a wide variety of problems.
-It is quite easy to see that \textbf{BFS} is an \textit{optimal} search strategy, in that when it does find a solution, that solution is of smallest possible cost.
-\textbf{DFS} is obvisouly not.
+\textbf{BFS} and \textbf{DFS} will only work for modest problems where the number of transitions and states are reasonably small; they can however work in a wide variety of situations as they make no assumptions and require no knowledge.
+It is quite easy to see that \textbf{BFS} is an \textit{optimal} search strategy, in that when it does find a solution, that solution is of smallest possible cost (i.e.
+optimal).
+\textbf{DFS} is obvisouly not optimal as it could very well find a solution very deep in the graphs while expanding the very first branch, not realising that a solution was 1 level away from the start on another branch!
 \\
 \\
-Informed strategies on the other hand exploit domain-specific knowledge.
+\textit{Informed} strategies, on the other hand, exploit domain-specific knowledge.
 One very popular such strategy is \textbf{A$^{*}$} (see \cite{DBLP:journals/jacm/DechterP85}), which always first examine the node of smallest expected total cost (\textbf{f}), itself the sum of the cost-so-far from initial to current node (\textbf{g}) plus the expected cost-to-go from current node to solution (\textbf{h}).
 The expected cost-to-go \textbf{h} is often called a \textit{heuristic}.
-The better the heuristic is, the better\textbf{A$^{*}$} usually performs.
+The better the heuristic is, the better \textbf{A$^{*}$} usually performs.
 An important property of heuristics is \textit{admissibility}.
-In short, admissible heuristics never over-estimate the real cost-to-go, and can easily be shown to render\textbf{A$^{*}$} optimal.
+In short, admissible heuristics never over-estimate the real cost-to-go, and can easily be shown to render \textbf{A$^{*}$} optimal.
 
@@
 Quite similarly to \textbf{RL}, Deep Learning (\textbf{DL}) has not initially had the success the \textbf{AI} community was hoping it would.
-Marvin Minsky, who often took the blame for having \textit{killed} funding into the field, even regretted writing his 1969 book (\cite{minskypapert69}) in which he had proved that three-layer feed forward perceptrons were not quite the universal functions approximators some had hypothesied them to be.
-Since around 2006 (see \cite{GoodBengCour16}for more history of the recent burgeoning of \textbf{DL}), the field has known a rebirth, probably due to combination of factors, from ever more powerful computers and larger storage, the availability of data sets large and small on the internet, the advances in many techniques and heuristics (backward propagation, normalisation, drop-outs, etc..) and the huge amount of experimentation with different architectures (from very deep feed forward fully connected networks, to convolutional, recurrent and more exoctic networks).
-It is not an exageration to say that \textbf{DL} has rendered obsolete entire fields of research and bodies of knowledge e.g.
-in \textbf{CV} and \textbf{NLP}.
-\textbf{DL} has often become one of the methods of choice to learn of approximate highly dimensional functions.
-The beauty of it is that the relevant features that go into the function are autodiscovered by the training of the network, via back-ward propagation and trial-and-error, rather than having to be postulated or handcrafted as is often the case with other \textbf{ML} techniques.
-\\
-\\
-For this project, I have implemented this in a class called DeepLearner (see section \ref{sec:codelearners} for details)
+Marvin Minsky, who often took the blame for having \textit{killed} funding into the field, even regretted writing his 1969 book (\cite{minskypapert69}) in which he had proved that three-layer feed forward perceptrons were not quite the universal functions approximators some had hypothesized them to be.
+Since around 2006 (see \cite{GoodBengCour16} for more history of the recent burgeoning of \textbf{DL}), the field has known a rebirth, probably due to combination of factors, from ever more powerful computers and larger storage, the availability of data sets large and small on the internet, the advances in many techniques and heuristics (backward propagation, normalisation, drop-outs, different optimisation schemes, etc...) and the huge amount of experimentation with different architectures (from very deep feed forward fully connected networks, to convolutional, recurrent and more exoctic networks).
+It is not an exageration to say that \textbf{DL} has rendered obsolete entire fields of research and bodies of knowledge, most notably in \textbf{CV}, robotics, computational biology and \textbf{NLP}.
+\textbf{DL} has often become the method of choice to learn or approximate highly dimensional functions or random processes.
+The beauty of it is that the relevant features are autodiscovered during the training of the network, via back-ward propagation and trial-and-error, rather than having to be postulated or handcrafted as is often the case with other \textbf{ML} techniques.
 
@@
 \\
-Sometimes we have ways of computing good solutions/heuristics to our puzzles, but cannot computationally do this for all possible states (maybe the state space is just too large!).
+Sometimes we have ways of computing good solutions/heuristics to our puzzles, but cannot computationally do this for all possible states (maybe the state space is simply too large!).
 One approach in that case is to compute the solution/heuristic for some states, and let a \textbf{DL} model learn how to extrapolate to the rest of the state space.
-This is a case of supervised learning in some sense, because we need a teacher-solver to tell us good solutions or heuristics.
+This is a case of \textit{supervised} learning in some sense, since we need a teacher-solver to tell us good solutions or heuristics to extrapolate from.
 \\
@@
 \begin{equation} \label{eq:DLVI}
-V_{t+1}(s) \leftarrow min \Big( V_{t}(s), 1 + min_{s' \in \mathcal{C}(s)}(V_{t}(s'))  \Big)
+\begin{split}
+& V_{t+1}(s) \leftarrow min \Big( V_{t}(s), 1 + min_{s' \in \mathcal{C}(s)}(V_{t}(s'))  \Big) \\
+& V_{t+1}(s) = 0 if s \in \mathcal{G}
+\end{split}
 \end{equation}
-where $\mathcal{C}(s)$ are the children (reachable states via 1 move) of $s$
+where $\mathcal{C}(s)$ are the children (reachable states via 1 move) of $s$ and $\mathcal{G}$ are the goal states.
+\\
 \\

### 2022-08-28 latex/Chapters/Implementation.tex
update

--- 2022-07-31 latex/Chapters/Code.tex
@@
 
-\chapter{Code} % Main chapter title
+\chapter{Implementation} % Main chapter title
 
@@
 \centering
-\includegraphics[scale=0.49]{./Figures/codebase}
+\includegraphics[scale=0.40]{./Figures/codebase}
 %\decoRule
@@
 \item \textbf{PerfectHeuristic}: this reads from a data base the optimal costs, pre-computed by the PerfectLearner (see below \ref{PLcode})
-\item \textbf{DeepLearningHeuristic}: this uses a network which has been trained using DRL by the DeepReinforcementLearner (see below \ref{DRLcode})
+\item \textbf{DeepLearningHeuristic}: this uses a network which has been trained using \textbf{DRL} by the DeepReinforcementLearner (see below \ref{DRLcode})
+\item \textbf{DeepQHeuristic}: this uses a network which has been trained using \textbf{DQL} by the DeepQLearner (see below \ref{DQLcode})
 \end{itemize}
@@
 \item \textbf{DeepLearner} tbd
-\item \textbf{DeepReinforcementLearner}: It instantiates a DeepLearning (network), and trains it using DRL.
+\item \textbf{DeepReinforcementLearner}: It instantiates a DeepLearning (network), and trains it using \textbf{DRL}.
 It then saves the trained network, which can then be used in the DeepLearningHeuristic we have seen earlier in section \ref{HSS}.
 The DeepReinforcementLearner runs for a number of epochs (or less if, based on other parameters discussed below, it is deemed to have converged).
+\item \textbf{DeepQLearner}: It instantiates a DeepLearning (network), and trains it using \textbf{DQL}.
+It then saves the trained network, which can then be used in the DeepLearningHeuristic as well as the DeepQHeuristic we have seen earlier in section \ref{HSS}.
+The DeepQLearner runs for a number of epochs (or less if, based on other parameters discussed below, it is deemed to have converged).
 
@@
 \section{rubiks.solvers}
+\label{sec:codesolvers}
 \begin{figure}[H]
@@
 \item \textbf{NaiveSlidingSolver} TBD
+\item \textbf{KociembaSolver} TBD
 \item \textbf{MonteCarloSearchTreeSolver} TBI I want to implement this in August when working on the Rubiks' ...

### 2022-08-28 latex/Chapters/SlidingPuzzle.tex
update

@@
 \label{sec:ResultsSP} % Change X to a consecutive number; for referencing this chapter elsewhere, use \ref{ChapterX}
+
+\begin{center}
+\begin{tabular}{l*{9}{c}r}
+\hline
+\textbf{solver}      & & \textbf{Sliding Puzzle} & \textbf{2x2} & \textbf{2x3} & \textbf{2x4} & \textbf{2x5}  & \textbf{3x3} & \textbf{3x4} & \textbf{4x4} \\ 
+\hline
+BFS   & & & & & & &  x  & & \\
+\hline
+A$^{*}$[Manhattan]   & & & & & &  x  &  x  & & \\
+\hline
+A$^{*}$[Manhattan++]   & & & & & &  x  &  x  & & \\
+\hline
+A$^{*}$[Perfect]   & & &  x  &  x  &  x  &  x  &  x  & & \\
+\hline
+Naive   & & & & & & &  x  & & \\
+\hline
+A$^{*}$[DL[A$^{*}$[Perfect]]]   & & & & & & &  x  & & \\
+\hline
+A$^{*}$[DRL]   & & & & & & &  x  & & \\
+\hline
+A$^{*}$[DQL]   & & & & & & &  x  & & \\
+\hline
+MCTS[DQL][c=TBD]   & & & & & & &  x  & & \\
+\end{tabular}
+\end{center}
 
@@
 In this section, I will discuss \textbf{full} results for these 5 puzzles.
-In order to fully solve them, one can simply use rubiks.scripts.learner, setting up the PerfectLearner with A* and manhattan heuristic, or instantiate directly a PerfectLearner as have seen in section \ref{PLSS}
+In order to fully solve them, one can simply use rubiks.scripts.learner, setting up the PerfectLearner with A* and manhattan heuristic, or instantiate directly a PerfectLearner as have seen in section \ref{PLSS}.
 I obtained the following God numbers for these puzzles:
@@
 \hline
-2              &   & 6 & 21 & 36 & 53* \\
+2              &   & 6 & 21 & 36 & 55* \\
 3              &   &   & 31 &  &    \\
@@
 \end{center}
-\underline{Most difficult 2x5 (53* moves)}:
+\underline{Most difficult 2x5 (55* moves)}:
 \begin{center}
 \begin{nine}
-\setrow{2}{,4,9,2,1}
-\setrow{1}{5,8,3,7,6}
+\setrow{2}{,9,3,7,1}
+\setrow{1}{5,4,8,2,6}
 \end{nine}
@@
 \centering
-\includegraphics[align=c, scale=0.55]{./Figures/33SPDeepReinforcementLearning.jpeg}
+\includegraphics[align=c, scale=0.55]{./Figures/33SPDeepReinforcementLearning}
 \caption[33SPDeepReinforcementLearning]{Deep reinforcement learner 3x3 SP}
@@
 
+\subsection{DRL vs DQL}
+
+\subsection{MCTS DQL}
+
 \subsection{Solvers' comparison}
@@
 The results are shown in figure \ref{fig:33SPPerformance}
+\\
+DL: 100 seq 15 to 31 shuffles 10k epochs connected 600 300 100
+\\
+Manhattan vs ++
+\\
+DQL vs DRL
+\\
+DxL vs Perfect
 
@@
 \hline
+AStarSolver[Manhattan++]   &   &      31  & 34,224  & 7,080  \\
+\hline
 AStarSolver[PerfectHeuristic]  &   & 31 & 1,585 & 202 \\
 \hline
-AStarSolver[DeepLearningHeuristic]  &   & 31 & 101 & 58 \\
+AStarSolver[DRLHeuristic]  &   & 31 & 101 & 58 \\
+\hline
+MCTSSolver[DQLHeuristic][c=0]  &   & 101 & 103 & 456 \\
+\hline
+MCTSSolver[DQLHeuristic][c=69]  &   & 35 & 2,244 & 8,873 \\
 \hline
@@
 
-TBD
+\label{ssec:34SPSC}
 
+\newgeometry{top=0mm, bottom=0mm, left=0mm, right=0mm}
+\begin{landscape}
+\centering\vspace*{\fill}
+\begin{figure}[H]
+\centering
+\includegraphics[scale=0.63]{./Figures/34SPPerformance.jpeg}
+%\decoRule
+\caption[34SPPerformance]{Solvers' performance comparison 3x4 SP}
+\label{fig:34SPPerformance}
+\end{figure}
+\vfill
+\end{landscape}
+\restoregeometry
 %-----------------------------------
@@
 
-TBD
+\label{ssec:44SPSC}
+
+\newgeometry{top=0mm, bottom=0mm, left=0mm, right=0mm}
+\begin{landscape}
+\centering\vspace*{\fill}
+\begin{figure}[H]
+\centering
+\includegraphics[scale=0.63]{./Figures/44SPPerformanceAll.jpeg}
+%\decoRule
+\caption[44SPPerformance]{Solvers' performance comparison 4x4 SP}
+\label{fig:44SPPerformance}
+\end{figure}
+\vfill
+\end{landscape}
+\restoregeometry
 

### 2022-08-28 latex/Chapters/DRLSearch.tex
update

@@
 
-In this section, I will succintly go through the different techniques that underly the algorithms I have implemented and compared in this project.
+In this section, I will succinctly go through the different techniques that underly the algorithms I have implemented and compared in this project.
 Along the way, I will give some references for the interested reader.
@@
 In the general theory, different transitions can have different costs associated with them.
-In this project however, we can reasonably consider that each move of a tile in the \textbf{SP} or each rotation of a face in the \textbf{RC} might are taking a constant and identical amount of time and effort to perform, and I will therefore consider all the graphs to be unit-cost (i.e.
+In this project however, we can reasonably consider that each move of a tile in the \textbf{SP} or each rotation of a face in the \textbf{RC} are taking a constant and identical amount of time and effort to perform, and I will therefore consider all the graphs to be unit-cost (i.e.
 all moves have cost 1).
@@
 optimal).
-\textbf{DFS} is obvisouly not optimal as it could very well find a solution very deep in the graphs while expanding the very first branch, not realising that a solution was 1 level away from the start on another branch!
+\textbf{DFS} is obvisouly not optimal as it could very well find a solution very deep in the graphs while expanding the very first branch, not realising that a solution was one level away from the start on another branch!
 \\
@@
 The beauty of it is that the relevant features are autodiscovered during the training of the network, via back-ward propagation and trial-and-error, rather than having to be postulated or handcrafted as is often the case with other \textbf{ML} techniques.
+\\
+\\
+For this project, I have implemented the procedure of training a \textbf{DL} network from the solutions of other solvers/heuristics in a class called DeepLearner (see section \ref{sec:codelearners} for details).
 
@@
 We can perform this value-iteration in an unsupervised way: we do not need a teacher to tell us the target value of the value function/cost-to-go but instead learn it simply from refining a state $s$'s value by replacing it at each iteration by the minimum of its current value, and the minimum value of all reachable states from $s$ polus one (since it takes one move to reach them).
-That is, for $t \in \mathbb{N}^{*}, s \in \mathcal{S}$, the state space, we update $V_{t}(s)$ via:
+That is, we can start with arbitrary (e.g.
+0) $V_{0}(s0$ and recursively update for $t \in \mathbb{N}^{*}, s \in \mathcal{S}$, the state space via the following iteration:
 \begin{equation} \label{eq:DLVI}
 \begin{split}
-& V_{t+1}(s) \leftarrow min \Big( V_{t}(s), 1 + min_{s' \in \mathcal{C}(s)}(V_{t}(s'))  \Big) \\
-& V_{t+1}(s) = 0 if s \in \mathcal{G}
+if \  s & \in \mathcal{G}: \\
+& \ V_{t+1}(s) = 0 \\
+else&: \\
+& \ V_{t+1}(s) \leftarrow min \Big( V_{t}(s), 1 + min_{s' \in \mathcal{C}(s)}(V_{t}(s'))  \Big)
 \end{split}
@@
 This is where \textbf{DL} comes again to the rescue to act as function approximator, combining with \textbf{RL} into what has become known as \textbf{DRL}.
-The subtelty is then that both the left-hand-side and right hand side in the value-iteration procedure above are given by a \textbf{DL} network which we are constantly tweaking by back-ward propagation.
-For my implementation of \textbf{DRL}, I followed the same approach as in \cite{Mnih2013}, where the right hand-side of \ref{eq:DLVI} is a fixed \textit{target-network}.
-The left hand side is another \textit{current-network} which alone is modified by the \textbf{DL} learning (backward propagation and optimization), until convergence is deemed reached.
-We then update the \textit{target-network} by a copy of the \textit{current-network}, and then keep iterating again, until the whole process is deemed to have converged.
+The subtelty is then that both the left-hand-side and right hand side in the value-iteration procedure \ref{eq:DLVI} are given by a \textbf{DL} network which we are constantly tweaking.
+For my implementation of \textbf{DRL}, I followed the same approach as in \cite{Mnih2013}, utilising two loops.
+During iteration over the first (inner) loop, the right hand-side $V$ of \ref{eq:DLVI} is computed using a fixed \textit{target-network}.
+The left hand side $V$ is using another \textit{current-network} which alone is modified by the \textbf{DL} learning (backward propagation and optimization), until convergence is deemed reached.
+When convergence in the inner loop is deemed obtained, we break out of it, and run the outer loop, which updates the \textit{target-network} via a copy of the \textit{current-network}.
+The outer loop is itself broken out of when deemed appropriate (some kind of convergence criteria).
 \\
@@
 
-Another concept of \textbf{RL} is that of Q-function $Q(s, a)$, the maximum expected value the agent can obtain from state $s$, taking action $a$ and acting optimally afterwards.
+Another basic concept of \textbf{RL} is that of the Q-function $Q(s, a)$, the maximum expected reward the agent can obtain from state $s$, taking action $a$ and acting optimally afterwards.
 Notice that the knowledge of $V$ everywhere is equivalent to that of $Q$, since obviously:
@@
 \\
-Similarly to our discussion about \textbf{DRL}, in cases where the state space is large, it can be useful to approximate $Q$, or some related quantity (such as e.g.
+Similarly to our discussion about \textbf{DRL}, in cases where the state space is large, it can be useful to approximate $Q$, or a somewhat related quantity (such as e.g.
 a probability distribution over the actions that can be taken from $s$) by a neural network, and use a similar procedure to that of the previous section to update the right-hand-side and left-hand-side of the iterations.
@@
 This leaves us with the question of how can we use the output of such a network to search for a solution.
-One simple answer (which I have tried) is that we could disregard the distribution and only take into account the cost-to-go by simply using $A^{*}$.
+One simple answer (which I have tried, see e.g.
+subsection \ref{sec:S33DRLDQL}) is that we could disregard the distribution and only take into account the cost-to-go by simply using $A^{*}$.
 The rationale for doing so is that we might hope that jointly learning the cost-to-go and the actions via a combination of MSE and cross-entropy-loss leads to better heuristics, since we use more information during the learning and iterations.
@@
 In short, it generates at each step (and possibly in a distributed manner) paths to a \textit{leaf} on the frontier of unexpanded nodes, expands that node and checks if it is a goal state (and keeps going until the goal is found).
-The paths followed from initial state to leaves are decided via a combination of the probability distribution over actions (following the most promising actions) as well as the value function.
+The paths followed from initial state to leaves are decided via a combination of the probability distribution over actions (following the most promising actions) as well as the value function, according to a trade-off which can be tuned via a hyper-parameter.
 In order to prevent always following the same paths (which might take us very far from the goal), \textbf{MCTS} algorithms usually keep track of actions they have followed already and underweight their probability so that subsequent paths explore other actions.

### 2022-08-28 latex/Chapters/SlidingPuzzle.tex
update

@@
 
+In this chapter, I show and discuss results on \textbf{SP} of various dimensions, going from \textit{small} dimensions (which I define as those I can reasonably \textit{fully} solve given my computing resources, the finite time I have to complete this project, and the choice of programming language), then I move to studying in much detail the case of the 3x3, and finally move on to slightly higher dimensions 3x4 and 4x4.
+Table \ref{tab:gridSP} below shows which solvers I have run on each of the dimensions attempted.
+
+\begin{table}[H]
 \begin{center}
@@
 \hline
-A$^{*}$[Manhattan]   & & & & & &  x  &  x  & & \\
+A$^{*}$[Manhattan]   & & & & & &  x  &  x  &  x  & \\
 \hline
-A$^{*}$[Manhattan++]   & & & & & &  x  &  x  & & \\
+A$^{*}$[Manhattan++]   & & & & & &  x  &  x  &  x  & x \\
 \hline
@@
 \hline
-Naive   & & & & & & &  x  & & \\
+Naive   & & & & & & &  x  &  x  & x \\
 \hline
-A$^{*}$[DL[A$^{*}$[Perfect]]]   & & & & & & &  x  & & \\
+A$^{*}$[DL[A$^{*}$[Manhattan++]]]   & & & & & & &  x  &  x  & x \\
 \hline
-A$^{*}$[DRL]   & & & & & & &  x  & & \\
+A$^{*}$[DRL]   & & & & & & &  x  &  x  & x \\
 \hline
@@
 \hline
-MCTS[DQL][c=TBD]   & & & & & & &  x  & & \\
+MCTS[DQL]        & & & & & & &  x  & & \\
 \end{tabular}
+\caption{\label{tab:gridSP} Solvers used vs \textbf{SP} dimension}
 \end{center}
+\end{table}
+
+ I am neither making claims of depth (each solver could surely be tuned and optimized/adapted, possibly differently for each dimension) nor breadth (I have not run all of the solvers versus each dimension).
+Running these experiments takes a lot of time, so I had to be somehow selective.
+I at least wanted to run all of the solvers I have implemented on the intermediary 3x3 dimension, and answer some of the following interesting questions:
+\begin{itemize}
+\item How hard is too hard for \textbf{BFS} to complete in \textit{finite} time?
+\item How much, if any, improvement do we get with Manhattan++ over Manhattan?
+\item Is there much loss between \textbf{DL} and \textbf{DRL}, i.e. is losing the teacher used by \textbf{DL} a deal breaker?
+\item Is \textbf{DQL}, all things being equal, able to learn a better cost-to-go heuristic than \textbf{DRL}?
+\item How does \textbf{MCTS} perform? How important is it to tune the hyper-parameter $c$? How much improvement, if any, does the post trim of the \textbf{MCTS} tree via \textbf{BFS} bring?
+\item Are any of the \textbf{D*L} techniques able to perform on par with A$^{*}$ with Manhattan++ or perfect heuristic?
+\end{itemize}
+All of these questions are answered (obviously not in generality, but in the context of the experiments I have been running) in this chapter.
 
@@
 As mentioned in chapter \ref{sec:Puzzles}, the state space cardinality for the SP grows very quickly with n and m.
-Here are the only dimensions which have less than 239.5 millions states.
-Note I am also only considering n $\leq$ m since (p, q) can always be solved if we know how to solve (q, p):
-\\
-\\
+The only dimensions which have less than 239.5 millions states are shown in table \ref{tab:smallSP}.
+Note I am also only considering n $\leq$ m since dimension (m, n) can always be solved by symmetry from the solutions (n, m):
+
+\begin{table}[H]
 \begin{center}
@@
 \end{tabular}
+\caption{\label{tab:smallSP}\# puzzles for \textit{small} dimensions}
 \end{center}
-In this section, I will discuss \textbf{full} results for these 5 puzzles.
+\end{table}
+In this section, I will discuss \textit{full} results for these 5 dimensions.
 In order to fully solve them, one can simply use rubiks.scripts.learner, setting up the PerfectLearner with A* and manhattan heuristic, or instantiate directly a PerfectLearner as have seen in section \ref{PLSS}.
-I obtained the following God numbers for these puzzles:
+I obtained the following God numbers (table \ref{tab:smallSPGN}) for these puzzles:
+\begin{table}[H]
 \begin{center}
@@
 3              &   &   & 31 &  &    \\
+* \tiny{provisional result}
 \end{tabular}
+\caption{\label{tab:smallSPGN}God numbers for \textit{small} dimensions}
 \end{center}
-* provisional result
+\end{table}
+Notice that among the above dimensions, (n=2, m=5) is the largest and hardest one.
+I had to run its PerfectLearner over a many weeks (whenever I had spare computing capacity and over a full week-end on a c5.18xlarge instance (72 cores) on Amazon EC2).
 \\
 \\
-Notice that among the above dimensions, (n=2, m=5) is the largest and hardest one. I had to run its PerfectLearner over a couple of days, on a c5.18xlarge instance (72 cores) on Amazon EC2.
-\\
-\\
-The perfectLearner also keeps track of (one of) the hardest puzzles it has encountered (i.e. requiring a number of steps equal to their respective God number to solves):
+The perfectLearner also keeps track of the hardest puzzle it has encountered (i.e.
+a configuration whose optimal solution has a cost equal to the God number).
+I obtained the following hardest puzzles for each of the 5 \textit{small} dimensions:
 \\
@@
 \setrow{1}{5,4,8,2,6}
+\tiny{* provisional result}
 \end{nine}
 \end{center}
-* provisional result
-\\
-\\
 \underline{Most difficult 3x3 (31 moves)}:
@@
 \hspace*{-2.5cm}
-\includegraphics[scale=0.45]{./Figures/25SPPerformanceManhattan}
+\includegraphics[scale=0.40]{./Figures/25SPPerformanceManhattan}
 %\decoRule
@@
 
-\subsection{DRL vs DQL}
+\subsection{MCTS DQL}
 
-\subsection{MCTS DQL}
+\newgeometry{top=0mm, bottom=0mm, left=0mm, right=0mm}
+\begin{landscape}
+\centering\vspace*{\fill}
+\begin{figure}[H]
+\centering
+\includegraphics[align=c, scale=0.55]{./Figures/33SPPerformanceMCTS}
+\caption[33SPPerformanceMCTS]{MCTS tuning 3x3 SP}
+\label{fig:33SPPerformanceMCTS}
+\end{figure}
+\vfill
+\end{landscape}
+\restoregeometry
 
@@
 \label{ssec:33SPSC}
+\label{sec:S33DRLDQL}
 Let me discuss a comparison of several algorithms on 1000 random puzzles generated for a number of random shuffling (with best-effort-no-backtracking) from 0 to 50 in step of 2, as well as for perfect shuffling (denoted by $\infty$) on the comparison graphs.
@@
 \centering
-\includegraphics[scale=0.63]{./Figures/33SPPerformance.jpeg}
+\includegraphics[scale=0.63]{./Figures/33SPPerformance}
 %\decoRule
@@
 \centering
-\includegraphics[scale=0.63]{./Figures/34SPPerformance.jpeg}
+\includegraphics[scale=0.63]{./Figures/34SPPerformance}
 %\decoRule
@@
 \centering
-\includegraphics[scale=0.63]{./Figures/44SPPerformanceAll.jpeg}
+\includegraphics[scale=0.63]{./Figures/44SPPerformanceAll}
 %\decoRule
@@
 \restoregeometry
-
-%-----------------------------------
-%	SECTION 5
-%-----------------------------------
-
-\section{5x5}
-
-TBD

### 2022-08-28 rubiks/scripts/draft.py

@@ -27,18 +27,18 @@ if '__main__' == __name__:
     """ What puzzle """
     puzzle_type = Puzzle.sliding_puzzle
     n = 3
-    m = None
+    m = 4
     dimension = Puzzle.factory(**globals()).dimension()
     """ How much to shuffle """
     nb_shuffles = 0
     """ For performance test """
-    nb_samples = 100
+    nb_samples = 1000
     min_nb_shuffles = 0
-    max_nb_shuffles = 50
-    step_nb_shuffles = 5
+    max_nb_shuffles = 0
+    step_nb_shuffles = 1
     add_perfect_shuffle = True
     nb_cpus = 10
-    chunk_size = 10
+    chunk_size = 0
     performance_file_name = get_performance_file_name(puzzle_type, dimension)
     shuffles_file_name = get_shuffles_file_name(puzzle_type, dimension)
     append = True
@@ -46,7 +46,7 @@ if '__main__' == __name__:
     do_not_reattempt_failed = False
     #skip = (36,39,85,145,146,161,174,190,) <- these 8 took more than 90 mins for 3x3x3 Rubiks'
     """ For plot """
-    loc = 'upper center'   # 'upper center'
+    loc = 'center'   # 'upper center'
     performance_metrics = [Solver.pct_solved,
                            #Solver.pct_optimal,
                            Solver.median_cost,
@@ -55,7 +55,9 @@ if '__main__' == __name__:
                            Solver.median_run_time,
                            Solver.median_expanded_nodes,
                            ]
-    plot_abbreviated_names = True
+    #markers = ['x', '.']
+    #colors = ['royalblue', 'royalblue', 'darkred', 'darkred', 'goldenrod', 'goldenrod', 'darkcyan', 'darkcyan']
+    plot_abbreviated_names = False
     fig_size = [20, 12]
     """ Which solver type {Solver.dfs,
                            Solver.bfs,
@@ -63,23 +65,25 @@ if '__main__' == __name__:
                            Solver.naive,
                            Solver.kociemba,
                            } """
-    solver_type = Solver.mcts
-    c = 1
-    trim_tree = False
+    solver_type = Solver.astar
+    c = 100
+    trim_tree = True
     limit = 12
-    time_out = 1800
+    time_out = 600
     log_solution = False
     check_optimal = False
-    max_consecutive_timeout = 100
+    max_consecutive_timeout = 1000
     """ Heuristic if a* {Heuristic.manhattan,
                          Heuristic.perfect,
                          Heuristic.deep_learning,
[…]

### 2022-08-28 rubiks/solvers/solver.py

@@ -231,6 +231,15 @@ class Solver(Factory, Puzzled, Loggable, metaclass=ABCMeta):
     fig_size = 'fig_size'
     loc = 'loc'
     plot_abbreviated_names = 'plot_abbreviated_names'
+    markers = 'markers'
+    colors = 'colors'
+    log_scale = 'log_scale'
+    default_log_scale = [avg_expanded_nodes,
+                         avg_run_time,
+                         median_expanded_nodes,
+                         median_run_time,
+                        ]
+    exclude_solver_names = 'exclude_solver_names'
 
     @classmethod
     def populate_parser(cls, parser):
@@ -239,6 +248,26 @@ class Solver(Factory, Puzzled, Loggable, metaclass=ABCMeta):
                          type=str,
                          nargs='+',
                          default=cls.default_performance_metrics)
+        cls.add_argument(parser,
+                         field=cls.log_scale,
+                         type=str,
+                         nargs='+',
+                         default=cls.default_log_scale)
+        cls.add_argument(parser,
+                         field=cls.exclude_solver_names,
+                         type=str,
+                         nargs='+',
+                         default=tuple())
+        cls.add_argument(parser,
+                         field=cls.markers,
+                         type=str,
+                         nargs='+',
+                         default=tuple())
+        cls.add_argument(parser,
+                         field=cls.colors,
+                         type=str,
+                         nargs='+',
+                         default=tuple())
         cls.add_argument(parser,
                          field=cls.time_out,
                          type=int,
@@ -624,8 +653,11 @@ class Solver(Factory, Puzzled, Loggable, metaclass=ABCMeta):
             return
         from rubiks.heuristics.deeplearningheuristic import DeepLearningHeuristic
         from rubiks.heuristics.deepqlearningheuristic import DeepQLearningHeuristic
+        for solver_name in self.exclude_solver_names:
+            performance = performance[performance[cls.solver_name].
+                apply(lambda sn: sn.find(solver_name) < 0)].reset_index(drop=True)
 
-        def short_name(solver_name):
+        def short_name(plot_abbreviated_names, solver_name):
             """ @Francois this function is a horrible hack ... abstract away by having short_name in Solver API """
             shorts = ['Naive', 'BFS', 'DFS', 'Kociemba']
             for short in shorts:
[…]

### 2022-08-29 latex/Chapters/SlidingPuzzle.tex
update

@@
 \centering
-\hspace*{-2.5cm}
+%\hspace*{-2.5cm}
 \includegraphics[scale=0.40]{./Figures/25SPPerformanceManhattan}
 %\decoRule
-\caption[SP]{Manhattan vs Manhattan++}
+\caption[SP]{Manhattan vs Manhattan++ for the 2x5 \textbf{SP}}
 \label{fig:25SPPerformanceManhattan}
@@
 However, for any non trivial case, Manhattan++ outperforms considerably (by a factor of about 2.5).
-For the 250 \textit{perfectly} shuffled instances, I got the following:
+Table \ref{tab:mppOutperformance} below summarizes the results for the 250 \textit{perfectly} shuffled instances:
 
+\begin{table}[H]
 \begin{center}
@@
 \end{tabular}
+\caption{\label{tab:mppOutperformance} Manhattan++ outperformance on perfectly scrambled 2x5 \textbf{SP}}
 \end{center}
+\end{table}
 
@@
 \subsection{Perfect learner}
-As discussed in the previous section section \ref{sec:SPLowDimension}, the 3 by 3 SP is one of the cases I have been able to solve perfectly, since it only has 181,440 possible configurations.
+As discussed in the previous section section \ref{sec:SPLowDimension}, the 3x3 \textbf{SP} is one of the cases I have been able to solve perfectly, since it only has 181,440 possible configurations.
 Its God number is only 31, which definitely makes it manageable.
@@
 %\decoRule
-\caption[33SPPerfectLearning]{Perfect Learning 3x3 SP}
+\caption[33SPPerfectLearning]{Perfect Learning 3x3 \textbf{SP}}
 \label{fig:33SPPerfectLearning}
@@
 \includegraphics[align=c, scale=0.55]{./Figures/33SPDeepReinforcementLearning}
-\caption[33SPDeepReinforcementLearning]{Deep reinforcement learner 3x3 SP}
+\caption[33SPDeepReinforcementLearning]{Deep reinforcement learner 3x3 \textbf{SP}}
 \label{fig:33SPDeepReinforcementLearning}
@@
 \end{center}
-On this specific configuration, there was obviously no chance that the BFS would complete, hence it timed out.
-It would have no matter what time out I set.
-Indeed, since it has no heuristic to guide its search, it would need to explore in the order of $3^{31}$ - roughly 617 trillions - nodes to reach the goal!
+On this specific configuration, \textbf{BFS} was unable to complete before the time-out of sixty seconds.
+In an implementation without duplicate pruning, \textbf{BFS} would time out no matter what the bound is set.
+Indeed, it would need to explore in the order of $3^{31}$ - roughly 617 trillions - nodes to reach the goal!
+Even with my implementation which does pruning, it would need to pretty much traverse the entire 3x3 \textbf{SP} tree.
 \\
-Rather interestingly, my DRL heuristic performs much better than the manhattan heuristic (not super suprising), but also outperforms the perfect heuristic quite significantly both in terms of run time and of nodes expansion.
+Rather interestingly, the \textbf{DRL} heuristic performs much better than the Manhattan heuristic (not super suprising), but also outperforms the perfect heuristic quite significantly both in terms of run time and of nodes expansion.
 Obviously there is no guarantee that the perfect heuristic will not be outperformed on some random configuration, and it does on this occasion.
 However, as we have seen in the previous subection \ref{ssec:33SPSC}, it is not the case on average.
+\\
+\textbf{MCTS} with $c$ = 0 performs rather poorly, finding the longest solution (even than my Naive solver).
+It does behave a bit like \textbf{DFS} when $c$ is very small, expect the direction of travel is a bit more informed.
+I increased $c$, and for values over 69, it always gave me a solution of cost 35, which is not bad at all.
 \\

### 2022-08-29 rubiks/solvers/solver.py

@@ -647,22 +647,28 @@ class Solver(Factory, Puzzled, Loggable, metaclass=ABCMeta):
         cls = self.__class__
         try:
             performance = read_pickle(self.performance_file_name)
+            self.log_info(performance)
         except FileNotFoundError:
             self.log_error('Cannot find \'%s\'. Did you really want to plot rather than solve?' %
                            self.performance_file_name)
             return
         from rubiks.heuristics.deeplearningheuristic import DeepLearningHeuristic
         from rubiks.heuristics.deepqlearningheuristic import DeepQLearningHeuristic
-        for solver_name in self.exclude_solver_names:
+        exclude_tag = 'exclude'
+        exclude_solver_names = list(self.exclude_solver_names) + [exclude_tag]
+        for _ in exclude_solver_names:
             performance = performance[performance[cls.solver_name].
-                apply(lambda sn: sn.find(solver_name) < 0)].reset_index(drop=True)
+                apply(lambda sn: sn.find(_) < 0)].reset_index(drop=True)
 
         def short_name(plot_abbreviated_names, solver_name):
             """ @Francois this function is a horrible hack ... abstract away by having short_name in Solver API """
-            shorts = ['Naive', 'BFS', 'DFS', 'Kociemba']
-            for short in shorts:
-                if solver_name.find(short) >= 0:
-                    return short
+            shorts = {'Naive': 'Naive',
+                      'BFS': 'BFS',
+                      'DFS': 'DFS',
+                      'KociembaSolver': 'Kociemba'}
+            for short_key, s_name in shorts.items():
+                if solver_name.find(short_key) >= 0:
+                    return s_name
             a_star = solver_name.find('AStarSolver') >= 0
             mcts = solver_name.find('MonteCarlo') >= 0
             if a_star:
@@ -675,6 +681,7 @@ class Solver(Factory, Puzzled, Loggable, metaclass=ABCMeta):
                     mcts += '[c=%d]' % c
                 if solver_name.find('[trim]') >= 0:
                     mcts += '[trim]'
+            solver_name = solver_name.replace('KociembaHeuristic', 'Kociemba')
             pos = solver_name.find(DeepLearningHeuristic.__name__)
             if pos < 0:
                 pos = solver_name.find(DeepQLearningHeuristic.__name__)
@@ -695,19 +702,21 @@ class Solver(Factory, Puzzled, Loggable, metaclass=ABCMeta):
                                                   dimension=self.get_puzzle_dimension(),
[…]

### 2022-09-02 latex/Chapters/SlidingPuzzle.tex
update

@@
 Table \ref{tab:gridSP} below shows which solvers I have run on each of the dimensions attempted.
+The reader is referred to the implementation chapter \ref{sec:Implementation} for a detailed explanation of each of these solvers, and to the appendix section \ref{sec:Examples} for code examples on how to run them.
 
@@
 \hline
-A$^{*}$[DL[A$^{*}$[Manhattan++]]]   & & & & & & &  x  &  x  & x \\
+A$^{*}$[DL[FullyConnected]]   & & & & & & &  x  &  x  & x \\
+\hline
+A$^{*}$[DL[Convolutional]]   & & & & & & &  x  &  & \\
 \hline
@@
 \hline
-Manhattan                   &  34.7  & 50 & 133,332 & 2,110,887 & 49,561 & 606,838 \\
-Manhattan++              & 34.7 &  50 & 53,637 & 962,324 & 19,723 & 239,468 \\
-Improvement               & n/a &  n/a & x2.5 & x2.2 & x2.5 & x2.5 \\
+Manhattan                   &  34.6  & 49 & 46,483 & 575,050 & 36,088 & 428,887 \\
+Manhattan++              & 34.6 &  49 & 18,780 & 213,557 & 14,665 & 165,830 \\
+Improvement               & n/a &  n/a & x2.5 & x2.7 & x2.5 & x2.6 \\
 \end{tabular}
@@
 
-\newgeometry{top=0mm, bottom=0mm, left=0mm, right=0mm}
-\begin{landscape}
-\centering\vspace*{\fill}
+%\newgeometry{top=0mm, bottom=0mm, left=0mm, right=0mm}
+%\begin{landscape}
+%\centering\vspace*{\fill}
 \begin{figure}[H]
 \centering
-\includegraphics[align=c, scale=0.55]{./Figures/33SPPerformanceMCTS}
-\caption[33SPPerformanceMCTS]{MCTS tuning 3x3 SP}
+\includegraphics[align=c, scale=0.6]{./Figures/33SPPerformanceMCTS}
+\caption[33SPPerformanceMCTS]{\textbf{MCTS} tuning 3x3 \textbf{SP}}
 \label{fig:33SPPerformanceMCTS}
 \end{figure}
-\vfill
-\end{landscape}
-\restoregeometry
+%\vfill
+%\end{landscape}
+%\restoregeometry
 
@@
 DxL vs Perfect
+\\ Put table for nb\_shuffle=inf and highlight best in diff categories (cost, optimality score, run time, nodes)
+\\ say how much of the puzzles the solvers have seen
 
-\newgeometry{top=0mm, bottom=0mm, left=0mm, right=0mm}
-\begin{landscape}
-\centering\vspace*{\fill}
 \begin{figure}[H]
 \centering
-\includegraphics[scale=0.63]{./Figures/33SPPerformance}
+\includegraphics[scale=0.65]{./Figures/33SPPerformance}
 %\decoRule
-\caption[33SPPerformance]{Solvers' performance comparison 3x3 SP}
+\caption[33SPPerformance]{Solvers' performance comparison 3x3 \textbf{SP}}
 \label{fig:33SPPerformance}
 \end{figure}
-\vfill
-\end{landscape}
-\restoregeometry
 
@@
 
-\newgeometry{top=0mm, bottom=0mm, left=0mm, right=0mm}
-\begin{landscape}
-\centering\vspace*{\fill}
 \begin{figure}[H]
@@
 %\decoRule
-\caption[34SPPerformance]{Solvers' performance comparison 3x4 SP}
+\caption[34SPPerformance]{Solvers' performance comparison 3x4 \textbf{SP}}
 \label{fig:34SPPerformance}
 \end{figure}
-\vfill
-\end{landscape}
-\restoregeometry
+
 %-----------------------------------
@@
 %\decoRule
-\caption[44SPPerformance]{Solvers' performance comparison 4x4 SP}
+\caption[44SPPerformance]{Solvers' performance comparison 4x4 \textbf{SP}}
 \label{fig:44SPPerformance}
 \end{figure}
-\vfill
-\end{landscape}
-\restoregeometry

### 2022-09-02 rubiks/scripts/draft3.py

@@ -10,8 +10,8 @@ from rubiks.solvers.solver import Solver
 if '__main__' == __name__:
     """ Just create a logger to print some stuff in this script """
     logger = Loggable(name=__file__)
-    sp = Puzzle.factory(puzzle_type='sliding_puzzle', tiles=[[8, 6, 7], [2, 5, 4], [3, 0, 1]])
-    solver = Solver.factory(solver_type='dfs', puzzle_type='sliding_puzzle', n=3, time_out=3600, limit=32)
+    sp = Puzzle.factory(puzzle_type='sliding_puzzle', n=3).apply_random_moves(2)
+    solver = Solver.factory(solver_type='bfs', puzzle_type='sliding_puzzle', n=3, time_out=3600, limit=2)
     solution = solver.solve(sp)
     logger.log_info(solution)
 

### 2022-09-02 rubiks/scripts/rubiks22/perf.py

@@ -1,0 +1,152 @@
+########################################################################################################################
+# Francois Berrier - Royal Holloway University London - MSc Project 2022                                               #
+########################################################################################################################
+from rubiks.core.loggable import Loggable
+from rubiks.deeplearning.deeplearning import DeepLearning
+from rubiks.heuristics.deeplearningheuristic import DeepLearningHeuristic
+from rubiks.heuristics.heuristic import Heuristic
+from rubiks.heuristics.perfectheuristic import PerfectHeuristic
+from rubiks.learners.learner import Learner
+from rubiks.learners.deepreinforcementlearner import DeepReinforcementLearner
+from rubiks.puzzle.puzzle import Puzzle
+from rubiks.solvers.solver import Solver
+from rubiks.utils.utils import get_model_file_name, get_shuffles_file_name, get_performance_file_name
+########################################################################################################################
+
+
+def performance(**kw_args):
+    action_type = Solver.do_performance_test
+    """ What puzzle """
+    puzzle_type = Puzzle.rubiks_cube
+    n = 2
+    dimension = Puzzle.factory(**locals()).dimension()
+    """ How much to shuffle """
+    nb_shuffles = 0
+    """ For performance test """
+    solver_type = Solver.bfs
+    nb_samples = 5
+    min_nb_shuffles = 0
+    max_nb_shuffles = 20
+    step_nb_shuffles = 2
+    add_perfect_shuffle = True
+    nb_cpus = 1
+    chunk_size = 0
+    performance_file_name = get_performance_file_name(puzzle_type, dimension)
+    shuffles_file_name = get_shuffles_file_name(puzzle_type, dimension)
+    append = True
+    verbose = True
+    do_not_reattempt_failed = False
+    time_out = 7200
+    log_solution = False
+    check_optimal = False
+    max_consecutive_timeout = 1000
+    if kw_args.get('heuristic_type', None) in [Heuristic.deep_learning,
+                                               Heuristic.deep_q_learning]:
+        model_file_name = Learner.factory(**{**locals(), **kw_args}).get_model_name()
+    loc = 'upper center'   # 'upper center'
+    performance_metrics = [Solver.pct_solved,
+                           Solver.optimality_score,
+                           Solver.median_cost,
+                           Solver.max_cost,
+                           Solver.median_run_time,
+                           Solver.median_expanded_nodes,
+                           ]
+    plot_abbreviated_names = True
+    kw_args = {**locals(), **kw_args}
+    Solver.factory(**kw_args).action()
+
[…]

### 2022-09-02 rubiks/scripts/trainingdata.py

@@ -12,14 +12,14 @@ from rubiks.puzzle.trainingdata import TrainingData
 if '__main__' == __name__:
     logger = Loggable(name=__file__)
     puzzle_type = Puzzle.sliding_puzzle
-    n = 3
+    n = 5
     nb_cpus = 1
     chunk_size=0
     verbose=False
-    nb_sequences = nb_cpus * 1000
+    nb_sequences = nb_cpus * 1
     training_data = TrainingData(**globals())
-    repeat = 10
-    for nb_shuffles in [inf]:
+    repeat = 1
+    for nb_shuffles in [10]:
         training_data.generate(nb_shuffles=nb_shuffles,
                                nb_sequences=nb_sequences,
                                repeat=repeat)

### 2022-09-02 rubiks/solvers/solution.py

@@ -83,6 +83,11 @@ class Solution:
                 string.pop(field, None)
         return '\n' + pformat(string)
 
+    def __getitem__(self, item):
+        if hasattr(self, item):
+            return getattr(self, item)
+        return self.additional_info.get(item)
+
     def to_str_light(self):
         return self.to_str(fields=[Solution.puzzle,
                                    Solution.cost,

### 2022-09-02 rubiks/solvers/solver.py

@@ -232,6 +232,7 @@ class Solver(Factory, Puzzled, Loggable, metaclass=ABCMeta):
     loc = 'loc'
     plot_abbreviated_names = 'plot_abbreviated_names'
     markers = 'markers'
+    marker_size = 'marker_size'
     colors = 'colors'
     log_scale = 'log_scale'
     default_log_scale = [avg_expanded_nodes,
@@ -240,6 +241,7 @@ class Solver(Factory, Puzzled, Loggable, metaclass=ABCMeta):
                          median_run_time,
                         ]
     exclude_solver_names = 'exclude_solver_names'
+    labels_at_top = 'labels_at_top'
 
     @classmethod
     def populate_parser(cls, parser):
@@ -272,6 +274,10 @@ class Solver(Factory, Puzzled, Loggable, metaclass=ABCMeta):
                          field=cls.time_out,
                          type=int,
                          default=0)
+        cls.add_argument(parser,
+                         type=int,
+                         field=cls.marker_size,
+                         default=70)
         cls.add_argument(parser,
                          type=int,
                          field=cls.max_consecutive_timeout,
@@ -367,6 +373,10 @@ class Solver(Factory, Puzzled, Loggable, metaclass=ABCMeta):
                          type=int,
                          nargs='+',
                          default=tuple())
+        cls.add_argument(parser,
+                         cls.labels_at_top,
+                         default=False,
+                         action=cls.store_true)
     
     def performance_test(self):
         """
@@ -561,7 +571,7 @@ class Solver(Factory, Puzzled, Loggable, metaclass=ABCMeta):
             res[cls.avg_run_time] = nan if isnan(avg_run_time) else int(avg_run_time * 1000)
             res[cls.median_run_time] = int(median_run_time * 1000)
             res[cls.max_run_time] = nan if isnan(max_run_time) else int(max_run_time * 1000)
-            res[cls.pct_solved] = int(100 * nb_solved / self.nb_samples)
+            res[cls.pct_solved] = int(100 * nb_solved / self.nb_samples) if not isnan(nb_solved) else 0
             res[cls.pct_optimal] = int(100 * (sample - nb_not_optimal) / self.nb_samples)
             optimality_score = [0, 0]
             for opt_c, c in zip(optimal_cost_list, cost_list):
@@ -693,7 +703,16 @@ class Solver(Factory, Puzzled, Loggable, metaclass=ABCMeta):
                           'DeepLearner']
                 for short in shorts:
[…]

### 2022-09-03 latex/Appendices/Kociemba.tex
update

% Chapter Template

\chapter{Kociemba Bug} % Main chapter title

\label{sec:Kociemba} % Change X to a consecutive number; for referencing this chapter elsewhere, use \ref{ChapterX}

Discuss annoyance with Kociemba 3x3x3 with respect to color scheme

%-----------------------------------
%	SECTION 1
%-----------------------------------

### 2022-09-03 latex/Chapters/Conclusion.tex
update

@@
 
-\chapter{Conclusion} % Main chapter title
+\chapter{Conclusion \& Professional Issues} % Main chapter title
 
@@
 
-TBW
+\textbf{Learning}
+Learnt about puzzles, and also even took the time to learn a couple of sequences of moves that are sufficient to solve \textbf{RC}, 2x2x2 in labout one minute, 3x3x3 in about 2 minutes.
+\\
+\\
+\textbf{Implementation}
+Implemented a bunch of solvers. Ideally should do in C++ for speed and multithreaded (esp for MCTS)
+\\
+\\
+\textbf{DRL \& DQL}
+Quite impressive, without too much effort, how well they perform.
+However, I feel to really get more out of them (higher dimensions in particular) would need careful tuning of the convergence criteria.
+\\
+\\
+\textbf{Professional issues}
+\\
+\\
+See \cite{ManyWorlds}
+\\
+\\
+See \cite{https://doi.org/10.1111/1758-5899.12718} and \cite{Bostrom2014}

### 2022-09-03 latex/Chapters/Puzzles.tex
update

@@
 %-----------------------------------
-\section{Rubiks' Cube}
+\section{Rubik's Cube}
 
 \subsection{History}
-The second puzzle I will focus on is the well known Rubiks' Cube (\textbf{RC}), invented in 1974 by Erno Rubik, originally known under the name of \textit{Magic Cube} (see \cite{RubiksWiki}).
+The second puzzle I will focus on is the well known Rubik's Cube (\textbf{RC}), invented in 1974 by Erno Rubik, originally known under the name of \textit{Magic Cube} (see \cite{RubiksWiki}).
 Since it was commericalised in 1980, it has kown a worldwide success: countless people are playing with the famous original 3x3x3 cube, as well as with countless variations, more or less difficult of it.
-Competitions nowadays bring together \textit{speedcubers} who have trained for years and memorized various algorithms to solve the Rubiks in astonishly effective and fast times (literally in seconds for the best ones).
+Competitions nowadays bring together \textit{speedcubers} who have trained for years and memorized various algorithms to solve the Rubik's in astonishly effective and fast times (literally in seconds for the best ones).
 Interestingly, some of the principles used by speedcubers generalise to different dimensions.
@@
 
+\cite{RubiksGodNumber}
+\\
 \cite{RubiksChicago}
@@
 \\
+
+%-----------------------------------
+%	SECTION 3
+%-----------------------------------
+\section{Solving puzzles \textit{without human knowledge}}
+
+From handcrafted (naive, kociemba) to supervised (DL) to unsupervised (DRL \& DQL) .... but sometimes subtle knowledge injected in e.g. representation or simplifications.

### 2022-09-03 latex/Chapters/SlidingPuzzle.tex
update

@@
 I am neither making claims of depth (each solver could surely be tuned and optimized/adapted, possibly differently for each dimension) nor breadth (I have not run all of the solvers versus each dimension).
-Running these experiments takes a lot of time, so I had to be somehow selective.
-I at least wanted to run all of the solvers I have implemented on the intermediary 3x3 dimension, and answer some of the following interesting questions:
+Running these experiments takes a lot of time, so I had to be somewhat selective.
+I wanted however to run all of the solvers I have implemented one particular dimension (the intermediary 3x3 seemed appropriate to do so), to answer some of the following interesting questions:
 \begin{itemize}
-\item How hard is too hard for \textbf{BFS} to complete in \textit{finite} time?
+\item How hard is too hard for \textbf{BFS} to complete in \textit{reasonable} time?
 \item How much, if any, improvement do we get with Manhattan++ over Manhattan?
@@
 \centering
-\includegraphics[scale=0.63]{./Figures/44SPPerformanceAll}
+\includegraphics[scale=0.63]{./Figures/44SPPerformance}
 %\decoRule

### 2022-09-03 latex/MScAIFrancoisBerrier.tex
update

@@
 \usepackage{geometry}
+\usepackage{amsmath}
+\usepackage{algorithm}
+\usepackage{algpseudocode}
+\usepackage{rubikcube}
 \newenvironment{three}[1][]{%
@@
 
-\thesistitle{Solving \\ the \\ 15-Puzzle \& Rubik's Cube \\by \\Deep Reinforcement Learning} % Your thesis title, this is used in the title and abstract, print it elsewhere with \ttitle
+\thesistitle{Solving \\ the \\ 15-Puzzle \\ \& \\ Rubik's Cube} % Your thesis title, this is used in the title and abstract, print it elsewhere with \ttitle
 \supervisor{Pr. Chris \textsc{Watkins}} % Your supervisor's name, this is used in the title page, print it elsewhere with \supname
@@
 \textbf{Motivation}
+\label{Abstract}
 \\
@@
 \addchaptertocentry{\acknowledgementname} % Add the acknowledgements to the table of contents
-\begin{comment}
+%\begin{comment}
 I am grateful to my employer Bank of America for sponsoring my part time 2-year MSc in Artificial Intelligence at Royal Holloway.
@@
 \\
-Last but not least, my wife Thanh Nh\~a is my biggest supporter in everything I do, and her continued encouragement and enthusiasm in my pursuit of this MSc has made it so much easier to juggle between a demanding banking job, running 50 to 60 miles a week in preparation of multiple marathons and half marathons, and the coursework, lectures and exam revisions.
-\end{comment}
+Last but not least, my wife Thanh Nh\~a is my biggest supporter in everything I do, and her continued encouragement and enthusiasm in my pursuit of this MSc has made it so much easier to juggle between a demanding banking job, running 50 to 70 miles a week in preparation for the Chicago marathon, and the coursework, lectures and exam revisions.
+%\end{comment}
 \end{acknowledgements}
@@
 \textbf{DRL} & \textbf{D}eep \textbf{R}einforcment \textbf{L}earning\\
+\textbf{FIFO} & \textbf{F}irst \textbf{I}n \textbf{F}irst \textbf{O}ut\\
+\textbf{LIFO} & \textbf{L}ast \textbf{I}n \textbf{F}irst \textbf{O}ut\\
+\textbf{MCTS} & \textbf{M}onte \textbf{C}arlo \textbf{T}ree \textbf{S}earch \\
 \textbf{ML} & \textbf{M}achine \textbf{L}earning\\
 \textbf{NLP} & \textbf{N}atural \textbf{L}anguage \textbf{P}rocessing\\
-\textbf{RC} & \textbf{R}ubiks' \textbf{C}ube\\
+\textbf{RC} & \textbf{R}ubik's \textbf{C}ube\\
 \textbf{RL} & \textbf{R}einforcement \textbf{L}earning\\
@@
 \tableofcontents % Prints the main table of contents
+
+\hypersetup{%
+  colorlinks = true,
+  linkcolor  = orange
+}
 
@@
 \include{Chapters/Implementation}
-\include{Chapters/Examples}
 \include{Chapters/SlidingPuzzle}
@@
 
-%\include{Appendices/AppendixA}
-%\include{Appendices/AppendixB}
-%\include{Appendices/AppendixC}
+\include{Appendices/Examples}
+\include{Appendices/Kociemba}
 
@@
 %----------------------------------------------------------------------------------------
-
-%\begin{thebibliography}{widest entry}
-
-%\bibitem{sliddingpuzzle} Whatever
-%{\href{https://en.wikipedia.org/wiki/Sliding\_puzzle}{wikipedia}} Sliding Puzzle Wikipedia article, accessed 19th June 2022
-
-%\end{thebibliography}
 

### 2022-09-03 rubiks/heuristics/kociembaheuristic.py

@@ -12,7 +12,7 @@ class KociembaHeuristic(Heuristic):
 
     @classmethod
     def known_to_be_admissible(cls):
-        return False
+        return True
 
     def __init__(self, **kw_args):
         Heuristic.__init__(self, **kw_args)

### 2022-09-03 rubiks/scripts/rubiks22/perf.py

@@ -52,6 +52,10 @@ def performance(**kw_args):
                            Solver.median_expanded_nodes,
                            ]
     plot_abbreviated_names = True
+    labels_at_top = True
+    show_title = True
+    lite_title = True
+    plot_abbreviated_names = True
     kw_args = {**locals(), **kw_args}
     Solver.factory(**kw_args).action()
 
@@ -59,18 +63,38 @@ def performance(**kw_args):
 
 
 if '__main__' == __name__:
-    action_type = Solver.do_plot
+    action_type = Solver.do_performance_test
     fig_size = (12, 12)
     if action_type == Solver.do_plot:
         performance(**locals())
     elif action_type == Solver.do_performance_test:
         nb_samples = 100
-        nb_cpus = 10
+        nb_cpus = 8
+        performance(solver_type=Solver.astar,
+                    heuristic_type=Heuristic.deep_learning,
+                    learner_type=Learner.deep_reinforcement_learner,
+                    time_out=600,
+                    add_perfect_shuffle=True,
+                    max_nb_shuffles=20,
+                    nb_sequences=1000,
+                    nb_shuffles=20,
+                    nb_epochs=100000,
+                    layers_description=(600, 300, 100),
+                    training_data_every_epoch=False,
+                    cap_target_at_network_count=True,
+                    one_hot_encoding=True,
+                    drop_out=0.,
+                    network_type=DeepLearning.fully_connected_net,
+                    learning_rate=1e-2,
+                    scheduler=DeepReinforcementLearner.exponential_scheduler,
+                    gamma_scheduler=0.9999,
+                    **locals())
+        exit()
         performance(solver_type=Solver.mcts,
                     c=10,
                     trim_tree=True,
                     add_perfect_shuffle=False,
-                    max_nb_shuffles=4,
+                    max_nb_shuffles=6,
                     heuristic_type=Heuristic.deep_q_learning,
                     learner_type=Learner.deep_q_learner,
                     nb_sequences=1000,
@@ -114,22 +138,6 @@ if '__main__' == __name__:
                     **locals())
         performance(solver_type=Solver.astar,
                     heuristic_type=Heuristic.kociemba,
-                    **locals())
-        performance(solver_type=Solver.astar,
-                    heuristic_type=Heuristic.deep_learning,
-                    learner_type=Learner.deep_reinforcement_learner,
-                    nb_sequences=1000,
-                    nb_shuffles=20,
-                    nb_epochs=100000,
[…]

### 2022-09-03 rubiks/scripts/rubiks333/perf.py

@@ -1,0 +1,150 @@
+########################################################################################################################
+# Francois Berrier - Royal Holloway University London - MSc Project 2022                                               #
+########################################################################################################################
+from rubiks.core.loggable import Loggable
+from rubiks.deeplearning.deeplearning import DeepLearning
+from rubiks.heuristics.deeplearningheuristic import DeepLearningHeuristic
+from rubiks.heuristics.heuristic import Heuristic
+from rubiks.heuristics.perfectheuristic import PerfectHeuristic
+from rubiks.learners.learner import Learner
+from rubiks.learners.deepreinforcementlearner import DeepReinforcementLearner
+from rubiks.puzzle.puzzle import Puzzle
+from rubiks.solvers.solver import Solver
+from rubiks.utils.utils import get_model_file_name, get_shuffles_file_name, get_performance_file_name
+########################################################################################################################
+
+
+def performance(**kw_args):
+    action_type = Solver.do_performance_test
+    """ What puzzle """
+    puzzle_type = Puzzle.rubiks_cube
+    n = 3
+    dimension = Puzzle.factory(**locals()).dimension()
+    """ How much to shuffle """
+    nb_shuffles = 0
+    """ For performance test """
+    solver_type = Solver.bfs
+    nb_samples = 50
+    min_nb_shuffles = 0
+    max_nb_shuffles = 40
+    step_nb_shuffles = 2
+    add_perfect_shuffle = True
+    nb_cpus = 5
+    chunk_size = 0
+    performance_file_name = get_performance_file_name(puzzle_type, dimension)
+    shuffles_file_name = get_shuffles_file_name(puzzle_type, dimension)
+    append = True
+    verbose = True
+    do_not_reattempt_failed = False
+    time_out = 7200
+    log_solution = False
+    check_optimal = False
+    max_consecutive_timeout = 1000
+    if kw_args.get('heuristic_type', None) in [Heuristic.deep_learning,
+                                               Heuristic.deep_q_learning]:
+        model_file_name = Learner.factory(**{**locals(), **kw_args}).get_model_name()
+    loc = 'upper center'   # 'upper center'
+    performance_metrics = [Solver.pct_solved,
+                           Solver.optimality_score,
+                           Solver.median_cost,
+                           Solver.max_cost,
+                           Solver.median_run_time,
+                           Solver.median_expanded_nodes,
+                           ]
+    show_title = True
+    lite_title = True
[…]

### 2022-09-04 latex/Appendices/Kociemba.tex
update

@@
 
-Discuss annoyance with Kociemba 3x3x3 with respect to color scheme
+Let me discuss in this appendix in more details the issue I mentioned in section \ref{sec:ResRubiks333} regarding the \textit{poor} interface of the kociemba (3x3x3) library.
+\\
+\\
+I personally like to follow the great Scott Meyers' advice (see item 18 of \cite{meyers2005effective}), which is to not merely make interfaces easy to use, but also difficult to misuse.
+In that instance, I would argue they have made it very easy to shoot oneself in the foot with their interface which assumes the cubes passed to the solver are in \textit{standard} form (my terminology), not only without making that explicit anywhere, but worse than that: sometimes non standard initial configuration raise an exception, sometimes not and returning a solution which is not really one!
+So what is the problem exactly?
+\\
+\\
+To start with, let us remember from chapter \ref{sec:Puzzles} that each \textbf{RC} configuration has 24 equivalent configurations under invariance by full rotation of the cube in space.
+Out of 24 equivalent configurations, and for an observer facing the cube, there is only one whose center front (F) cubie is red and center up (U) cubie is white.
+\\
+\\
+In hkociemba (the third parties 2x2x2 implementation which I used under the hood of my KociembaSolver), the problem of these equivalent puzzles is irrelevant (since full cube rotation can always be achieved by 2 rotations of 2 opposite faces in opposite directions, e.g.
+FB' is equivalent to full cube rotation around F (CF in my jargon)).
+That being said, the hkociemba is smarter than just using the equivalence between full cube rotation and two faces' rotations.
+It does automatic color scheme recognition when passed a \textit{cube string}, and gets a smart solution given that coloring scheme.
+For instance, if we pass it a solved configuration, but not in standard form, it knows there is no need to do anything.
+In the general case of a scrambled cube, it will also endeavour to find solutions to the closest among the 24 equivalent solved configurations.
+Let's look at the following concerete example:
+
+\afblue
+\paragraph{}{\textbf{Kociemba bug -- 2x2x2 example of good interface design}}
+\begin{python}
+#####################################################################
+from rubiks.puzzle.puzzle import Puzzle
+from rubiks.solvers.solver import Solver
+from rubiks.solvers.kociembasolver import KociembaSolver
+from_kociemba = KociembaSolver.from_kociemba
+to_kociemba = KociembaSolver.to_kociemba
+#####################################################################
+puzzle_type = Puzzle.rubiks_cube
+n=2
+init_from_random_goal=False
+cube = Puzzle.factory(**globals()).get_equivalent()[-1].apply_random_moves(nb_moves=1)
+solver_type = Solver.kociemba
+solver = Solver.factory(**globals())
+print(solver.solve(cube))
+#####################################################################
+\end{python}
+\black
+
+which, subject to randomness, gave me the following scrambled cube and solution:
+
+\begin{figure}[H]
+\centering
+\includegraphics[scale=0.6]{./Figures/KociembaBug222}
+%\decoRule
+\caption[Kociemba Bug]{2x2x2}
+\label{fig:KociembaBug222}
+\end{figure}
+As we can see, hkociemba happily returned a 1-move solution to green F, white U, instead of the costlier solution it would have taken to get red F white U (clearly at a cost of 3 with solution RUD').
+\\
+\\
+In the 3x3x3 case, things are not quite that simple.
+There is no way to reproduce full cube rotation from faces rotations alone, since the center cubies will never move via the latter, but do via the former.
+If we run the equivalent of the above problem with the (3x3x3) kociemba library, we silently get to a false solution as is apparent from printing the resulting cube.
+\afblue
+\paragraph{}{\textbf{Kociemba bug -- 3x3x3 example of \textbf{bad} interface design}}
+\begin{python}
+#####################################################################
+from kociemba import solve
+#####################################################################
+from rubiks.puzzle.puzzle import Puzzle
+from rubiks.solvers.kociembasolver import KociembaSolver
+from_kociemba = KociembaSolver.from_kociemba
+to_kociemba = KociembaSolver.to_kociemba
+#####################################################################
+puzzle_type = Puzzle.rubiks_cube
+n=3
+init_from_random_goal=False
+cube = Puzzle.factory(**globals()).get_equivalent()[-1].apply_random_moves(nb_moves=1)
+print(cube)
+cube_string = to_kociemba(cube)
+solution = solve(cube_string)
+print('Kociemba Solution:', solution)
+print(cube.apply_moves(from_kociemba(solution)))
+###################################################################
+\end{python}
+\black
+
+\begin{figure}[H]
+\centering
+\includegraphics[scale=0.6]{./Figures/KociembaBug333}
+%\decoRule
+\caption[Kociemba Bug]{3x3x3 Buggy}
+\label{fig:KociembaBug333}
+\end{figure}
+In order to counteract that, and make the comparisons between Kociemba 3x3x3 and other solvers and methods fairer, I have had to implement a layer on top of kociemba.
+My Kociemba solver basically will check if a cube is in \textit{standard} form, and if not, it will find the sequence of full cube rotations that bring us to an equivalent configuration in \textit{standard} form, pass that to equivalent \textit{standard} cube to kociemba, and stich the cube rotation to the kociemba solution (treating cube rotation as 0 cost everywhere in the code).
+That way, we seamlessly get to an answer that makes sense from kociemba, without the user having to worry about whether or not the cube they are passing as input is in \textit{standard} form or not!
+Running the same example as above via my KociembaSolver, we get the expected result:
+
+\afblue
+\paragraph{}{\textbf{Kociemba bug -- 3x3x3 massage to fix the broken interface}}
+\begin{python}
+#####################################################################
+from rubiks.puzzle.puzzle import Puzzle
+from rubiks.solvers.solver import Solver
+from rubiks.solvers.kociembasolver import KociembaSolver
+from_kociemba = KociembaSolver.from_kociemba
+to_kociemba = KociembaSolver.to_kociemba
+#####################################################################
+puzzle_type = Puzzle.rubiks_cube
+n=3
+init_from_random_goal=False
+cube = Puzzle.factory(**globals()).get_equivalent()[-1].apply_random_moves(nb_moves=1)
+solver_type = Solver.kociemba
+solver = Solver.factory(**globals())
+print(solver.solve(cube))
+###################################################################
+\end{python}
+\black
+
+\begin{figure}[H]
+\centering
+\includegraphics[scale=0.5]{./Figures/KociembaBug333Fix1}
+%
+\includegraphics[scale=0.5]{./Figures/KociembaBug333Fix2}
+%\decoRule
+\caption[Kociemba Bug Fix]{3x3x3 Fixed}
+\label{fig:KociembaBug333Fix}
+\end{figure}
 

### 2022-09-04 rubiks/learners/deepreinforcementlearner.py

@@ -523,7 +523,7 @@ class DeepReinforcementLearner(Learner):
                 self.latency_tag: latency,
                 }
         to_pickle(data, self.learning_file_name)
-        count = self.target_network_count if self.__class__ is DeepReinforcementLearner else self.epoch
+        count = self.target_network_count if hasattr(self, 'target_network_count') else self.epoch
         versioned_learning_file_name = self.learning_file_name.replace('.pkl', '.%d.pkl' % count)
         to_pickle(data, versioned_learning_file_name)
         self.log_info('Saved learner state & convergence data in ',

### 2022-09-04 rubiks/puzzle/rubikscube.py

@@ -102,7 +102,7 @@ rubiks_int_to_face_map = {1: Face.F,
 
 
 rubiks_color_to_print_color_map = {Color.r: '\033[91m',
-                                   Color.w: '\033[90m',
+                                   Color.w: '\033[97m', #'\033[90m',
                                    Color.g: '\033[92m',
                                    Color.b: '\033[94m',
                                    Color.o: '\033[38;2;255;165;0m',

### 2022-09-04 rubiks/scripts/learner.py

@@ -34,7 +34,7 @@ if '__main__' == __name__:
                             Learner.deep_reinforcement_learner,
                             }
     """
-    learner_type = Learner.deep_reinforcement_learner
+    learner_type = Learner.deep_q_learner
     """ for plot """
     plot_metrics = DeepReinforcementLearner.default_plot_metrics \
         if learner_type is Learner.deep_reinforcement_learner \
@@ -67,7 +67,7 @@ if '__main__' == __name__:
     nb_shuffles = 15
     cap_target_at_network_count = False
     update_target_network_frequency = 1000
-    update_target_network_threshold = 1e-2
+    update_target_network_threshold = 2e-2
     max_nb_target_network_update = 10
     max_target_not_increasing_epochs_pct = 0.5
     max_target_uptick = 0.01

### 2022-09-04 rubiks/scripts/rubiks333/perf.py

@@ -64,10 +64,35 @@ if '__main__' == __name__:
     action_type = Solver.do_performance_test
     fig_size = (12, 12)
     if action_type == Solver.do_plot:
+        exclude_solver_names = ['122seq_40shf', '500seq_12shf']
         performance(**locals())
     elif action_type == Solver.do_performance_test:
         nb_samples = 50
-        nb_cpus = 1
+        nb_cpus = 10
+        performance(solver_type=Solver.astar,
+                    heuristic_type=Heuristic.deep_learning,
+                    learner_type=Learner.deep_learner,
+                    time_out=1200,
+                    add_perfect_shuffle=False,
+                    min_nb_shuffles=12,
+                    max_nb_shuffles=12,             #  ->  one more to re-try for nb_shuffle=10 #  -> try up to 12 to start with
+                    step_nb_shuffles=2,
+                    nb_sequences=100,
+                    nb_shuffles=32,
+                    nb_epochs=100000,
+                    layers_description=(600, 300, 100),
+                    training_data_every_epoch=True,
+                    cap_target_at_network_count=True,
+                    one_hot_encoding=True,
+                    drop_out=0.,
+                    network_type=DeepLearning.fully_connected_net,
+                    learning_rate=1e-3,
+                    scheduler=DeepReinforcementLearner.exponential_scheduler,
+                    gamma_scheduler=0.9999,
+                    nb_shuffles_min=1,
+                    nb_shuffles_max=32,
+                    **locals())
+        exit()
         performance(solver_type=Solver.astar,
                     heuristic_type=Heuristic.deep_learning,
                     learner_type=Learner.deep_reinforcement_learner,
@@ -87,30 +112,6 @@ if '__main__' == __name__:
                     learning_rate=1e-3,
                     scheduler=DeepReinforcementLearner.exponential_scheduler,
                     gamma_scheduler=0.9999,
-                    **locals())
-        exit()
-        performance(solver_type=Solver.astar,
-                    heuristic_type=Heuristic.deep_learning,
-                    learner_type=Learner.deep_learner,
-                    time_out=1800,
-                    add_perfect_shuffle=False,
-                    min_nb_shuffles=10,
-                    max_nb_shuffles=10,             #  -> try up to 12 to start with
-                    step_nb_shuffles=2,
-                    nb_sequences=100,
-                    nb_shuffles=32,
-                    nb_epochs=100000,
-                    layers_description=(600, 300, 100),
[…]

### 2022-09-05 latex/Appendices/Examples.tex
update

--- 2022-08-06 latex/Chapters/Examples.tex
@@
 
-In this chapter, I will go through some examples, illustrating how to use the code base to learn and solve various puzzles.
-For each section, simple examples of code will be indicated in \afblue \textbf{python code} \black paragraphs, and can easily be run from command line or copied into a script and run from your favourite IDE.
+In this appendix, I will go through many examples, illustrating how to use the code base to create, learn and solve various puzzles.
+For each section, a snippet of code will be indicated in \afblue \textbf{python code} \black paragraphs, and can easily be run from command line or copied into a script and run from your favourite Python IDE.
 
@@
 Let me start by showing how to construct puzzles, using the Puzzle factory.
-Notice that in order to run a learner or solver of any kind (assuming of course that they are meant to work on the puzzle type in question), we can just use the exact same code, but just specify \textit{puzzle\_type}  and the expected parameters to construct a puzzle.
+Notice that in order to run a learner or solver of any kind (assuming of course that they are meant to work on the puzzle type in question), we can just use the exact same code, simply specifying \textit{puzzle\_type} and the expected parameters to construct a puzzle.
+The factories will just pick up the relevant parameters indicating the puzzle type and dimension and everything should work seamlessly.
 \\
 \\
-For instance, let us create a (n=5, m=6) SP, shuffle it a number of times, and print it:
+For instance, let us create a 5x6 \textbf{SP}, shuffle it a thousand times, and print it:
 \afblue
@@
 %\decoRule
-\caption[Examples]{sliding puzzle construction example}
+\caption[Examples]{scrambled 5x6 \textbf{SP} example}
 \label{fig:examplespconstruction}
 \end{figure}
-.. similarly to construct a (n=2) RC and shuffle it perfectly:
+.. similarly to construct a perfectly scrambled 2x2x2 \textbf{RC}:
 \afblue
-\paragraph{}{\textbf{python code -- rubiks cube construction}}
+\paragraph{}{\textbf{python code -- Rubik's cube construction}}
 \begin{python}
@@
 The perfect learner has been discussed in details in \ref{sec:codelearners}.
-We simply show here how to run it to learn the value function for the (n=2, m=3) SP, that is, to solve all 360 possible configurations via an optimal solver (A{*} with Manhattan heuristic).
+We simply show here how to run it to learn the value function for the 2x3 \textbf{SP}, that is, to solve all 360 possible configurations via an optimal solver (A{*} with Manhattan heuristic).
 
@@
 \black
-The above snippet of code will solve the 2 by 3 SP and then display the results, showing the distribution of optimal costs, as well as the most difficult puzzle.
+The above snippet of code solves the 2x3 \textbf{SP} and then displays the results, showing the distribution of optimal costs, as well as the most difficult puzzle:
 
@@
 \centering
-\includegraphics[scale=0.45]{./Figures/exampleperfectlearner}
+\includegraphics[scale=0.41]{./Figures/exampleperfectlearner}
 %\decoRule
@@
 generated by an optimal solver).
-Here however, I make use of the TrainingData class to generate 100 sequences of fully solved (n=5, m=2) SP via A{*} with Manhattan++.
+Here however, I make use of the TrainingData class to generate 100 sequences of fully solved 5x2 \textbf{SP} via A{*} with Manhattan++.
 
@@
 \subsubsection{Deep Reinforcement Learning Heuristic}
+\subsubsection{Deep Q Learning Heuristic}

### 2022-09-05 latex/Appendices/Kociemba.tex
update

@@
 \afblue
-\paragraph{}{\textbf{Kociemba bug -- 2x2x2 example of good interface design}}
+\paragraph{}{\textbf{Kociemba bug -- 2x2x2 \textbf{RC} -- example of good interface design}}
 \begin{python}
@@
 %\decoRule
-\caption[Kociemba Bug]{2x2x2}
+\caption[Kociemba Bug]{2x2x2 \textbf{RC} -- hkociemba finds path to closest of the 24 equivalent goals}
 \label{fig:KociembaBug222}
 \end{figure}
-As we can see, hkociemba happily returned a 1-move solution to green F, white U, instead of the costlier solution it would have taken to get red F white U (clearly at a cost of 3 with solution RUD').
+As we can see, hkociemba happily returned a 1-move solution R to green F, white U, instead of the costlier solution it would have taken to get to \textit{standard} form red F white U (clearly at a cost of 3 with solution RUD').
 \\
@@
 \afblue
-\paragraph{}{\textbf{Kociemba bug -- 3x3x3 example of \textbf{bad} interface design}}
+\paragraph{}{\textbf{Kociemba bug -- 3x3x3 \textbf{RC} -- example of \textbf{bad} interface design}}
 \begin{python}
@@
 %\decoRule
-\caption[Kociemba Bug]{3x3x3 Buggy}
+\caption[Kociemba Bug]{3x3x3 \textbf{RC} -- kociemba does not check center cubies and returns wrong solution}
 \label{fig:KociembaBug333}
@@
 That way, we seamlessly get to an answer that makes sense from kociemba, without the user having to worry about whether or not the cube they are passing as input is in \textit{standard} form or not!
-Running the same example as above via my KociembaSolver, we get the expected result:
+Running the same example as above via my KociembaSolver, we get the expected result, composed of 3 cube rotations to bring it to \textit{standard} form, followed by one face rotation to solve it:
 
 \afblue
-\paragraph{}{\textbf{Kociemba bug -- 3x3x3 massage to fix the broken interface}}
+\paragraph{}{\textbf{Kociemba bug -- 3x3x3 \textbf{RC} --  massage to fix the broken interface}}
 \begin{python}
@@
 \includegraphics[scale=0.5]{./Figures/KociembaBug333Fix1}
-%
+%\decoRule
+\label{fig:KociembaBug333Fix1}
+\end{figure}
+
+\begin{figure}[H]
+\centering
 \includegraphics[scale=0.5]{./Figures/KociembaBug333Fix2}
 %\decoRule
-\caption[Kociemba Bug Fix]{3x3x3 Fixed}
-\label{fig:KociembaBug333Fix}
+\caption[Kociemba Bug Fix]{3x3x3 \textbf{RC} -- Fixed kociemba interface}
+\label{fig:KociembaBug333Fix2}
 \end{figure}

### 2022-09-05 latex/Chapters/DRLSearch.tex
update

@@
 
-In this section, I will succinctly go through the different techniques that underly the algorithms I have implemented and compared in this project.
+In this chapter, I will succinctly go through the different techniques that underly the algorithms I have implemented and compared in this project and present them in (very simplified) pseudo code.
 Along the way, I will give some references for the interested reader.
@@
 
-\section{Graph Search \& Heuristics}
+\section{Graphs Search \& Heuristics}
 
@@
 \\
+\afblue
+\paragraph{}{\textbf{pseudo code -- BFS \& DFS}}
+\begin{python}
+###############################################################################
+def blind_search(initial_node, time_out):
+    """ pseudo code for BFS/DFS """
+    initialize_time_out(time_out)
+    if initial_node.is_goal():
+        return initial_node
+    explored = set()
+    frontier = [initial_node]
+    while True:
+        check_time_out() # -> raise TimeoutError if appropriate
+        if frontier.empty():
+            return None # -> Failure to find a solution
+        node = pop(frontier) # pop as FIFO/LIFO for BFS/DFS
+        explored.add(node)
+        for child in node.children() and not in frontier.union(explored):
+            if child.is_goal():
+                return child
+            frontier.add(child)
+###############################################################################
+\end{python}
+\black
 \textit{Informed} strategies, on the other hand, exploit domain-specific knowledge.
@@
 In short, admissible heuristics never over-estimate the real cost-to-go, and can easily be shown to render \textbf{A$^{*}$} optimal.
+\afblue
+\paragraph{}{\textbf{pseudo code -- A$^{*}$}}
+\begin{python}
+###############################################################################
+def A*(initial_node, time_out, heuristic):
+    """ pseudo code for A*
+    g = cost from initial_node to a current node
+    h = heuristic (expected remaining cost from current node to goal)
+    """
+    initialize_time_out(time_out)
+    node = initial_node
+    cost = heuristic(node) # g = 0
+    frontier = sorted_multi_container({cost: {node}})
+    explored = set()
+    while True:
+        check_time_out()
+        if frontier.empty():
+            return None # -> Failure to find a solution
+        (cost, node) = frontier.pop_smallest() # smallest total expected cost
+        if node.is_goal():
+            return node
+        explored.add(node)
+        for child in node.children(): # -> children have g = node.g + 1
+            child_cost = child.g + heuristic(child)
+            if child in frontier and child_cost < cost:
+                frontier.update_cost(child, child_cost)
+            elif child not in explored:
+                frontier.add(child, child_cost)
+###############################################################################
+\end{python}
+\black
 
@@
 \section{Reinforcement Learning}
+\label{sec:RLTheory}
 
@@
 \textbf{RL} is concerned with how intelligent agents can learn optimal decisions from observing/experiencing rewards while interacting in their environment.
-One of the early concept in the field is that of value function $s \to \textbf{V}(s)$, which tells us the maximum expected reward the agent can obtain from a given state $s$ (if it takes the optimal decisions).
-In discrete states and transitions spaces, the value function can be remarkably computed (or at least approximated) by a rather mechanical and magic-like procedure called value-iteration (which uses the equivalent of the Bellman equation from optimal control), where each state $s$'s value is iteratively refined by combining the value of $s$'s reachable states.
+One of the fundamental concept in the field is that of value function $s \to \textbf{V}(s)$, which tells us the maximum expected reward - or equivalently and better suited to our puzzle solving task, the minimum cost - the agent can obtain from a given state $s$ (if it takes optimal decisions).
+In finite state and transitions spaces, the value function can be remarkably computed by a rather mechanical and magic-like procedure called value-iteration (which uses the equivalent of the Bellman equation from optimal control), where each state $s$'s value is iteratively refined by combining the value of $s$'s reachable children states.
+\afblue
+\paragraph{}{\textbf{pseudo code -- \textbf{RL} Value Iteration}}
+\begin{python}
+###############################################################################
+def value_iteration(states):
+    """ pseudo code for RL value iteration """
+    V = {state: inf for state in states}
+    change = True
+    while change:
+        change = False
+        for state, cost in V:
+            V[state] = min(V[child] + t_cost for child, t_cost in state.children())
+            change |= V[state] < cost
+    return V 
+###############################################################################
+\end{python}
+\black
 
@@
 Marvin Minsky, who often took the blame for having \textit{killed} funding into the field, even regretted writing his 1969 book (\cite{minskypapert69}) in which he had proved that three-layer feed forward perceptrons were not quite the universal functions approximators some had hypothesized them to be.
-Since around 2006 (see \cite{GoodBengCour16} for more history of the recent burgeoning of \textbf{DL}), the field has known a rebirth, probably due to combination of factors, from ever more powerful computers and larger storage, the availability of data sets large and small on the internet, the advances in many techniques and heuristics (backward propagation, normalisation, drop-outs, different optimisation schemes, etc...) and the huge amount of experimentation with different architectures (from very deep feed forward fully connected networks, to convolutional, recurrent and more exoctic networks).
+Since around 2006 (see \cite{GoodBengCour16} for more history of the recent burgeoning of \textbf{DL}), the field has known a rebirth, probably due to combination of factors, from ever more powerful computers and larger storage, the availability of data sets large and small on the internet, the advances in many techniques and heuristics (backward propagation, normalisation, drop-outs, different optimisation schemes, etc \dots) and the huge amount of experimentation with different architectures (from very deep feed forward fully connected networks, to convolutional, recurrent and more exoctic networks).
 It is not an exageration to say that \textbf{DL} has rendered obsolete entire fields of research and bodies of knowledge, most notably in \textbf{CV}, robotics, computational biology and \textbf{NLP}.
@@
 The beauty of it is that the relevant features are autodiscovered during the training of the network, via back-ward propagation and trial-and-error, rather than having to be postulated or handcrafted as is often the case with other \textbf{ML} techniques.
-\\
-\\
-For this project, I have implemented the procedure of training a \textbf{DL} network from the solutions of other solvers/heuristics in a class called DeepLearner (see section \ref{sec:codelearners} for details).
 
@@
 
-So what is the link between all of these - A$^{*}$'s heuristics, \textbf{DL}, \textbf{RL}'s value iteration?
+So what is the link between all of these: A$^{*}$'s heuristics, \textbf{DL}, \textbf{RL}'s value iteration?
 \\
@@
 This is a case of \textit{supervised} learning in some sense, since we need a teacher-solver to tell us good solutions or heuristics to extrapolate from.
+In very simplified pseudo code, the procedure looks like the following:
+\afblue
+\paragraph{}{\textbf{pseudo code -- \textbf{DL} heuristics}}
+\begin{python}
+###############################################################################
+def deep_learning(puzzle_type,
+                  teacher_heuristic,
+                  loss_function,
+                  network_architecture):
+    """ pseudo code for learning a puzzle via deep learning and a teacher heuristic """
+    target_value_function = {state: teacher_heuristic(state)
+                             for state in generate_random_states(puzzle_type)}
+    dl_heuristic = train_neural_network(target_value_function,
+                                        loss_function,
+                                        network_architecture)
+    return dl_heuristic
+###############################################################################
+\end{python}
+\black
+
+For this project, I have implemented the above procedure of training a \textbf{DL} network from the solutions of other solvers/heuristics in a class called DeepLearner (see section \ref{sec:codelearners} for details).
 \\
 \\
-An alternative approach in order to compute the cost-to-go is via \textbf{RL} value-iteration.
-Indeed, minimizing the cost-to-go (heuristic) is equivalent to maximising the cumulative reward given by -1 for each transition not reaching the goal state and +1 for the last transition to the goal.
-Clearly these two quantities are equivalent (modulo a trivial linear transformation).
-We can perform this value-iteration in an unsupervised way: we do not need a teacher to tell us the target value of the value function/cost-to-go but instead learn it simply from refining a state $s$'s value by replacing it at each iteration by the minimum of its current value, and the minimum value of all reachable states from $s$ polus one (since it takes one move to reach them).
-That is, we can start with arbitrary (e.g.
-0) $V_{0}(s0$ and recursively update for $t \in \mathbb{N}^{*}, s \in \mathcal{S}$, the state space via the following iteration:
-\begin{equation} \label{eq:DLVI}
-\begin{split}
-if \  s & \in \mathcal{G}: \\
-& \ V_{t+1}(s) = 0 \\
-else&: \\
-& \ V_{t+1}(s) \leftarrow min \Big( V_{t}(s), 1 + min_{s' \in \mathcal{C}(s)}(V_{t}(s'))  \Big)
-\end{split}
-\end{equation}
-where $\mathcal{C}(s)$ are the children (reachable states via 1 move) of $s$ and $\mathcal{G}$ are the goal states.
-\\
-\\
-The state space is often so large however, that we cannot practically perform this procedure for all states.
+An alternative approach in order to compute the cost-to-go is via \textbf{RL} value-iteration as described in the previous section (\ref{sec:RLTheory}) The state space is often so large however - and this certainly is the case for pretty much all \textbf{SP} and \textbf{RC}, except maybe the smallest dimensional \textbf{SP}s - that we cannot practically perform the value iteration procedure for all states.
 This is where \textbf{DL} comes again to the rescue to act as function approximator, combining with \textbf{RL} into what has become known as \textbf{DRL}.
-The subtelty is then that both the left-hand-side and right hand side in the value-iteration procedure \ref{eq:DLVI} are given by a \textbf{DL} network which we are constantly tweaking.
+The subtelty is then that both the left-hand-side and right hand side in the value-iteration procedure are given by a \textbf{DL} network which we are constantly tweaking.
 For my implementation of \textbf{DRL}, I followed the same approach as in \cite{Mnih2013}, utilising two loops.
@@
 The outer loop is itself broken out of when deemed appropriate (some kind of convergence criteria).
+In simplified pseudo-code, this looks like:
 \\
 \\
+\afblue
+\paragraph{}{\textbf{pseudo code -- \textbf{DRL} heuristics}}
+\begin{python}
+###############################################################################
+def deep_reinforcement_learning(puzzle_type,
+                                loss_function,
+                                max_epochs,
+                                target_network_update_criteria,
+                                puzzles_generation_criteria,
+                                network_architecture):
+    """ pseudo code for learning a puzzle via deep reinforcement learning """
+    network = get_network(puzzle_type, network_architecture)
+    target_network = get_network(puzzle_type, network_architecture)
+    epoch = 0
+    puzzles = list()
+    while epoch < max_epochs and other_convergence_criteria(network):
+        epoch += 1
+        """ Generate a new bunch of puzzles """
+        if puzzles_generation_criteria(puzzles):
+            puzzles = generate_random_states(puzzle_type)
+        V = dict()
+        """ Compute their updated cost via value iteration ... 
+             all moves are assumed to have cost 1 """
+        for puzzle in puzzles:
+            if puzzle.is_goal():
+                V[puzzle] = 0
+            else:
+                V[puzzle] = min(target_network(puzzle),
+                                1 + min(target_network(child) for child in puzzle.children()))
+        """ Train left-hand-side network to approximate right-hand-side target network better
+        i.e. perform a forward / backward-propagation update of the network
+        """
+        network = train_neural_network(V,
+                                       loss_function,
+                                       network_architecture)
+        """ Update the target network if criteria are met 
+        (e.g. epochs, convergence, no progress, etc...) """
+        if target_network_update_criteria(network, target_network):
+            target_network = copy(network)
+    return network
+###############################################################################
+\end{python}
+\black
 For this project, I have implemented this procedure in a class called DeepReinforcementLearner (see section \ref{sec:codelearners} for details)
@@
 \\
+I have for this project followed the exact same procedure described in \cite{https://doi.org/10.48550/arxiv.1805.07470} (and also used by alpha-go \cite{AlphaGo}): the output of the neural network that is learnt by \textbf{DQL}, which actually is surprisingly similar to the approach and pseudo code shown for \textbf{DRL} earlier.
+The only 3 differences are that:
+\begin{itemize}
+\item the networks (target and current) now output a vector of size $a$ + 1, where $a$ is the number of possible actions/moves.
+That is, one dimension for each of the possible actions, representing the probability/desirability of that action, and one dimension for the value-function's value.
+\item the target update by Q-iteration now needs to update both the value-function and the optimal decision ($a$-dimensional vector of 0s, with a 1 for the optimal decision only)
+\item the loss function used need to be compatible with the size of the output and target vectors.
+I have used as loss the average of the mean square error of the value function difference (current network and target) and the cross entropy loss of the actions probability versus the actual optimal move.
+\item
+Since in my \textbf{DRL} pseudo-code above (as well as in my actual code), the loss function is abstracted, and the network architecture (including the size of the output) configurable, the only modification to get \textbf{DQL} in lieu of \textbf{DRL} are extremely minimal, and essentially only to do with the Q-iteration-update section:
+
+\afblue
+\paragraph{}{\textbf{pseudo code -- \textbf{DQL} heuristics}}
+\begin{python}
+###############################################################################
+def deep_q_learning(... same as DRL ...):
+    """ pseudo code for learning a puzzle via deep q learning """
+    # ... same as DRL ...
+    nb_actions = puzzle_type.get_nb_actions()
+    while epoch < max_epochs and other_convergence_criteria(network):
+        # ... same as DRL ...
+        Q_and_V = dict()
+        """ Compute updated cost and actions via V & Q iteration ... 
+             all moves are assumed to have cost 1 """
+        for puzzle in puzzles:
+            value = target_network(puzzle)[-1]
+            actions = [0] * nb_actions
+            best_action_id = 0
+            if puzzle.is_goal():
+                value = 0
+            else:
+                for action_id, child in puzzle.children():
+                    child_value = target_network(child)[-1]
+                    if child_value < value:
+                        value = child_value
+                        best_action_id = action_id
+            # We update both the value function and the best action
+            actions[best_action_id] = 1
+            Q_and_V[puzzle] = actions + [value]
+        # ... same as DRL ...
+###############################################################################
+\end{python}
+\black
+
+\end{itemize}
 For this project, I have implemented the procedure described in \cite{https://doi.org/10.48550/arxiv.1805.07470} in a class called DeepQLearner (see section \ref{sec:codelearners} for details).
@@
 
-In my implementation of \textbf{DQL}, which follows \cite{https://doi.org/10.48550/arxiv.1805.07470} (and is also the same procedure followed by alpha-go \cite{AlphaGo}), the output of the neural network that is learnt by \textbf{DQL} is no longer just a cost-to-go estimate as with \textbf{DRL}, but a joint cost-to-go and distribution over actions (to be concrete, that means a 5-dimensional vector for the \textbf{SP} and a 13-dimensional vector for the \textbf{RC}).
+Since my implementation of \textbf{DQL} is no longer just a cost-to-go estimate as with \textbf{DRL}, but a joint cost-to-go and distribution over actions (to be concrete, that means a 5-dimensional vector for the \textbf{SP} and a 13-dimensional vector for the \textbf{RC}).
 This leaves us with the question of how can we use the output of such a network to search for a solution.

### 2022-09-05 latex/MScAIFrancoisBerrier.tex
update

@@
 \\
-I would also like to thank Professor Chris Watkins, inventor of QL :), for agreeing to supervise my project.
+I would also like to thank Professor Chris Watkins, for agreeing to supervise my project.
 It was an honour to attend his very lively lectures on Deep Learning during the first year of the MSc, as well as to absorb some of his wisdom and enthusiasm at our meetings during my work on this project.
+I am also glad to have followed his advice at our last meeting: I stopped shaving for a month, which did free enough time for me to implement DQL (joint policy and value learning), as well as a Monte Carlo Tree Search algorithm, to nicely complement my assortment of sliding-puzzle and Rubik's cube solvers.
 \\
 \\
-Last but not least, my wife Thanh Nh\~a is my biggest supporter in everything I do, and her continued encouragement and enthusiasm in my pursuit of this MSc has made it so much easier to juggle between a demanding banking job, running 50 to 70 miles a week in preparation for the Chicago marathon, and the coursework, lectures and exam revisions.
+Last but not least, my wife Thanh Nh\~a is my biggest supporter in everything I do, and her continued encouragement in my pursuit of this MSc has made it so much easier to juggle between a demanding banking job, running 50 to 70 miles a week in preparation for the Chicago marathon, and the coursework, lectures and exam revisions.
 %\end{comment}

### 2022-09-05 rubiks/scripts/pseudocode/bfs.py

@@ -1,0 +1,23 @@
+##########################################################################################
+def blind_search(initial_node, time_out):
+    initialize_time_out(time_out)
+    """ pseudo code for BFS/DFS """
+    if initial_node.is_goal():
+        return initial_node
+    explored = set()
+    frontier = [initial_node]
+    while True:
+        check_time_out() # -> raise TimeoutError if appropriate
+        if frontier.empty():
+            """ Failure to find a solution """
+            return None
+        node = pop_from_frontier(frontier) # pop_from_frontier is FIFO/LIFO for BFS/DFS
+        explored.add(node)
+        for chile in node.children():
+            if child in frontier.union(explored):
+                continue
+            if child.state.is_goal():
+                return child
+            frontier.add(child)
+##########################################################################################
+

### 2022-09-05 rubiks/scripts/pseudocode/dql.py

@@ -1,0 +1,31 @@
+##########################################################################################
+def deep_q_learning(... same as DRL ...):
+    """ pseudo code for learning a puzzle via deep q learning """
+    # ... same as DRL ...
+    nb_actions = puzzle_type.get_nb_actions()
+    while epoch < max_epochs and other_convergence_criteria(network):
+        # ... same as DRL ...
+        Q_and_V = dict()
+        """ Compute their updated cost via value iteration ... 
+             all moves are assumed to have cost 1 """
+        for puzzle in puzzles:
+            value = target_network(puzzle)
+            actions = [0] * nb_actions
+            best_action_id = 0
+            if puzzle.is_goal():
+                value = 0
+            else:
+                for action_id, child in puzzle.children():
+                    child_value = target_network(child)[-1]
+                    if child_value < value:
+                        value = min(value, child_value)
+                        best_action_id = action_id
+            # We update both the value function and the best action
+            actions[best_action_id] = 1
+            Q_and_V[puzzle] = actions + [value]
+        """ Train left-hand-side network to approximate right-hand-side target network better
+        i.e. perform a forward / backward-propagation update of the network
+        """
+        # ... same as before ...
+##########################################################################################
+

### 2022-09-05 rubiks/scripts/rubiks22/perf.py

@@ -69,7 +69,28 @@ if '__main__' == __name__:
         performance(**locals())
     elif action_type == Solver.do_performance_test:
         nb_samples = 100
-        nb_cpus = 8
+        nb_cpus = 1
+        performance(solver_type=Solver.astar,
+                    heuristic_type=Heuristic.deep_learning,
+                    learner_type=Learner.deep_reinforcement_learner,
+                    time_out=600,
+                    add_perfect_shuffle=False,
+                    min_nb_shuffles=12,
+                    max_nb_shuffles=12,
+                    nb_sequences=10000,
+                    nb_shuffles=15,
+                    nb_epochs=10000,
+                    layers_description=(600, 300, 100),
+                    training_data_every_epoch=False,
+                    cap_target_at_network_count=False,
+                    one_hot_encoding=True,
+                    drop_out=0.,
+                    network_type=DeepLearning.fully_connected_net,
+                    learning_rate=1e-2,
+                    scheduler=DeepReinforcementLearner.exponential_scheduler,
+                    gamma_scheduler=0.9999,
+                    **locals())
+        exit()
         performance(solver_type=Solver.astar,
                     heuristic_type=Heuristic.deep_learning,
                     learner_type=Learner.deep_reinforcement_learner,

### 2022-09-05 latex/Chapters/Puzzles.tex
update

@@
 %-----------------------------------
-\subsection{History - The 15 Puzzle}
+\subsection{History - the 15-Puzzle}
 
-The first puzzle I will focus on is the sliding puzzle (see \cite{SlidingPuzzleWiki}).The 15-puzzle seems to have been invented in the late 19th century by Noyes Chapman (see \cite{SlidingPuzzleWolfram}), who applied in 1880 for a patent on what was then called "Block Solitaire Puzzle".
+The first puzzle I will focus on in this thesis is the sliding puzzle (see \cite{SlidingPuzzleWiki}) whose most common variant is the 15-puzzle, seemingly invented in the late 19th century by Noyes Chapman (see \cite{SlidingPuzzleWolfram}), who applied in 1880 for a patent on what was then called "Block Solitaire Puzzle".
 In 1879, a couple of interesting notes (\cite{Johnson1879}), published in the American Journal of Mathematics proved that exactly half of the $16!$ possible ways of setting up the 16 tiles on the board lead to solvable puzzles.
@@
 
-\subsection{History}
+\subsection{History -- from Erno Rubik to speed cubing competitions}
 The second puzzle I will focus on is the well known Rubik's Cube (\textbf{RC}), invented in 1974 by Erno Rubik, originally known under the name of \textit{Magic Cube} (see \cite{RubiksWiki}).
-Since it was commericalised in 1980, it has kown a worldwide success: countless people are playing with the famous original 3x3x3 cube, as well as with countless variations, more or less difficult of it.
+Since it was commercialized in 1980, it has kown a worldwide success: countless people are playing with the famous original 3x3x3 cube, as well as with many variants of it, more or less difficult and contrived.
 Competitions nowadays bring together \textit{speedcubers} who have trained for years and memorized various algorithms to solve the Rubik's in astonishly effective and fast times (literally in seconds for the best ones).
 Interestingly, some of the principles used by speedcubers generalise to different dimensions.
-As an example, it is quite impressive to watch \textit{Cubastic} solve a 15x15x15 RC in two and a half hours (\cite{151515Rubiks}).
+As an example, it is quite impressive to watch \textit{Cubastic} solve a 15x15x15 cube in (only) two and a half hours!
+(see \cite{151515Rubiks}).
+The key insight to solving high dimensional \textbf{RC}s (short obviously of coming up with specific and sophisticated strategies) is the following: if we manage to group all the inside (i.e.
+not located on a corner or an edge of the cube) cubies by colour, and then all the non-corner edges on a given tranche of a face by colour, we are essentially reducing the cube to a 3x3x3, since no face rotation will ever undo the inside cubies and non-corner-edge cubies' grouping.
+For instance, the following picture shows faces of my own 4x4x4 and 3x3x3 cubes which I have arranged to be in identical configurations.
+\\
+\\
+It is quite interesting to note that the difficuly of hand-solving cubes does not grow nearly as much with increasing dimensionality (most of the time difference comes rather from the difficulty of handling and manipulating the cubes of high dimensions at speed).
 
-\subsection{Search Space \& Solvability}
-In my code as well as in this write-up, as is very customary in the RC literature, I will refer to the faces as F (front), B (back), U (up), D (down), L (left) and R (right).
-The RC's tiles are of 6 different colors, and without loss of generality, we can consider that these colors are also called F, B, U, D, L and R, and consider the goal state as one where, up to rotations of the whole cube, the color's name match the faces (color F on face F, etc).
-The RC has, as one would expect, an extremely large state space.
-If we ignore the equivalence via full-cube rotations, there are 24 goal states, since any of the colors can be placed on the F face, and then any of the 4 adjacent colors can be placed on the U face (say).
-Once these 2 are chosen, the remaining faces are determined.
+\begin{figure}[H]
+\centering
+\includegraphics[scale=0.3]{./Figures/RC333444}
+\decoRule
+\caption[Reducing RC]{Reducing a 4x4x4 \textbf{RC} to a 3x3x3 -- home made picture}
+\end{figure}
+
+ It is quite clear from that picture above that if we can bring the 4x4x4 cube to such a configuration on all faces (which is not super hard), and we know how to solve a 3x3x3, we are essentially done.
+(well, almost, since the additional degrees of freedom coming from the higher dimension of the 4x4x4 mean that certain arrangements of the colors which are possible on the 4x4x4 are not on the 3x3x3 (for instance we can never have the center red and white on opposite faces of a 3x3x3, but we easily can on a 4x4x4!
+Fortunately that is nothing that a couple of easily learnable algorithms of a few moves cannot fix).
+
+\subsection{Conventions: faces, moves and configurations}
+In my code as well as in this write-up, I follow the conventions and notations of the \textbf{RC} literature and community as much as possible, I will refer to the faces as F (front), B (back), U (up), D (down), L (left) and R (right).
+The cubes' tiles are of 6 different colors, and without loss of generality (though the interested reader is referred to appendix \ref{sec:Kociemba} for some subtlety around that), we can consider that these colors are also called F, B, U, D, L and R, and consider the goal state as one where, up to rotations of the whole cube, the color's name match the faces (color F on face F, etc).
+The Rubik's has, as one would expect, an extremely large state space.
+I will compute the state space size for the 3x3x3 and 2x2x2 cubes in the next two subsections and discuss conditions under which  a cube is solvable.
+\\
+It is common in \textbf{RC} jargon to refer to moves  by the name of the face that is rotated clock-wise (when facing it), and by adding a prime \textit{'} for counter-clock-wise rotations.
+Sometimes people use a number after the face to indicate repeated similar moves, though I will count in my code that as 2 moves, since obviously e.g.
+U2 = U U = U' U'.
+As another example: F B' F2 means rotating the front face clock-wise, the back face counter-clock-wise, followed by the front face twice.
+\\
+Finally, let me make a remark on what I consider in this project to be distinct configurations of the Rubik's cube.
+Since I am interested in experimenting with, among other methods, \textbf{DRL} and \textbf{DQL}, and that they are usually presented as methods which learn \textit{without-human-knowledge}, I thought it would be a bit akin to cheating to consider that two cubes equivalent by full space rotation are one and the same and represent them as such in my code.
+Ideally, we would rather let the algorithms learn (if they are able to) about symmetries.
+For instance, the two following cubes (which are clearly just a rotation away or equivalently a change-in-observer away from one another) are the same for us humans.
+I however represent them as different tensors in my code, and this is up to the solvers to make sense of things.
+
+\begin{figure}[H]
+\centering
+\includegraphics[scale=0.4]{./Figures/equivalentcubes}
+\caption[Reducing RC]{Two equivalent configurations by full-cube-rotation}
+\label{fig:configurations}
+\end{figure}
+
+\subsubsection{3x3x3 Rubik's -- search space, solvability \& God's number}
+
+Firstly, it is important to realise that each configuration can be arranged in (only) 24 different ways via full cube rotation in space.
+In particular that means that there are 24 possible goal states (and not 6!
+as we might naively think at first), so for instance in figure \ref{fig:configurations} I could only have picked 2 out of 24 different configurations.
+Why is this?
+\\
+\\
+Any of the six colors can be placed on the F face, and then any of the four adjacent colors can be placed on the U face (say).
+Once these two colors are chosen, the remaining faces are determined.
 This is not immediately obvious, but not difficult to convince oneself that this is the case with the following two observations about the structure of corner \textit{cubies}: first observation is that they have 3 determined colors, which are fixed once and for all.
-In particular, the fact that there is no corner with colors F and B, means that once the RC is in solved state, we have to have colors F and B on opposite faces, and similarly for colors U and D as well as colors L and R.
+In particular, the fact that there is no corner with colors F and B, means that once the cube is in solved state, we have to have colors F and B on opposite faces, and similarly for colors U and D as well as colors L and R.
 Hence, having fixed the color on face F, the color on face B is fixed too.
@@
 For instance, consider the corner cubie with colors (F, R, U).
-There is no scrambling of the cube that will produce clock-wise order of e.g.
+No scrambling of the cube will ever produce clock-wise order of e.g.
 (F, U, R).
@@
 \\
-Finally a further word of notation.
-It is common in RC jargon to refer to moves  by the name of the face that is rotated clock-wise (when facing it), and by adding a prime \textit{'} for counter-clock-wise rotations.
-Sometimes people use a number after the face to indicate repeated similar moves, though I will count in my code that as 2 moves, since obviously e.g.
-U2 = U U = U' U'.
-As another example: F B' F2 means rotating the front face clock-wise, the back face counter-clock-wise, followed by the front face twice.
+\\
 
-\subsubsection{2 x 2 x 2}
+\cite{RubiksGodNumber}
+\\
+\cite{RubiksChicago}
+\\
+\cite{RubiksRadu}
+\\
+
+\subsubsection{2x2x2 Rubik's -- search space, solvability \& God's number}
 
@@
 
-\subsubsection{3 x 3 x 3}
-
-TBD
-
-\subsection{Optimal Cost \& God's Number}
-
-\cite{RubiksGodNumber}
-\\
-\cite{RubiksChicago}
-\\
-\cite{RubiksRadu}
-\\
-
 %-----------------------------------

### 2022-09-05 latex/Chapters/DRLSearch.tex
update

@@
 \\
-\afblue
+\teal
 \paragraph{}{\textbf{pseudo code -- BFS \& DFS}}
-\begin{python}
+\begin{pseudocode}
 ###############################################################################
-def blind_search(initial_node, time_out):
+def blind_search(initial_node, time_out, search_type=Search.BFS):
    """ pseudo code for BFS/DFS """
@@
             return None # -> Failure to find a solution
-        node = pop(frontier) # pop as FIFO/LIFO for BFS/DFS
+       # FIFO/LIFO for BFS/DFS
+        node = pop_front(frontier) if search_type is Search.BFS else pop_back(frontier)
         explored.add(node)
@@
 ###############################################################################
-\end{python}
+\end{pseudocode}
 \black
@@
 In short, admissible heuristics never over-estimate the real cost-to-go, and can easily be shown to render \textbf{A$^{*}$} optimal.
-\afblue
+\teal
 \paragraph{}{\textbf{pseudo code -- A$^{*}$}}
-\begin{python}
+\begin{pseudocode}
 ###############################################################################
@@
 One of the fundamental concept in the field is that of value function $s \to \textbf{V}(s)$, which tells us the maximum expected reward - or equivalently and better suited to our puzzle solving task, the minimum cost - the agent can obtain from a given state $s$ (if it takes optimal decisions).
-In finite state and transitions spaces, the value function can be remarkably computed by a rather mechanical and magic-like procedure called value-iteration (which uses the equivalent of the Bellman equation from optimal control), where each state $s$'s value is iteratively refined by combining the value of $s$'s reachable children states.
-\afblue
+In finite state and transitions spaces, the value function can be remarkably computed by a rather mechanical and magic-like procedure called value-iteration (kind of the equivalent of the Bellman equation from optimal control), where each state $s$'s value is iteratively refined by combining the value of $s$'s reachable children states.
+\teal
 \paragraph{}{\textbf{pseudo code -- \textbf{RL} Value Iteration}}
-\begin{python}
+\begin{pseudocode}
 ###############################################################################
@@
 Marvin Minsky, who often took the blame for having \textit{killed} funding into the field, even regretted writing his 1969 book (\cite{minskypapert69}) in which he had proved that three-layer feed forward perceptrons were not quite the universal functions approximators some had hypothesized them to be.
-Since around 2006 (see \cite{GoodBengCour16} for more history of the recent burgeoning of \textbf{DL}), the field has known a rebirth, probably due to combination of factors, from ever more powerful computers and larger storage, the availability of data sets large and small on the internet, the advances in many techniques and heuristics (backward propagation, normalisation, drop-outs, different optimisation schemes, etc \dots) and the huge amount of experimentation with different architectures (from very deep feed forward fully connected networks, to convolutional, recurrent and more exoctic networks).
+Since around 2006 (see \cite{GoodBengCour16} for more history of the recent burgeoning of \textbf{DL}), the field has known a rebirth, probably due to a combination of factors, from ever more powerful computers and larger storage, the availability of data sets large and small on the internet, the advances in many techniques and heuristics (backward propagation, normalisation, drop-outs, different optimisation schemes, etc \dots) and the huge amount of experimentation with different architectures (from very deep feed forward fully connected networks, to convolutional, recurrent and more exoctic networks).
 It is not an exageration to say that \textbf{DL} has rendered obsolete entire fields of research and bodies of knowledge, most notably in \textbf{CV}, robotics, computational biology and \textbf{NLP}.
@@
 Sometimes we have ways of computing good solutions/heuristics to our puzzles, but cannot computationally do this for all possible states (maybe the state space is simply too large!).
-One approach in that case is to compute the solution/heuristic for some states, and let a \textbf{DL} model learn how to extrapolate to the rest of the state space.
+One approach in that case is to compute the solution/heuristic for \textit{some manageable number of} states, and let a \textbf{DL} model learn how to extrapolate to the rest of the state space.
 This is a case of \textit{supervised} learning in some sense, since we need a teacher-solver to tell us good solutions or heuristics to extrapolate from.
 In very simplified pseudo code, the procedure looks like the following:
-\afblue
+\teal
 \paragraph{}{\textbf{pseudo code -- \textbf{DL} heuristics}}
-\begin{python}
+\begin{pseudocode}
 ###############################################################################
@@
 \\
-An alternative approach in order to compute the cost-to-go is via \textbf{RL} value-iteration as described in the previous section (\ref{sec:RLTheory}) The state space is often so large however - and this certainly is the case for pretty much all \textbf{SP} and \textbf{RC}, except maybe the smallest dimensional \textbf{SP}s - that we cannot practically perform the value iteration procedure for all states.
+An alternative approach in order to compute the cost-to-go is via \textbf{RL} value-iteration as described in the previous section (\ref{sec:RLTheory}).
+The state space is often so large however - and this certainly is the case for pretty much all \textbf{SP} and \textbf{RC}, except maybe the smallest dimensional \textbf{SP}s - that we cannot practically perform the value iteration procedure for all states.
 This is where \textbf{DL} comes again to the rescue to act as function approximator, combining with \textbf{RL} into what has become known as \textbf{DRL}.
@@
 For my implementation of \textbf{DRL}, I followed the same approach as in \cite{Mnih2013}, utilising two loops.
-During iteration over the first (inner) loop, the right hand-side $V$ of \ref{eq:DLVI} is computed using a fixed \textit{target-network}.
+During iteration over the first (inner) loop, the right hand-side $V$ of the value-function update equation is computed using a fixed \textit{target-network}.
 The left hand side $V$ is using another \textit{current-network} which alone is modified by the \textbf{DL} learning (backward propagation and optimization), until convergence is deemed reached.
@@
 \\
-\afblue
+\teal
 \paragraph{}{\textbf{pseudo code -- \textbf{DRL} heuristics}}
-\begin{python}
+\begin{pseudocode}
 ###############################################################################
@@
     """ pseudo code for learning a puzzle via deep reinforcement learning """
-    network = get_network(puzzle_type, network_architecture)
-    target_network = get_network(puzzle_type, network_architecture)
+    net = get_network(puzzle_type, network_architecture)
+    target_net = get_network(puzzle_type, network_architecture)
     epoch = 0
@@
             else:
-                V[puzzle] = min(target_network(puzzle),
-                                1 + min(target_network(child) for child in puzzle.children()))
-        """ Train left-hand-side network to approximate right-hand-side target network better
-        i.e. perform a forward / backward-propagation update of the network
+                V[puzzle] = 1 + min(target_net(child) for child in puzzle.children()))
+        """ Train lhs network to approximate rhs target network better
+        i.e. perform a forward / backward-propagation update
         """
-        network = train_neural_network(V,
+        network = train_neural_net(V,
                                    loss_function,
@@
         (e.g. epochs, convergence, no progress, etc...) """
-        if target_network_update_criteria(network, target_network):
-            target_network = copy(network)
-    return network
+        if target_network_update_criteria(net, target_net):
+            target_net = copy(net)
+    return net
 ###############################################################################
-\end{python}
+\end{pseudocode}
 \black
-For this project, I have implemented this procedure in a class called DeepReinforcementLearner (see section \ref{sec:codelearners} for details)
+For this project, I have implemented the above algorithm in a class called DeepReinforcementLearner which I shall describe in details in section \ref{sec:codelearners}.
 
@@
 
-Another basic concept of \textbf{RL} is that of the Q-function $Q(s, a)$, the maximum expected reward the agent can obtain from state $s$, taking action $a$ and acting optimally afterwards.
+Another fundamental concept of \textbf{RL} is that of the Q-function $Q(s, a)$, the maximum expected reward the agent can obtain from state $s$, taking action $a$ and acting optimally afterwards.
 Notice that the knowledge of $V$ everywhere is equivalent to that of $Q$, since obviously:
@@
 \end{equation}
-where $\mathcal{A}(s)$ are the transitions/actions we can take from state $s$. Conversely, we can recover $Q$ from $V$ since:
+where $\mathcal{A}(s)$ are the transitions/actions available from state $s$. Conversely, we can recover $Q$ from $V$ since:
 \begin{equation} \label{eq:QI2}
@@
 where $R(a)$ is the reward of performing action $a$ and $a(s)$ is the state we reach by performing action $a$ from state $s$.
-Similarly to the way one can compute $V$ via value-iteration (at least in finite settings), $Q$ can also be computed by iteration (See \cite{Watkins1992} for details of convergence of that procedure).
+Similarly to the way one can compute $V$ via value-iteration (at least in finite state and action spaces), $Q$ can also be computed by iteration (See \cite{Watkins1992} for details of convergence of that procedure).
 \\
@@
 I have used as loss the average of the mean square error of the value function difference (current network and target) and the cross entropy loss of the actions probability versus the actual optimal move.
-\item
+\end{itemize}
 Since in my \textbf{DRL} pseudo-code above (as well as in my actual code), the loss function is abstracted, and the network architecture (including the size of the output) configurable, the only modification to get \textbf{DQL} in lieu of \textbf{DRL} are extremely minimal, and essentially only to do with the Q-iteration-update section:
 
-\afblue
+\teal
 \paragraph{}{\textbf{pseudo code -- \textbf{DQL} heuristics}}
-\begin{python}
+\begin{pseudocode}
 ###############################################################################
@@
 ###############################################################################
-\end{python}
+\end{pseudocode}
 \black
 
-\end{itemize}
-For this project, I have implemented the procedure described in \cite{https://doi.org/10.48550/arxiv.1805.07470} in a class called DeepQLearner (see section \ref{sec:codelearners} for details).
-The network jointly learns the value function (heuristic, or cost-to-go) as well as a probability distribution over the actions.
+For this project, I have implemented the above algorithm in a class called DeepQLearner which I shall describe in details in section \ref{sec:codelearners}.
 
@@
 subsection \ref{sec:S33DRLDQL}) is that we could disregard the distribution and only take into account the cost-to-go by simply using $A^{*}$.
-The rationale for doing so is that we might hope that jointly learning the cost-to-go and the actions via a combination of MSE and cross-entropy-loss leads to better heuristics, since we use more information during the learning and iterations.
+The rationale for doing so is that we might hope that jointly learning the cost-to-go and the actions via a combination of MSE and cross-entropy-loss produces a better network and leads to better heuristics, since we use more information during the learning and iterations.
 \\
@@
 In order to prevent always following the same paths (which might take us very far from the goal), \textbf{MCTS} algorithms usually keep track of actions they have followed already and underweight their probability so that subsequent paths explore other actions.
+A pseudo code description of it looks like this:
 \\
-I have implemented the exact \textbf{MCTS} procedure described in \cite{https://doi.org/10.48550/arxiv.1805.07470} in a class called MonteCarloTreeSearch (see section \ref{sec:codesolvers} for details).
+\teal
+\paragraph{}{\textbf{pseudo code -- \textbf{MCTS}}}
+\begin{pseudocode}
+############################################################
+def MCTS(initial_node, time_out, q_v_network):
+    """ pseudo code for Monte Carlo Tree Search """
+    initialize_time_out(time_out)
+    tree = Tree(initial_node)
+    while True:
+        check_time_out()
+        leaf = add_new_leaf(tree, q_v_network)
+        if leaf.is_goal():
+            return BFS(tree) # path to leaf can usually be improved by BFS
+############################################################
+def add_new_leaf(tree, q_v_network):
+    leaf = tree.initial_node
+    while leaf.expanded:
+        pass
+    leaf.expand() # add children of leaf to tree
+    leaf.update_penalties()
+    return leaf
+############################################################
+\end{pseudocode}
+\black
+
+I have implemented the exact \textbf{MCTS} procedure described in \cite{https://doi.org/10.48550/arxiv.1805.07470} and outlined in pseudo code above in a class called MonteCarloTreeSearch (see section \ref{sec:codesolvers} for details).

### 2022-09-05 latex/Chapters/Puzzles.tex
update

@@
 As another example: F B' F2 means rotating the front face clock-wise, the back face counter-clock-wise, followed by the front face twice.
+Some people refer to the convention I have chosen as the \textit{quarter turn metric}, and obvioulsy the length of solutions is affected by which convention/metric we use.
 \\
@@
 
-Firstly, it is important to realise that each configuration can be arranged in (only) 24 different ways via full cube rotation in space.
+A good strategy to compute the state space size of a the Rubik's of a given dimension is to compute:
+\begin{itemize}
+\item $N$: the number of possible arrangements of the \textit{cubies} (individual small cubes that constitute the Rubik's) in space, taking into account physical constraints.
+\item $D$: the number of possible configurations which are equivalent via full cube rotation in space.
+\end{itemize}
+The state space can then be considered to be either $\frac{N}{D}$ or $N$, depending on our representation (as mentioned earlier my code will treat each rotation of the cube in space as a different representation, so I shall not be dividing by $D$ below).
+
+In order to compute $D$, we just need to realise that each configuration can be arranged in (only) 24 different ways via full cube rotation in space.
 In particular that means that there are 24 possible goal states (and not 6!
@@
 Why is this?
-\\
 \\
@@
 \\
+Computing $N$ is a bit tricker.
+For the 3x3x3 Rubik's, each of the 8 corner cubies could be placed at one of the physical corner locations and oriented three different ways (rather than six, since as discussed earlier, the clock-wise orientation of a given corner cubie can never be changed).
+We have therefore $8!
+.
+3^{8}$ corner arrangements.
+Similarly, the 12 edge cubies can be placed at 12 physical edge locations and oriented 2 different ways, so we have another multiplicative factor of $12!
+.
+2^{12}$.
+There are however some well know parities (see \cite{Schoenert} for details) which are invariant under legal moves of the Rubik's cube and which reduce the number of arrangements that are physically attainable via face rotations.
+The \textit{corner orientation parity} (\textbf{COP}) for instance sums over the corner cubies the number of 120$^{o}$ rotations it would take to bring a given cubie to its standard orientation (i.e.
+compared to the top cube in figure \ref{fig:configurations}).
+It is easy to observe that the \textbf{COP} is invariant (\% 3) since any face rotation will displace four corner cubies, two of which see their individual \textbf{COP} increase by 2, and the other two by 1 (so a total increase of 6).
+Similarly, there are two additional parities called \textit{edge orientation parity} (\textbf{EOP}) and \textit{total permutation parity} (\textbf{TPP}) which are both invariant (\% 2) under face rotation.
+Overall, these three parities mean that only $1/12$ of the $8!
+.
+3^{8} .
+12!
+.
+2^{12}$ arrangements are physically attainable, leaving us with a grand total of $43,252,003,274,489,856,000$ different cubes (still a rather staggering number!).
+\\
+\\
+Finally, let me mention that there is a long history of research into finding tighter and tighter bounds on the 3x3x3 God's number (usually, proofs of upper bounds were obtained and refined by decomposing the Rubik's moves group into subgroups, then finding upper bounds of distance between two elements of a given subgroup, and finally summing these upper bounds over the different subgroups).
+See e.g.
+\cite{RubiksChicago} for an exposition of this type of approach.
+In 2007, an upper bound of 34 in the quarter turn metric was published (\cite{RubiksRadu}) but the question got properly settled in \cite{RubiksGodNumber} by Google's researchers which used clever symmetry arguments to reduce the search space to the maximum, followed by Google's vast computing power capabilities to brute force solve all of the configurations they were left with.
+They concluded that the 3x3x3 Rubik's God number is in fact a mere 26.
 
-\cite{RubiksGodNumber}
+\subsubsection{2x2x2 Rubik's -- search space, solvability}
+
+For the 2x2x2 cube, there are eight possible ways of choosing the physical position of the eight corner cubies (and since there are known algorithms, such as R’ U R’ D2 R U’ R’ D2 R2, to swap exactly two and only two adjacent corners we know all can indeed be obtained).
+Each corner cubie can be oriented in 3 different ways, so we get a total of $8!
+.
+3^{*}$ configurations.
+However, due to the COP parity discussed in the previous section, we conclude that only one third of these are attainable (namely those for which the COP is equal to 0 \% 3).
+Overall, this gives $8!
+* 3^{7}$ = 88,179,840 possible configurations (probably a larger number than most people would expect).
 \\
-\cite{RubiksChicago}
+The insight and knowledge of this section has been useful for me to implement \textit{perfect scrambling} for the 2x2x2 RC.
+Indeed, the way I generate perfectly shuffled RCs is by randomly placing the 8 corners, then randomly choosing the orientation of the first 7, and finally fixing the $8^{th}$ corner's orientationto make sure COP is equal to 0 \% 3.
 \\
-\cite{RubiksRadu}
-\\
-
-\subsubsection{2x2x2 Rubik's -- search space, solvability \& God's number}
-
-Up to rotations of the full RC, there are $\frac{8!
-* 3^{7}}{24}$ = 3,674,160 possible combinations.
-Indeed, there are 8!
-possible ways of choosing the position of the eight corner cubies (since there are known algorithms, such as R’ U R’ D2 R U’ R’ D2 R2, to swap exactly two and only two adjacent corners), each of which can be oriented in 3 different ways (not 6, since as discussed earlier, the clock-wise order of a given corner cubie can never be changed).
-That gives us $3^8$ permutations of the corners orientation.
-However, the \textit{corner orientation parity} (\textbf{COP}) (see \cite{Schoenert} for details) is invariant (\% 3) by any legal move of the RC.
-Since rotating by $120^{o}$ a given corner cubie increases the COP by 1, we conclude that only one third of all these $3^8$ permutations are attainable (namely those for which the COP is equal to 0 \% 3).
-Finally, the denominator of 24 is due to the equivalence by whole cube rotation as discussed in the previous section.
-\\
-The insight and knowledge of this section will be useful for me to implement \textit{perfect scrambling} for the 2x2x2 RC.
-Indeed, the way I generate perfectly shuffled RCs is by randomly placing the 8 corners, then randomly choosing the orientation of the first 7, and finally fixing the $8^{th}$ corner's orientationto make sure COP is equal to 0 \% 3.
+Some quick googling suggests the 2x2x2 Rubik's God number to be 11 in the quarter turn metric, but I could not find any clear and definitive paper on this.
 
@@
 
-From handcrafted (naive, kociemba) to supervised (DL) to unsupervised (DRL \& DQL) .... but sometimes subtle knowledge injected in e.g. representation or simplifications.
+In the upcoming implementation (\ref{sec:Implementation}) and results (\ref{sec:ResultsSP}, \ref{sec:ResRubiks}) sections, I will be discussing various solvers.
+Some of them, such as my implementation of a naive \textbf{SP} solver, or the Kociemba open source implementation for the \textbf{RC}, make heavy use of specific knowledge about these two puzzles.
+Then there are the A$^{*}$ with handcrafted heuristics (e.g.
+Manhattan heuristic for the \textbf{SP}) algorithms which .
+
+, which I have taken from third parties (kociemba) to supervised (DL) to unsupervised (DRL \& DQL) .... but sometimes subtle knowledge injected in e.g. representation or simplifications.

### 2022-09-06 latex/Chapters/DRLSearch.tex
update

@@
 \begin{pseudocode}
-############################################################
+#################################################################
 def MCTS(initial_node, time_out, q_v_network):
@@
         check_time_out()
-        leaf = add_new_leaf(tree, q_v_network)
+        leaf = construct_path_to_new_leaf(tree, q_v_network)
         if leaf.is_goal():
             return BFS(tree) # path to leaf can usually be improved by BFS
-############################################################
-def add_new_leaf(tree, q_v_network):
+#################################################################
+def construct_path_to_new_leaf(tree, q_v_network):
     leaf = tree.initial_node
     while leaf.expanded:
-        pass
-    leaf.expand() # add children of leaf to tree
-    leaf.update_penalties()
+        actions_proba, value = q_v_network(leaf)
+        historical_actions_count = leaf.actions_taken
+        leaf = leaf.choose_child(actions_proba,
+                                 value,
+                                 historical_actions_count)
+    # add children of leaf to tree
+    leaf.expand()
+    # penalize path taken to favour exploration
+    leaf.increment_historical_actions_count()
     return leaf
-############################################################
+#################################################################
 \end{pseudocode}
@@
 
-I have implemented the exact \textbf{MCTS} procedure described in \cite{https://doi.org/10.48550/arxiv.1805.07470} and outlined in pseudo code above in a class called MonteCarloTreeSearch (see section \ref{sec:codesolvers} for details).
+I have implemented the exact \textbf{MCTS} procedure described in \cite{https://doi.org/10.48550/arxiv.1805.07470} and outlined in the above pseudo code in  a class called MonteCarloTreeSearch (see section \ref{sec:codesolvers} for details).

### 2022-09-06 latex/Chapters/Puzzles.tex
update

@@
 %-----------------------------------
-\section{Solving puzzles \textit{without human knowledge}}
+\section{Solving puzzles without human knowledge}
+\label{sec:WHK}
 
 In the upcoming implementation (\ref{sec:Implementation}) and results (\ref{sec:ResultsSP}, \ref{sec:ResRubiks}) sections, I will be discussing various solvers.
-Some of them, such as my implementation of a naive \textbf{SP} solver, or the Kociemba open source implementation for the \textbf{RC}, make heavy use of specific knowledge about these two puzzles.
-Then there are the A$^{*}$ with handcrafted heuristics (e.g.
-Manhattan heuristic for the \textbf{SP}) algorithms which .
+I would put them broadly in four different categories, by decreasing level of human knowledge that has been assumed and handcrafted in them.
+Notice that this human-knowledge-required scale is obviously somewhat arbitrary, but good to keep in mind in later (and in particular results) sections.
 
-, which I have taken from third parties (kociemba) to supervised (DL) to unsupervised (DRL \& DQL) .... but sometimes subtle knowledge injected in e.g. representation or simplifications.
+\paragraph{Domain-specific solvers}
+At 4 on the human-knowledge-required scale, I would put my NaiveSolver (handcrafted to solve the \textbf{SP}) as well as the hkociemba (2x2x2) and kociemba (3x3x3) open source implementations which I have integrated in my KociembaSolver.
+They make heavy use of specific knowledge about the two puzzles, not only about their mechanics, but also about how to actually (provably) solve them.
+
+\paragraph{Domain-specific heuristics}
+At 3 on the human-knowledge-required scale, I put A$^{*}$-with-handcrafted-heuristics (e.g.
+Manhattan heuristic for the \textbf{SP}) algorithms.
+Constructing the heuristics usually require some knowledge about the puzzle's mechanics, but not necessarily about how to solve them efficiently (though of course knowing the latter might help come up with more efficient heuristics).
+
+\paragraph{Supervised learning}
+At 2 on the human-knowledge-required scale, I put my DeepLearner solver (also using A$^{*}$ but whose heuristic is learnt via \textbf{DL}).
+This solver does not require any knowledge of the structure or mechanics of the puzzle, nor of how to solve them, but just need a teacher algorithm to learn/extrapolate from.
+
+\paragraph{Blind search\& unsupervised learning}
+Finally at 1 on the human-knowledge-required scale, I put the blind solvers such as \textbf{BFS} and \textbf{DFS}.
+Also in that group I put A$^{*}$ with heuristics learnt using my DeepReinforcementLearner and DeepQLearner, and my MonteCarloSearchTree which all know (almost) nothing about the puzzles and do totally work from generic interfaces alone (states, transitions, etc \dots) and are operating in an unsupervised way.
+\\
+\\
+One important comment I would make to end this section is that it is easy to inject, without realising or acknowledging it, specific knowledge into the solvers, or in the tensor representation of the puzzles that the solvers take as input.
+Pretty much all the papers about \textbf{RC} that I have surveyed chose a more compact representation for the cube than I have.
+They remove the center cubies of the 3x3x3 cube since, they argue, those never move via face-turns.
+Some of them go much further and make use of various symmetry arguments to reduce the search space and reduce the \textbf{RC} to having one goal state (instead of 24).
+One of the main papers I have inspired myself for this project, \cite{DBLP:journals/corr/abs-1805-07470}, reduces the representation of the Rubik's considerably more, by making use of arguments about the physical structure of the edge and corner cubies: knowing the color of one of their face allows you do deduct the other faces, and therefore, they only need to track one face.
+I of course understand that they do that for the sake of reducing the size of the tensors representing the cube, and in order to remove redundant information.
+This is still \textit{cheating} a bit in my opinion, especially if you are claiming to solve these puzzles \textit{without-human-knowledge}, as we can only operate these dimensionality reduction tricks if we have external world knowledge about the physical structure of the puzzles.
+In this project, I have chosen as pure an approach as I could, and not only did not reduce the dimensionality of the tensor representation of the puzzles, but also, specifically for the Rubik's cube, I did not make use of invariance by spacial rotation of the full cube.

### 2022-09-06 latex/MScAIFrancoisBerrier.tex
update

@@
 \newcommand{\afblue}{\color{airforceblue}}
+\newcommand{\teal}{\color{teal}}
 
@@
 I might throw a few random (but hopefully interesting) observations in there and give some references for the keen reader.
-I will then give an overview in chapter 4 of the code base I have developed - and put in the open on my github page (\cite{FB}) - to complete this project before detailing a few examples in chapter 5.
+I will then give an overview in chapter 4 of the code base I have developed - and put in the open on my github page (\cite{FB}) - to complete this project.
 Finally, in chapters 6 and 7, I will present all my various results on respectively the sliding puzzle and the Rubik's cube before offering some conclusion in chapter 8.
+\\
+A first appendix details how to use the code base, from constructing puzzles, to learning and solving them using all the different methods I have implemented for this project.
+A second appendix describes a limitation I bumped into while using the kociemba library (an open source 3x3x3 \textbf{RC} solver), and how I circumvented it.
 

### 2022-09-06 rubiks/scripts/pseudocode/mcts.py

@@ -1,20 +1,25 @@
-##########################################################################################
+#################################################################
 def MCTS(initial_node, time_out, q_v_network):
     """ pseudo code for Monte Carlo Tree Search """
     initialize_time_out(time_out)
     tree = Tree(initial_node)
     while True:
         check_time_out()
-        leaf = add_new_leaf(tree, q_v_network)
+        leaf = construct_path_to_new_leaf(tree, q_v_network)
         if leaf.is_goal():
             return BFS(tree) # path to leaf can usually be improved by BFS
-##########################################################################################
-def add_new_leaf(tree, q_v_network):
+#################################################################
+def construct_path_to_new_leaf(tree, q_v_network):
     leaf = tree.initial_node
     while leaf.expanded:
-        pass
-    leaf.expand() # add children of leaf to tree
-    leaf.update_penalties()
+        actions_proba, value = q_v_network(leaf)
+        historical_actions_count = leaf.actions_taken
+        leaf = leaf.choose_child(actions_proba,
+                                 value,
+                                 historical_actions_count)
+    # add children of leaf to tree
+    leaf.expand()
+    # penalize path taken to favour exploration
+    leaf.increment_historical_actions_count()
     return leaf
-##########################################################################################
-
+#################################################################

### 2022-09-06 latex/Chapters/DRLSearch.tex
update

@@
 \paragraph{}{\textbf{pseudo code -- BFS \& DFS}}
+\label{sec:TheoryBFSDFS}
 \begin{pseudocode}
@@
 \paragraph{}{\textbf{pseudo code -- A$^{*}$}}
+\label{sec:TheoryAStar}
 \begin{pseudocode}
@@
 \section{DL and DLR Heuristics}
-
+\label{sec:TheoryDLDRL}
 So what is the link between all of these: A$^{*}$'s heuristics, \textbf{DL}, \textbf{RL}'s value iteration?
@@
 \section{Deep Q-Learning}
-
+\label{sec:TheoryDQL}
 Another fundamental concept of \textbf{RL} is that of the Q-function $Q(s, a)$, the maximum expected reward the agent can obtain from state $s$, taking action $a$ and acting optimally afterwards.
@@
     while leaf.expanded:
-        actions_proba, value = q_v_network(leaf)
-        historical_actions_count = leaf.actions_taken
-        leaf = leaf.choose_child(actions_proba,
-                                 value,
-                                 historical_actions_count)
+        actions_probas, actions_values = q_v_network(leaf)
+        """ choose next node based on joint actions probas & values
+        as well as history of actions taken
+        """
+        leaf = leaf.choose_child(actions_probas,
+                                 actions_values)
     # add children of leaf to tree

### 2022-09-06 latex/Chapters/Implementation.tex
update

--- 2022-07-31 latex/Chapters/Code.tex
@@
 
-\chapter{Code} % Main chapter title
+\chapter{Implementation} % Main chapter title
 
-\label{Chapter3} % Change X to a consecutive number; for referencing this chapter elsewhere, use \ref{ChapterX}
+\label{sec:Implementation} % Change X to a consecutive number; for referencing this chapter elsewhere, use \ref{ChapterX}
 
@@
 We can quickly see that this heuristic is admissible.
-Indeeed, each move in the SP moves one (and one only) tile by one position.
+Indeeed, each move displaces one (and one only) tile by one position.
 In other puzzles, it is generally the case that a move will affect the position of many tiles (e.g.
 RC).
-Therefore, if tiles could somehow freely move on top of one-another (that is, we remove the constraint that there can be at most one tile per compartment, the number of moves necessary to solve the SP would exactly be the Manhattan distance.
+Therefore, if tiles could somehow freely move on top of one-another (that is, we remove the constraint that there can be at most one tile per compartment, the number of moves necessary to solve the \textbf{SP} would exactly be the Manhattan distance.
 \\
@@
 \\
-I have also implemented an improvement to the Manhattan distance, which I shall call Manhattan++ in here (and in code logs and graphs) and can be activated by simply passing \textit{plus=True} to the Manhattan Heuristic (see example \ref{ASSS} as well as a thorougher performance comparison on the (n=2, m=5) SP in \ref{MHComp}).
+I have also implemented an improvement to the Manhattan distance, which I shall call Manhattan++ in here (and in code logs and graphs) and can be activated by simply passing \textit{plus=True} to the Manhattan Heuristic (see example \ref{ASSS} as well as a thorougher performance comparison on the 2x5 \textbf{SP} in \ref{MHComp}).
 It is based on the concept of linear constraints that I read about in lecture notes from Washington University (\cite{SlidingPuzzleLectureNotes}).
 The idea is simply that when tiles are already placed on their target row or column, but not in the expected order, we will need to get the tiles out of the way of one another to get to the goal state, and this is not accounted for by the Manhattan distance.
-For instance, in the following (n=2, m=3) SP, the Manhattan distance does not account for the fact that, at best, tile \red 3 \black needs to get out of the way for tiles \blue 1 \black and \blue 2 \black to move, and then needs to get back in its row, adding a cost of 2 to the Manhattan distance.
+For instance, in the following 2x3 puzzle, the Manhattan distance does not account for the fact that, at best, tile \red 3 \black needs to get out of the way for tiles \blue 1 \black and \blue 2 \black to move, and then needs to get back in its row, adding a cost of 2 to the Manhattan distance.
 
@@
 
-Two important things to notice are that linear constraints across rows and columns (which I might generically refer to as \textit{lines} in the following) can to be added without breaking admissibility (hence giving more pessimistic, or accurate, cost estimates than simple Manhattan).
+Two important things to notice are that linear constraints across rows and columns (which I might generically refer to as \textit{lines} in the following) can be added without breaking admissibility (hence giving more pessimistic, or accurate, cost estimates than simple Manhattan).
 This is because if a tile is involved in two linear constraints, it will need to get out of the way both horizontally and vertically.
@@
 Indeed if, say, tile 3 got somehow out of the way at the back of the SP (imagine just another dimension there where we can move tiles) and tile 2 got out of the way by moving down to let tile 1 pass across, we could be done by simply adding 4 to the Manhattan distance (2 to move tile 3 out and back, and 2 to move tile 2 out and back).
-The correct way to compute the penalty cost for linear constraints is therefore to do it recursively, taking the minimum additional cost of moving either left-most or right-most tile of the line under consideration out of the way (that additional cost to move these left-most or right-most tile is 2 if not at their expected order in the line, 0 otherwise) plus the penalty of reordering the rest of the line.
+The correct way to compute the penalty cost for linear constraints is therefore to do it recursively, taking the minimum additional cost of moving either the left-most or the right-most tile of the line under consideration out of the way (that additional cost to move these left-most or right-most tile is 2 if not at their expected order in the line, 0 otherwise) plus the penalty of reordering the rest of the line.
 \\
@@
 For memory efficiency, I also only saved penalties which are non-zero.
-The very first time any call to Manhattan++ is made, for a given dimension (n, m), the appropriate data base is computed and populated.
+The very first time any call to Manhattan++ is made, for a given dimension nxm, the appropriate linear constraint penalties are computed and populated in a database.
 \\
 \\
-Notice that for an (n, m) SP, there are n rows, each of which can have $\frac{(n * m)!}{(n * m - m)!}$ different ordering of tiles and m columns which can have n columns, each can have $\frac{(n * m)!}{(n * m - n)!}$ different ordering of tiles.
+Notice that for an nxm \textbf{SP}, there are n rows, each of which can have $\frac{(n * m)!}{(n * m - m)!}$ different ordering of tiles and m columns which can each have $\frac{(n * m)!}{(n * m - n)!}$ different ordering of tiles.
 This means the pre-computations and data-base sizes for the Manhattan++ heuristic are actually manageable, as it grows much slower than the number of possible puzzles.
-The maximum number of penalties to compute for $(n, m) \leq (5, 5)$ are:
+The maximum number of penalties to compute for $nxm \leq 5x5$ are:
 
@@
 \item \textbf{PerfectHeuristic}: this reads from a data base the optimal costs, pre-computed by the PerfectLearner (see below \ref{PLcode})
-\item \textbf{DeepLearningHeuristic}: this uses a network which has been trained using DRL by the DeepReinforcementLearner (see below \ref{DRLcode})
+\item \textbf{DeepLearningHeuristic}: this uses a network which has been trained using \textbf{DL} (DeepLearner) \textbf{DRL} (DeepReinforcementLearner) or \textbf{DQL} (DeepQLearner).
+See \ref{DRLcode} for a discussion of these three learners.
+\item \textbf{DeepQLearningHeuristic}: this uses a network which has been trained using \textbf{DQL} by the DeepQLearner (see below \ref{DQLcode}).
+The DeepQLearningHeuristic follows the same interface as the other heuristics, and therefore can be used by A$^{*}$ for instance.
+It also has an additional method \textit{optimal\_actions} that returns the learnt probability distribution over actions for a given puzzle, for use in search algorithms (e.g.
+\textbf{MCTS}) which make use of this to inform their search.
 \end{itemize}
@@
 There are some params as well to add drop out, and to indicate whether or not the inputs are one hot encoding (in which case the first layer is automatically adjusted in size, using information from the puzzle dimension).
+There is also importantly a \textit{joint\_policy} parameter to indicate whether we want the output to be 1 dimensional (e.g.
+for value function learning) or $a+1$-dimensional, where $a$ is the size of the actions space (for joint policy-value learning).
 \item \textbf{Convolutional}: similar wrapper to FullyConnected, but with the ability to add some parallel convolutional layers to complement fully connected layers.
@@
 \label{DRLcode}
+\label{DQLcode}
 This module implements learners, which learn something from a puzzle, store what they learnt, and can display interesting things about what they learnt.
@@
 Concrete derived implementation define what and how they learn, and what interesting they can display about this learning process.
-Currently the two implemented learners are:
+The four implemented learners are:
 \item \textbf{PerfectLearner}: It instantiates an optimal solver ($A^{*}$ with a configurable heuristic - but will only accept heuristic that advertise themselves as optimal.
@@
 
-\item \textbf{DeepLearner} tbd
-\item \textbf{DeepReinforcementLearner}: It instantiates a DeepLearning (network), and trains it using DRL.
-It then saves the trained network, which can then be used in the DeepLearningHeuristic we have seen earlier in section \ref{HSS}.
-The DeepReinforcementLearner runs for a number of epochs (or less if, based on other parameters discussed below, it is deemed to have converged).
+\item \textbf{DeepLearner} DeepLearner instanciates a DeepLarning (neural network), and trains it on training data by minimizing the mean square error between the cost-to-go from the target solver versus the neural network.
+The main parameters driving this learner are:
+\begin{itemize}
+\item \textit{training\_data\_from\_data\_base} which controls whether the training samples are generated on the fly via another solver (defaults to A$^{*}[Manhattan++]$ for \textbf{SP} and A$^{*}[Kociemba]$ for \textbf{RC}) or read from a data base (similarly constructed from other solvers, but has the advantage of being done offline)
+\item \textit{nb\_epochs} and \textit{threshold} which control exit criteria for the network training.
+\item \textit{nb\_sequences}, \textit{nb\_shuffles\_min}, \textit{nb\_shuffles\_max}, \textit{training\_data\_freq} and \textit{training\_data\_every\_epoch} which control how often we regenerate new training data, how many sequences of puzzles it contais, how scrambled the puzzles should be.
+\item \textit{learning\_rate}, \textit{scheduler}, \textit{optimizer} and assorted parameters which control the optimisation at each backward propagation. 
+\end{itemize}
+Once it has completed (or on a regular basis), DeepLearner saves the trained network, which can then be used with DeepLearningHeuristic to guide A$^{*}$.
+\item \textbf{DeepReinforcementLearner}: It instantiates a DeepLearning (network), and trains it using \textbf{DRL} essentially following the pseudo-code from \ref{sec:TheoryDLDRL}.
+Unlike DeepLearner, the targets are here generated via value iteration rather than via another solver (teacher).
+It has very similar parameters to the DeepLearner, in particular some parameters to control exit criteria (same as DeepLearner's plus a few that help it exit early when the value function's range has not been increasing anymore for a while), some parameters to control and configure the optimiser, some to control the training puzzles (how many, how often, how scrambled).
+
+\item \textbf{DeepQLearner}: The implementation of the DeepQLearner is actually only a few lines of code, for it inherits pretty much all of its behaviour from the DeepReinforcementLearner.
+The only overwritten functions are:
+
+\begin{itemize}
+\item \textit{get\_loss\_function} which constructs the loss function to be used by the network's training.
+Instead of returning a simple torch.nn.MSELoss, it returns a function which computes the average of torch.nn.MSELoss on the value-function and torch.nn.CrossEntropyLoss on the actions.
+\item \textit{\_\_construct\_target\_\_} which performs the the Q-V-iteration to update the targets, as described in pseudo code form in \ref{sec:TheoryDQL}.
+This function updates not only the value function of a node given its children nodes, but also updates the optimal action vector (0 for non optimal actions, 1 for the optimal action leading to the child node of highest value function).
+\end{itemize}
 
@@
 This module implements solvers, which solve puzzles.
-The base class Solver is a Factory of solvers, and in addition to being able to instantiating the following types of solvers, can run different solvers through a similal sequences of random puzzles (for various increasing degrees of difficulty (shuffling), and/or perfectly shuffled ones) and display a comparison of how they perform in a number of metrics.
+The base class Solver is a Factory of solvers, and in addition to being able to instantiating the following types of solvers, can run different solvers through a similal sequences of random puzzles (for various increasing degrees of difficulty (shuffling), and/or perfectly shuffled ones) and display a comparison of how they perform in a number of metrics (chosable from percentage of puzzles solved before time out, percentage of optimal puzzle (if optimality can be ascertained), mean-median-maximum run time, mean-median-maximum expanded nodes (for solvers where it makes sense), mean-median-max cost, and an optimality score which compares the total sum of cost over all puzzles versus a benchmark solver).
+Ideally the benchmark solver to compute the optimality score should be one which is provably optimal, but in the case of the \textbf{RC} that was not possible so I used A$^{*}$ with KociembaHeuristic as my benchmark.
+For the \textbf{SP} I used the optimal A$^{*}$ with Manhattan++ heuristic.
 
 \begin{itemize}
-\item \textbf{DFSSolver} TBD
-\item \textbf{BFSSolver} TBD
-\item \textbf{AStarSolver} TBD
-\item \textbf{NaiveSlidingSolver} TBD
+\item \textbf{DFSSolver} This solver is literally a few lines of code, extremely close to the pseudo code detailed in \ref{sec:TheoryBFSDFS}
+\item \textbf{BFSSolver} Ditto, see \ref{sec:TheoryBFSDFS}
+\item \textbf{AStarSolver} Ditto, see \ref{sec:TheoryAStar}
+\item \textbf{NaiveSlidingSolver} The NaiveSolver is conceptually simple, but was far from trivial to implement and debug.
+Its functioning is detailed in the appendix section \ref{sec:Examples}.
+Basically, it does what a beginner player would do (certainly what I do when I play the sliding puzzle), solving the top row first, then the left column, hence reducing the puzzle to a smaller dimensional one, and keep iterating until done.
+\item \textbf{KociembaSolver} This solver is a wrapper around two libraries.
+The first one, hkociemba (see \cite{HKociemba}) is a 2x2x2 implementation written by Kociemba himself, the second one is a Python pip-installable which solves the 3x3x3 Rubik's.
+My  wrapper simply translates my cube representation in that accepted by these 2 solvers, and their result string into moves usable by my code.
+I also do some additional massaging for the 3x3x3 case, due to a shortcoming/bug in the 3x3x3 implementation, as I have detailed in appendix \ref{sec:Kociemba}.
 \item \textbf{MonteCarloSearchTreeSolver} TBI I want to implement this in August when working on the Rubiks' ...

### 2022-09-06 latex/Chapters/Puzzles.tex
update

@@
 \\
-A favourite past time among computer scientists around the glove is therefore to search for more refined lower and upper bounds for $\mathcal{G}(n, m)$, for ever increasing values of n and m.
+A favourite past time among computer scientists around the globe is therefore to search for more refined lower and upper bounds for $\mathcal{G}(n, m)$, for ever increasing values of n and m.
 For moderate n and m, we can actually solve optimally all possible configurations of the puzzle and compute exactly $\mathcal{G}(n, m)$  (using for instance $A^{*}$ and an admissible heuristic (recall \ref{GSH}, and we shall see modest examples of that in the results section later).
@@
  It is quite clear from that picture above that if we can bring the 4x4x4 cube to such a configuration on all faces (which is not super hard), and we know how to solve a 3x3x3, we are essentially done.
-(well, almost, since the additional degrees of freedom coming from the higher dimension of the 4x4x4 mean that certain arrangements of the colors which are possible on the 4x4x4 are not on the 3x3x3 (for instance we can never have the center red and white on opposite faces of a 3x3x3, but we easily can on a 4x4x4!
-Fortunately that is nothing that a couple of easily learnable algorithms of a few moves cannot fix).
+(well, almost, since the additional degrees of freedom coming from the higher dimension of the 4x4x4 mean that certain arrangements of the colors which are possible on the 4x4x4 are not on the 3x3x3.
+For instance we can never have the center red and white on opposite faces of a 3x3x3, but we easily can on a 4x4x4!
+Fortunately that is nothing that a couple of easily learnable algorithms cannot fix).
 
@@
 \begin{itemize}
-\item $N$: the number of possible arrangements of the \textit{cubies} (individual small cubes that constitute the Rubik's) in space, taking into account physical constraints.
+\item $N$: the number of possible arrangements of the \textit{cubies} (individual small cubes that constitute the Rubik's) in space, taking into account physical constraints and some invariants.
 \item $D$: the number of possible configurations which are equivalent via full cube rotation in space.
@@
 .
-2^{12}$ arrangements are physically attainable, leaving us with a grand total of $43,252,003,274,489,856,000$ different cubes (still a rather staggering number!).
+2^{12}$ arrangements are physically attainable, leaving us with a grand total of $43,252,003,274,489,856,000$ different cubes (still a rather staggeringly large number!).
 \\
@@
 .
-3^{*}$ configurations.
-However, due to the COP parity discussed in the previous section, we conclude that only one third of these are attainable (namely those for which the COP is equal to 0 \% 3).
+3^{8}$ configurations.
+However, due to the \textbf{COP} parity discussed in the previous section, we conclude that only one third of these are attainable (namely those for which the \textbf{COP} is equal to 0 \% 3).
 Overall, this gives $8!
-* 3^{7}$ = 88,179,840 possible configurations (probably a larger number than most people would expect).
+* 3^{7}$ = 88,179,840 possible configurations (probably a larger number than most people would expect for a mere 2x2x2 cube).
 \\
-The insight and knowledge of this section has been useful for me to implement \textit{perfect scrambling} for the 2x2x2 RC.
-Indeed, the way I generate perfectly shuffled RCs is by randomly placing the 8 corners, then randomly choosing the orientation of the first 7, and finally fixing the $8^{th}$ corner's orientationto make sure COP is equal to 0 \% 3.
+The insight and knowledge of this section has been useful for me to implement \textit{perfect scrambling} for the 2x2x2 \textbf{RC}.
+Indeed, the way I generate perfectly shuffled cubes is by randomly placing the 8 corners, then randomly choosing the orientation of the first 7, and finally fixing the $8^{th}$ corner's orientation to make sure COP is equal to 0 \% 3.
 \\

### 2022-09-06 rubiks/learners/deeplearner.py

@@ -261,11 +261,9 @@ class DeepLearner(Learner):
             if puzzle.is_goal():
                 target = 0
             else:
-                """ We solve using A* and admissible """
                 config = self.get_config()
-                config[Solver.solver_type] = Solver.astar
-                config[Heuristic.heuristic_type] = Heuristic.manhattan
-                config['plus'] = True
+                """ @todo Francois: probably should make that configurable """
+                config.update(puzzle.optimal_solver_config())
                 solution = Solver.factory(**config).solve(puzzle)
                 if solution.success:
                     target = solution.cost

### 2022-09-06 rubiks/learners/deepqlearner.py

@@ -11,6 +11,7 @@ from rubiks.deeplearning.deeplearning import DeepLearning
 dql_cross_entropy_loss = CrossEntropyLoss()
 dql_mse_loss = MSELoss()
 
+
 def dql_loss(output, target):
     return dql_mse_loss(output[:, 0], target[:, 0]) + \
            dql_cross_entropy_loss(output[:, 1:], target[:, 1].long())

### 2022-09-06 latex/Chapters/DRLSearch.tex
update

@@
 The Monte Carlo Tree Search (\textbf{MCTS}) is such an algorithm.
-In short, it generates at each step (and possibly in a distributed manner) paths to a \textit{leaf} on the frontier of unexpanded nodes, expands that node and checks if it is a goal state (and keeps going until the goal is found).
+In short, it generates at each step (and possibly in a distributed or multithreaded manner) paths to a \textit{leaf} on the frontier of unexpanded nodes, expands that node and checks if it is a goal state (and keeps going until the goal is found).
 The paths followed from initial state to leaves are decided via a combination of the probability distribution over actions (following the most promising actions) as well as the value function, according to a trade-off which can be tuned via a hyper-parameter.

### 2022-09-06 latex/Chapters/Implementation.tex
update

@@
 Its functioning is detailed in the appendix section \ref{sec:Examples}.
-Basically, it does what a beginner player would do (certainly what I do when I play the sliding puzzle), solving the top row first, then the left column, hence reducing the puzzle to a smaller dimensional one, and keep iterating until done.
+Basically, it does what a beginner player would do, certainly what I do when I play the sliding puzzle: solving the top row first, then the left column, hence reducing the puzzle to a smaller dimensional one, and keep iterating until done.
 \item \textbf{KociembaSolver} This solver is a wrapper around two libraries.
-The first one, hkociemba (see \cite{HKociemba}) is a 2x2x2 implementation written by Kociemba himself, the second one is a Python pip-installable which solves the 3x3x3 Rubik's.
+The first one, hkociemba (see \cite{HKociemba}) is a 2x2x2 implementation written by Kociemba himself, the second one is a Python pip-installable library which solves the 3x3x3 Rubik's.
 My  wrapper simply translates my cube representation in that accepted by these 2 solvers, and their result string into moves usable by my code.
-I also do some additional massaging for the 3x3x3 case, due to a shortcoming/bug in the 3x3x3 implementation, as I have detailed in appendix \ref{sec:Kociemba}.
-\item \textbf{MonteCarloSearchTreeSolver} TBI I want to implement this in August when working on the Rubiks' ...
-following the Rubik's paper fromMcAleer \& Agostilenni et al \cite{https://doi.org/10.48550/arxiv.1805.07470}
+I also do some additional massaging for the 3x3x3 case, due to a shortcoming/bug in the 3x3x3 Python library, as I have detailed in appendix \ref{sec:Kociemba}.
+\item \textbf{MonteCarloSearchTreeSolver} I have implemented the algorithm described in the paper from McAleer \& Agostilenni et al \cite{https://doi.org/10.48550/arxiv.1805.07470}.
+My implementation is single-threaded however (in part due to my using Python for this project, which would have made using multiprocessing not efficient at all for their algorithm).
+A consequence of that is that I dropped the $\nu$ parameter from their implementation, since it only introduces a penalty to discourage parallel threads from generating the same paths from initial node to leaves (in single threaded mode, this parameter is inconsequential since the generation of a path decrements it, only to reincrement it once the leaf is generated).
+\\
+This solver basically generates (pseudo-random) paths from initial node to a new leaf at each iteration, until we obtain a goal leaf.
+I use the word pseudo-random since the paths are actually completely deterministic.
+They indeed result from a trade-off between the probability of actions (penalised over time by discounting actions which have been taken already) and the value function, but there is no random draws performed anywhere in there.
+\\ As is described in McAleer \& Agostilenni's paper, I have implemented a final pruning via BFS of the tree resulting from the above \textit{Monte-Carlo} process consisting in generating random paths until a goal state is reached.
+Indeed, it is clear that the paths generated have very little chance of being optimal since there is no correction mechanism to avoid long and unecessary loops.
+I have however made this tree trimming by BFS optional via a \textit{trim\_tree} parameter as I wanted to study its effect on solutions' optimality and run time (see results section \ref{sec:MTCSHyperParamsEffect}).
+\\
+Similarly to their implementation, the trade off between optimal actions and value function is driven by a hyper-parameter \textit{c}, as I wanted to study its effect as well, and confirm my intuitions as to how it would affect solutions' quality and run time (again see results section \ref{sec:MTCSHyperParamsEffect}).
+
 \end{itemize}
 
-\section{rubiks.scripts}
+\section{data}
 
-\begin{figure}[H]
-\centering
-\includegraphics[scale=0.25]{./Figures/codebasescripts}
-%\decoRule
-\caption[Codebase]{rubiks.scripts}
-\label{fig:Codebasescripts}
-\end{figure}
-
-Finally it is worth noting that the code will save on disk a lot of data (e.g.
-the learners will save what they have learnt, e.g.
-a Pytorch network or a data base of optimal costs, the performance comparison will run solvers versus very many configurations of puzzles and save the results for later being able to display) etc...
-The base of the tree to save all this data can be chosen by setting up the "RUBIKSDATA" environment variable.
-If not, it will go somewhere in you HOME :)
+Finally it is worth noting that the code can/will save on disk a lot of data, including:
+\begin{itemize}
+\item Manhattan++ generates databases of linear constraint penalties (see \ref{HSS})
+\item TrainingData: (see e.g. \ref{DLSS}) generates databases of puzzles and their associated costs for later training of a DeepLearner.
+\item Learners: models generated by the DeepLearner, DeepReinforcementLearner, DeepQLearner as well as some convergence data useful to understand how the networks learnt.
+Similarly PerfectLearner saves its results for later reuse (in e.g.
+PerfectHeuristic) or display.
+\item Solver: Sequences of scrambled puzzles which we use to make the performance tests (comparing different solvers) fair.
+\item Solver.performance\_test: Results of the performance tests (all the metrics discussed earlier in \ref{sec:codesolvers}, for each solver and each difficulty level)
+\item etc \dots
+\end{itemize}
+The data is saved according to a file system tree of the form:
+$$root/data\_type/puzzle\_type/dimension/file\_name.pkl$$
+where the root is driven by the "RUBIKSDATA" environment variable or, if not set, will go somewhere in your HOME.

### 2022-09-06 rubiks/scripts/sp25learner.py

@@ -42,7 +42,7 @@ if '__main__' == __name__:
         else DeepLearner.default_plot_metrics
     """ If it's a perfect learner config here """
     time_out = 3600
-    nb_cpus = 12
+    nb_cpus = 4
     cpu_multiplier = 100
     max_puzzles = nb_cpus * cpu_multiplier * 10000
     regular_save = nb_cpus * cpu_multiplier * 1

### 2022-09-06 latex/Chapters/Conclusion.tex
update

--- 2022-08-26 latex/Chapters/Conclusion.tex
@@
 
-TBW
+\section{Learning}
+Learnt about puzzles, and also even took the time to learn a couple of sequences of moves that are sufficient to solve \textbf{RC}, 2x2x2 in labout one minute, 3x3x3 in about 2 minutes.
+
+\section{Implementation}
+Implemented a bunch of solvers. Ideally should do in C++ for speed and multithreaded (esp for MCTS) but time was limited so I prioritized breadth over depth.
+
+\section{Deep Reinforcement/Q Learning}
+Quite impressive, without too much effort, how well they perform.
+However, I feel to really get more out of them (higher dimensions in particular) would need careful tuning of the convergence criteria.
+
+\section{Professional issues}
+See \cite{ManyWorlds}
+See \cite{https://doi.org/10.1111/1758-5899.12718} and \cite{Bostrom2014}

### 2022-09-06 latex/Chapters/Puzzles.tex
update

@@
 Some of them go much further and make use of various symmetry arguments to reduce the search space and reduce the \textbf{RC} to having one goal state (instead of 24).
-One of the main papers I have inspired myself for this project, \cite{DBLP:journals/corr/abs-1805-07470}, reduces the representation of the Rubik's considerably more, by making use of arguments about the physical structure of the edge and corner cubies: knowing the color of one of their face allows you do deduct the other faces, and therefore, they only need to track one face.
+One of the main papers I have taken inspiration from for this project, \cite{DBLP:journals/corr/abs-1805-07470}, reduces the representation of the Rubik's considerably more, by making use of arguments about the physical structure of the edge and corner cubies: knowing the color of one of their face allows you do deduct the other faces, and therefore, they only need to track one face.
 I of course understand that they do that for the sake of reducing the size of the tensors representing the cube, and in order to remove redundant information.
-This is still \textit{cheating} a bit in my opinion, especially if you are claiming to solve these puzzles \textit{without-human-knowledge}, as we can only operate these dimensionality reduction tricks if we have external world knowledge about the physical structure of the puzzles.
+This is still \textit{cheating} a bit in my opinion, especially if you are claiming to solve these puzzles \textit{without-human-knowledge}, as we can only perform these dimensionality reduction tricks if we have external world knowledge about the physical structure of the puzzles.
 In this project, I have chosen as pure an approach as I could, and not only did not reduce the dimensionality of the tensor representation of the puzzles, but also, specifically for the Rubik's cube, I did not make use of invariance by spacial rotation of the full cube.

### 2022-09-06 latex/Chapters/SlidingPuzzle.tex
update

@@
-% Chapter Template
-
 \chapter{Results - Sliding Puzzle} % Main chapter title
@@
 I am neither making claims of depth (each solver could surely be tuned and optimized/adapted, possibly differently for each dimension) nor breadth (I have not run all of the solvers versus each dimension).
-Running these experiments takes a lot of time, so I had to be somewhat selective.
-I wanted however to run all of the solvers I have implemented one particular dimension (the intermediary 3x3 seemed appropriate to do so), to answer some of the following interesting questions:
+Running these experiments takes a lot of computing time (even though I have very decent computing resources), so I had to be somewhat selective.
+I wanted however to run all of the solvers I have implemented one one particular dimension (the intermediary 3x3 seemed appropriate to do so) and try to answer as many of the following interesting questions as possible:
 \begin{itemize}
@@
 The only dimensions which have less than 239.5 millions states are shown in table \ref{tab:smallSP}.
-Note I am also only considering n $\leq$ m since dimension (m, n) can always be solved by symmetry from the solutions (n, m):
+Note I am also only considering n $\leq$ m since dimension mxn can always be solved by symmetry from the solutions mxn:
 
@@
 \end{table}
-Notice that among the above dimensions, (n=2, m=5) is the largest and hardest one.
-I had to run its PerfectLearner over a many weeks (whenever I had spare computing capacity and over a full week-end on a c5.18xlarge instance (72 cores) on Amazon EC2).
+Notice that among the above dimensions, 2x5 is the largest and hardest one.
+My current displayed result in table \ref{tab:smallSPGN} is provisional, as I have only solved a bit over 1.1 million puzzles out of the 1.8 million possible, so it is well possible that I haven't yet bumped into the hardest configuration.
+I ran a combined 600 hours, much of which run over 16 cores, and even over a full week-end on a c5.18xlarge instance (72 cores) on Amazon EC2.
 \\
@@
 In this section, I verify empirically that, as expected, the overhead of adding penalty in Manhattan++ for the linear constraint (which have all been precomputed and stored in a database) is more than compensated for by the reduction in nodes expansion.
-I have run my solver script for (n=2, m=5) in performance test mode, for both Manhattan and Manhattan++, with 250 randomly shuffled puzzles with \textit{nb\_shuffles} from 0 to 60 by increment of 5, as well as with \textit{nb\_shuffles = $\inf$}.
+I have run my solver script for 2x5 in performance test mode, for both Manhattan and Manhattan++, with 250 randomly shuffled puzzles with \textit{nb\_shuffles} from 0 to 60 by increment of 5, as well as with \textit{nb\_shuffles = $\inf$}.
 The resulting run time and nodes expansions are as follows:
@@
 
+In this section, which focusses on the 3x3 \textbf{SP}, I detail some results from the PerfectLearner (\ref{sec:S33PL}), from the DeepReinforcementLearner (\ref{sec:S33DRL}), before moving to a discussion of the \textbf{DQL} \textbf{MCTS} solver and the effect of its trade-off hyper-parameter \textit{c} and of its optional \textbf{BFS} tree-trimming on the quality of the solutions (\ref{sec:MTCSHyperParamsEffect}).
+I then run a comprehensive comparison of my 10 different solvers on the 3x3 \textbf{SP} in \ref{ssec:33SPSC}.
+Finally, I end the section by running a few select solvers on one of the two hardest 3x3 puzzle in \ref{ssec:33SPHard}
+
 \subsection{Perfect learner}
+\label{sec:S33PL}
 As discussed in the previous section section \ref{sec:SPLowDimension}, the 3x3 \textbf{SP} is one of the cases I have been able to solve perfectly, since it only has 181,440 possible configurations.
 Its God number is only 31, which definitely makes it manageable.
-However, this is already an intermediary size, large enough that it is worth trying and comparing a few different methods, including deep reinforcement learning.
-To start with, I ran the PerfectLearner with n=m=3, and the results are shown below in figure \ref{fig:33SPPerfectLearning}.
-It is interesting to note that there are only two hardest configurations (cost 31) and 221 configurations of cost 30.
+However, this is already an intermediary size, large enough that it is worth trying and comparing a few different methods, including \textbf{DL}, \textbf{DRL} and \textbf{DQL}.
+To start with, I ran the PerfectLearner and the results are shown below in figure \ref{fig:33SPPerfectLearning}.
+It is interesting to note that there are only two hardest configurations (of cost 31) and 221 configurations of cost 30.
 
-\newgeometry{top=0mm, bottom=0mm, left=0mm, right=0mm}
-\begin{landscape}
-\centering\vspace*{\fill}
 \begin{figure}[H]
 \centering
-\includegraphics[scale=0.7]{./Figures/33SPPerfectLearning.jpeg}
-%\decoRule
-\caption[33SPPerfectLearning]{Perfect Learning 3x3 \textbf{SP}}
+\includegraphics[scale=0.42]{./Figures/33SPPerfectLearning}
+\caption[33SPPerfectLearning]{Perfect Learning of the 3x3 \textbf{SP}}
 \label{fig:33SPPerfectLearning}
 \end{figure}
-\vfill
-\end{landscape}
-\restoregeometry
 
 \subsection{Deep reinforcement learner}
- The DeepReinforcementLearner's learning is shown in figure \ref{fig:33SPDeepReinforcementLearning}:
+\label{sec:S33DRL}
+Next I have run the DeepReinforcementLearner with the following main parameters:
+\begin{itemize}
+\item Update the target network every 1,000 epochs (\textit{update\_target\_network\_frequency=1000}) or when the MSE falls below 10 basis points of the maximum target cost-to-go (\textit{update\_target\_network\_threshold=1e-3})
+\item Run for a maximum of 11,000 epochs (\textit{nb\_epochs=11000}) or when the maximum target so far has not been increasing by more than 1\% in the last 5,500 epochs (\textit{max\_target\_uptick=1e-2} and \textit{max\_target\_not\_increasing\_epochs\_pct=0.5})
+\item At every update of the target network (\textit{training\_data\_every\_epoch=False}), generate 100 sequences of puzzles (\textit{nb\_sequences=100}), each comprised of a 50-scramble puzzle back to goal state, with all intermediary puzzles along the path (\textit{nb\_shuffles=50}).
+\item Use a fully connected network (\textit{network\_type=fully\_connected\_net}) with three hidden layers of size 600, 300, 100 (\textit{layers\_description=(600,300,100)}) and use torch.optim.RMSProp as optimiser (\textit{optimiser=rms\_prop}) with an exponential scheduler (\textit{scheduler=exponential\_scheduler}) with gamma 0.999 (\textit{gamma\_scheduler=0.9999}).
+\end{itemize}
+Below I show some of the DeepReinforcementLearner's quantities tracked over the epochs (see \ref{fig:33SPDeepReinforcementLearning}). I have shown only the following quantities:
+\begin{itemize}
+\item learning rate, which decreases due to the scheduler, but gets reset at each network update.
+\item max target.
+One interesting dynamic during learning is that due to the way targets are constructed using value-iteration update, and starting from an all 0 network, the max target only increases by about 1 at best at every network update.
+That is, the first network only learns how to differentiate goal from non-goal, then the second one learns how to differentiate goals, cost 1 and cost 2+ in some sense.
+This is why I have introduced the first set of parameters discussed just above.
+The first one makes sure we do not waste time initially training the left-hand-side network when training is easy (as consist of just goal versus not goal).
+The second ones make sure that we stop training altogether when the max target does not grow anymore.
+This is because we do not know a-priori what the God number of a given puzzle/dimension is, but let's assume for the sake of the argument that it is 15.
+Then that means that after 15 updates of the target network we might not really get so much benefit by keeping training and value-iterating.
+\item the MSE loss
+\item the MSE loss divided by the max target.
+The exit criterion \textit{update\_target\_network\_threshold} is based on that quantity, since a loss of e.g.
+0.01 means little if the we do not compare it to the possible range that the cost-to-go can take.
+\item The percentage of all the possible puzzles of that dimension \textit{seen} by the DeepReinforcementLearner.
+This is an interesting quantity, as it tells us later on when testing out of sample and comparing with other solvers, how well the solvers are able to generalise.
+In this case (3x3) we can see that the DeepReinforcementLearner quickly has seen a very large percentage of all puzzles.
+That being said, it is good to keep in mind that \textit{seen} does not mean really much here, since the learning is unsupervised.
+This is quite in contrast to the DeepLearner, for which puzzles seen come along with the cost-to-go computed by a \textit{teacher}, usually an optimal or at least efficient solver.
+\end{itemize}
 
-\newgeometry{top=0mm, bottom=0mm, left=0mm, right=0mm}
-\begin{landscape}
-\centering\vspace*{\fill}
 \begin{figure}[H]
 \centering
-\includegraphics[align=c, scale=0.55]{./Figures/33SPDeepReinforcementLearning}
-\caption[33SPDeepReinforcementLearning]{Deep reinforcement learner 3x3 \textbf{SP}}
+\includegraphics[align=c, scale=0.42]{./Figures/33SPDeepReinforcementLearning}
+\caption[33SPDeepReinforcementLearning]{\textbf{DRL} of the 3x3 \textbf{SP}}
 \label{fig:33SPDeepReinforcementLearning}
 \end{figure}
-\vfill
-\end{landscape}
-\restoregeometry
+
+Notice that the DeepLearner and DeepQLearner have very similar sets of parameters and track similar metrics during their learning for later display, but I shall not detail it here since there is not much extra insight to be had compared to the DeepReinforcementLearning example we just went through.
 
 \subsection{MCTS DQL}
+\label{sec:MTCSHyperParamsEffect}
 
-%\newgeometry{top=0mm, bottom=0mm, left=0mm, right=0mm}
-%\begin{landscape}
-%\centering\vspace*{\fill}
 \begin{figure}[H]
@@
 \end{figure}
-%\vfill
-%\end{landscape}
-%\restoregeometry
 
@@
 \subsection{Solving the hardest 3x3 problem}
+\label{ssec:33SPHard}
 To finish with the 3x3 SP, let me try to throw one of the two hardest 3x3 configurations (see subsection \ref{sec:SPLowDimension}) at the different solvers to see how they fare.

### 2022-09-07 latex/Abstract/Abstract.tex
make both RHUL template and my own thesis format coexist .... I d like to have both formats

--- 2022-09-06 latex/MScAIFrancoisBerrier.tex
@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Masters/Doctoral Thesis 
-% LaTeX Template
-% Version 2.5 (27/8/17)
-%
-% This template was downloaded from:
-% http://www.LaTeXTemplates.com
-%
-% Version 2.x major modifications by:
-% Vel (vel@latextemplates.com)
-%
-% This template is based on a template by:
-% Steve Gunn (http://users.ecs.soton.ac.uk/srg/softwaretools/document/templates/)
-% Sunil Patel (http://www.sunilpatel.co.uk/thesis-template/)
-%
-% Template license:
-% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%----------------------------------------------------------------------------------------
-%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
-%----------------------------------------------------------------------------------------
-
-\documentclass[
-11pt, % The default document font size, options: 10pt, 11pt, 12pt
-%oneside, % Two side (alternating margins) for binding by default, uncomment to switch to one side
-english, % ngerman for German
-singlespacing, % Single line spacing, alternatives: onehalfspacing or doublespacing
-%draft, % Uncomment to enable draft mode (no pictures, no links, overfull hboxes indicated)
-%nolistspacing, % If the document is onehalfspacing or doublespacing, uncomment this to set spacing in lists to single
-%liststotoc, % Uncomment to add the list of figures/tables/etc to the table of contents
-%toctotoc, % Uncomment to add the main table of contents to the table of contents
-%parskip, % Uncomment to add space between paragraphs
-%nohyperref, % Uncomment to not load the hyperref package
-headsepline, % Uncomment to get a line under the header
-%chapterinoneline, % Uncomment to place the chapter title next to the number on one line
-%consistentlayout, % Uncomment to change the layout of the declaration, abstract and acknowledgements pages to match the default layout
-]{MastersDoctoralThesis} % The class file specifying the document structure
-
-\usepackage[utf8]{inputenc} % Required for inputting international characters
-\usepackage[T1]{fontenc} % Output font encoding for international characters
-\usepackage{mathpazo} % Use the Palatino font by default
-\usepackage[T5,T1]{fontenc}
-\usepackage{verbatim}
-\usepackage[backend=bibtex,style=authoryear,natbib=true]{biblatex} % Use the bibtex backend with the authoryear citation style (which resembles APA)
-\usepackage{float}
-\usepackage{graphbox}
-\usepackage{lscape}
-\usepackage{pdflscape}
-\usepackage{rotating}
-\usepackage[paper=portrait,pagesize]{typearea}
-\usepackage{logicpuzzle}
-\usepackage{geometry}
-\usepackage{amsmath}
-\usepackage{algorithm}
-\usepackage{algpseudocode}
-\usepackage{rubikcube}
-\newenvironment{three}[1][]{%
-\begin{logicpuzzle}[rows=2,columns=2,#1]
-\begin{puzzleforeground}
-\framepuzzle
-\end{puzzleforeground}
-}{\end{logicpuzzle}}
-
-\newenvironment{five}[1][]{%
-\begin{logicpuzzle}[rows=2,columns=3,#1]
-\begin{puzzleforeground}
-\framepuzzle
-\end{puzzleforeground}
-}{\end{logicpuzzle}}
-
-\newenvironment{seven}[1][]{%
-\begin{logicpuzzle}[rows=2,columns=4,#1]
-\begin{puzzleforeground}
-\framepuzzle
-\end{puzzleforeground}
-}{\end{logicpuzzle}}
-
-\newenvironment{nine}[1][]{%
-\begin{logicpuzzle}[rows=2,columns=5,#1]
-\begin{puzzleforeground}
-\framepuzzle
-\end{puzzleforeground}
-}{\end{logicpuzzle}}
-
-\newenvironment{eight}[1][]{%
-\begin{logicpuzzle}[rows=3,columns=3,#1]
-\begin{puzzleforeground}
-\framepuzzle
-\end{puzzleforeground}
-}{\end{logicpuzzle}}
-
-\newenvironment{fifteen}[1][]{%
-\begin{logicpuzzle}[rows=4,columns=4,#1]
-\begin{puzzleforeground}
-\framepuzzle
-\end{puzzleforeground}
-}{\end{logicpuzzle}}
-
-\newenvironment{twenty}[1][]{%
-\begin{logicpuzzle}[rows=4,columns=5,#1]
-\begin{puzzleforeground}
-\framepuzzle
-\end{puzzleforeground}
-}{\end{logicpuzzle}}
-
-\newenvironment{twentyfour}[1][]{%
-\begin{logicpuzzle}[rows=5,columns=5,#1]
-\begin{puzzleforeground}
-\framepuzzle
-\end{puzzleforeground}
-}{\end{logicpuzzle}}
-
-\newenvironment{thityfive}[1][]{%
-\begin{logicpuzzle}[rows=6,columns=6,#1]
-\begin{puzzleforeground}
-\framepuzzle
-\end{puzzleforeground}
-}{\end{logicpuzzle}}
-
-\newcommand{\red}{\color{red}}
-\newcommand{\black}{\color{black}}
-\newcommand{\blue}{\color{blue}}
-\definecolor{airforceblue}{rgb}{0.36, 0.54, 0.66}
-\newcommand{\afblue}{\color{airforceblue}}
-\newcommand{\teal}{\color{teal}}
-
-\newenvironment{titlemize}[1]{%
-  \paragraph{#1}
-  \begin{enumerate}}
-  {\end{enumerate}}
-
-\usepackage{pythonhighlight}
-
-\addbibresource{references.bib} % The filename of the bibliography
-
-\usepackage[autostyle=true]{csquotes} % Required to generate language-dependent quotes in the bibliography
-\usepackage{hyperref}
-
-%----------------------------------------------------------------------------------------
-%	MARGIN SETTINGS
-%----------------------------------------------------------------------------------------
-
-\geometry{
-	paper=a4paper, % Change to letterpaper for US letter
-	inner=1.5cm, % Inner margin
-	outer=1.5cm, %3.8cm, % Outer margin
-	bindingoffset=.5cm, % Binding offset
-	top=1.5cm, % Top margin
-	bottom=1.5cm, % Bottom margin
-	%showframe, % Uncomment to show how the type block is set on the page
-}
-
-%----------------------------------------------------------------------------------------
-%	THESIS INFORMATION
-%----------------------------------------------------------------------------------------
-
-\thesistitle{Solving \\ the \\ 15-Puzzle \\ \& \\ Rubik's Cube} % Your thesis title, this is used in the title and abstract, print it elsewhere with \ttitle
-\supervisor{Pr. Chris \textsc{Watkins}} % Your supervisor's name, this is used in the title page, print it elsewhere with \supname
-\examiner{} % Your examiner's name, this is not currently used anywhere in the template, print it elsewhere with \examname
-\degree{\textbf{MSc in Artificial Intelligence}} % Your degree name, this is used in the title page and abstract, print it elsewhere with \degreename
-\author{Fran\c{c}ois \textsc{Berrier}} % Your name, this is used in the title page and abstract, print it elsewhere with \authorname
-\addresses{} % Your address, this is not currently used anywhere in the template, print it elsewhere with \addressname
-
-\subject{Artificial Intelligence} % Your subject area, this is not currently used anywhere in the template, print it elsewhere with \subjectname
-\keywords{} % Keywords for your thesis, this is not currently used anywhere in the template, print it elsewhere with \keywordnames
-\university{\href{https://www.royalholloway.ac.uk/}{Royal Holloway, University of London}} % Your university's name and URL, this is used in the title page and abstract, print it elsewhere with \univname
-\department{\href{https://www.royalholloway.ac.uk/research-and-teaching/departments-and-schools/computer-science/}{Computer Science Department}} % Your department's name and URL, this is used in the title page and abstract, print it elsewhere with \deptname
-%\group{\href{http://researchgroup.university.com}{Research Group Name}} % Your research group's name and URL, this is used in the title page, print it elsewhere with \groupname
-%\faculty{\href{http://faculty.university.com}{Faculty Name}} % Your faculty's name and URL, this is used in the title page and abstract, print it elsewhere with \facname
-
-\AtBeginDocument{
-\hypersetup{pdftitle=\ttitle} % Set the PDF's title to your title
-\hypersetup{pdfauthor=\authorname} % Set the PDF's author to your name
-\hypersetup{pdfkeywords=\keywordnames} % Set the PDF's keywords to your keywords
-}
-
-\begin{document}
-
-\frontmatter % Use roman page numbering style (i, ii, iii, iv...) for the pre-content pages
-
-\pagestyle{plain} % Default to the plain heading style until the thesis style is called for the body content
-
-%----------------------------------------------------------------------------------------
-%	TITLE PAGE
-%----------------------------------------------------------------------------------------
-
-\begin{titlepage}
-\begin{center}
-
-\vspace*{.06\textheight}
-{\scshape\LARGE \univname\par}\vspace{1cm} % University name
-\textsc{\Large MSc Thesis}\\[0.5cm] % Thesis type
-
-\HRule \\[0.4cm] % Horizontal line
-{\huge \bfseries \ttitle\par}\vspace{0.4cm} % Thesis title
-\HRule \\[1.5cm] % Horizontal line
-
-\begin{minipage}[t]{0.4\textwidth}
-\begin{flushleft} \large
-\emph{Author:}\\
-\href{francois.berrier@cantab.net}{\authorname} % Author name - remove the \href bracket to remove the link
-\end{flushleft}
-\end{minipage}
-\begin{minipage}[t]{0.5\textwidth}
-\begin{flushright} \large
-\emph{Supervisor:} \\
-\href{https://www.cs.rhul.ac.uk/~chrisw/}{\supname} % Supervisor name - remove the \href bracket to remove the link  
-\end{flushright}
-\end{minipage}
-
-\vfill
-
-\begin{figure}[H]
-\centering
-\includegraphics[scale=0.3]{./Figures/RHUL}
-%\decoRule
-\end{figure}
-\large \textit{A thesis submitted in fulfillment of the requirements\\ for the degree of \degreename}\\[0.3cm] % University requirement text
-%\textit{in the}\\[0.4cm]
-%\groupname\\\deptname\\[2cm] % Research group name and department name
-
-\vfill
-
-{\large \today}\\[4cm] % Date
-%\includegraphics{Logo} % University/department logo - uncomment to place it
-
-\vfill
-\end{center}
-\end{titlepage}
-
-%----------------------------------------------------------------------------------------
-%	QUOTATION PAGE
-%----------------------------------------------------------------------------------------
-
-\vspace*{0.2\textheight}
-
-\noindent\enquote{\itshape The best advice I've ever received is 'No one else knows what they're doing either'}\bigbreak
-
-\hfill Ricky Gervais
-
-%----------------------------------------------------------------------------------------
-%	ABSTRACT PAGE
-%----------------------------------------------------------------------------------------
-
-\begin{abstract}
-\addchaptertocentry{\abstractname} % Add the abstract to the table of contents
+\noindent
 \textbf{Motivation}
@@
 A second appendix describes a limitation I bumped into while using the kociemba library (an open source 3x3x3 \textbf{RC} solver), and how I circumvented it.
-
-\end{abstract}
-
-%----------------------------------------------------------------------------------------
-%	ACKNOWLEDGEMENTS
-%----------------------------------------------------------------------------------------
-
-\begin{acknowledgements}
-\addchaptertocentry{\acknowledgementname} % Add the acknowledgements to the table of contents
-%\begin{comment}
-I am grateful to my employer Bank of America for sponsoring my part time 2-year MSc in Artificial Intelligence at Royal Holloway.
-I entered in the study of AI with a heavy dose of skepticism and am delighted to have learnt so much and revised my judgment of this field.
-In particular I would like to thank my managers Mitrajit Dutta and Stephen Thompson for supporting me in pursuing this MSc.
-\\
-\\
-I have been fortunate to have had the company of my colleague Saurabh Kumar, as he decided to follow me in this MSc journey.
-We have had countless opportunities to debate about all the new cool stuff we learnt during the MSc, and that has definitely made the whole process much more enjoyable.
-\\
-\\
-I would also like to thank Professor Chris Watkins, for agreeing to supervise my project.
-It was an honour to attend his very lively lectures on Deep Learning during the first year of the MSc, as well as to absorb some of his wisdom and enthusiasm at our meetings during my work on this project.
-I am also glad to have followed his advice at our last meeting: I stopped shaving for a month, which did free enough time for me to implement DQL (joint policy and value learning), as well as a Monte Carlo Tree Search algorithm, to nicely complement my assortment of sliding-puzzle and Rubik's cube solvers.
-\\
-\\
-Last but not least, my wife Thanh Nh\~a is my biggest supporter in everything I do, and her continued encouragement in my pursuit of this MSc has made it so much easier to juggle between a demanding banking job, running 50 to 70 miles a week in preparation for the Chicago marathon, and the coursework, lectures and exam revisions.
-%\end{comment}
-\end{acknowledgements}
-%----------------------------------------------------------------------------------------
-%	LIST OF CONTENTS/FIGURES/TABLES PAGES
-%----------------------------------------------------------------------------------------
-
-%\listoffigures % Prints the list of figures
-
-%\listoftables % Prints the list of tables
-
-%----------------------------------------------------------------------------------------
-%	ABBREVIATIONS
-%----------------------------------------------------------------------------------------
-
-\begin{abbreviations}{ll} % Include a list of abbreviations (a table of two columns)
-
-\textbf{AIPnT} & \textbf{A}rtificial \textbf{I}ntelligence \textbf{P}rinciples and \textbf{T}echniques\\
-\textbf{BFS} & \textbf{B}readth \textbf{F}irst \textbf{S}earch\\
-\textbf{CS} & \textbf{C}omputer \textbf{S}cience\\
-\textbf{CV} & \textbf{C}omputer \textbf{V}ision\\
-\textbf{DFS} & \textbf{D}epth \textbf{F}irst \textbf{S}earch\\
-\textbf{DL} & \textbf{D}eep \textbf{L}earning\\
-\textbf{DQL} & \textbf{D}eep \textbf{Q}-\textbf{L}earning\\
-\textbf{DRL} & \textbf{D}eep \textbf{R}einforcment \textbf{L}earning\\
-\textbf{FIFO} & \textbf{F}irst \textbf{I}n \textbf{F}irst \textbf{O}ut\\
-\textbf{LIFO} & \textbf{L}ast \textbf{I}n \textbf{F}irst \textbf{O}ut\\
-\textbf{MCTS} & \textbf{M}onte \textbf{C}arlo \textbf{T}ree \textbf{S}earch \\
-\textbf{ML} & \textbf{M}achine \textbf{L}earning\\
-\textbf{NLP} & \textbf{N}atural \textbf{L}anguage \textbf{P}rocessing\\
-\textbf{RC} & \textbf{R}ubik's \textbf{C}ube\\
-\textbf{RL} & \textbf{R}einforcement \textbf{L}earning\\
-\textbf{RHUL} & \textbf{R}oyal \textbf{H}olloway, \textbf{U}niversity of \textbf{L}ondon\\
-\textbf{SP} & \textbf{S}liding \textbf{P}uzzle\\
-
-\end{abbreviations}
-
-\hypersetup{%
-  colorlinks = true,
-  linkcolor  = black
-}
-\tableofcontents % Prints the main table of contents
-
-\hypersetup{%
-  colorlinks = true,
-  linkcolor  = orange
-}
-
-%----------------------------------------------------------------------------------------
-%	PHYSICAL CONSTANTS/OTHER DEFINITIONS
-%----------------------------------------------------------------------------------------
-
-%\begin{constants}{lr@{${}={}$}l} % The list of physical constants is a three column table
-
-% The \SI{}{} command is provided by the siunitx package, see its documentation for instructions on how to use it
-
-%Speed of Light & $c_{0}$ & \SI{2.99792458e8}{\meter\per\second} (exact)\\
-%Constant Name & $Symbol$ & $Constant Value$ with units\\
-
-%\end{constants}
-
-%----------------------------------------------------------------------------------------
-%	SYMBOLS
-%----------------------------------------------------------------------------------------
-
-%\begin{symbols}{lll} % Include a list of Symbols (a three column table)
-
-%$a$ & distance & \si{\meter} \\
-%$P$ & power & \si{\watt} (\si{\joule\per\second}) \\
-%Symbol & Name & Unit \\
-
-%\addlinespace % Gap to separate the Roman symbols from the Greek
-
-%$\omega$ & angular frequency & \si{\radian} \\
-
-%\end{symbols}
-
-%----------------------------------------------------------------------------------------
-%	DEDICATION
-%----------------------------------------------------------------------------------------
-
-%\dedicatory{For/Dedicated to/To my\ldots}
-
-%----------------------------------------------------------------------------------------
-%	THESIS CONTENT - CHAPTERS
-%----------------------------------------------------------------------------------------
-
-\mainmatter % Begin numeric (1,2,3...) page numbering
-
-\pagestyle{thesis} % Return the page headers back to the "thesis" style
-
-% Include the chapters of the thesis as separate files from the Chapters folder
-% Uncomment the lines as you write the chapters
-
-\include{Chapters/Objectives}
-\include{Chapters/DRLSearch}
-\include{Chapters/Puzzles}
-\include{Chapters/Implementation}
-\include{Chapters/SlidingPuzzle}
-\include{Chapters/Rubiks}
-\include{Chapters/Conclusion}
-
-%----------------------------------------------------------------------------------------
-%	THESIS CONTENT - APPENDICES
-%----------------------------------------------------------------------------------------
-
-\appendix % Cue to tell LaTeX that the following "chapters" are Appendices
-
-% Include the appendices of the thesis as separate files from the Appendices folder
-% Uncomment the lines as you write the Appendices
-
-\include{Appendices/Examples}
-\include{Appendices/Kociemba}
-
-%----------------------------------------------------------------------------------------
-%	BIBLIOGRAPHY
-%----------------------------------------------------------------------------------------
-
-%\printbibliography[heading=bibintoc]
-\printbibheading
-\printbibliography[type=article,heading=subbibliography,title={Journal Papers}]
-\printbibliography[type=book,heading=subbibliography,title={Books}]
-\printbibliography[type=misc,heading=subbibliography,title={Misc}]
-
-%----------------------------------------------------------------------------------------
-
-\end{document}

### 2022-09-07 latex/Acknowledgements/Acknowledgements.tex
make both RHUL template and my own thesis format coexist .... I d like to have both formats

--- 2022-09-05 latex/MScAIFrancoisBerrier.tex
@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Masters/Doctoral Thesis 
-% LaTeX Template
-% Version 2.5 (27/8/17)
-%
-% This template was downloaded from:
-% http://www.LaTeXTemplates.com
-%
-% Version 2.x major modifications by:
-% Vel (vel@latextemplates.com)
-%
-% This template is based on a template by:
-% Steve Gunn (http://users.ecs.soton.ac.uk/srg/softwaretools/document/templates/)
-% Sunil Patel (http://www.sunilpatel.co.uk/thesis-template/)
-%
-% Template license:
-% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%----------------------------------------------------------------------------------------
-%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
-%----------------------------------------------------------------------------------------
-
-\documentclass[
-11pt, % The default document font size, options: 10pt, 11pt, 12pt
-%oneside, % Two side (alternating margins) for binding by default, uncomment to switch to one side
-english, % ngerman for German
-singlespacing, % Single line spacing, alternatives: onehalfspacing or doublespacing
-%draft, % Uncomment to enable draft mode (no pictures, no links, overfull hboxes indicated)
-%nolistspacing, % If the document is onehalfspacing or doublespacing, uncomment this to set spacing in lists to single
-%liststotoc, % Uncomment to add the list of figures/tables/etc to the table of contents
-%toctotoc, % Uncomment to add the main table of contents to the table of contents
-%parskip, % Uncomment to add space between paragraphs
-%nohyperref, % Uncomment to not load the hyperref package
-headsepline, % Uncomment to get a line under the header
-%chapterinoneline, % Uncomment to place the chapter title next to the number on one line
-%consistentlayout, % Uncomment to change the layout of the declaration, abstract and acknowledgements pages to match the default layout
-]{MastersDoctoralThesis} % The class file specifying the document structure
-
-\usepackage[utf8]{inputenc} % Required for inputting international characters
-\usepackage[T1]{fontenc} % Output font encoding for international characters
-\usepackage{mathpazo} % Use the Palatino font by default
-\usepackage[T5,T1]{fontenc}
-\usepackage{verbatim}
-\usepackage[backend=bibtex,style=authoryear,natbib=true]{biblatex} % Use the bibtex backend with the authoryear citation style (which resembles APA)
-\usepackage{float}
-\usepackage{graphbox}
-\usepackage{lscape}
-\usepackage{pdflscape}
-\usepackage{rotating}
-\usepackage[paper=portrait,pagesize]{typearea}
-\usepackage{logicpuzzle}
-\usepackage{geometry}
-\usepackage{amsmath}
-\usepackage{algorithm}
-\usepackage{algpseudocode}
-\usepackage{rubikcube}
-\newenvironment{three}[1][]{%
-\begin{logicpuzzle}[rows=2,columns=2,#1]
-\begin{puzzleforeground}
-\framepuzzle
-\end{puzzleforeground}
-}{\end{logicpuzzle}}
-
-\newenvironment{five}[1][]{%
-\begin{logicpuzzle}[rows=2,columns=3,#1]
-\begin{puzzleforeground}
-\framepuzzle
-\end{puzzleforeground}
-}{\end{logicpuzzle}}
-
-\newenvironment{seven}[1][]{%
-\begin{logicpuzzle}[rows=2,columns=4,#1]
-\begin{puzzleforeground}
-\framepuzzle
-\end{puzzleforeground}
-}{\end{logicpuzzle}}
-
-\newenvironment{nine}[1][]{%
-\begin{logicpuzzle}[rows=2,columns=5,#1]
-\begin{puzzleforeground}
-\framepuzzle
-\end{puzzleforeground}
-}{\end{logicpuzzle}}
-
-\newenvironment{eight}[1][]{%
-\begin{logicpuzzle}[rows=3,columns=3,#1]
-\begin{puzzleforeground}
-\framepuzzle
-\end{puzzleforeground}
-}{\end{logicpuzzle}}
-
-\newenvironment{fifteen}[1][]{%
-\begin{logicpuzzle}[rows=4,columns=4,#1]
-\begin{puzzleforeground}
-\framepuzzle
-\end{puzzleforeground}
-}{\end{logicpuzzle}}
-
-\newenvironment{twenty}[1][]{%
-\begin{logicpuzzle}[rows=4,columns=5,#1]
-\begin{puzzleforeground}
-\framepuzzle
-\end{puzzleforeground}
-}{\end{logicpuzzle}}
-
-\newenvironment{twentyfour}[1][]{%
-\begin{logicpuzzle}[rows=5,columns=5,#1]
-\begin{puzzleforeground}
-\framepuzzle
-\end{puzzleforeground}
-}{\end{logicpuzzle}}
-
-\newenvironment{thityfive}[1][]{%
-\begin{logicpuzzle}[rows=6,columns=6,#1]
-\begin{puzzleforeground}
-\framepuzzle
-\end{puzzleforeground}
-}{\end{logicpuzzle}}
-
-\newcommand{\red}{\color{red}}
-\newcommand{\black}{\color{black}}
-\newcommand{\blue}{\color{blue}}
-\definecolor{airforceblue}{rgb}{0.36, 0.54, 0.66}
-\newcommand{\afblue}{\color{airforceblue}}
-
-\newenvironment{titlemize}[1]{%
-  \paragraph{#1}
-  \begin{enumerate}}
-  {\end{enumerate}}
-
-\usepackage{pythonhighlight}
-
-\addbibresource{references.bib} % The filename of the bibliography
-
-\usepackage[autostyle=true]{csquotes} % Required to generate language-dependent quotes in the bibliography
-\usepackage{hyperref}
-
-%----------------------------------------------------------------------------------------
-%	MARGIN SETTINGS
-%----------------------------------------------------------------------------------------
-
-\geometry{
-	paper=a4paper, % Change to letterpaper for US letter
-	inner=1.5cm, % Inner margin
-	outer=1.5cm, %3.8cm, % Outer margin
-	bindingoffset=.5cm, % Binding offset
-	top=1.5cm, % Top margin
-	bottom=1.5cm, % Bottom margin
-	%showframe, % Uncomment to show how the type block is set on the page
-}
-
-%----------------------------------------------------------------------------------------
-%	THESIS INFORMATION
-%----------------------------------------------------------------------------------------
-
-\thesistitle{Solving \\ the \\ 15-Puzzle \\ \& \\ Rubik's Cube} % Your thesis title, this is used in the title and abstract, print it elsewhere with \ttitle
-\supervisor{Pr. Chris \textsc{Watkins}} % Your supervisor's name, this is used in the title page, print it elsewhere with \supname
-\examiner{} % Your examiner's name, this is not currently used anywhere in the template, print it elsewhere with \examname
-\degree{\textbf{MSc in Artificial Intelligence}} % Your degree name, this is used in the title page and abstract, print it elsewhere with \degreename
-\author{Fran\c{c}ois \textsc{Berrier}} % Your name, this is used in the title page and abstract, print it elsewhere with \authorname
-\addresses{} % Your address, this is not currently used anywhere in the template, print it elsewhere with \addressname
-
-\subject{Artificial Intelligence} % Your subject area, this is not currently used anywhere in the template, print it elsewhere with \subjectname
-\keywords{} % Keywords for your thesis, this is not currently used anywhere in the template, print it elsewhere with \keywordnames
-\university{\href{https://www.royalholloway.ac.uk/}{Royal Holloway, University of London}} % Your university's name and URL, this is used in the title page and abstract, print it elsewhere with \univname
-\department{\href{https://www.royalholloway.ac.uk/research-and-teaching/departments-and-schools/computer-science/}{Computer Science Department}} % Your department's name and URL, this is used in the title page and abstract, print it elsewhere with \deptname
-%\group{\href{http://researchgroup.university.com}{Research Group Name}} % Your research group's name and URL, this is used in the title page, print it elsewhere with \groupname
-%\faculty{\href{http://faculty.university.com}{Faculty Name}} % Your faculty's name and URL, this is used in the title page and abstract, print it elsewhere with \facname
-
-\AtBeginDocument{
-\hypersetup{pdftitle=\ttitle} % Set the PDF's title to your title
-\hypersetup{pdfauthor=\authorname} % Set the PDF's author to your name
-\hypersetup{pdfkeywords=\keywordnames} % Set the PDF's keywords to your keywords
-}
-
-\begin{document}
-
-\frontmatter % Use roman page numbering style (i, ii, iii, iv...) for the pre-content pages
-
-\pagestyle{plain} % Default to the plain heading style until the thesis style is called for the body content
-
-%----------------------------------------------------------------------------------------
-%	TITLE PAGE
-%----------------------------------------------------------------------------------------
-
-\begin{titlepage}
-\begin{center}
-
-\vspace*{.06\textheight}
-{\scshape\LARGE \univname\par}\vspace{1cm} % University name
-\textsc{\Large MSc Thesis}\\[0.5cm] % Thesis type
-
-\HRule \\[0.4cm] % Horizontal line
-{\huge \bfseries \ttitle\par}\vspace{0.4cm} % Thesis title
-\HRule \\[1.5cm] % Horizontal line
-
-\begin{minipage}[t]{0.4\textwidth}
-\begin{flushleft} \large
-\emph{Author:}\\
-\href{francois.berrier@cantab.net}{\authorname} % Author name - remove the \href bracket to remove the link
-\end{flushleft}
-\end{minipage}
-\begin{minipage}[t]{0.5\textwidth}
-\begin{flushright} \large
-\emph{Supervisor:} \\
-\href{https://www.cs.rhul.ac.uk/~chrisw/}{\supname} % Supervisor name - remove the \href bracket to remove the link  
-\end{flushright}
-\end{minipage}
-
-\vfill
-
-\begin{figure}[H]
-\centering
-\includegraphics[scale=0.3]{./Figures/RHUL}
-%\decoRule
-\end{figure}
-\large \textit{A thesis submitted in fulfillment of the requirements\\ for the degree of \degreename}\\[0.3cm] % University requirement text
-%\textit{in the}\\[0.4cm]
-%\groupname\\\deptname\\[2cm] % Research group name and department name
-
-\vfill
-
-{\large \today}\\[4cm] % Date
-%\includegraphics{Logo} % University/department logo - uncomment to place it
-
-\vfill
-\end{center}
-\end{titlepage}
-
-%----------------------------------------------------------------------------------------
-%	QUOTATION PAGE
-%----------------------------------------------------------------------------------------
-
-\vspace*{0.2\textheight}
-
-\noindent\enquote{\itshape The best advice I've ever received is 'No one else knows what they're doing either'}\bigbreak
-
-\hfill Ricky Gervais
-
-%----------------------------------------------------------------------------------------
-%	ABSTRACT PAGE
-%----------------------------------------------------------------------------------------
-
-\begin{abstract}
-\addchaptertocentry{\abstractname} % Add the abstract to the table of contents
-\textbf{Motivation}
-\label{Abstract}
-\\
-Reinforcement Learning (\textbf{RL}), exposed with brilliant clarity in the Sutton book (\cite{Sutton1998}), has until recently known less success than we might have hoped for.
-Its framework is very appealing and intuitive.
-In particular, the mathematical beauty of Value iteration and Q iteration (\cite{WatkinsThesis}) for discrete state and action spaces, blindly iterating from \textit{any} initial value, is quite profound.
-Sadly, it had until recently proven hard to achieve practical success with these methods.
-\\
-Inspired however by the seminal success of Deep Mind's team in using Deep Reinforcement Learning (\textbf{DRL}) to play Atari games (\cite{Mnih2013})  and to master the game of Go (\cite{AlphaGo}), researchers have in recent years made a lot of progress towards designing algorithms capable of learning and solving, \textit{without human knowledge}, the Rubik's Cube (\textbf{RC}) - as well as similar single player puzzles - by using Deep Q-Learning (\textbf{DQL})  (\cite{DBLP:journals/corr/abs-1805-07470}) or search and (\textbf{DRL}) value iteration (\cite{https://doi.org/10.48550/arxiv.1805.07470}).
-\\
-In this project, I will attempt to implement a variety of solvers combining $A^{*}$ search and heuristics, some of which will be handcrafted, others which I will train on randomly generated sequences of puzzles using (\textbf{DL}) or (\textbf{DRL}), to solve the 15-puzzle (and variations of different dimensions), as well as the Rubik's cube.
-\\
-\\
-\textbf{Organisation of this thesis}
-\\
-In the first chapter, I will quickly describe what I hope to get out of this project, both in terms of personal learning, as well as in terms of tangible results (solving some puzzles!).
-In chapter 2, I will do a quick recap of the different methods that I will use in the project.
-Chapter 3 will be dedicated to discussing the mathematics of the sliding puzzle (\textbf{SP}) and the (\textbf{RC}).
-I might throw a few random (but hopefully interesting) observations in there and give some references for the keen reader.
-I will then give an overview in chapter 4 of the code base I have developed - and put in the open on my github page (\cite{FB}) - to complete this project before detailing a few examples in chapter 5.
-Finally, in chapters 6 and 7, I will present all my various results on respectively the sliding puzzle and the Rubik's cube before offering some conclusion in chapter 8.
-
-\end{abstract}
-
-%----------------------------------------------------------------------------------------
-%	ACKNOWLEDGEMENTS
-%----------------------------------------------------------------------------------------
-
-\begin{acknowledgements}
-\addchaptertocentry{\acknowledgementname} % Add the acknowledgements to the table of contents
-%\begin{comment}
 I am grateful to my employer Bank of America for sponsoring my part time 2-year MSc in Artificial Intelligence at Royal Holloway.
@@
 Last but not least, my wife Thanh Nh\~a is my biggest supporter in everything I do, and her continued encouragement in my pursuit of this MSc has made it so much easier to juggle between a demanding banking job, running 50 to 70 miles a week in preparation for the Chicago marathon, and the coursework, lectures and exam revisions.
-%\end{comment}
-\end{acknowledgements}
-%----------------------------------------------------------------------------------------
-%	LIST OF CONTENTS/FIGURES/TABLES PAGES
-%----------------------------------------------------------------------------------------
-
-%\listoffigures % Prints the list of figures
-
-%\listoftables % Prints the list of tables
-
-%----------------------------------------------------------------------------------------
-%	ABBREVIATIONS
-%----------------------------------------------------------------------------------------
-
-\begin{abbreviations}{ll} % Include a list of abbreviations (a table of two columns)
-
-\textbf{AIPnT} & \textbf{A}rtificial \textbf{I}ntelligence \textbf{P}rinciples and \textbf{T}echniques\\
-\textbf{BFS} & \textbf{B}readth \textbf{F}irst \textbf{S}earch\\
-\textbf{CS} & \textbf{C}omputer \textbf{S}cience\\
-\textbf{CV} & \textbf{C}omputer \textbf{V}ision\\
-\textbf{DFS} & \textbf{D}epth \textbf{F}irst \textbf{S}earch\\
-\textbf{DL} & \textbf{D}eep \textbf{L}earning\\
-\textbf{DQL} & \textbf{D}eep \textbf{Q}-\textbf{L}earning\\
-\textbf{DRL} & \textbf{D}eep \textbf{R}einforcment \textbf{L}earning\\
-\textbf{FIFO} & \textbf{F}irst \textbf{I}n \textbf{F}irst \textbf{O}ut\\
-\textbf{LIFO} & \textbf{L}ast \textbf{I}n \textbf{F}irst \textbf{O}ut\\
-\textbf{MCTS} & \textbf{M}onte \textbf{C}arlo \textbf{T}ree \textbf{S}earch \\
-\textbf{ML} & \textbf{M}achine \textbf{L}earning\\
-\textbf{NLP} & \textbf{N}atural \textbf{L}anguage \textbf{P}rocessing\\
-\textbf{RC} & \textbf{R}ubik's \textbf{C}ube\\
-\textbf{RL} & \textbf{R}einforcement \textbf{L}earning\\
-\textbf{RHUL} & \textbf{R}oyal \textbf{H}olloway, \textbf{U}niversity of \textbf{L}ondon\\
-\textbf{SP} & \textbf{S}liding \textbf{P}uzzle\\
-
-\end{abbreviations}
-
-\hypersetup{%
-  colorlinks = true,
-  linkcolor  = black
-}
-\tableofcontents % Prints the main table of contents
-
-\hypersetup{%
-  colorlinks = true,
-  linkcolor  = orange
-}
-
-%----------------------------------------------------------------------------------------
-%	PHYSICAL CONSTANTS/OTHER DEFINITIONS
-%----------------------------------------------------------------------------------------
-
-%\begin{constants}{lr@{${}={}$}l} % The list of physical constants is a three column table
-
-% The \SI{}{} command is provided by the siunitx package, see its documentation for instructions on how to use it
-
-%Speed of Light & $c_{0}$ & \SI{2.99792458e8}{\meter\per\second} (exact)\\
-%Constant Name & $Symbol$ & $Constant Value$ with units\\
-
-%\end{constants}
-
-%----------------------------------------------------------------------------------------
-%	SYMBOLS
-%----------------------------------------------------------------------------------------
-
-%\begin{symbols}{lll} % Include a list of Symbols (a three column table)
-
-%$a$ & distance & \si{\meter} \\
-%$P$ & power & \si{\watt} (\si{\joule\per\second}) \\
-%Symbol & Name & Unit \\
-
-%\addlinespace % Gap to separate the Roman symbols from the Greek
-
-%$\omega$ & angular frequency & \si{\radian} \\
-
-%\end{symbols}
-
-%----------------------------------------------------------------------------------------
-%	DEDICATION
-%----------------------------------------------------------------------------------------
-
-%\dedicatory{For/Dedicated to/To my\ldots}
-
-%----------------------------------------------------------------------------------------
-%	THESIS CONTENT - CHAPTERS
-%----------------------------------------------------------------------------------------
-
-\mainmatter % Begin numeric (1,2,3...) page numbering
-
-\pagestyle{thesis} % Return the page headers back to the "thesis" style
-
-% Include the chapters of the thesis as separate files from the Chapters folder
-% Uncomment the lines as you write the chapters
-
-\include{Chapters/Objectives}
-\include{Chapters/DRLSearch}
-\include{Chapters/Puzzles}
-\include{Chapters/Implementation}
-\include{Chapters/SlidingPuzzle}
-\include{Chapters/Rubiks}
-\include{Chapters/Conclusion}
-
-%----------------------------------------------------------------------------------------
-%	THESIS CONTENT - APPENDICES
-%----------------------------------------------------------------------------------------
-
-\appendix % Cue to tell LaTeX that the following "chapters" are Appendices
-
-% Include the appendices of the thesis as separate files from the Appendices folder
-% Uncomment the lines as you write the Appendices
-
-\include{Appendices/Examples}
-\include{Appendices/Kociemba}
-
-%----------------------------------------------------------------------------------------
-%	BIBLIOGRAPHY
-%----------------------------------------------------------------------------------------
-
-%\printbibliography[heading=bibintoc]
-\printbibheading
-\printbibliography[type=article,heading=subbibliography,title={Journal Papers}]
-\printbibliography[type=book,heading=subbibliography,title={Books}]
-\printbibliography[type=misc,heading=subbibliography,title={Misc}]
-
-%----------------------------------------------------------------------------------------
-
-\end{document}

### 2022-09-07 latex/MScAIFrancoisBerrier.tex
make both RHUL template and my own thesis format coexist .... I d like to have both formats

--- 2022-09-03 latex/MScAIFrancoisBerrier.tex
@@
 
-\usepackage[utf8]{inputenc} % Required for inputting international characters
-\usepackage[T1]{fontenc} % Output font encoding for international characters
-\usepackage{mathpazo} % Use the Palatino font by default
-\usepackage[T5,T1]{fontenc}
-\usepackage{verbatim}
-\usepackage[backend=bibtex,style=authoryear,natbib=true]{biblatex} % Use the bibtex backend with the authoryear citation style (which resembles APA)
-\usepackage{float}
-\usepackage{graphbox}
-\usepackage{lscape}
-\usepackage{pdflscape}
-\usepackage{rotating}
-\usepackage[paper=portrait,pagesize]{typearea}
-\usepackage{logicpuzzle}
-\usepackage{geometry}
-\usepackage{amsmath}
-\usepackage{algorithm}
-\usepackage{algpseudocode}
-\usepackage{rubikcube}
-\newenvironment{three}[1][]{%
-\begin{logicpuzzle}[rows=2,columns=2,#1]
-\begin{puzzleforeground}
-\framepuzzle
-\end{puzzleforeground}
-}{\end{logicpuzzle}}
+\input{Packages/Packages}
 
-\newenvironment{five}[1][]{%
-\begin{logicpuzzle}[rows=2,columns=3,#1]
-\begin{puzzleforeground}
-\framepuzzle
-\end{puzzleforeground}
-}{\end{logicpuzzle}}
-
-\newenvironment{seven}[1][]{%
-\begin{logicpuzzle}[rows=2,columns=4,#1]
-\begin{puzzleforeground}
-\framepuzzle
-\end{puzzleforeground}
-}{\end{logicpuzzle}}
-
-\newenvironment{nine}[1][]{%
-\begin{logicpuzzle}[rows=2,columns=5,#1]
-\begin{puzzleforeground}
-\framepuzzle
-\end{puzzleforeground}
-}{\end{logicpuzzle}}
-
-\newenvironment{eight}[1][]{%
-\begin{logicpuzzle}[rows=3,columns=3,#1]
-\begin{puzzleforeground}
-\framepuzzle
-\end{puzzleforeground}
-}{\end{logicpuzzle}}
-
-\newenvironment{fifteen}[1][]{%
-\begin{logicpuzzle}[rows=4,columns=4,#1]
-\begin{puzzleforeground}
-\framepuzzle
-\end{puzzleforeground}
-}{\end{logicpuzzle}}
-
-\newenvironment{twenty}[1][]{%
-\begin{logicpuzzle}[rows=4,columns=5,#1]
-\begin{puzzleforeground}
-\framepuzzle
-\end{puzzleforeground}
-}{\end{logicpuzzle}}
-
-\newenvironment{twentyfour}[1][]{%
-\begin{logicpuzzle}[rows=5,columns=5,#1]
-\begin{puzzleforeground}
-\framepuzzle
-\end{puzzleforeground}
-}{\end{logicpuzzle}}
-
-\newenvironment{thityfive}[1][]{%
-\begin{logicpuzzle}[rows=6,columns=6,#1]
-\begin{puzzleforeground}
-\framepuzzle
-\end{puzzleforeground}
-}{\end{logicpuzzle}}
-
-\newcommand{\red}{\color{red}}
-\newcommand{\black}{\color{black}}
-\newcommand{\blue}{\color{blue}}
-\definecolor{airforceblue}{rgb}{0.36, 0.54, 0.66}
-\newcommand{\afblue}{\color{airforceblue}}
-
-\newenvironment{titlemize}[1]{%
-  \paragraph{#1}
-  \begin{enumerate}}
-  {\end{enumerate}}
-
-\usepackage{pythonhighlight}
+\newcommand{\Section}{\section}
+\newcommand{\Subsection}{\subsection}
 
 \addbibresource{references.bib} % The filename of the bibliography
-
-\usepackage[autostyle=true]{csquotes} % Required to generate language-dependent quotes in the bibliography
-\usepackage{hyperref}
 
@@
 \addchaptertocentry{\abstractname} % Add the abstract to the table of contents
-\textbf{Motivation}
-\label{Abstract}
-\\
-Reinforcement Learning (\textbf{RL}), exposed with brilliant clarity in the Sutton book (\cite{Sutton1998}), has until recently known less success than we might have hoped for.
-Its framework is very appealing and intuitive.
-In particular, the mathematical beauty of Value iteration and Q iteration (\cite{WatkinsThesis}) for discrete state and action spaces, blindly iterating from \textit{any} initial value, is quite profound.
-Sadly, it had until recently proven hard to achieve practical success with these methods.
-\\
-Inspired however by the seminal success of Deep Mind's team in using Deep Reinforcement Learning (\textbf{DRL}) to play Atari games (\cite{Mnih2013})  and to master the game of Go (\cite{AlphaGo}), researchers have in recent years made a lot of progress towards designing algorithms capable of learning and solving, \textit{without human knowledge}, the Rubik's Cube (\textbf{RC}) - as well as similar single player puzzles - by using Deep Q-Learning (\textbf{DQL})  (\cite{DBLP:journals/corr/abs-1805-07470}) or search and (\textbf{DRL}) value iteration (\cite{https://doi.org/10.48550/arxiv.1805.07470}).
-\\
-In this project, I will attempt to implement a variety of solvers combining $A^{*}$ search and heuristics, some of which will be handcrafted, others which I will train on randomly generated sequences of puzzles using (\textbf{DL}) or (\textbf{DRL}), to solve the 15-puzzle (and variations of different dimensions), as well as the Rubik's cube.
-\\
-\\
-\textbf{Organisation of this thesis}
-\\
-In the first chapter, I will quickly describe what I hope to get out of this project, both in terms of personal learning, as well as in terms of tangible results (solving some puzzles!).
-In chapter 2, I will do a quick recap of the different methods that I will use in the project.
-Chapter 3 will be dedicated to discussing the mathematics of the sliding puzzle (\textbf{SP}) and the (\textbf{RC}).
-I might throw a few random (but hopefully interesting) observations in there and give some references for the keen reader.
-I will then give an overview in chapter 4 of the code base I have developed - and put in the open on my github page (\cite{FB}) - to complete this project before detailing a few examples in chapter 5.
-Finally, in chapters 6 and 7, I will present all my various results on respectively the sliding puzzle and the Rubik's cube before offering some conclusion in chapter 8.
-
+\input{Abstract/Abstract}
 \end{abstract}
@@
 %\begin{comment}
-I am grateful to my employer Bank of America for sponsoring my part time 2-year MSc in Artificial Intelligence at Royal Holloway.
-I entered in the study of AI with a heavy dose of skepticism and am delighted to have learnt so much and revised my judgment of this field.
-In particular I would like to thank my managers Mitrajit Dutta and Stephen Thompson for supporting me in pursuing this MSc.
-\\
-\\
-I have been fortunate to have had the company of my colleague Saurabh Kumar, as he decided to follow me in this MSc journey.
-We have had countless opportunities to debate about all the new cool stuff we learnt during the MSc, and that has definitely made the whole process much more enjoyable.
-\\
-\\
-I would also like to thank Professor Chris Watkins, inventor of QL :), for agreeing to supervise my project.
-It was an honour to attend his very lively lectures on Deep Learning during the first year of the MSc, as well as to absorb some of his wisdom and enthusiasm at our meetings during my work on this project.
-\\
-\\
-Last but not least, my wife Thanh Nh\~a is my biggest supporter in everything I do, and her continued encouragement and enthusiasm in my pursuit of this MSc has made it so much easier to juggle between a demanding banking job, running 50 to 70 miles a week in preparation for the Chicago marathon, and the coursework, lectures and exam revisions.
+\input{Acknowledgements/Acknowledgements}
 %\end{comment}

### 2022-09-07 latex/RHULMScAIFrancoisBerrier.tex
make both RHUL template and my own thesis format coexist .... I d like to have both formats

\documentclass[11pt]{article} % do not change this line
\input{BigDataStyle.txt}      % do not change this line
\input{Packages/Packages}

\addbibresource{references.bib} % The filename of the bibliography
\newcommand{\chapter}{\section}
\newcommand{\Section}{\subsection}
\newcommand{\Subsection}{\subsubsection}

\hypersetup{%
  colorlinks = true,
  linkcolor  = black
}

\emergencystretch=5mm
\tolerance=400
\allowdisplaybreaks[4]

\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{problem}[theorem]{Problem}

\theoremstyle{definition}
\newtheorem*{remark}{Remark}

\title{Solving \\ the \\ 15-Puzzle \\ \& \\ Rubik's Cube}
\author{Fran\c{c}ois Berrier}

\newcommand{\Programme}{Artificial Intelligence}

\hypersetup{%
  citecolor=mdtRed
}

\begin{document}
\maketitle

\begin{abstract}
\input{Abstract/Abstract}
\end{abstract}

\noindent
\textbf{Acknowledgements}
\newline
\newline
\input{Acknowledgements/Acknowledgements}

%\tableofcontents % Prints the main table of contents

\newpage

\hypersetup{%
  colorlinks = true,
  linkcolor  = orange
}
\input{Chapters/Objectives}
\include{Chapters/DRLSearch}

\printbibheading
\hypersetup{pdfpagemode={UseOutlines},
bookmarksopen=true,
bookmarksopenlevel=0,
hypertexnames=false,
colorlinks=true,% Set to false to disable coloring links
citecolor=magenta,% The color of citations
linkcolor=red,% The color of references to document elements (sections, figures, etc)
urlcolor=mdtRed,% The color of hyperlinks (URLs)
pdfstartview={FitV},
unicode,
breaklinks=true,
}
\printbibliography[type=article,heading=subbibliography,title={Journal Papers}]
\printbibliography[type=book,heading=subbibliography,title={Books}]
\printbibliography[type=misc,heading=subbibliography,title={Misc}]
\end{document}

### 2022-09-08 latex/Appendices/Examples.tex
update

@@
 
-\chapter{Examples} % Main chapter title
+\Appendix{Examples} % Main chapter title
 
@@
 %-----------------------------------
-\section{Puzzles}
+\Section{Puzzles}
 Let me start by showing how to construct puzzles, using the Puzzle factory.
@@
 %-----------------------------------
-\section{Learners}
+\Section{Learners}
 
@@
 %-----------------------------------
-\subsection{Perfect Learner}
+\Subsection{Perfect Learner}
 \label{PLSS}
@@
 %-----------------------------------
-\subsection{Deep Learner}
+\Subsection{Deep Learner}
 \label{DLSS}
@@
 %-----------------------------------
-\subsection{Deep Reinforcement Learner}
+\Subsection{Deep Reinforcement Learner}
 \label{DRLSS}
@@
 
-\section{Solvers}
+\Section{Solvers}
 
@@
 %-----------------------------------
-\subsection{Blind search}
+\Subsection{Blind search}
 First let me start with some of the blind search algorithms, which I only have been able to use on small dimension SP.
@@
 %-----------------------------------
-\subsubsection{BFS}
+\Subsubsection{BFS}
 \label{BFSSS}
@@
 %-----------------------------------
-\subsubsection{DFS}
+\Subsubsection{DFS}
 \label{DFSSS}
@@
 %-----------------------------------
-\subsection{Naive Sliding Puzzle Solver}
+\Subsection{Naive Sliding Puzzle Solver}
 As a comparison point for otherSP solvers, I have implemented a naive sliding puzzle solver, which does what most beginner players would intuitively do when solving the sliding puzzle by hand: solve the top row, then the left column, and keep iterating until done.
@@
 %-----------------------------------
-\subsection{Kociemba}
+\Subsection{Kociemba}
 
@@
 %-----------------------------------
-\subsection{A*}
+\Subsection{A*}
 \label{ASSS}
 
-\subsubsection{Manhattan heuristic}
+\Subsubsection{Manhattan heuristic}
 
@@
 
-\subsubsection{Perfect Heuristic}
+\Subsubsection{Perfect Heuristic}
 To run A{*} with a perfect heuristic, we just need to specify the heuristic type as such, and set the parameter \textit{model\_file\_name} to point to a pre-recorded database populated by the PerfectLearner (see earlier section \ref{PLSS}).
@@
 
-\subsubsection{Deep Learning Heuristic}
-\subsubsection{Deep Reinforcement Learning Heuristic}
-\subsubsection{Deep Q Learning Heuristic}
+\Subsubsection{Deep Learning Heuristic}
+\Subsubsection{Deep Reinforcement Learning Heuristic}
+\Subsubsection{Deep Q Learning Heuristic}

### 2022-09-08 latex/Chapters/Conclusion.tex
update

--- 2022-08-26 latex/Chapters/Conclusion.tex
@@
 
-TBW
+\Section{Learning}
+Learnt about puzzles, and also even took the time to learn a couple of sequences of moves that are sufficient to solve \textbf{RC}, 2x2x2 in labout one minute, 3x3x3 in about 2 minutes.
+
+\Section{Implementation}
+Implemented a bunch of solvers. Ideally should do in C++ for speed and multithreaded (esp for MCTS) but time was limited so I prioritized breadth over depth.
+
+\Section{Deep Reinforcement/Q Learning}
+Quite impressive, without too much effort, how well they perform.
+However, I feel to really get more out of them (higher dimensions in particular) would need careful tuning of the convergence criteria.
+
+\Section{Professional issues}
+See \cite{ManyWorlds}
+See \cite{https://doi.org/10.1111/1758-5899.12718} and \cite{Bostrom2014}

### 2022-09-08 latex/Chapters/Puzzles.tex
update

@@
 
-\section{Sliding Puzzle}
+\Section{Sliding Puzzle}
 
@@
 %-----------------------------------
-\subsection{History - the 15-Puzzle}
+\Subsection{History - the 15-Puzzle}
 
@@
 \includegraphics[scale=0.3]{./Figures/coiled15puzzle2.png}
-\decoRule
 \caption[Coiled 15-puzzle]{Coiled 15-puzzle}
@@
 %-----------------------------------
-\subsection{Search Space \& Solvability}
+\Subsection{Search Space \& Solvability}
 
@@
 %-----------------------------------
-\subsection{Optimal Cost \& God's Number}
+\Subsection{Optimal Cost \& God's Number}
 
@@
 %-----------------------------------
-\section{Rubik's Cube}
+\Section{Rubik's Cube}
 
-\subsection{History -- from Erno Rubik to speed cubing competitions}
+\Subsection{History -- from Erno Rubik to speed cubing competitions}
 The second puzzle I will focus on is the well known Rubik's Cube (\textbf{RC}), invented in 1974 by Erno Rubik, originally known under the name of \textit{Magic Cube} (see \cite{RubiksWiki}).
@@
 \includegraphics[scale=0.3]{./Figures/RC333444}
-\decoRule
 \caption[Reducing RC]{Reducing a 4x4x4 \textbf{RC} to a 3x3x3 -- home made picture}
@@
 
-\subsection{Conventions: faces, moves and configurations}
+\Subsection{Conventions: faces, moves and configurations}
 In my code as well as in this write-up, I follow the conventions and notations of the \textbf{RC} literature and community as much as possible, I will refer to the faces as F (front), B (back), U (up), D (down), L (left) and R (right).
@@
 
-\subsubsection{3x3x3 Rubik's -- search space, solvability \& God's number}
+\Subsection{3x3x3 Rubik's -- search space, solvability \& God's number}
 
@@
 
-\subsubsection{2x2x2 Rubik's -- search space, solvability}
+\Subsection{2x2x2 Rubik's -- search space, solvability}
 
@@
 %-----------------------------------
-\section{Solving puzzles without human knowledge}
+\Section{Solving puzzles without human knowledge}
 \label{sec:WHK}

### 2022-09-10 latex/Abstract/Abstract.tex
update

@@
 \\
-Reinforcement Learning (\textbf{RL}), exposed with brilliant clarity in the Sutton book (\cite{Sutton1998}), has until recently known less success than we might have hoped for.
+Reinforcement Learning (\textbf{RL}), exposed with brilliant clarity in \cite{Sutton1998}, has until recently known less success than we might have hoped for.
 Its framework is very appealing and intuitive.
@@
 \\
-In this project, I will attempt to implement a variety of solvers combining $A^{*}$ search and heuristics, some of which will be handcrafted, others which I will train on randomly generated sequences of puzzles using (\textbf{DL}) or (\textbf{DRL}), to solve the 15-puzzle (and variations of different dimensions), as well as the Rubik's cube.
+In this project, I will attempt to implement a variety of solvers combining $A^{*}$ search and heuristics, some of which will be handcrafted, others which I will train on randomly generated sequences of puzzles using \textbf{DL}, \textbf{DRL} or \textbf{DQL}, to solve the 15-puzzle (and variations of different dimensions), as well as the Rubik's cube.
 \\
@@
 \\
-In the first chapter, I will quickly describe what I hope to get out of this project, both in terms of personal learning, as well as in terms of tangible results (solving some puzzles!).
-In chapter 2, I will do a quick recap of the different methods that I will use in the project.
-Chapter 3 will be dedicated to discussing the mathematics of the sliding puzzle (\textbf{SP}) and the (\textbf{RC}).
+In the first section, I will quickly describe what I hope to get out of this project, both in terms of personal learning, as well as in terms of tangible results (solving some puzzles!).
+In section 2, I will do a quick recap of the different methods that I will use in the project.
+Section 3 will be dedicated to discussing the mathematics of the sliding puzzle \textbf{SP} and the \textbf{RC}.
 I might throw a few random (but hopefully interesting) observations in there and give some references for the keen reader.
-I will then give an overview in chapter 4 of the code base I have developed - and put in the open on my github page (\cite{FB}) - to complete this project.
-Finally, in chapters 6 and 7, I will present all my various results on respectively the sliding puzzle and the Rubik's cube before offering some conclusion in chapter 8.
+I will then give an overview in section 4 of the code base I have developed - and put in the open on my github page (\cite{FB}) - to complete this project.
+Finally, in sections 6 and 7, I will present all my various results on respectively the sliding puzzle and the Rubik's cube before offering some conclusion in section 8.
 \\
-A first appendix details how to use the code base, from constructing puzzles, to learning and solving them using all the different methods I have implemented for this project.
-A second appendix describes a limitation I bumped into while using the kociemba library (an open source 3x3x3 \textbf{RC} solver), and how I circumvented it.
+A couple of appendix sections complement the main text.
+The first one details how to use the code base, from constructing puzzles, to learning and solving them using all the different methods I have implemented for this project.
+The second describes a limitation I bumped into while using the kociemba library (a well-known open source 3x3x3 \textbf{RC} solver), and how I circumvented it.

### 2022-09-10 latex/Chapters/Rubiks.tex
update

@@
 
-\label{Chapter5} % Change X to a consecutive number; for referencing this chapter elsewhere, use \ref{ChapterX}
+\label{sec:ResRubiks} % Change X to a consecutive number; for referencing this chapter elsewhere, use \ref{ChapterX}
 
@@
 %----------------------------------------------------------------------------------------
+
+In this second results section, I discuss my experiments on the Rubik's cube.
+I have run reasonably comprehensive experiments on the 2x2x2, throwing 6 of my solvers at the problem.
+As mentioned in \ref{sec:Puzzles}, the 2x2x2 \textbf{RC}, in spite of its deceivingly simple physical appearance, is actually not that much easier to solve.
+For instance, it takes me usually just under 1 minute, whereras the 3x3x3 takes me only double that time, just under 2 minutes.
+I am obviously a beginner in Rubik's cube handsolving, but my personal ratio of time-to-solve and perceived difficulty is in line with that of experienced speedcubers who might take and average of, say, 2-3 seconds for the 2x2x2 and 10 seconds for the 3x3x3.
+\\
+Computing wise, the 2x2x2 was of decent size for my purpose, with its 88 million configurations, so in particular it 
+\\
+\\
+Let me start with a table summarizing which solvers I have attempted to run on the 2 dimensions I have looked at:
 
@@
 \hline
-BFS   &   &        & x   &   \\
+BFS   &   &        &  x  &  x  \\
 \hline
-Kociemba   &   &      & x  &   \\
+Kociemba   &   &      &  x  &  x  \\
 \hline
-A$^{*}$[Kociemba]  &   &  & x & \\
+A$^{*}$[Kociemba]  &   &  &  x  &  x  \\
 \hline
-A$^{*}$[DL[A$^{*}$[Kociemba]]]  &   &  & x & \\
+A$^{*}$[DL[A$^{*}$[Kociemba]]]  &   &  &  x  &  x  \\
 \hline
@@
 \hline
-MCTS[DQL][c=0]  &   &  & x &  \\
-\hline
-MCTS[DQL][c=69]  &   & & x &  \\
-\hline
+%MCTS[DQL][c=0]  &   &  & x &  \\
+%\hline
+%MCTS[DQL][c=69]  &   & & x &  \\
+%\hline
 \end{tabular}
@@
 
-\section{2x2x2}
+\Section{2x2x2}
 
@@
 %\decoRule
-\caption[222RCPerformance]{Solvers' performance comparison 2x2x2 RC}
+\caption[222RCPerformance]{Solvers' performance comparison 2x2x2 \textbf{RC}}
 \label{fig:222RCPerformance}
 \end{figure}
-\vfill
-\end{landscape}
-\restoregeometry
 
@@
 %-----------------------------------
-\section{3x3x3}
+\Section{3x3x3}
+\label{sec:ResRubiks333}
 
@@
 
-\newgeometry{top=0mm, bottom=0mm, left=0mm, right=0mm}
-\begin{landscape}
-\centering\vspace*{\fill}
 \begin{figure}[H]
 \centering
-\includegraphics[scale=0.50]{./Figures/333RCPerformance.jpeg}
+\includegraphics[scale=0.50]{./Figures/333RCPerformance}
 %\decoRule
-\caption[333RCPerformance]{Solvers' performance comparison 3x3x3 RC}
+\caption[333RCPerformance]{Solvers' performance comparison 3x3x3 \textbf{RC}}
 \label{fig:333RCPerformance}
 \end{figure}
-\vfill
-\end{landscape}
-\restoregeometry

### 2022-09-10 rubiks/learners/deepreinforcementlearner.py

@@ -533,7 +533,7 @@ class DeepReinforcementLearner(Learner):
 
     @staticmethod
     def axis_label_format(what):
-        return what.upper().replace('_', ' ')
+        return what.upper().replace('_', ' ').replace('OVER', '/').replace(' PCT', '')
 
     def plot_learning(self):
         cls = self.__class__

### 2022-09-10 rubiks/scripts/rubiks333/perf.py

@@ -67,36 +67,15 @@ if '__main__' == __name__:
         exclude_solver_names = ['122seq_40shf', '500seq_12shf']
         performance(**locals())
     elif action_type == Solver.do_performance_test:
-        nb_samples = 20
-        nb_cpus = 4
-        performance(solver_type=Solver.astar,
-                    heuristic_type=Heuristic.deep_learning,
-                    learner_type=Learner.deep_reinforcement_learner,
-                    time_out=1200,
-                    add_perfect_shuffle=False,
-                    max_nb_shuffles=4,
-                    step_nb_shuffles=2,
-                    nb_sequences=10000, #500,
-                    nb_shuffles=12,
-                    nb_epochs=10000,
-                    layers_description=(600, 300, 100),
-                    training_data_every_epoch=False, #True,
-                    cap_target_at_network_count=False,
-                    one_hot_encoding=True,
-                    drop_out=0.,
-                    network_type=DeepLearning.fully_connected_net,
-                    learning_rate=1e-3,
-                    scheduler=DeepReinforcementLearner.exponential_scheduler,
-                    gamma_scheduler=0.9999,
-                    **locals())
-        exit()
+        nb_samples = 50
+        nb_cpus = 1
         performance(solver_type=Solver.astar,
                     heuristic_type=Heuristic.deep_learning,
                     learner_type=Learner.deep_learner,
-                    time_out=3600 + 7200,
+                    time_out=8 * 3600,
                     add_perfect_shuffle=False,
                     min_nb_shuffles=10,
-                    max_nb_shuffles=12,             #  ->  one more to re-try for nb_shuffle=10 #  -> try up to 12 to start with
+                    max_nb_shuffles=10,             #  ->  one more to re-try for nb_shuffle=10 #  -> try up to 12 to start with
                     step_nb_shuffles=2,
                     nb_sequences=100,
                     nb_shuffles=32,
@@ -112,6 +91,27 @@ if '__main__' == __name__:
                     gamma_scheduler=0.9999,
                     nb_shuffles_min=1,
                     nb_shuffles_max=32,
+                    **locals())
+        exit()
+        performance(solver_type=Solver.astar,
+                    heuristic_type=Heuristic.deep_learning,
+                    learner_type=Learner.deep_reinforcement_learner,
+                    time_out=7200,
+                    add_perfect_shuffle=False,
+                    max_nb_shuffles=6,
+                    step_nb_shuffles=2,
+                    nb_sequences=10000, #500,
[…]

### 2022-09-10 latex/Chapters/Puzzles.tex
update

@@
 \Subsection{2x2x2 Rubik's -- search space, solvability}
+\label{Theory:222RCSSS}
 
@@
 
+\Subsection{Rubik's -- multi-stage search}
+\label{Theory:RCMSS}
+Modern Rubik's domain-specific solvers (such as the Kociemba solver which I have used in this project as a comparison point) make use of a multi-stage search approach.
+Essentially, they just do what speedcubers do by hand, just much faster and better.
+The general idea is as follows: let us suppose we have a puzzle with a state space $\mathcal{S}$ of size $N$ (some large number) and one goal state $g$.
+Now imagine that we have discovered a much smaller subset of configurations $\mathcal{S}_{1} \in \mathcal{S}$ (say of size $n \ll N$) such that $g \in \mathcal{S}_{1}$ and such that it is \textit{easy} to check whether or not a given configuration $p \in \mathcal{S}$ belongs to $\mathcal{S}_{1}$.
+In addition, suppose we also have discovered a set of algorithms $\mathcal{A}_{1}$ (sequence of actions) under which elements of $\mathcal{S}_{1}$ are stable and which spans $\mathcal{S}_{1}$.
+Finally, suppose that we have a set of algorithms $\mathcal{A}_{2}$ which allow us to (provably) take any given configuration $p \in \mathcal{S}$ into $\mathcal{S}_{1}$ in a bounded number of steps.
+A viable strategy to solve any puzzle then becomes:
+\begin{titlemize}{Multi-stage search}
+\item If $p \in \mathcal{S}_{1}$, set $a_{2}$ to be the identity move in $\mathcal{S}$, and jump to step 3.
+\item Apply a search algorithm (e.g.
+A$^{*}$ - or some variant such as IDA$^{*}$ (\cite{IDAWiki}) - restricting eligible actions to $\mathcal{A}_{2}$, until we find a sequence of actions $a_{2}$ such that $a_{2}(p) \in \mathcal{S}_{1}$
+\item Apply our search algorithm to $a_{2}(p)$, restricting eligible actions to $\mathcal{A}_{1}$, until we find a sequence of actions $a_{1}$ such that $a_{1}(a_{2}(p)) = g$.
+\item $a_{1} \circ a_{2}$ is our multi-stage solution
+\end{titlemize}
+Notice that this is exactly how beginners and advanced speedcubers solve the Rubik's cube, except they usually have more than 2 stages as described above.
+$S_{1}$ might for instance be the set of cubes where both lower layers are solved, the cross on the top layer as well as all corners are in their respective places (but maybe not oriented as expected), and then they search within a well known and small set of algorithms how to get to the goal state.
+A previous stage might consist in $S_{2}$, the set of configurations where the two layers are solved, but the top layer is random, and then they will search for a sequence of moves to get into $S_{1}$, etc \dots
+
 %-----------------------------------
@@
 I would put them broadly in four different categories, by decreasing level of human knowledge that has been assumed and handcrafted in them.
-Notice that this human-knowledge-required scale is obviously somewhat arbitrary, but good to keep in mind in later (and in particular results) sections.
+Notice that my human-knowledge-required scale is obviously somewhat arbitrary, but good to keep in mind in later (and in particular results) sections.
 
@@
 At 4 on the human-knowledge-required scale, I would put my NaiveSolver (handcrafted to solve the \textbf{SP}) as well as the hkociemba (2x2x2) and kociemba (3x3x3) open source implementations which I have integrated in my KociembaSolver.
-They make heavy use of specific knowledge about the two puzzles, not only about their mechanics, but also about how to actually (provably) solve them.
+They make heavy use of specific knowledge about the two puzzles (see \ref{Theory:RCMSS}), not only about their mechanics, but also about how to actually (and provably) solve them.
 

### 2022-09-10 latex/Chapters/Rubiks.tex
update

@@
 \Section{2x2x2}
+All the following experiments on the 2x2x2 \textbf{RC} have been run using my Solver's \textit{performance\_test} (see \ref{sec:codesolvers}): for each level of difficulty, the test present the solvers with the same 100 random configurations (for fairness and to increase the significance of the results).
+The levels of difficulty (number of scrambles from goal) used here go from 2 to 20 in step of 2.
+The tests were also run with perfect shuffling (difficulty = $\infty$ on the graphs), corresponding to taking scrambling to $\infty$ as explained in \ref{Theory:222RCSSS}).
 
-1000 seq 30 shuffles training each target network update seen 1.1\% of all puzzles
-199 of the 200 perfectly shuffled got solved well under 1h (median 83 seconds), only one outlier took about 67 minutes.
+\Subsection{BFS}
+BFS is the only one for which I used a step of 1 on the difficuly and stopped at 5 as it was clear it would not go any further.
+Given that the run time and expanded nodes are clearly going to increase linearly, and give where difficulty of 5 took this solver already, there was no point going any further!
+
+\Subsection{Kociemba, A$^{*}$[Kociemba], A$^{*}$[DL[A$^{*}$[Kociemba]]]}
+Next, let me discuss Kociemba (\ref{}) and associated experiments.
+We obviously expect Kociemba to be by far the fastest solver since it is handcrafted with much specific knowledge about the Rubik's cube.
+
+\Subsection{Deep reinforcement learner}
+\label{sec:S33DRL}
+Next I have run the DeepReinforcementLearner with the following main parameters:
+\begin{itemize}
+\item Update the target network every 1,000 epochs (\textit{update\_target\_network\_frequency=1000}) or when the MSE falls below 10 basis points of the maximum target cost-to-go (\textit{update\_target\_network\_threshold=1e-3})
+\item Run for a maximum of 11,000 epochs (\textit{nb\_epochs=11000}) or when the maximum target so far has not been increasing by more than 1\% in the last 5,500 epochs (\textit{max\_target\_uptick=1e-2} and \textit{max\_target\_not\_increasing\_epochs\_pct=0.5})
+\item At every update of the target network (\textit{training\_data\_every\_epoch=False}), generate 100 sequences of puzzles (\textit{nb\_sequences=100}), each comprised of a 50-scramble puzzle back to goal state, with all intermediary puzzles along the path (\textit{nb\_shuffles=50}).
+\item Use a fully connected network (\textit{network\_type=fully\_connected\_net}) with three hidden layers of size 600, 300, 100 (\textit{layers\_description=(600,300,100)}) and use torch.optim.RMSProp as optimiser (\textit{optimiser=rms\_prop}) with an exponential scheduler (\textit{scheduler=exponential\_scheduler}) with gamma 0.999 (\textit{gamma\_scheduler=0.9999}).
+\end{itemize}
+Below I show some of the DeepReinforcementLearner's quantities tracked over the epochs (see \ref{fig:33SPDeepReinforcementLearning}). I have shown only the following quantities:
+\begin{itemize}
+\item learning rate, which decreases due to the scheduler, but gets reset at each network update.
+\item max target.
+One interesting dynamic during learning is that due to the way targets are constructed using value-iteration update, and starting from an all 0 network, the max target only increases by about 1 at best at every network update.
+That is, the first network only learns how to differentiate goal from non-goal, then the second one learns how to differentiate goals, cost 1 and cost 2+ in some sense.
+This is why I have introduced the first set of parameters discussed just above.
+The first one makes sure we do not waste time initially training the left-hand-side network when training is easy (as consist of just goal versus not goal).
+The second ones make sure that we stop training altogether when the max target does not grow anymore.
+This is because we do not know a-priori what the God number of a given puzzle/dimension is, but let's assume for the sake of the argument that it is 15.
+Then that means that after 15 updates of the target network we might not really get so much benefit by keeping training and value-iterating.
+\item the MSE loss
+\item the MSE loss divided by the max target.
+The exit criterion \textit{update\_target\_network\_threshold} is based on that quantity, since a loss of e.g.
+0.01 means little if the we do not compare it to the possible range that the cost-to-go can take.
+\item The percentage of all the possible puzzles of that dimension \textit{seen} by the DeepReinforcementLearner.
+This is an interesting quantity, as it tells us later on when testing out of sample and comparing with other solvers, how well the solvers are able to generalise.
+In this case (3x3) we can see that the DeepReinforcementLearner quickly has seen a very large percentage of all puzzles.
+That being said, it is good to keep in mind that \textit{seen} does not mean really much here, since the learning is unsupervised.
+This is quite in contrast to the DeepLearner, for which puzzles seen come along with the cost-to-go computed by a \textit{teacher}, usually an optimal or at least efficient solver.
+\end{itemize}
+
+\begin{figure}[H]
+\centering
+\includegraphics[align=c, scale=0.42]{./Figures/33SPDeepReinforcementLearning}
+\caption[33SPDeepReinforcementLearning]{\textbf{DRL} of the 3x3 \textbf{SP}}
+\label{fig:33SPDeepReinforcementLearning}
+\end{figure}
+
+Notice that the DeepLearner and DeepQLearner have very similar sets of parameters and track similar metrics during their learning for later display, but I shall not detail it here since there is not much extra insight to be had compared to the DeepReinforcementLearning example we just went through.
+
+\Subsection{Solvers' comparison}
 

### 2022-09-10 rubiks/scripts/rubiks22/perf.py

@@ -71,13 +71,13 @@ if '__main__' == __name__:
         performance(**locals())
     elif action_type == Solver.do_performance_test:
         nb_samples = 100
-        nb_cpus = 8
-        for counter in [14]: # range(14, 23, 2):
+        nb_cpus = 10
+        for counter in range(14, 23, 2):
             logger.log_info('counter: ', counter)
             performance(solver_type=Solver.astar,
                         heuristic_type=Heuristic.deep_q_learning,
                         learner_type=Learner.deep_q_learner,
-                        time_out=300,
+                        time_out=600,
                         add_perfect_shuffle=counter == 22,
                         min_nb_shuffles=0 if counter > 20 else counter,
                         max_nb_shuffles=0 if counter > 20 else counter,

### 2022-09-10 rubiks/solvers/solver.py

@@ -839,6 +839,9 @@ class Solver(Factory, Puzzled, Loggable, metaclass=ABCMeta):
                                    cls.median_cost,
                                    cls.max_cost,
                                    cls.median_run_time,
+                                   cls.pct_solved,
+                                   cls.pct_optimal,
+                                   cls.nb_timeout,
                                    ]].groupby(cls.solver_name).mean().\
             reset_index(drop=False).sort_values(cls.optimality_score)
         self.log_info(pformat(performance, floatfmt='.2f', showindex=True))

### 2022-09-10 latex/Chapters/Conclusion.tex
update

--- 2022-09-03 latex/Chapters/Conclusion.tex
@@
 
-\textbf{Learning}
-Learnt about puzzles, and also even took the time to learn a couple of sequences of moves that are sufficient to solve \textbf{RC}, 2x2x2 in labout one minute, 3x3x3 in about 2 minutes.
-\\
-\\
-\textbf{Implementation}
-Implemented a bunch of solvers. Ideally should do in C++ for speed and multithreaded (esp for MCTS)
-\\
-\\
-\textbf{DRL \& DQL}
-Quite impressive, without too much effort, how well they perform.
-However, I feel to really get more out of them (higher dimensions in particular) would need careful tuning of the convergence criteria.
-\\
-\\
-\textbf{Professional issues}
-\\
-\\
+\Section{Learning}
+Working on this project has been overall a lot of fun.
+Even setting aside implementation and computer-experimentation for a moment, I have learnt a great deal about puzzles in general and how to approach thinking about them in a more structured fashion.
+Specifically about the Rubik's cube, I have even spent the time to learn a few algorithms that are sufficient to solve them by hand.
+I am now able to solve the 2x2x2 under one minute, 3x3x3 usually around 2 minutes and the 4x4x4 in about 5 minutes.
+More specifically regarding the techniques discussed in this project, the principles that underly uninformed and informed search, A$^{*}$ and heuristics, multi-stage solvers, and of course the various deep learning methods I have played with (\textbf{DL}, \textbf{DRL} and \textbf{DQL}) all extend to many other puzzle and dimensions and are a great new tool-set to add to my knowledge and experience.
+
+\Section{Implementation}
+I am quite satistified to have had the time to implement and play with so many solvers, from handcrafting my own \textbf{SP} \textit{naive} solver, to implementing \textbf{BFS}, \textbf{DFS}, A$^{*}$, admissible heuristics such as Manhattan for the \textbf{SP} and improving on it via linear constraints penalty, and of course all the machinery necessary to fit, train and exploit some \textbf{DL}, \textbf{DRL} and \textbf{DQL} heuristics to guide search algorithms such as  A$^{*}$ and \textbf{MCTS}.
+Ideally I would have liked to implement things in C++ for speed and multithreading (especially for \textbf{MCTS}) but time was limited so I decided to prioritize breadth and ease of experimentation over depth and efficiency of implementation.
+The choice of Python definitely made it much easier to integrate with many of the ecosystem's libraries (matplotlib, pandas and pytorch in particular).
+
+\Section{Deep Reinforcement-Q Learning}
+Even though the Deep Reinforcement Learning and Deep Q Learning models were quite capricious at time to train.
+Choosing the convergence mechanisms and criteria, the interplay of generation of new training data with the inner and outer loop of the \textbf{DRL} and \textbf{DQL} algorithms (see \ref{sec:TheoryDLDRL} and \ref{sec:TheoryDQL}) seemed more of an art than a sicence at times).
+Obviously a lot more tuning, as well as much longer training time and computing resources would have been necessary to make these techniques more competitive at higher dimensions, but overall I remain extremely impressed by how, all things considered - and given how daunting the project seemed at first, it took me little effort to train successful \textbf{DRL} and \textbf{DQL} models capable of solving puzzles of such large state and action spaces.
+
+\Section{Professional issues}
+Bug in Kociemba and software design
 See \cite{ManyWorlds}
-\\
-\\
 See \cite{https://doi.org/10.1111/1758-5899.12718} and \cite{Bostrom2014}

### 2022-09-10 latex/Chapters/SlidingPuzzle.tex
update

@@
 
-\section{Low dimension}
+\Section{Low dimension}
 \label{sec:SPLowDimension}
 
-\subsection{God numbers and hardest puzzles}
+\Subsection{God numbers and hardest puzzles}
 
@@
 
-\subsection{Manhattan heuristic}
+\Subsection{Manhattan heuristic}
 \label{MHComp}
@@
 %-----------------------------------
-\section{Intermediary case - 3x3}
+\Section{Intermediary case - 3x3}
 \label{sec:S33}
@@
 
-\subsection{Perfect learner}
+\Subsection{Perfect learner}
 \label{sec:S33PL}
@@
 \begin{figure}[H]
-\centering
-\includegraphics[scale=0.42]{./Figures/33SPPerfectLearning}
+  \noindent
+  \makebox[\textwidth]{\includegraphics[scale=0.4]{./Figures/33SPPerfectLearning}}
   \caption[33SPPerfectLearning]{Perfect Learning of the 3x3 \textbf{SP}}
@@
 
-\subsection{Deep reinforcement learner}
+\Subsection{Deep reinforcement learner}
 \label{sec:S33DRL}
+
+\begin{comment}
+
 Next I have run the DeepReinforcementLearner with the following main parameters:
@@
 
-\subsection{MCTS DQL}
+\end{comment}
+
+\Subsection{MCTS DQL}
 \label{sec:MTCSHyperParamsEffect}
+In this subsection I explore the effect on \textbf{MCTS} of the trade-off hyper-parameter $c$ as well as of trimming the resulting tree via \textbf{BFS}.
+(see the threory section \ref{sec:TheoryMCTS} as well as the implementation section \ref{sec:codesolvers} for more details).
+To that effect, I have run my Solver performance test with four different values of $c$ (0, 1, 10 and 100), and in each case, with and without \textbf{BFS} trimming.
+A$^{*}[Manhattan++]$ is also presented to have an optimal solver benchmark.
+For each level of difficulty, the test present the solvers with the same 150 random configurations (for fairness and to increase the significance of the results).
+The levels of difficulty (number of scrambles from goal) used here go from 5 to 50 in step of 5.
+The tests were also run with perfect shuffling (difficulty = $\infty$ on the graphs), corresponding to taking scrambling to $\infty$ as explained in \ref{Theory:SPSSS}).
+I used a \textit{time\_out} of 5 minutes, which none of the solvers hit on any of the (1,650) configurations they were presented with.
+The detailed results (percentage of configurations solved, optimality score, median and max cost, run time and expanded nodes) are shown on the next page \ref{fig:33SPPerformanceMCTS}, but let me make here some remarks on those results:
+\begin{itemize}
+\item $c=0$ makes no exploration as it simply follows the path of lowest cost-to-go.
+As $c$ increases, we expect more exploration, and therefore higher run time, more expanded nodes, and higher quality solutions.
+This is what we observe, where $c=0$ converges (for difficulty=$\infty$) to solutions on average three times longer than the optimal solver but take not much longer to obtain.
+At the other extreme, $c=100$ takes almost 2 orders of magnitude more time to obtain a solution, but these solutions are only 20-25\% longer than that of the optimal solver
+\item trimming via \textbf{BFS} takes hardly any time, and improves solutions' lengths by about 10\% at high values of $c$.
+At low value of $c$ there is little benefit as expected (since the trees are deep but not bushy at all and have much fewer leaves and branches).
+\item The quality of solutions is not amazing, but then given the algorithm has not really any mechanism to guarantee optimality this is not very surprising. 
+\item Notice that the number of expanded nodes can be compared within \textbf{MCTS} but given the algorithm is totally different to A$^{*}$ it makes little sense to compare between the different families of solvers.
+Also one point to note is that \textbf{MCTS} would be fairly easy to multithread (within the same process) using an appropriate language such as C++, and this is what \cite{https://doi.org/10.48550/arxiv.1805.07470} have done, and probably why they get considerably faster and better results.
+\item Despite the name \textit{Monte Carlo}, there is no randomness (or even peudo-randomness) in this algorithm, since the paths generated at each iteration are really just a deterministic function of the output of the \textbf{DQL} network (or QHeuristic in my code-jargon).
+\end{itemize}
 
 \begin{figure}[H]
-\centering
-\includegraphics[align=c, scale=0.6]{./Figures/33SPPerformanceMCTS}
+  \noindent
+  \makebox[\textwidth]{\includegraphics[scale=0.6]{./Figures/33SPPerformanceMCTS}}
 \caption[33SPPerformanceMCTS]{\textbf{MCTS} tuning 3x3 \textbf{SP}}
@@
 
-\subsection{Solvers' comparison}
+\Subsection{Solvers' comparison}
 \label{ssec:33SPSC}
@@
 \begin{figure}[H]
-\centering
-\includegraphics[scale=0.65]{./Figures/33SPPerformance}
-%\decoRule
+  \noindent
+  \makebox[\textwidth]{\includegraphics[scale=0.6]{./Figures/33SPPerformance}}
   \caption[33SPPerformance]{Solvers' performance comparison 3x3 \textbf{SP}}
@@
 
-\subsection{Solving the hardest 3x3 problem}
+\Subsection{Solving the hardest 3x3 problem}
 \label{ssec:33SPHard}
@@
 %-----------------------------------
-\section{3x4}
+\Section{3x4}
 
@@
 
-\section{4x4}
+\Section{4x4}
 

### 2022-09-10 latex/Chapters/Conclusion.tex
update

@@
 \Section{Implementation}
-I am quite satistified to have had the time to implement and play with so many solvers, from handcrafting my own \textbf{SP} \textit{naive} solver, to implementing \textbf{BFS}, \textbf{DFS}, A$^{*}$, admissible heuristics such as Manhattan for the \textbf{SP} and improving on it via linear constraints penalty, and of course all the machinery necessary to fit, train and exploit some \textbf{DL}, \textbf{DRL} and \textbf{DQL} heuristics to guide search algorithms such as  A$^{*}$ and \textbf{MCTS}.
+I am quite satistified to have had the time to implement and play with so many solvers and methods, spanning so many different \textbf{CS} and \textbf{AI} paradigms, from handcrafting my own \textbf{SP} \textit{naive} solver, playing with Kociemba, to implementing \textbf{BFS}, \textbf{DFS}, A$^{*}$, admissible heuristics such as Manhattan for the \textbf{SP} and improving on it via linear constraints penalty, and of course all the machinery necessary to fit, train and exploit some \textbf{DL}, \textbf{DRL} and \textbf{DQL} heuristics to guide search algorithms such as  A$^{*}$ and \textbf{MCTS}.
 Ideally I would have liked to implement things in C++ for speed and multithreading (especially for \textbf{MCTS}) but time was limited so I decided to prioritize breadth and ease of experimentation over depth and efficiency of implementation.
@@
 \Section{Deep Reinforcement-Q Learning}
-Even though the Deep Reinforcement Learning and Deep Q Learning models were quite capricious at time to train.
-Choosing the convergence mechanisms and criteria, the interplay of generation of new training data with the inner and outer loop of the \textbf{DRL} and \textbf{DQL} algorithms (see \ref{sec:TheoryDLDRL} and \ref{sec:TheoryDQL}) seemed more of an art than a sicence at times).
-Obviously a lot more tuning, as well as much longer training time and computing resources would have been necessary to make these techniques more competitive at higher dimensions, but overall I remain extremely impressed by how, all things considered - and given how daunting the project seemed at first, it took me little effort to train successful \textbf{DRL} and \textbf{DQL} models capable of solving puzzles of such large state and action spaces.
+It will not come as a surprise to anyone who has played with them that the Deep Reinforcement Learning and Deep Q Learning models can be quite capricious to train: choosing the convergence mechanisms and criteria, the interplay of generation of new training data with the inner and outer loop of the \textbf{DRL} and \textbf{DQL} algorithms (see \ref{sec:TheoryDLDRL} and \ref{sec:TheoryDQL}) seemed more of an art than a sicence at times.
+Obviously a lot more tuning, as well as much longer training time and computing resources would have been necessary to make these techniques more competitive at higher dimensions, but overall I remain extremely impressed by how, all things considered - and given how daunting the project seemed at first - it took me little effort to train successful \textbf{DRL} and \textbf{DQL} models capable of solving puzzles with such large state and action spaces.
 
 \Section{Professional issues}
-Bug in Kociemba and software design
+Bug in Kociemba and software design\\
+AI safety 
 See \cite{ManyWorlds}

### 2022-09-10 latex/Chapters/Conclusion.tex
update

@@
 \Section{Deep Reinforcement-Q Learning}
-It will not come as a surprise to anyone who has played with them that the Deep Reinforcement Learning and Deep Q Learning models can be quite capricious to train: choosing the convergence mechanisms and criteria, the interplay of generation of new training data with the inner and outer loop of the \textbf{DRL} and \textbf{DQL} algorithms (see \ref{sec:TheoryDLDRL} and \ref{sec:TheoryDQL}) seemed more of an art than a sicence at times.
+It will not come as a surprise to anyone who has played with them that Deep Reinforcement Learning and Deep Q Learning models can be quite capricious to train: choosing the convergence mechanisms and criteria, the interplay of generation of new training data with the inner and outer loop of the \textbf{DRL} and \textbf{DQL} algorithms (see \ref{sec:TheoryDLDRL} and \ref{sec:TheoryDQL}) seemed more of an art than a sicence at times.
 Obviously a lot more tuning, as well as much longer training time and computing resources would have been necessary to make these techniques more competitive at higher dimensions, but overall I remain extremely impressed by how, all things considered - and given how daunting the project seemed at first - it took me little effort to train successful \textbf{DRL} and \textbf{DQL} models capable of solving puzzles with such large state and action spaces.

### 2022-09-11 latex/Chapters/DRLSearch.tex
make both RHUL template and my own thesis format coexist .... I d like to have both formats
update

@@
 
-\section{Graphs Search \& Heuristics}
+\Section{Graphs Search \& Heuristics}
 
 \label{GSH}
+\label{sec:TheoryBFSDFS}
+\label{sec:TheoryAStar}
 
@@
 optimal).
-\textbf{DFS} is obvisouly not optimal as it could very well find a solution very deep in the graphs while expanding the very first branch, not realising that a solution was one level away from the start on another branch!
+\textbf{DFS} is obvisouly not optimal as it could very well find a solution very deep in the graph while expanding the very first branch, not realising that a solution was one level away from the root on the next branch!
 \\
@@
 \teal
-\paragraph{}{\textbf{pseudo code -- BFS \& DFS}}
-\label{sec:TheoryBFSDFS}
-\begin{pseudocode}
-###############################################################################
-def blind_search(initial_node, time_out, search_type=Search.BFS):
-   """ pseudo code for BFS/DFS """
-    initialize_time_out(time_out)
-    if initial_node.is_goal():
-        return initial_node
-    explored = set()
-    frontier = [initial_node]
-    while True:
-        check_time_out() # -> raise TimeoutError if appropriate
-        if frontier.empty():
-            return None # -> Failure to find a solution
-       # FIFO/LIFO for BFS/DFS
-        node = pop_front(frontier) if search_type is Search.BFS else pop_back(frontier)
-        explored.add(node)
-        for child in node.children() and not in frontier.union(explored):
-            if child.is_goal():
-                return child
-            frontier.add(child)
-###############################################################################
-\end{pseudocode}
+\begin{algorithm}[H]
+\caption{Uninformed Graphs Search -- BFS \& DFS}\label{alg:TheoryBFSDFS}
+\begin{algorithmic}
+\Function{BlindSearch}{$root, time\_out, search\_type=Search.BFS$}
+\State\Call{InitialiseTimeOut}{$time\_out$}
+\If{\Call{IsGoal}{$root$}}
+  \Return{$root$}
+\EndIf
+\State $explored \gets \{\}$
+\State $frontier \gets [root]$
+\While{true}
+\State\Call{CheckTimeOut}{void}
+ \If{\Call{Empty}{frontier}}
+  \Return{$failure$}
+ \EndIf
+ \If{\Call{Equal}{$search\_type, Search.BFS$}}
+ \State $node \gets \Call{PopFront}{$frontier$}$
+ \Else
+ \State $node \gets \Call{PopBack}{$frontier$}$ \teal \algorithmiccomment{$Search.DFS$} \black
+ \EndIf
+\State \Call{Add}{$explored, node$}
+\For{$child \in $ \Call{Children}{$node$}}
+\If{\Call{IsGoal}{$child$}}
+\Return{$child$}
+\EndIf
+\State \Call{Add}{$frontier, child$}
+\EndFor
+\EndWhile
+\EndFunction
+\end{algorithmic}
+\end{algorithm}
 \black
-\textit{Informed} strategies, on the other hand, exploit domain-specific knowledge.
-One very popular such strategy is \textbf{A$^{*}$} (see \cite{DBLP:journals/jacm/DechterP85}), which always first examine the node of smallest expected total cost (\textbf{f}), itself the sum of the cost-so-far from initial to current node (\textbf{g}) plus the expected cost-to-go from current node to solution (\textbf{h}).
+\noindent \textit{Informed} strategies, on the other hand, exploit domain-specific knowledge.
+One popular such strategy is \textbf{A$^{*}$} (see \cite{DBLP:journals/jacm/DechterP85}), which first examines the node of smallest expected total cost (\textbf{f}), itself the sum of the cost-so-far from initial to current node (\textbf{g}) plus the expected cost-to-go from current node to goal (\textbf{h}).
 The expected cost-to-go \textbf{h} is often called a \textit{heuristic}.
 The better the heuristic is, the better \textbf{A$^{*}$} usually performs.
-An important property of heuristics is \textit{admissibility}.
-In short, admissible heuristics never over-estimate the real cost-to-go, and can easily be shown to render \textbf{A$^{*}$} optimal.
+An important property of heuristics is \textit{admissibility}: admissible heuristics never over-estimate the real cost-to-go, and can easily be shown to render \textbf{A$^{*}$} optimal.
 \teal
-\paragraph{}{\textbf{pseudo code -- A$^{*}$}}
-\label{sec:TheoryAStar}
-\begin{pseudocode}
-###############################################################################
-def A*(initial_node, time_out, heuristic):
-    """ pseudo code for A*
-    g = cost from initial_node to a current node
-    h = heuristic (expected remaining cost from current node to goal)
-    """
-    initialize_time_out(time_out)
-    node = initial_node
-    cost = heuristic(node) # g = 0
-    frontier = sorted_multi_container({cost: {node}})
-    explored = set()
-    while True:
-        check_time_out()
-        if frontier.empty():
-            return None # -> Failure to find a solution
-        (cost, node) = frontier.pop_smallest() # smallest total expected cost
-        if node.is_goal():
-            return node
-        explored.add(node)
-        for child in node.children(): # -> children have g = node.g + 1
-            child_cost = child.g + heuristic(child)
-            if child in frontier and child_cost < cost:
-                frontier.update_cost(child, child_cost)
-            elif child not in explored:
-                frontier.add(child, child_cost)
-###############################################################################
-\end{pseudocode}
+\begin{algorithm}[H]
+\caption{Informed Graphs Search -- A$^{*}$}\label{alg:TheoryAStar}
+\begin{algorithmic}
+\Function{A$^{*}$}{$root, time\_out, H$} \teal \algorithmiccomment{H -- cost-to-go heuristic} \black
+\State\Call{InitialiseTimeOut}{$time\_out$}
+\State $f \gets 0 + \Call{H}{$node$}$
+\State $explored \gets \{\}$
+\State $frontier \gets \Call{SortedMultiMap}{\{cost: node\}}$
+\While{true}
+\State\Call{CheckTimeOut}{void}
+ \If{\Call{Empty}{frontier}}
+  \Return{$failure$}
+ \EndIf
+\State $(cost, node) = \Call{PopSmallest}{frontier}$
+\If{\Call{IsGoal}{$node$}}
+\Return{$node$}
+\EndIf
+\State \Call{Add}{$explored, node$}
+\For{$transition, child \in $ \Call{Children}{$node$}}
+	\State $child\_cost \gets \Call{Cost}{transition} + \Call{H}{child}$
+            \If{$child \in frontier \ and \ child\_cost \le cost$} \\
+             \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \Call{UpdateCost}{$frontier, child, child\_cost$}
+            \ElsIf{$child \notin explored$}
+                \Call{Add}{$frontier, child, child\_cost$}
+	\EndIf
+\EndFor
+\EndWhile
+\EndFunction
+\end{algorithmic}
+\end{algorithm}
 \black
@@
 
-\section{Reinforcement Learning}
+\Section{Reinforcement Learning}
 \label{sec:RLTheory}
@@
 In finite state and transitions spaces, the value function can be remarkably computed by a rather mechanical and magic-like procedure called value-iteration (kind of the equivalent of the Bellman equation from optimal control), where each state $s$'s value is iteratively refined by combining the value of $s$'s reachable children states.
+
 \teal
-\paragraph{}{\textbf{pseudo code -- \textbf{RL} Value Iteration}}
-\begin{pseudocode}
-###############################################################################
-def value_iteration(states):
-    """ pseudo code for RL value iteration """
-    V = {state: inf for state in states}
-    change = True
-    while change:
-        change = False
-        for state, cost in V:
-            V[state] = min(V[child] + t_cost for child, t_cost in state.children())
-            change |= V[state] < cost
-    return V 
-###############################################################################
-\end{pseudocode}
+\begin{algorithm}[H]
+\caption{Reinforcement Learning -- Value Iteration}\label{alg:RLTheory}
+\begin{algorithmic}
+\Function{ValueIteration}{$states$} \teal \algorithmiccomment{states -- generator of states} \black
+\State $V \gets \{state: \infty \ for \ state \in states\}$
+\State $change \gets true$
+\While{$change$}
+\State $change \gets False$
+\For{$state, cost \in states$}
+	\State $V[state] \gets \infty$
+	\For{$child, t\_cost \in \Call{Children}{state}$}
+	\State $V[state] \gets \Call{Min}{V[state], V[child] + t\_cost}$
+	\EndFor
+           \If{$V[state] \neq cost$}
+		\State{$change \gets true$}
+	\EndIf
+\EndFor
+\EndWhile
+\Return{$V$}
+\EndFunction
+\end{algorithmic}
+\end{algorithm}
 \black
@@
 
-\section{Deep Learning}
+\Section{Deep Learning}
 
@@
 
-\section{DL and DLR Heuristics}
+\Section{DL and DLR Heuristics}
 \label{sec:TheoryDLDRL}
@@
 In very simplified pseudo code, the procedure looks like the following:
+
+\newlength\mylen
+\settowidth\mylen{DeepLearningHeuristic(}
 \teal
-\paragraph{}{\textbf{pseudo code -- \textbf{DL} heuristics}}
-\begin{pseudocode}
-###############################################################################
-def deep_learning(puzzle_type,
-                  teacher_heuristic,
-                  loss_function,
-                  network_architecture):
-    """ pseudo code for learning a puzzle via deep learning and a teacher heuristic """
-    target_value_function = {state: teacher_heuristic(state)
-                             for state in generate_random_states(puzzle_type)}
-    dl_heuristic = train_neural_network(target_value_function,
-                                        loss_function,
-                                        network_architecture)
-    return dl_heuristic
-###############################################################################
-\end{pseudocode}
+\begin{algorithm}[H]
+\caption{Deep Learning -- Heuristic}\label{alg:TheoryDLDRL}
+\begin{algorithmic}
+\Function{DeepLearningHeuristic}{**args}
+\State{$puzzle\_type \gets \Call{FromParams}{**args}$}
+\State{$n\_puzzles \gets$ \Call{FromParams}{**args}}
+\State{$H \gets \Call{FromParams}{**args}$} \teal{\algorithmiccomment{H -- teacher heuristic to extrapolate via \textbf{DL}}}\black
+\State{$loss\_function \gets \Call{FromParams}{**args}$}
+\State{$net\_architecture \gets \Call{FromParams}{**args}$}
+\State $V = \{state: \Call{H}{state} \ for \ state \in \Call{GenerateRandomStates}{puzzle\_type, n\_puzzles}\}$
+\ \ \ \ \Return{$\Call{TrainNeuralNetwork}{V, loss\_function, net\_architecture}$}
+\EndFunction
+\end{algorithmic}
+\end{algorithm}
 \black
@@
 \\
+
 \teal
-\paragraph{}{\textbf{pseudo code -- \textbf{DRL} heuristics}}
-\begin{pseudocode}
-###############################################################################
-def deep_reinforcement_learning(puzzle_type,
-                                loss_function,
-                                max_epochs,
-                                target_network_update_criteria,
-                                puzzles_generation_criteria,
-                                network_architecture):
-    """ pseudo code for learning a puzzle via deep reinforcement learning """
-    net = get_network(puzzle_type, network_architecture)
-    target_net = get_network(puzzle_type, network_architecture)
-    epoch = 0
-    puzzles = list()
-    while epoch < max_epochs and other_convergence_criteria(network):
-        epoch += 1
-        """ Generate a new bunch of puzzles """
-        if puzzles_generation_criteria(puzzles):
-            puzzles = generate_random_states(puzzle_type)
-        V = dict()
-        """ Compute their updated cost via value iteration ... 
-             all moves are assumed to have cost 1 """
-        for puzzle in puzzles:
-            if puzzle.is_goal():
-                V[puzzle] = 0
-            else:
-                V[puzzle] = 1 + min(target_net(child) for child in puzzle.children()))
-        """ Train lhs network to approximate rhs target network better
-        i.e. perform a forward / backward-propagation update
-        """
-        network = train_neural_net(V,
-                                   loss_function,
-                                   network_architecture)
-        """ Update the target network if criteria are met 
-        (e.g. epochs, convergence, no progress, etc...) """
-        if target_network_update_criteria(net, target_net):
-            target_net = copy(net)
-    return net
-###############################################################################
-\end{pseudocode}
+\begin{algorithm}[H]
+\caption{Deep Reinforcement Learning -- Heuristic}\label{alg:TheoryDRL}
+\begin{algorithmic}
+\Function{DeepReinforcementLearningHeuristic}{**args}
+\State{$puzzle\_type \gets \Call{FromParams}{**args}$}
+\State{$ puzzles\_gen\_params \gets \Call{FromParams}{**args}$}
+\State{$puzzles\_generation\_criteria \gets$ \Call{FromParams}{**args}}
+\State{$n\_puzzles \gets$ \Call{FromParams}{**args}}
+\State{$target\_network\_update\_params \gets \Call{FromParams}{**args}$}
+\State{$max\_epochs \gets \Call{FromParams}{**args}$}
+\State{$loss\_function \gets \Call{FromParams}{**args}$}
+\State{$net\_architecture \gets \Call{FromParams}{**args}$}
+\\ \teal{\algorithmiccomment{initial network and target network}}\black
+\State {$net \gets \Call{GetNetwork}{puzzle\_type, network\_architecture}$}
+\State {$target\_net \gets \Call{GetNetwork}{puzzle\_type, network\_architecture}$}
+\State{$epoch \gets 0$}
+\State{$puzzles \gets null$}
+
+\While{epoch < max\_epochs \&\& other\_convergence\_criteria(network)}
+        \State{$epoch \gets epoch + 1$}
+\\ \teal{\algorithmiccomment{generate \textit{training} puzzles if criteria met}}\black
+        \If{puzzles\_generation\_criteria($puzzles$)}
+            \State{$puzzles \gets generate\_random\_states(puzzle\_type)$}
+        \EndIf
+        \State{$V \gets dict()$}
+        \For{$puzzle \in puzzles}$
+            \State{$V[puzzle] \gets \infty$}
+            \If \Call{IsGoal}{$puzzle$}:
+                \State{$V[puzzle] \gets 0$}
+            \Else:
+	     \For{$transition, child \in \Call{Children}{$puzzle$}$}
+\\ \teal{\algorithmiccomment{perform value iteration}}\black
+		\State{$child_cost \gets \Call{Cost}{transition} + target\_net(child)$}
+                	\State{$V[puzzle] \gets \Call{Min}{V[puzzle], child\_cost}$}
+	     \EndFor
+	\EndIf
+       \EndFor
+\\ \teal{\algorithmiccomment{iterate through network training running feed-forward and backward-prop}}\black
+	\State{$net \gets \Call{TrainNeuralNet}{V, loss\_function, net\_architecture}$}
+\\ \teal{\algorithmiccomment{update target network if criteria met}}\black
+	\If{$target\_network\_update\_criteria(net, target\_net)$}
+		\State{$target\_net \gets \Call{Copy}{net}$}
+	\EndIf
+    \Return{$net$}
+\EndWhile
+
+\EndFunction
+\end{algorithmic}
+\end{algorithm}
 \black
+
 For this project, I have implemented the above algorithm in a class called DeepReinforcementLearner which I shall describe in details in section \ref{sec:codelearners}.
@@
 
-\section{Deep Q-Learning}
+\Section{Deep Q-Learning}
 \label{sec:TheoryDQL}
@@
 
-\section{Monte Carlo Tree Search}
+\Section{Monte Carlo Tree Search}
+\label{sec:TheoryMCTS}
 

### 2022-09-11 rubiks/scripts/rubiks22/perf.py

@@ -71,13 +71,13 @@ if '__main__' == __name__:
         performance(**locals())
     elif action_type == Solver.do_performance_test:
         nb_samples = 100
-        nb_cpus = 2
-        for counter in [12]:
+        nb_cpus = 1
+        for counter in [22]:
             logger.log_info('counter: ', counter)
             performance(solver_type=Solver.astar,
                         heuristic_type=Heuristic.deep_q_learning,
                         learner_type=Learner.deep_q_learner,
-                        time_out=7200,
+                        time_out=3600,
                         add_perfect_shuffle=counter == 22,
                         min_nb_shuffles=0 if counter > 20 else counter,
                         max_nb_shuffles=0 if counter > 20 else counter,

### 2022-09-11 latex/Chapters/DRLSearch.tex
update

@@
 \Function{DeepReinforcementLearningHeuristic}{**args}
-\State{$puzzle\_type \gets \Call{FromParams}{**args}$}
-\State{$ puzzles\_gen\_params \gets \Call{FromParams}{**args}$}
+\State{$puzzle\_type \gets$ \Call{FromParams}{**args}}
+\State{$ puzzles\_gen\_params \gets$ \Call{FromParams}{**args}}
 \State{$puzzles\_generation\_criteria \gets$ \Call{FromParams}{**args}}
-\State{$n\_puzzles \gets$ \Call{FromParams}{**args}}
-\State{$target\_network\_update\_params \gets \Call{FromParams}{**args}$}
-\State{$max\_epochs \gets \Call{FromParams}{**args}$}
-\State{$loss\_function \gets \Call{FromParams}{**args}$}
-\State{$net\_architecture \gets \Call{FromParams}{**args}$}
+\State{$n\_puzzles \gets \gets$ \Call{FromParams}{**args}}
+\State{$target\_network\_update\_params \gets$ \Call{FromParams}{**args}}
+\State{$max\_epochs \gets$ \Call{FromParams}{**args}}
+\State{$loss\_function \gets$ \Call{FromParams}{**args}}
+\State{$net\_architecture \gets$ \Call{FromParams}{**args}}
 \\ \teal{\algorithmiccomment{initial network and target network}}\black
@@
 \teal
-\paragraph{}{\textbf{pseudo code -- \textbf{DQL} heuristics}}
-\begin{pseudocode}
-###############################################################################
-def deep_q_learning(... same as DRL ...):
-    """ pseudo code for learning a puzzle via deep q learning """
-    # ... same as DRL ...
-    nb_actions = puzzle_type.get_nb_actions()
-    while epoch < max_epochs and other_convergence_criteria(network):
-        # ... same as DRL ...
-        Q_and_V = dict()
-        """ Compute updated cost and actions via V & Q iteration ... 
-             all moves are assumed to have cost 1 """
-        for puzzle in puzzles:
-            value = target_network(puzzle)[-1]
-            actions = [0] * nb_actions
-            best_action_id = 0
-            if puzzle.is_goal():
-                value = 0
-            else:
-                for action_id, child in puzzle.children():
-                    child_value = target_network(child)[-1]
-                    if child_value < value:
-                        value = child_value
-                        best_action_id = action_id
-            # We update both the value function and the best action
-            actions[best_action_id] = 1
-            Q_and_V[puzzle] = actions + [value]
-        # ... same as DRL ...
-###############################################################################
-\end{pseudocode}
+\begin{algorithm}[H]
+\caption{Deep Q Learning -- Heuristic}\label{alg:TheoryDQL}
+\begin{algorithmic}
+
+\Function{DeepReinforcementLearningHeuristic}{**args}
+
+\\ \teal{\algorithmiccomment{ ... same params and init as for \textbf{DRL} ...}}\black
+ \State{$nb\_actions \gets$ \Call{GetNbActions}{puzzle\_type}}
+\While{epoch ...}
+\\ \teal{\algorithmiccomment{ ... same while epoch loop as for \textbf{DRL} ...}}\black
+\\ \teal{\algorithmiccomment{ ... update cost \textbf{and actions} via V \& Q iteration ...}}\black
+	\State{$Q\_and\_V \gets dict()$}
+	\For {$puzzle \in puzzles$}
+            \State{$value \gets target\_network(puzzle).value$}
+            \State{$actions = [0] * nb\_actions$}
+            \State{$best\_action\_id \gets 0$}
+	 \If{\Call{IsGoal}{$puzzle$}}
+                \State{$value \gets 0$}
+            \Else:
+                \For{$action\_id, child \in \Call{Children}{puzzle}$}
+                    \State{$child\_value \gets target\_network(child).value$}
+                    \If{$child\_value \le value$}:
+                        \State{$value \gets child\_value$}
+                        \State{$best\_action\_id \gets action\_id$}
+		\EndIf
+		\EndFor
+	\EndIf
+\\ \teal{\algorithmiccomment{ ... update cost \textbf{and best action} ...}}\black
+            \State{$actions[best\_action\_id] \gets 1$}
+            \State{$Q\_and\_V[puzzle] \gets actions + [value]$}
+	\EndFor
+\EndWhile
+\EndFunction
+\end{algorithmic}
+\end{algorithm}
 \black
@@
 In order to prevent always following the same paths (which might take us very far from the goal), \textbf{MCTS} algorithms usually keep track of actions they have followed already and underweight their probability so that subsequent paths explore other actions.
-A pseudo code description of it looks like this:
+Let me first present some pseudo-code for the paths generation, in particular the trade-off between actions-values (exploration-exploitation in some sense):
 \\
+
+\teal
+\begin{algorithm}[H]
+\caption{Monte Carlo Tree Search -- Child Node Choice}\label{alg:TheoryDQLNodeChoice}
+\begin{algorithmic}
+
+\Function{ChooseChildNode}{$node, actions\_probas, actions\_values, actions\_count, c$}
+
+\\ \teal{\algorithmiccomment{c is the actions-values trade-off hyper-parameter}}\black
+
+\State{$N \gets \sqrt{\sum_{n \in actions\_count}{n}}$}
+
+\\ \teal{\algorithmiccomment{... frequent actions penalized as denominator grows faster than numerator N ... }}\black
+\State{$Q \gets N . \frac{actions\_probas}{actions\_count}$}
+\State{$V \gets actions\_values$}
+\\ \teal{\algorithmiccomment{... high c gives weight to actions to favour exploration ... }}\black
+\\ \teal{\algorithmiccomment{... return child for which cQ + V is maximized ... }}\black
+\State{$action\_id \gets \Call{IndexMax}{Q + V}$}
+\\
+\Return{\Call{Child}{$node, action\_id$}}
+\EndFunction
+\end{algorithmic}
+\end{algorithm}
+\black
+
 \teal

### 2022-09-11 latex/Chapters/DRLSearch.tex
update

@@
 \Function{BlindSearch}{$root, time\_out, search\_type=Search.BFS$}
-\State\Call{InitialiseTimeOut}{$time\_out$}
+\State{\Call{InitialiseTimeOut}{$time\_out$}}
 \If{\Call{IsGoal}{$root$}}
@@
 In order to prevent always following the same paths (which might take us very far from the goal), \textbf{MCTS} algorithms usually keep track of actions they have followed already and underweight their probability so that subsequent paths explore other actions.
-Let me first present some pseudo-code for the paths generation, in particular the trade-off between actions-values (exploration-exploitation in some sense):
+Let me first present some pseudo-code for the paths generation, in particular the trade-off between actions-values (exploration-exploitation):
 \\
@@
 \\ \teal{\algorithmiccomment{... return child for which cQ + V is maximized ... }}\black
-\State{$action\_id \gets \Call{IndexMax}{Q + V}$}
+\State{$action\_id \gets \Call{IndexMax}{c.Q + V}$}
 \\
@@
 
+Then the actual \textbf{MCTS} algorithm is simply generating a new path from root to a new leaf at each iteration, until it finds the goal:
+
 \teal
-\paragraph{}{\textbf{pseudo code -- \textbf{MCTS}}}
-\begin{pseudocode}
-#################################################################
-def MCTS(initial_node, time_out, q_v_network):
-    """ pseudo code for Monte Carlo Tree Search """
-    initialize_time_out(time_out)
-    tree = Tree(initial_node)
-    while True:
-        check_time_out()
-        leaf = construct_path_to_new_leaf(tree, q_v_network)
-        if leaf.is_goal():
-            return BFS(tree) # path to leaf can usually be improved by BFS
-#################################################################
-def construct_path_to_new_leaf(tree, q_v_network):
-    leaf = tree.initial_node
-    while leaf.expanded:
-        actions_probas, actions_values = q_v_network(leaf)
-        """ choose next node based on joint actions probas & values
-        as well as history of actions taken
-        """
-        leaf = leaf.choose_child(actions_probas,
-                                 actions_values)
-    # add children of leaf to tree
-    leaf.expand()
-    # penalize path taken to favour exploration
-    leaf.increment_historical_actions_count()
-    return leaf
-#################################################################
-\end{pseudocode}
+\begin{algorithm}[H]
+\caption{Monte Carlo Tree Search}\label{alg:TheoryMCTS}
+\begin{algorithmic}
+\Function{MCTS}{$root, time\_out, q\_v\_network, c$}
+\State{\Call{InitialiseTimeOut}{$time\_out$}}
+\State{$tree \gets \Call{Tree}{root}$}
+\If{\Call{IsGoal}{$root$}}
+  \Return{$root$}
+\EndIf
+\While{true}
+	\State\Call{CheckTimeOut}{void}
+	\State{$leaf \gets root$}
+	\While{$leaf.expanded$}
+        		\State{$actions\_probas, actions\_values \gets q\_v\_network(leaf)$}
+	           \State{$leaf \gets\Call{ChooseChildNode}{leaf,  actions\_probas, actions\_values, leaf.actions\_count, c}$}
+		 \State{\Call{Expand}{$leaf, tree$}}
+	\EndWhile
+           \Call{IncrementActionsFromLeafToRoot}{$leaf$}
+       \\ \teal{\algorithmiccomment{path to leaf can usually be improved by re-running \textbf{BFS} on the resulting tree}}\black
+       \If{\Call{IsGoal}{$leaf$}}
+            \Return{\Call{BlindSearch}{$tree, search\_type=Search.BFS$}}
+       \EndIf
+\EndWhile
+\EndFunction
+\end{algorithmic}
+\end{algorithm}
 \black

### 2022-09-11 latex/Chapters/DRLSearch.tex
update

@@
 
-For this project, I have implemented the above algorithm in a class called DeepQLearner which I shall describe in details in section \ref{sec:codelearners}.
+\noindent For this project, I have implemented the above algorithm in a class called DeepQLearner which I shall describe in details in section \ref{sec:codelearners}.
 
@@
 
-Since my implementation of \textbf{DQL} is no longer just a cost-to-go estimate as with \textbf{DRL}, but a joint cost-to-go and distribution over actions (to be concrete, that means a 5-dimensional vector for the \textbf{SP} and a 13-dimensional vector for the \textbf{RC}).
+The \textbf{DQL} heuristics from \ref{alg:TheoryDQL} no longer only produce cost-to-go estimates as \textbf{DRL}, but rather a joint cost-to-go and distribution over actions (i.e.
+a 5-dimensional vector for the \textbf{SP} and a 13-dimensional vector for the \textbf{RC}).
 This leaves us with the question of how can we use the output of such a network to search for a solution.
@@
 \\
-Another possibility is to use an algorithm which combines both the actions distribution and the value function to inform its search.
+Another possibility is to use an algorithm combining the actions distribution and the value function to inform its search.
 The Monte Carlo Tree Search (\textbf{MCTS}) is such an algorithm.
-In short, it generates at each step (and possibly in a distributed or multithreaded manner) paths to a \textit{leaf} on the frontier of unexpanded nodes, expands that node and checks if it is a goal state (and keeps going until the goal is found).
-The paths followed from initial state to leaves are decided via a combination of the probability distribution over actions (following the most promising actions) as well as the value function, according to a trade-off which can be tuned via a hyper-parameter.
-In order to prevent always following the same paths (which might take us very far from the goal), \textbf{MCTS} algorithms usually keep track of actions they have followed already and underweight their probability so that subsequent paths explore other actions.
-Let me first present some pseudo-code for the paths generation, in particular the trade-off between actions-values (exploration-exploitation):
+It generates at each iteration (possibly in a distributed or multithreaded manner) paths to a \textit{leaf} on the frontier of unexpanded nodes, expands that node and checks if it is a goal state (and keeps going until the goal is found).
+The paths followed from initial state to leaves are decided via a combination of the probability distribution over actions (allowing exploration) and value function (exploitation), according to a trade-off which can be tuned via a hyper-parameter $c$.
+To prevent following the same paths (which might lead the search far from the goal), the \textbf{MCTS} keeps track of actions already followed and downweight their probability so that subsequent paths explore different actions.
+Let me first present in pseudo-code the paths generation, in particular the trade-off between actions-values (exploration-exploitation):
 \\
@@
 
-Then the actual \textbf{MCTS} algorithm is simply generating a new path from root to a new leaf at each iteration, until it finds the goal:
+\noindent The actual \textbf{MCTS} algorithm is simply generating a new path from root to a new leaf at each iteration, until the goal is found:
 
@@
 
-I have implemented the exact \textbf{MCTS} procedure described in \cite{https://doi.org/10.48550/arxiv.1805.07470} and outlined in the above pseudo code in a class called MonteCarloTreeSearch (see section \ref{sec:codesolvers} for details).
+\noindent I have implemented the exact \textbf{MCTS} procedure described in \cite{https://doi.org/10.48550/arxiv.1805.07470} and outlined above (\ref{alg:TheoryMCTS}) in a class called MonteCarloTreeSearch (see section \ref{sec:codesolvers} for details).

### 2022-09-11 latex/Chapters/Implementation.tex
update

--- 2022-09-06 latex/Chapters/Implementation.tex
@@
 
-\begin{landscape}
 \begin{figure}[H]
-\centering
-\includegraphics[scale=0.40]{./Figures/codebase}
-%\decoRule
+  \noindent
+  \makebox[\textwidth]{\includegraphics[scale=0.5]{./Figures/codebase}}
   \caption[Codebase]{Code base}
@@
 \end{figure}
-\end{landscape}
 
-Let me now describe what each submodule does in more details:
+\noindent Let me now describe what each submodule does in more details:
 
-\section{rubiks.core}
+\Section{rubiks.core}
 \begin{figure}[H]
 \centering
-\includegraphics[scale=0.25]{./Figures/codebasecore}
+\includegraphics[scale=0.22]{./Figures/codebasecore}
 %\decoRule
@@
 \begin{itemize}
-\item \textbf{Loggable}: a wrapper around Python's logger which automatically picks up classes' names at init and format things (dict, series and dataframes in particular) in a nicer way.
+\item \textbf{Loggable}: a wrapper around Python's logger to automatically picks up classes' names at init and format things nicely (dict, series and dataframes in particular).
 \item \textbf{Parsable}: a wrapper around ArgumentParser, which allows to construct objects in the project from command line, to define dependencies between object's configurations and to help a bit with typing of configs.
@@
 
-\section{rubiks.puzzle}
+\Section{rubiks.puzzle}
 \begin{figure}[H]
 \centering
-\includegraphics[scale=0.25]{./Figures/codebasepuzzle}
+\includegraphics[scale=0.22]{./Figures/codebasepuzzle}
 %\decoRule
@@
 
-\section{rubiks.search}
+\Section{rubiks.search}
 \begin{figure}[H]
 \centering
-\includegraphics[scale=0.25]{./Figures/codebasesearch}
+\includegraphics[scale=0.22]{./Figures/codebasesearch}
 %\decoRule
@@
 
-\section{rubiks.heuristics}
+\Section{rubiks.heuristics}
 \begin{figure}[H]
 \centering
-\includegraphics[scale=0.25]{./Figures/codebaseheuristics}
+\includegraphics[scale=0.22]{./Figures/codebaseheuristics}
 %\decoRule
@@
 
-\section{rubiks.deeplearning}
+\Section{rubiks.deeplearning}
 \begin{figure}[H]
 \centering
-\includegraphics[scale=0.25]{./Figures/codebasedeeplearning}
+\includegraphics[scale=0.22]{./Figures/codebasedeeplearning}
 %\decoRule
@@
 \item \textbf{FullyConnected}: wrapper around a Pytorch fully connected network, with configurable hidden layers and size.
-There are some params as well to add drop out, and to indicate whether or not the inputs are one hot encoding (in which case the first layer is automatically adjusted in size, using information from the puzzle dimension).
+Parameters allow to add drop out, to indicate whether or not the inputs are one hot encoding (in which case the first layer is automatically adjusted in size, using information from the puzzle dimension).
 There is also importantly a \textit{joint\_policy} parameter to indicate whether we want the output to be 1 dimensional (e.g.
@@
 
-\section{rubiks.learners}
+\Section{rubiks.learners}
 \label{sec:codelearners}
@@
 \centering
-\includegraphics[scale=0.25]{./Figures/codebaselearners}
+\includegraphics[scale=0.22]{./Figures/codebaselearners}
 %\decoRule
@@
 
-\section{rubiks.solvers}
+\Section{rubiks.solvers}
 \label{sec:codesolvers}
@@
 \centering
-\includegraphics[scale=0.25]{./Figures/codebasesolvers}
+\includegraphics[scale=0.22]{./Figures/codebasesolvers}
 %\decoRule
@@
 \end{figure}
-This module implements solvers, which solve puzzles.
-The base class Solver is a Factory of solvers, and in addition to being able to instantiating the following types of solvers, can run different solvers through a similal sequences of random puzzles (for various increasing degrees of difficulty (shuffling), and/or perfectly shuffled ones) and display a comparison of how they perform in a number of metrics (chosable from percentage of puzzles solved before time out, percentage of optimal puzzle (if optimality can be ascertained), mean-median-maximum run time, mean-median-maximum expanded nodes (for solvers where it makes sense), mean-median-max cost, and an optimality score which compares the total sum of cost over all puzzles versus a benchmark solver).
-Ideally the benchmark solver to compute the optimality score should be one which is provably optimal, but in the case of the \textbf{RC} that was not possible so I used A$^{*}$ with KociembaHeuristic as my benchmark.
+This module implements the actual puzzles solvers.
+The base class Solver is a Factory, and in addition to being able to instantiating the following types of solvers, can run different solvers through a similal sequences of random puzzles (for various increasing degrees of difficulty (scrambling), and/or perfectly shuffled ones) and display a comparison of how they perform in a number of metrics (configurable from percentage of puzzles solved before time out, percentage of optimal puzzle (if optimality can be ascertained), mean-median-maximum run time, mean-median-maximum expanded nodes (for solvers where it makes sense), mean-median-max cost, and an optimality score which compares the total sum of cost over all puzzles versus a benchmark solver).
+Ideally the benchmark solver to compute the optimality score should provably optimal, but in the case of the \textbf{RC} that was not possible so I used A$^{*}$ with KociembaHeuristic as my benchmark.
 For the \textbf{SP} I used the optimal A$^{*}$ with Manhattan++ heuristic.
@@
 \begin{itemize}
-\item \textbf{DFSSolver} This solver is literally a few lines of code, extremely close to the pseudo code detailed in \ref{sec:TheoryBFSDFS}
-\item \textbf{BFSSolver} Ditto, see \ref{sec:TheoryBFSDFS}
-\item \textbf{AStarSolver} Ditto, see \ref{sec:TheoryAStar}
+\item \textbf{DFSSolver} This solver is literally a few lines of code, extremely close to the pseudo code detailed in \ref{alg:TheoryBFSDFS}
+\item \textbf{BFSSolver} Ditto, see \ref{alg:TheoryBFSDFS}
+\item \textbf{AStarSolver} Ditto, see \ref{alg:TheoryAStar}
 \item \textbf{NaiveSlidingSolver} The NaiveSolver is conceptually simple, but was far from trivial to implement and debug.
 Its functioning is detailed in the appendix section \ref{sec:Examples}.
-Basically, it does what a beginner player would do (certainly what I do when I play the sliding puzzle), solving the top row first, then the left column, hence reducing the puzzle to a smaller dimensional one, and keep iterating until done.
+Basically, it does what a beginner player would do, certainly what I do when I play the sliding puzzle: solving the top row first, then the left column, hence reducing the puzzle to a smaller dimensional one, and keep iterating until done.
 \item \textbf{KociembaSolver} This solver is a wrapper around two libraries.
-The first one, hkociemba (see \cite{HKociemba}) is a 2x2x2 implementation written by Kociemba himself, the second one is a Python pip-installable which solves the 3x3x3 Rubik's.
+The first one, hkociemba (see \cite{HKociemba}) is a 2x2x2 implementation written by Kociemba himself, the second one is a Python pip-installable library which solves the 3x3x3 Rubik's.
 My  wrapper simply translates my cube representation in that accepted by these 2 solvers, and their result string into moves usable by my code.
-I also do some additional massaging for the 3x3x3 case, due to a shortcoming/bug in the 3x3x3 implementation, as I have detailed in appendix \ref{sec:Kociemba}.
-\item \textbf{MonteCarloSearchTreeSolver} TBI I want to implement this in August when working on the Rubiks' ...
-following the Rubik's paper fromMcAleer \& Agostilenni et al \cite{https://doi.org/10.48550/arxiv.1805.07470}
+I also do some additional massaging for the 3x3x3 case, due to a shortcoming/bug in the 3x3x3 Python library, as I have detailed in appendix \ref{sec:Kociemba}.
+\item \textbf{MonteCarloSearchTreeSolver} My implementation follows McAleer \& Agostilenni et al \cite{https://doi.org/10.48550/arxiv.1805.07470} but is however single-threaded.
+As a consequence, I dropped the $\nu$ parameter from their implementation, since it only introduces a penalty to discourage parallel threads from generating the same paths (in single threaded mode, this parameter is inconsequential).
+\\
+This solver basically generates \textit{pseudo-random} paths from initial node to a new leaf at each iteration (see \ref{alg:TheoryMCTS}), until we obtain a goal leaf.Paths are, despite the name, completely deterministic: they indeed result from a trade-off between the probability of actions (penalised for repeated use) and the value function, but there are no random draws being performed.
+\\ As per McAleer \& Agostilenni's paper, I have implemented a final pruning via \textbf{BFS} of the tree resulting from the above \textit{Monte-Carlo} process.It is inded clear that the paths generated by this \textbf{MCTS} algorithm have little chance of being optimal since there is no correction mechanism to favour shortcuts over long loops.
+I have however made this \textbf{BFS} trimming optional via a \textit{trim\_tree} parameter to study its effect on solutions' optimality and run time (see results section \ref{sec:MTCSHyperParamsEffect}).
+\\
+I will study and report on the effect of the trade-off hyper-parameter \textit{c}, and confirm how it affects solutions' quality and run time in section \ref{sec:MTCSHyperParamsEffect}.
+
 \end{itemize}
 
-\section{rubiks.scripts}
+\Section{data}
 
-\begin{figure}[H]
-\centering
-\includegraphics[scale=0.25]{./Figures/codebasescripts}
-%\decoRule
-\caption[Codebase]{rubiks.scripts}
-\label{fig:Codebasescripts}
-\end{figure}
-
-Finally it is worth noting that the code will save on disk a lot of data (e.g.
-the learners will save what they have learnt, e.g.
-a Pytorch network or a data base of optimal costs, the performance comparison will run solvers versus very many configurations of puzzles and save the results for later being able to display) etc...
-The base of the tree to save all this data can be chosen by setting up the "RUBIKSDATA" environment variable.
-If not, it will go somewhere in you HOME :)
+The code generates a fair amount of data, in particular:
+\begin{itemize}
+\item Manhattan++ generates databases of linear constraint penalties (see \ref{HSS})
+\item TrainingData: (see e.g. \ref{DLSS}) generates databases of puzzles and their associated costs for later training of a DeepLearner.
+\item Learners: DeepLearner, DeepReinforcementLearner, DeepQLearner generate models as well as some convergence data useful to understand how the networks learn.
+Similarly PerfectLearner saves its results for later reuse (in e.g.
+PerfectHeuristic) or display.
+\item Solver: Generates sequences of scrambled puzzles used to make the performance test (comparing different solvers) fair.
+\item Solver.performance\_test: Generates a data frame containing the metrics discussed earlier in \ref{sec:codesolvers}, for each solver and each level of difficulty.
+\end{itemize}
+That data is saved under a tree following the schema below:
+$$\$\{ "RUBIKSDATA"\}/data\_type/puzzle\_type/dimension/file\_name.pkl$$

### 2022-09-11 latex/Chapters/Puzzles.tex
update

@@
 
-It is rather easy to see why this is the case, let us discuss why: given a configuration c of the tiles, let us define the permutation p(c) of this configuration according to the following schema: we enumerate the tiles row by row (top to bottom), left to right for odd rows and right to left for the even rows, ignoring the empty compartment.
+\noindent It is rather easy to see why this is the case, let us discuss why: given a configuration c of the tiles, let us define the permutation p(c) of this configuration according to the following schema: we enumerate the tiles row by row (top to bottom), left to right for odd rows and right to left for the even rows, ignoring the empty compartment.
 For instance, the following 15-puzzle c:
@@
 \Subsection{Search Space \& Solvability}
+\label{Theory:SPSSS}
 
@@
 
- It is quite clear from that picture above that if we can bring the 4x4x4 cube to such a configuration on all faces (which is not super hard), and we know how to solve a 3x3x3, we are essentially done.
+\noindent It is quite clear from that picture above that if we can bring the 4x4x4 cube to such a configuration on all faces (which is not super hard), and we know how to solve a 3x3x3, we are essentially done.
 (well, almost, since the additional degrees of freedom coming from the higher dimension of the 4x4x4 mean that certain arrangements of the colors which are possible on the 4x4x4 are not on the 3x3x3.

### 2022-09-11 latex/Chapters/SlidingPuzzle.tex
update

@@
 
-I am neither making claims of depth (each solver could surely be tuned and optimized/adapted, possibly differently for each dimension) nor breadth (I have not run all of the solvers versus each dimension).
+\noindent I am neither making claims of depth (each solver could surely be tuned and optimized/adapted, possibly differently for each dimension) nor breadth (I have not run all of the solvers versus each dimension).
 Running these experiments takes a lot of computing time (even though I have very decent computing resources), so I had to be somewhat selective.
@@
 
-As mentioned in chapter \ref{sec:Puzzles}, the state space cardinality for the SP grows very quickly with n and m.
+As mentioned in chapter \ref{sec:Puzzles}, the state space cardinality for the \textbf{SP} grows very quickly with n and m.
 The only dimensions which have less than 239.5 millions states are shown in table \ref{tab:smallSP}.
@@
 \end{table}
-In this section, I will discuss \textit{full} results for these 5 dimensions.
+\noindent In this section, I will discuss \textit{full} results for these 5 dimensions.
 In order to fully solve them, one can simply use rubiks.scripts.learner, setting up the PerfectLearner with A* and manhattan heuristic, or instantiate directly a PerfectLearner as have seen in section \ref{PLSS}.
@@
 \end{table}
-Notice that among the above dimensions, 2x5 is the largest and hardest one.
-My current displayed result in table \ref{tab:smallSPGN} is provisional, as I have only solved a bit over 1.1 million puzzles out of the 1.8 million possible, so it is well possible that I haven't yet bumped into the hardest configuration.
-I ran a combined 600 hours, much of which run over 16 cores, and even over a full week-end on a c5.18xlarge instance (72 cores) on Amazon EC2.
+\noindent Among the above dimensions, 2x5 has the largest state space.
+My current result in table \ref{tab:smallSPGN} for that dimension is provisional, as I have only solved about 1.2 million puzzles out of the 1.8 million possible configurations, despite running a combined 600 hours (most of the time over 16 cores, except for a full week-end run on a 72-core c5.18xlarge Amazon EC2 instance).
+It is well possible that I haven't yet bumped into the hardest configuration though!
 \\
@@
 \\
-\\
-\underline{Most difficult 2x2 (6 moves)}:
-\begin{center}
 \begin{three}
+\centering
 \setrow{2}{,3}
@@
 \end{three}
-\end{center}
-\underline{Most difficult 2x3 (21 moves)}:
-\begin{center}
 \begin{five}
@@
 \end{five}
-\end{center}
-\underline{Most difficult 2x4 (36 moves)}:
-\begin{center}
 \begin{seven}
@@
 \end{seven}
-\end{center}
-\underline{Most difficult 2x5 (55* moves)}:
-\begin{center}
 \begin{nine}
@@
 \setrow{1}{5,4,8,2,6}
-\tiny{* provisional result}
 \end{nine}
-\end{center}
-\underline{Most difficult 3x3 (31 moves)}:
-\begin{center}
 \begin{eight}
@@
 \end{eight}
-\end{center}
 
@@
 \centering
-%\hspace*{-2.5cm}
 \includegraphics[scale=0.40]{./Figures/25SPPerformanceManhattan}
-%\decoRule
 \caption[SP]{Manhattan vs Manhattan++ for the 2x5 \textbf{SP}}
@@
 
-As can be seen, for low difficulty (up to \textit{nb\_shuffles = 20}), the node expansions are about the same in both cases, and the overhead of adding the linear constraints penalty increases the run time.
+\noindent As can be seen, for low difficulty (up to \textit{nb\_shuffles = 20}), the node expansions are about the same in both cases, and the overhead of adding the linear constraints penalty increases the run time.
 However, for any non trivial case, Manhattan++ outperforms considerably (by a factor of about 2.5).
@@
 Manhattan++              & 34.6 &  49 & 18,780 & 213,557 & 14,665 & 165,830 \\
-Improvement               & n/a &  n/a & x2.5 & x2.7 & x2.5 & x2.6 \\
+Improvement               & n/a &  n/a & \textbf{x2.5} & \textbf{x2.7} & \textbf{x2.5} & \textbf{x2.6} \\
 \end{tabular}
@@
 
+blablabla
+\\
+blablabla
+\\
+blablabla
+
 \label{ssec:34SPSC}
@@
 \begin{figure}[H]
-\centering
-\includegraphics[scale=0.63]{./Figures/34SPPerformance}
-%\decoRule
+  \noindent
+  \makebox[\textwidth]{\includegraphics[scale=0.6]{./Figures/34SPPerformance}}
   \caption[34SPPerformance]{Solvers' performance comparison 3x4 \textbf{SP}}
@@
 \begin{figure}[H]
-\centering
-\includegraphics[scale=0.63]{./Figures/44SPPerformance}
-%\decoRule
+  \noindent
+  \makebox[\textwidth]{\includegraphics[scale=0.55]{./Figures/44SPPerformance}}
   \caption[44SPPerformance]{Solvers' performance comparison 4x4 \textbf{SP}}

### 2022-09-11 rubiks/scripts/solver.py

@@ -26,8 +26,8 @@ if '__main__' == __name__:
     action_type = Solver.do_plot
     """ What puzzle """
     puzzle_type = Puzzle.sliding_puzzle
-    n = 3
-    m = 3
+    n = 2
+    m = 5
     dimension = Puzzle.factory(**globals()).dimension()
     """ How much to shuffle """
     nb_shuffles = 0
@@ -55,7 +55,7 @@ if '__main__' == __name__:
                            Solver.median_expanded_nodes,
                            ]
     plot_abbreviated_names = True
-    labels_at_top = True
+    labels_at_top = False
     show_title = True
     lite_title = True
     plot_abbreviated_names = True
@@ -64,7 +64,7 @@ if '__main__' == __name__:
     #colors = ['olive', 'royalblue', 'royalblue', 'darkred', 'darkred', 'goldenrod', 'goldenrod', 'darkcyan', 'darkcyan']
     #exclude_solver_names = ['hattan']
     exclude_solver_names = ['seen=1.4', '1.2e-07', '1e-07', '9.5']   #  <- for SP 4, 4
-    fig_size = (12.5, 12)
+    fig_size = (10, 6)
     """ Which solver type {Solver.dfs,
                            Solver.bfs,
                            Solver.astar,

### 2022-09-11 latex/Chapters/SlidingPuzzle.tex
update

@@
 However, for any non trivial case, Manhattan++ outperforms considerably (by a factor of about 2.5).
-Table \ref{tab:mppOutperformance} below summarizes the results for the 250 \textit{perfectly} shuffled instances:
+Table \ref{tab:mppOutperformance} below summarizes the results for the 250 \textit{perfectly} shuffled configurations:
 
@@
 
-In this section, which focusses on the 3x3 \textbf{SP}, I detail some results from the PerfectLearner (\ref{sec:S33PL}), from the DeepReinforcementLearner (\ref{sec:S33DRL}), before moving to a discussion of the \textbf{DQL} \textbf{MCTS} solver and the effect of its trade-off hyper-parameter \textit{c} and of its optional \textbf{BFS} tree-trimming on the quality of the solutions (\ref{sec:MTCSHyperParamsEffect}).
-I then run a comprehensive comparison of my 10 different solvers on the 3x3 \textbf{SP} in \ref{ssec:33SPSC}.
+In this section, which focusses on the 3x3 \textbf{SP}, I detail some results from the PerfectLearner (\ref{sec:S33PL}) before moving to a discussion of the \textbf{DQL} \textbf{MCTS} solver and the effect of its trade-off hyper-parameter \textit{c} and of its optional \textbf{BFS} tree-trimming on the quality of the solutions (\ref{sec:MTCSHyperParamsEffect}).
+I then run a comprehensive comparison of my 10 different solvers on the 3x3 \textbf{SP} in \ref{ssec:33SPSC} including \textbf{DL},  \textbf{DRL} and  \textbf{DQL}.
 Finally, I end the section by running a few select solvers on one of the two hardest 3x3 puzzle in \ref{ssec:33SPHard}
@@
 
-\Subsection{Deep reinforcement learner}
-\label{sec:S33DRL}
+%\Subsection{Deep reinforcement learner}
+%\label{sec:S33DRL}
 

### 2022-09-11 latex/Chapters/SlidingPuzzle.tex
update

@@
 \label{sec:S33DRLDQL}
-Let me discuss a comparison of several algorithms on 1000 random puzzles generated for a number of random shuffling (with best-effort-no-backtracking) from 0 to 50 in step of 2, as well as for perfect shuffling (denoted by $\infty$) on the comparison graphs.
-The results are shown in figure \ref{fig:33SPPerformance}
-\\
-DL: 100 seq 15 to 31 shuffles 10k epochs connected 600 300 100
-\\
-Manhattan vs ++
-\\
-DQL vs DRL
-\\
-DxL vs Perfect
-\\ Put table for nb\_shuffle=inf and highlight best in diff categories (cost, optimality score, run time, nodes)
-\\ say how much of the puzzles the solvers have seen
+As promised earlier in table \ref{tab:gridSP}, I have run the Solver's comparison test with 10 different solvers (Naive, \textbf{MCTS}, \textbf{BFS}, A$^{*}$ with the following heuristics: Manhattan, Manhattan++, Perfect, \textbf{DRL}, \text{DQL} and \text{DL}), to get an idea of their respective success rate, strengths and weaknesses.
+The \textbf{DRL} and \textbf{DQL} heuristics were trained with the exact same parameters and a fully connected network of 3 hidden layers with 600, 300 and 100 nodes.
+For the \textbf{DL} heuristic (trained on perfect solutions from A$^{*}$[Perfect]) I tried with the same fully connected network as well as with a combination of parallel fully connected and convolutional layers (see figure \ref{fig:33SPnets} for the exact architecture I used).
+The detailed results (percentage of configurations solved, optimality score, median and max cost, run time and expanded nodes) can be found on the next page  in figure \ref{fig:33SPPerformance}.
+Each solver had to solve 1,000 random puzzles for each level of difficulty (scrambling from goal) going from 5 to 50 in step of 5, as well as for perfect shuffling.
+Let me list some of the important findings from this rather comprehensive experiment:
+
+\begin{itemize}
+\item As expected, \textbf{BFS} becomes quickly hopeless.
+\item The naive solver is obvioulsy by far the fastest but finds poor solutions (about 2.5 longer on average than optimal).
+\item As discussed in the previous section, \textbf{MCTS} is slow, and even with much exploration and \textbf{BFS} tree trimming found solution about 25\% longer than optimal.
+\item I originally was naively hoping a pure convolutional network would do well but did not manage to get the loss to converge.
+I later realized that this is quite different from the typical image patterns recognition as there is less direct translation and scale invariants at play here, and decided instead to try combining convolutional with fully connected layers in parallel.
+That improved things, but overall I saw little difference in performance from my simpler 3-hidden-layer fully connected network.
+\item \textbf{DRL} and \textbf{DQL} heuristics, though not as efficient in terms node-expansion as the perfect heuristic and the \textbf{DL} heuristics, were rather surprisingly doing better than even Manhattan++.
+They ended up \textit{seeing} around 25\% of all the possible configurations during training (in an \textit{unsupervised} way though) so did show some reasonable amount of generalization out of sample.
+Their optimality score was an impressive 99\%+ at all levels of difficuly for \textbf{DRL} and 98\%+ at all levels of difficuly for \textbf{DQL}.
+Even more interestingly, the \textbf{DQL} heuristic expanded almost 4 times less nodes than \textbf{DRL}, suggesting some benefit from the joint learning of the policy and value function.
+\item The \textbf{DL} heuristics (fully connected and convolutional) performed almost on par with the perfect heuristic (optimality score of 99\%+ at all levels of difficulty and nodes expansion only increased by about 12\%), despite seeing only 8\% of the configurations by the end of their training.
+\end{itemize}
+
+\begin{figure}
+  \noindent
+  \makebox[\textwidth]{\includegraphics[scale=0.45]{./Figures/33SPnets}}
+  \caption[33SPnet]{Architecture used for the \textbf{DL} heuristics training on the 3x3 \textbf{SP}}
+  \label{fig:33SPnets}
+\end{figure}
 
@@
   \noindent
-  \makebox[\textwidth]{\includegraphics[scale=0.6]{./Figures/33SPPerformance}}
+  \makebox[\textwidth]{\includegraphics[scale=0.55]{./Figures/33SPPerformance}}
   \caption[33SPPerformance]{Solvers' performance comparison 3x3 \textbf{SP}}

### 2022-09-11 latex/Chapters/SlidingPuzzle.tex
update

@@
 \item I originally was naively hoping a pure convolutional network would do well but did not manage to get the loss to converge.
-I later realized that this is quite different from the typical image patterns recognition as there is less direct translation and scale invariants at play here, and decided instead to try combining convolutional with fully connected layers in parallel.
-That improved things, but overall I saw little difference in performance from my simpler 3-hidden-layer fully connected network.
-\item \textbf{DRL} and \textbf{DQL} heuristics, though not as efficient in terms node-expansion as the perfect heuristic and the \textbf{DL} heuristics, were rather surprisingly doing better than even Manhattan++.
-They ended up \textit{seeing} around 25\% of all the possible configurations during training (in an \textit{unsupervised} way though) so did show some reasonable amount of generalization out of sample.
-Their optimality score was an impressive 99\%+ at all levels of difficuly for \textbf{DRL} and 98\%+ at all levels of difficuly for \textbf{DQL}.
+I later realized that this is quite different from the typical image patterns recognition as there is less direct translation and scale invariants at play here, and decided instead to try combining convolutional with fully connected layers in parallel in order to appropriately value 2x2 patterns together with their position on the sliding-puzzle.
+That did improve things, but overall I saw little difference in performance from my simpler 3-hidden-layer fully connected network.
+\item \textbf{DRL} and \textbf{DQL} heuristics, though not as efficient in terms node-expansion as the perfect and the \textbf{DL} heuristics, were rather surprisingly doing better than Manhattan++.
+They ended up \textit{seeing} around 25\% of all the possible configurations during training (in an \textit{unsupervised} way though) so did show some reasonable amount of out-of-sample generalization.
+Their optimality score was an impressive 99\%+ at all levels of difficuly for \textbf{DRL} and 98\%+ for \textbf{DQL}.
 Even more interestingly, the \textbf{DQL} heuristic expanded almost 4 times less nodes than \textbf{DRL}, suggesting some benefit from the joint learning of the policy and value function.
@@
 
-blablabla
-\\
-blablabla
-\\
-blablabla
+For the 3x4 \textbf{SP}, I ran the performance test (100 random puzzles for difficulty going from 1 to 45 as well as $\infty$ shuffling) on the Naive solver and A$^{*}$ with Manhattan, Manhattan++, \textbf{DRL}, \textbf{DQL} and \textbf{DL} heuristics.
+Given the 239,500,800 total possible configurations there was no chance for me to compute the perfect heuristic as I did with smaller dimensions, so \textbf{DL} got trained on randomly generated configurations solved by A$^{*}$[Manhattan++].
+The network architecture for all \textbf{DxL} was the fully-connected (left-hand-side of figure \ref{fig:33SPnets}), mutatis-mutandis for the input size, which in the general case is $(n.m)^{2}$ when using one-hot-encoding (I always did after a few unsuccessful attempts without it).
+Detailed results can be found at \ref{fig:34SPPerformance}.
+Key take-aways are that all \textbf{DxL} outperformed A$^{*}$[Manhattan++] in terms of node expansion and got optimality scores of 99\%+ for all levels of difficulty, despite only seeing 0.009\% of all configurations for \textbf{DL}, and 0.27\% for \textbf{DRL} and \textbf{DQL}.
+Again this shows very impressive generalisation and performance from the\textbf{DxL} methods.
 
@@
   \noindent
-  \makebox[\textwidth]{\includegraphics[scale=0.6]{./Figures/34SPPerformance}}
+  \makebox[\textwidth]{\includegraphics[scale=0.55]{./Figures/34SPPerformance}}
   \caption[34SPPerformance]{Solvers' performance comparison 3x4 \textbf{SP}}
@@
 
+In the 4x4 case, due to time constraints, I ran the performance test with 100 random puzzles for difficulty going from 5 to 60 except for $\infty$ shuffling where I only ran a handful of configurations (5).
+I ran the Naive solver as well as A$^{*}$ with Manhattan++, \textbf{DRL} and \textbf{DL} heuristics, again using the same fully-connected network architecture (left-hand-side of figure \ref{fig:33SPnets}).
+The 4x4 \textbf{SP} has 10,461,394,944,000 total possible configurations so it does start being quite a large state space!
+Again \textbf{DL} got trained on randomly generated configurations solved by A$^{*}$[Manhattan++].
+\textbf{DL} obtained an optimality scores of 97\%+ across all levels of difficulty and \textbf{DRL} managed 99\%+.
+\textbf{DL} and \textbf{DRL} saw of the order of respectively $10^{-7}$\% and $10^{-5}$\% of all configurations during training, so again showed impressive performance and generalization!
+
 \label{ssec:44SPSC}
@@
   \noindent
-  \makebox[\textwidth]{\includegraphics[scale=0.55]{./Figures/44SPPerformance}}
+  \makebox[\textwidth]{\includegraphics[scale=0.6]{./Figures/44SPPerformance}}
   \caption[44SPPerformance]{Solvers' performance comparison 4x4 \textbf{SP}}

### 2022-09-11 latex/Chapters/Rubiks.tex
update

@@
 I have run reasonably comprehensive experiments on the 2x2x2, throwing 6 of my solvers at the problem.
-As mentioned in \ref{sec:Puzzles}, the 2x2x2 \textbf{RC}, in spite of its deceivingly simple physical appearance, is actually not that much easier to solve.
-For instance, it takes me usually just under 1 minute, whereras the 3x3x3 takes me only double that time, just under 2 minutes.
+As mentioned in \ref{sec:Puzzles}, the 2x2x2 \textbf{RC}, in spite of its deceivingly simple physical appearance, is actually not that much easier to solve (by hand) than the 3x3x3.
+For instance, it takes me usually just under 1 minute, wherras the 3x3x3 takes me only double that time, just under 2 minutes.
 I am obviously a beginner in Rubik's cube handsolving, but my personal ratio of time-to-solve and perceived difficulty is in line with that of experienced speedcubers who might take and average of, say, 2-3 seconds for the 2x2x2 and 10 seconds for the 3x3x3.
 \\
-Computing wise, the 2x2x2 was of decent size for my purpose, with its 88 million configurations, so in particular it 
-\\
-\\
-Let me start with a table summarizing which solvers I have attempted to run on the 2 dimensions I have looked at:
+Computing wise, the 2x2x2 was of decent size for my purpose, with its 88 million configurations.
+Let me start with a table summarizing which solvers I have attempted to run on the 2 dimensions I have attempted to experiment with:
 
@@
 BFS is the only one for which I used a step of 1 on the difficuly and stopped at 5 as it was clear it would not go any further.
-Given that the run time and expanded nodes are clearly going to increase linearly, and give where difficulty of 5 took this solver already, there was no point going any further!
+Given that the run time and expanded nodes are clearly going to increase linearly, and given where 5 scrambling from goal took this solver already, there was no point going any further!
 
 \Subsection{Kociemba, A$^{*}$[Kociemba], A$^{*}$[DL[A$^{*}$[Kociemba]]]}
-Next, let me discuss Kociemba (\ref{}) and associated experiments.
-We obviously expect Kociemba to be by far the fastest solver since it is handcrafted with much specific knowledge about the Rubik's cube.
+Next, let me discuss Kociemba (see \cite{HKociemba} and \cite{Kociemba}) and associated experiments.
+We obviously expect Kociemba to be by far the fastest solver since it is handcrafted with much specific knowledge about the Rubik's cube group theory and multi-stage solving which I discussed in \ref{Theory:RCMSS}.
+I found the 2x2x2 library I used to not be optimal however, despite claims to the contrary.
+I have not had the time to investigate why this was not the case, but empirically I found that for instance running A$^{*}$ with Kociemba's cost as a heuristic was resulting in higher quality (i.e.
+of lower cost) solutions, clearly showing the non-optimality of Kociemba to begin with.
+\\
+\noindent Similarly to the higher dimensions \textbf{SP} from the previous section, there was no hope for me to run a perfect solver on the 2x2x2 \textbf{RC}.
+For starter, I have no admissible heuristic to guarantee optimality!
+Even if I had found one such heuristic and managed to implement it, it is doubtful I could have solved all 88 millions configurations without considerable work and time to reduce the configurations via symmetry arguments.
+As a result, I used solutions from A$^{*}$[Kociemba] on (perfectly) randomly generated configurations to train my \textbf{DL} heuristic.
+Thinking of it, it is rather amusing since Kociemba is itself using a two-stage IDA$^{*}$ solver, so we end-up with 3 levels of nested A$^{*}$ algorithm, namely A$^{*}$[DL[A$^{*}$[IDA$^{*}$[2-stage-subgroups]]]].
 
@@
 \begin{figure}[H]
-\centering
-\includegraphics[scale=0.60]{./Figures/222RCPerformance}
-%\decoRule
+  \noindent
+  \makebox[\textwidth]{\includegraphics[scale=0.55]{./Figures/222RCPerformance}}
   \caption[222RCPerformance]{Solvers' performance comparison 2x2x2 \textbf{RC}}

### 2022-09-11 rubiks/scripts/rubiks22/perf.py

@@ -64,8 +64,8 @@ def performance(**kw_args):
 
 if '__main__' == __name__:
     logger = Loggable(name='Perf22RC')
-    action_type = Solver.do_performance_test
-    fig_size = (12, 12)
+    action_type = Solver.do_plot
+    fig_size = (12, 11)
     if action_type == Solver.do_plot:
         exclude_solver_names = ['1000seq_20shf']
         performance(**locals())

### 2022-09-11 latex/Chapters/Rubiks.tex
update

@@
 As a result, I used solutions from A$^{*}$[Kociemba] on (perfectly) randomly generated configurations to train my \textbf{DL} heuristic.
-Thinking of it, it is rather amusing since Kociemba is itself using a two-stage IDA$^{*}$ solver, so we end-up with 3 levels of nested A$^{*}$ algorithm, namely A$^{*}$[DL[A$^{*}$[IDA$^{*}$[2-stage-subgroups]]]].
+Thinking of it, it is rather amusing since Kociemba is itself using a two-stage IDA$^{*}$ solver, so I end-up here with 3 levels of nested A$^{*}$ algorithms, namely A$^{*}$[DL[A$^{*}$[IDA$^{*}$[2-stage-subgroups]]]].
+\\
+Notice finally that, since I have no optimal solver for the \textbf{RC}, I have set Kociemba as my benchmark in the performance test for the purpose of computing the optimality score.
+This is why some (actually all) of the other solvers will show scores larger than 100\% (top-right graph of figure \ref{fig:222RCPerformance}).
 
-\Subsection{Deep reinforcement learner}
+\Subsection{Deep Reinforcement Learner \& Deep Q Learner}
 \label{sec:S33DRL}
-Next I have run the DeepReinforcementLearner with the following main parameters:
+To get my \textbf{DRL} and \textbf{DQL} heuristics, I have run the DeepReinforcementLearner and DeepQLearner with the following main parameters:
 \begin{itemize}
@@
 \item Run for a maximum of 11,000 epochs (\textit{nb\_epochs=11000}) or when the maximum target so far has not been increasing by more than 1\% in the last 5,500 epochs (\textit{max\_target\_uptick=1e-2} and \textit{max\_target\_not\_increasing\_epochs\_pct=0.5})
-\item At every update of the target network (\textit{training\_data\_every\_epoch=False}), generate 100 sequences of puzzles (\textit{nb\_sequences=100}), each comprised of a 50-scramble puzzle back to goal state, with all intermediary puzzles along the path (\textit{nb\_shuffles=50}).
-\item Use a fully connected network (\textit{network\_type=fully\_connected\_net}) with three hidden layers of size 600, 300, 100 (\textit{layers\_description=(600,300,100)}) and use torch.optim.RMSProp as optimiser (\textit{optimiser=rms\_prop}) with an exponential scheduler (\textit{scheduler=exponential\_scheduler}) with gamma 0.999 (\textit{gamma\_scheduler=0.9999}).
+\item At every update of the target network (\textit{training\_data\_every\_epoch=False}), generate 10,00 sequences of puzzles (\textit{nb\_sequences=10,000}), each comprised of a 15-scramble puzzle back to goal state, with all intermediary puzzles along the path (\textit{nb\_shuffles=15}).
+\item Use a fully connected network (\textit{network\_type=fully\_connected\_net}) with three hidden layers of size 600, 300, 100 (\textit{layers\_description=(600,300,100)}) and use torch.optim.RMSProp as optimiser (\textit{optimiser=rms\_prop}) with an exponential scheduler (\textit{scheduler=exponential\_scheduler}) with gamma 0.999 (\textit{gamma\_scheduler=0.9999}) and \textit{learning\_rate=0.001}.
+The networks architecture is shown in figure \ref{fig:222RCnets}.
 \end{itemize}
-Below I show some of the DeepReinforcementLearner's quantities tracked over the epochs (see \ref{fig:33SPDeepReinforcementLearning}). I have shown only the following quantities:
+Graphs \ref{fig:222RCDRL} and \ref{fig:222RCDQL} show both learners' following metrics tracked over the epochs:
 \begin{itemize}
@@
 Then that means that after 15 updates of the target network we might not really get so much benefit by keeping training and value-iterating.
-\item the MSE loss
-\item the MSE loss divided by the max target.
+\item the loss (MSE for \textbf{DRL}, MSE + CrossEntropyLoss for \textbf{DQL})
+\item the loss divided by the max target.
 The exit criterion \textit{update\_target\_network\_threshold} is based on that quantity, since a loss of e.g.
@@
 
-\begin{figure}[H]
-\centering
-\includegraphics[align=c, scale=0.42]{./Figures/33SPDeepReinforcementLearning}
-\caption[33SPDeepReinforcementLearning]{\textbf{DRL} of the 3x3 \textbf{SP}}
-\label{fig:33SPDeepReinforcementLearning}
+\begin{figure}
+  \noindent
+  \makebox[\textwidth]{\includegraphics[scale=0.6]{./Figures/222RCnets}}
+  \caption[222RCnets]{2x2x2 \textbf{RC} architecture for \textbf{DRL} (left) \& \textbf{DQL} (right) heuristics training}
+  \label{fig:222RCnets}
+\end{figure}
+
+\begin{figure}
+  \noindent
+  \makebox[\textwidth]{\includegraphics[scale=0.5]{./Figures//222RCDRL}}
+  \caption[222RCDRL]{\textbf{DRL} of the 2x2x2 \textbf{RC}}
+  \label{fig:222RCDRL}
+\end{figure}
+
+\begin{figure}
+  \noindent
+  \makebox[\textwidth]{\includegraphics[scale=0.5]{./Figures//222RCDQL}}
+  \caption[222RCDQL]{\textbf{DQL} of the 2x2x2 \textbf{RC}}
+  \label{fig:222RCDQL}
 \end{figure}

### 2022-09-11 rubiks/learners/deepreinforcementlearner.py

@@ -77,6 +77,8 @@ class DeepReinforcementLearner(Learner):
                             puzzles_seen_pct]
     training_data_every_epoch = 'training_data_every_epoch'
     cap_target_at_network_count = 'cap_target_at_network_count'
+    fig_size = 'fig_size'
+    show_title = 'show_title'
 
     @classmethod
     def populate_parser_impl(cls, parser):
@@ -151,6 +153,15 @@ class DeepReinforcementLearner(Learner):
         cls.add_argument(parser,
                          field=cls.cap_target_at_network_count,
                          default=False,
+                         action=cls.store_true)
+        cls.add_argument(parser,
+                         cls.fig_size,
+                         type=int,
+                         nargs='+',
+                         default=[15, 12])
+        cls.add_argument(parser,
+                         field=cls.show_title,
+                         default=True,
                          action=cls.store_true)
 
     def more_model_name(self):
@@ -567,12 +578,13 @@ class DeepReinforcementLearner(Learner):
             title[field] = config[field]
         title = pformat(title)
         title = cls.__name__ + '\n' + title
-        fig = plt.figure(self.learning_file_name, figsize=(15, 10))
+        fig = plt.figure(self.learning_file_name, figsize=self.fig_size)
         plt.tick_params(axis='x', which='both', bottom=False, top=False, labelbottom=False)
         plt.tick_params(axis='y', which='both', left=False, right=False, labelleft=False)
         plt.axis('off')
         title = pformat(title).upper().replace('_', ' ')
-        plt.title(title, fontname='Consolas', fontdict={'weight': 'bold'})
+        if self.show_title:
+            plt.title(title, fontname='Consolas', fontdict={'weight': 'bold'})
         x = drl.epoch
         gs = fig.add_gridspec(len(self.plot_metrics), 1)
         axes = list()

### 2022-09-12 latex/Chapters/Rubiks.tex
update

@@
 
-Notice that the DeepLearner and DeepQLearner have very similar sets of parameters and track similar metrics during their learning for later display, but I shall not detail it here since there is not much extra insight to be had compared to the DeepReinforcementLearning example we just went through.
+\Subsection{Solvers' comparison}
 
-\Subsection{Solvers' comparison}
+We summarize as usual the 2x2x2 \textbf{RC} performance tests by our usual metrics \ref{fig:222RCPerformance}.
+Overall, the \textbf{DL} and \textbf{DRL} managed to solve all cubes (though \textbf{DL} took over an houd and \textbf{DRL} up to 6 hours for the slowest cube they encountered.
+Both methods outperformed in terms of optimality score Kociemba and got on par with A$^{*}[Kociemba]$.
+Their max cost over all cubes encountered was 12 for up to 20 shuffles, and 13 for $\infty$ shuffling.
+Considering God's number for the 2x2x2 \textbf{RC} of 11, this is rather impressive!
+\\
+\\
+An very interesting remark is that if we check the learning graph \ref{fig:222RCDRL} of the \textbf{DRL}, we can see it only ever assigned a cost-to-go (value function) of 7.6 at most.
+Despite that fact, it was still able to solve cubes requiring 13 moves.
+How is that possible?
+Notice that the loss of 0.2 towards which it converged at the end of its training is misleading, since it is not a loss versus the actual true cost-to-go, but rather versus that of the target network.
+My theory is that while the value the \textbf{DRL} assigned to hard-to-solve cubes is off (by at least 4 or 5), it is likely the case that the network gets the relative value of neighbouring cubes about right, and this is sufficient to guide A$^{*}$ extremely well.
+\\
+\\
+Finally, we can see that \textbf{DQL}, despite showing seemingly/visually cleaner and more monotonic loss convergence than \textbf{DRL} (see \ref{fig:222RCDQL} versus \ref{fig:222RCDRL}), struggled a lot more to solve difficult cubes (above 10 shuffles).
+Clearly this shows that my earlier findings on the 3x3 \textbf{SP} are likely not universal, and very much depend on the parameters set during training, as well as on the problem at hand.
 
@@
 
-For now just showing Kociemba as a baseline.
-Will first try out DL trained on data generated from Kociemba, then will try out DRL.
-If unsuccessful or too slow, might try similar to paper using DQL/MC search.
-\\
-Notice Kociemba 3, unlike the 2x2x2 implementation that I found on github, does not do automatic color mapping.
-So if you present it with a cube which is not with the \textit{standard} centers (facing red and white up), it will actually not solve the cube.
-I have therefore added a bit of logic in Kociemba solver to look for the equivalent cube among the 24 equivalent cubes to the one being solved, with standard colors.
-We then solve this one using Kociemba and add the whole cube rotations to the solution (which in my code are deemed to have 0 cost anyway).
-This way, not only can I use Kociemba 3 to solve any (solvable) 3x3x3 cube irrespective of rotations, but I can also use that in A* or to train a DL network (also for A*)
+I did not get to run very many experiments on the 3x3x3 \textbf{RC}.
+The results of the performance test (50 random configurations for levels of difficulty from 2 to 40 by increment of 2 as well as $\infty$ shuffling) can be found on figure \ref{fig:333RCPerformance}.
+\begin{itemize}
+\item Here again, Kociemba does not show optimality, since we can see that my A$^{*}[Kociemba]$, that is A$^{*}$ using Kociemba as a heuristic, performs much better than Kociemba (solutions about 20\% shorter at $\infty$ shuffling).
+This is not surprising though since Kociemba 3x3x3 does not claim to be optimal and is meant to find solution in the 40 moves at most (which is what I have observed indeed).
+A$^{*}[Kociemba]$ never took more than 30 moves on any of the cubes presented to it, which is not bad given the known God number of 26 in the quarter-turn-metric (as discussed earlier in \ref{Theory:333RCSSS})
+\item A$^{*}$[\textbf{DL}] managed to solve all puzzles scrambled less than 10 times, and found shorter worst case solutions than even A$^{*}[Kociemba]$ on which it was trained.
+Within the 2 hours time out, it failed to solve one of the 50 cubes for difficulty 10, and two of the 50 cubes of difficulty 12, so i did not try further than that.
+\item A$^{*}$[\textbf{DRL}] only managed to solve all puzzles of difficulty less than 6 but at least did find only optimal solutions (compared with \textbf{BFS}).
+\end{itemize}
 
 \begin{figure}[H]
-\centering
-\includegraphics[scale=0.50]{./Figures/333RCPerformance}
-%\decoRule
+  \noindent
+  \makebox[\textwidth]{\includegraphics[scale=0.55]{./Figures/333RCPerformance}}
   \caption[333RCPerformance]{Solvers' performance comparison 3x3x3 \textbf{RC}}

### 2022-09-12 latex/Chapters/Conclusion.tex
update

@@
 I am now able to solve the 2x2x2 under one minute, 3x3x3 usually around 2 minutes and the 4x4x4 in about 5 minutes.
-More specifically regarding the techniques discussed in this project, the principles that underly uninformed and informed search, A$^{*}$ and heuristics, multi-stage solvers, and of course the various deep learning methods I have played with (\textbf{DL}, \textbf{DRL} and \textbf{DQL}) all extend to many other puzzle and dimensions and are a great new tool-set to add to my knowledge and experience.
+More specifically regarding the techniques discussed in this project, the principles that underly uninformed and informed search, A$^{*}$ and heuristics, multi-stage solvers, and of course the various deep learning methods I have played with (\textbf{DL}, \textbf{DRL} and \textbf{DQL}) all extend to many other puzzles and dimensions and are a great new addition to my tool-set and knowledge.
 
@@
 \Section{Professional issues}
-Bug in Kociemba and software design\\
-AI safety 
-See \cite{ManyWorlds}
-See \cite{https://doi.org/10.1111/1758-5899.12718} and \cite{Bostrom2014}
+My favourite principle within the BCS code of conduct (\cite{BCS}) is to `show what you know, learn what you don't'.
+The whole point of this MSc in AI, after a 20 year career, most of which spent in quantitative trading, was indeed for me to revise old skills, learn some new ones and further my knowledge in \textbf{ML} and \textbf{AI}.
+That has the `learn what you don't' covered.
+As for the `show what you know', I have endeavoured to show honest, and therefore in some cases modest, results.
+Ideally I would have liked to be able to try more methods and go further with the \textbf{DxL} methods on the 3x3x3 \textbf{RC} as well as higher dimensional \textbf{SP} like the 5x5, but this was sadly not possible within the time imparted.
+\\
+\\
+As for issues related to ethics and safety, some of which are outined in the ACM code of ethics (\cite{ACM}), anyone working in \textbf{AI} should in my opinion be acutely aware of them.
+There is clearly some divide in the community regarding the dangers of \textbf{AI}, but I myself stand firmly in the Bostrom camp (see in particular \cite{https://doi.org/10.1111/1758-5899.12718} and \cite{Bostrom2014}).
+His paper-clip-maximiser is one of the most thought provoking ideas and does show that the most benign looking \textbf{AI} agent, if sufficiently capable, could create disproportionate harm.
+To conclude on a humorous (?) note, we probably would not want either to see a Rubik's solver engulf all of the earth's and galaxy's resources in order to compute the nxnxn Rubik's God number for ever growing values of n.

### 2022-09-12 latex/Abstract/Abstract.tex
update

@@
-\noindent
-\textbf{Motivation}
 \label{Abstract}
-\\
-Reinforcement Learning (\textbf{RL}), exposed with brilliant clarity in \cite{Sutton1998}, has until recently known less success than we might have hoped for.
+\noindent \textbf{Motivation} -- Reinforcement Learning (\textbf{RL}), exposed with brilliant clarity in \cite{Sutton1998}, has until recently known less success than we might have hoped for.
 Its framework is very appealing and intuitive.
@@
 Sadly, it had until recently proven hard to achieve practical success with these methods.
+Inspired however by the seminal success of Deep Mind's team in using Deep Reinforcement Learning (\textbf{DRL}) to play Atari games (\cite{Mnih2013})  and to master the game of Go (\cite{AlphaGo}), researchers have in recent years made a lot of progress towards designing algorithms capable of learning and solving, \textit{without human knowledge}, the Rubik's Cube (\textbf{RC}) - as well as similar single player puzzles - by using Deep Q-Learning (\textbf{DQL})  (\cite{DBLP:journals/corr/abs-1805-07470}) or search and (\textbf{DRL}) value iteration (\cite{https://doi.org/10.48550/arxiv.1805.07470}).
 \\
-Inspired however by the seminal success of Deep Mind's team in using Deep Reinforcement Learning (\textbf{DRL}) to play Atari games (\cite{Mnih2013})  and to master the game of Go (\cite{AlphaGo}), researchers have in recent years made a lot of progress towards designing algorithms capable of learning and solving, \textit{without human knowledge}, the Rubik's Cube (\textbf{RC}) - as well as similar single player puzzles - by using Deep Q-Learning (\textbf{DQL})  (\cite{DBLP:journals/corr/abs-1805-07470}) or search and (\textbf{DRL}) value iteration (\cite{https://doi.org/10.48550/arxiv.1805.07470}).
 \\
@@
 \\
-\textbf{Organisation of this thesis}
+\textbf{Organisation of this thesis} -- In the first section, I describe what I hope to get out of this project, both in terms of personal learning, as well as in terms of tangible results (solving some puzzles!).
+In section 2, I will do a recap of the different methods used in the project.
+Section 3 is dedicated to discussing the mathematics of the sliding puzzle \textbf{SP} and the \textbf{RC}.
+I might make interesting observations in there and give some references for the keen reader.
+I give then an overview in section 4 of the code base I have developed (see my github page \cite{FB}) - for this project.
+Finally, sections 6 and 7 present my results on respectively the \textbf{SP} and the \textbf{RC} and section 8 concludes the thesis.
 \\
-In the first section, I will quickly describe what I hope to get out of this project, both in terms of personal learning, as well as in terms of tangible results (solving some puzzles!).
-In section 2, I will do a quick recap of the different methods that I will use in the project.
-Section 3 will be dedicated to discussing the mathematics of the sliding puzzle \textbf{SP} and the \textbf{RC}.
-I might throw a few random (but hopefully interesting) observations in there and give some references for the keen reader.
-I will then give an overview in section 4 of the code base I have developed - and put in the open on my github page (\cite{FB}) - to complete this project.
-Finally, in sections 6 and 7, I will present all my various results on respectively the sliding puzzle and the Rubik's cube before offering some conclusion in section 8.
 \\
 A couple of appendix sections complement the main text.
-The first one details how to use the code base, from constructing puzzles, to learning and solving them using all the different methods I have implemented for this project.
-The second describes a limitation I bumped into while using the kociemba library (a well-known open source 3x3x3 \textbf{RC} solver), and how I circumvented it.
+The first one details how to use the code base, from constructing puzzles, to learning and solving them using the various solvers I have developed.
+The second one describes a limitation I bumped into while using the kociemba library (a well-known open source 3x3x3 \textbf{RC} solver), and how I circumvented it.
+\newpage
+\thispagestyle{empty}
+\noindent \textbf{Acknowledgements} -- 
+\input{Acknowledgements/Acknowledgements}

### 2022-09-12 latex/Chapters/Rubiks.tex
update

@@
 One interesting dynamic during learning is that due to the way targets are constructed using value-iteration update, and starting from an all 0 network, the max target only increases by about 1 at best at every network update.
-That is, the first network only learns how to differentiate goal from non-goal, then the second one learns how to differentiate goals, cost 1 and cost 2+ in some sense.
-This is why I have introduced the first set of parameters discussed just above.
-The first one makes sure we do not waste time initially training the left-hand-side network when training is easy (as consist of just goal versus not goal).
-The second ones make sure that we stop training altogether when the max target does not grow anymore.
-This is because we do not know a-priori what the God number of a given puzzle/dimension is, but let's assume for the sake of the argument that it is 15.
-Then that means that after 15 updates of the target network we might not really get so much benefit by keeping training and value-iterating.
+That is, the first network only learns how to differentiate goal from non-goal, then the second one learns how to differentiate goals, cost 1 and cost 2+ \dots.
+This is why I have introduced the \textit{update\_target\_network\_threshold} discussed above.
+It makes sure we do not waste time initially training the left-hand-side network when training is easy.
+\textit{max\_target\_uptick} and \textit{max\_target\_not\_increasing\_epochs\_pct} on the other hand make sure that we stop training altogether when the max target does not grow anymore.
+This is because we do not know a-priori what the God number of a given puzzle/dimension is.
+Let us assume for the sake of the argument that it is 15: that means that beyond 15 updates of the target network we might not get much benefit to additional training and value-iterating.
 \item the loss (MSE for \textbf{DRL}, MSE + CrossEntropyLoss for \textbf{DQL})
@@
 0.01 means little if the we do not compare it to the possible range that the cost-to-go can take.
-\item The percentage of all the possible puzzles of that dimension \textit{seen} by the DeepReinforcementLearner.
-This is an interesting quantity, as it tells us later on when testing out of sample and comparing with other solvers, how well the solvers are able to generalise.
-In this case (3x3) we can see that the DeepReinforcementLearner quickly has seen a very large percentage of all puzzles.
-That being said, it is good to keep in mind that \textit{seen} does not mean really much here, since the learning is unsupervised.
-This is quite in contrast to the DeepLearner, for which puzzles seen come along with the cost-to-go computed by a \textit{teacher}, usually an optimal or at least efficient solver.
+\item The percentage of all the possible puzzles of that dimension \textit{seen} by the learner.
+This is an interesting quantity, as it tells us how well the solvers are able to generalise.
+From both graphs (\ref{fig:222RCDRL} \& \ref{fig:222RCDQL}) we can see that the learners saw less than 1\% of all puzzles.
+It is good to keep in mind however that \textit{seen} does not mean really much here, since the learning is \textit{unsupervised}.
+This is quite in contrast to the DeepLearner, whose traiing data consist of puzzles with their cost-to-go computed by a \textit{teacher} (optimal or other efficient solver).
 \end{itemize}
@@
 Notice that the loss of 0.2 towards which it converged at the end of its training is misleading, since it is not a loss versus the actual true cost-to-go, but rather versus that of the target network.
-My theory is that while the value the \textbf{DRL} assigned to hard-to-solve cubes is off (by at least 4 or 5), it is likely the case that the network gets the relative value of neighbouring cubes about right, and this is sufficient to guide A$^{*}$ extremely well.
+My theory is that while the value the \textbf{DRL} assigned to hard-to-solve cubes is off (by at least 4 or 5), it is likely the case that the network gets the relative value of neighbouring cubes about right (given that the value-iteration teaches the solver that, it is not super surprising either), and this is sufficient to guide A$^{*}$ extremely well.
+If both the absolute value of the cost-to-go and the relative (among neighbouring cubes) was totally off, we would indeed not expect the heuristic to make A$^{*}$ any better than, say, \textbf{BFS} and we would therefore not expect it to be able to solve \textit{costly} cubes in 10-13 moves.
 \\

### 2022-09-12 latex/Chapters/Objectives.tex
make both RHUL template and my own thesis format coexist .... I d like to have both formats
trimming objectives

--- 2022-06-26 latex/Chapters/Objectives.tex
@@
 
-\section{Learning Objectives}
+\Section{Learning Objectives}
 
-Back when I studied Financial Mathematics, almost 2 decades ago, it was all about probability theory, stochastic calculus and asset (in particular derivatives) pricing.
-These skills were of course very sought after in the field of options trading, but were also often enough to get a job in algorithmic or systematic trading.
-By the middle of the 2010s, with the constant advances in computing power and storage, the better availability of off-the-shelves libraries and data sets, I witnessed a first revolution: the field of machine learning became more and more prominent and pretty much overshadowed other (more traditional maths) skills.
-More recently, a  second revolution has taken not only the world of finance, but that of pretty much every science and industry, by storm: we are now in the artificial intelligence age.
-In 2019-2020, I decided it was time to see by myself what this was all about, and if the hype was justified. What better way to do that than embark on a proper MSc in Artificial Intelligence?
+Back when I studied Financial Mathematics, almost 2 decades ago, it was all about probability theory, stochastic calculus and asset pricing.
+These skills were sought after in the field of derivatives trading, and often enough to get a job in algorithmic or systematic trading.
+By the middle of the 2010s, with the constant advances in computing power and storage, the better availability of off-the-shelves libraries and data sets, I witnessed a first revolution: machine learning (\textbf{ML}) became more prominent and started overshadowing more traditional maths skills.
+In recent years, a second revolution has taken not only the world of finance, but that of every science and industry, by storm: we are now in the artificial intelligence (\textbf{AI}) age.
+In 2019, I decided to see by myself what this was all about, and if the hype was justified.
+What better way to do that than embark on a proper MSc in Artificial Intelligence?
 \\
-Of all the modules I have studied over the last two years of the Royal Holloway MSc in AI, I have been the most impressed by DL and  NLP (itself arguably largely an application of DL) and particularly interested in AIPnT, especially our excursion in the field of graphs search (a very traditional CS topic, but which somehow I had not yet had a chance to study in much details).
-Even though I still believe there is a tremendous amount of malinvestment  everywhere, due in good part to the inability of the average investor to distinguish between serious and scammy AI applications and startups (the same obviously goes for blockchain applications, which might warrant another MSc?), I have totally changed my mind around the potential of DL, DRL and NLP and think they are incredibly promising.
-I have been astonished to see by myself, through several of the courseworks we have done during the MSc, how incredibly efficient sophisticated ML, DL and DLR algorithms can be, when applied well on the right problems.
-Sometimes they just vastly outperform more naive and traditional approaches to the point of rendering older approaches entirely obsolete (e.g CV, NLP, game solvers, etc...).
 \\
-For the project component of the MSc, I thought it would be interesting (and fun) for me to try and apply some of the DL, DRL and search techniques (from AIPnT) to a couple of single-player games, such as the sliding puzzle (of which some variations are well known under different names, e.g.
-the 8-puzzle and 15-puzzle) and of course the Rubiks' cube.
-I am in particular looking to solidify my understanding of DRL by  implementing and experimenting with concrete (though arguably of limited practical use) problems.
+Of the modules I have studied over the last two years, I have been most impressed by \textbf{DL},  Natural Language Processing (\textbf{NLP}) and particularly interested in AI Principles \& Techniques (\textbf{AIPnT}), especially our excursion in graphs search.
+I have totally changed my mind around the potential of \textbf{DL}, \textbf{DRL} and \textbf{NLP} and think they are incredibly promising.
+I have been astonished to see by myself, through several of the MSc courseworks, how incredibly efficient sophisticated \textbf{ML}, \textbf{DL} and \textbf{DRL} algorithms can be, when applied well on the right problems, often vastly outperforming more naive and traditional approaches.
+\\
+\\
+For the project component of the MSc, I thought it would be interesting (and fun) for me to apply some of the \textbf{DL}, \textbf{DRL} and search techniques from \textbf{AIPnT} to the sliding puzzle (\textbf{SP}) and of course the famous Rubik's cube (\textbf{RC}).
+I am in particular looking to solidify my understanding of \textbf{DRL} and \textbf{DQL} by  implementing and experimenting with concrete (though arguably of limited practical use) problems.
 
@@
 
-\newpage
-\section{Project's Objectives}
+\Section{Project's Objectives}
 
-I am hoping with this project to implement and compare a few different methods to solve the SP and the RC.
-Both these puzzles have tremendously large state spaces (see section \hyperref[sec:Games]{Games} for details) and only one goal state.
-I am therefore likely to only succeed with reasonably small dimensional puzzles, especially since I have chosen for simplicity to implement things in Python.
-Depending on the progress I will be able to make in the imparted time, I am hoping to try a mix of simple searches (depth first search, breadth first search, A* with simple admissible heuristics), then more advanced ones such as A* informed by heuristics learnt via DL and DRL, as well as try different architectures and network sizes and designs for the DL and DRL heuristics.
-Time permitting I would like to give a go at DQL, and maybe also compare things with some open-source domain-specific implementations (for instance a Kociemba Rubik's algorithm implementation, see e.g.
+I am hoping to try a mix of simple searches such as depth first search (\textbf{DFS}), breadth first search (\textbf{BFS}) and A* with simple admissible heuristics, then move to more advanced ones such as A* informed by heuristics learnt via \textbf{DL} and \textbf{DRL}, as well as try different network architectures.
+Time permitting I would like to give a go at \textbf{DQL}, and maybe also compare things with some open-source domain-specific implementations (for instance a Kociemba Rubik's algorithm implementation, see e.g.
 \cite{Kociemba}).
 \\
-Along the way, I am also hoping to learn a bit about these two games that I have chosen to work on, and maybe make a couple of remarks on them that the reader of this thesis might find interesting.
+\\
+Along the way, I am also hoping to learn a bit about the mathematics of the \textbf{SP} and the \textbf{RC}.

### 2022-09-12 latex/Chapters/DRLSearch.tex
trimming theory

@@
 
-In this chapter, I will succinctly go through the different techniques that underly the algorithms I have implemented and compared in this project and present them in (very simplified) pseudo code.
+In this chapter, I succinctly go through the different techniques underlying the algorithms I have implemented for this project and present them in simplified pseudo-code.
 Along the way, I will give some references for the interested reader.
@@
 
-It is quite fruitful to think of single-player, deterministic, fully-observable puzzles such as the \textbf{SP} and \textbf{RC} as unit-cost graphs.
+It is fruitful to think of single-player, deterministic, fully-observable puzzles such as the \textbf{SP} and \textbf{RC} as unit-cost graphs.
 We start from an initial \textit{node} (usually some scrambled configuration of the \textbf{SP} or \textbf{RC}), and via legal moves, we transition to new nodes, until hopefully we manage to get to the goal in the shortest number of moves possible.
@@
 \\
-Search strategies typically maintain a frontier, which is the collection of unexpanded nodes so far.
-They keep expanding the frontier, by choosing the next node to expand (expanding a node simply means adding all of its children nodes - i.e.
+Search strategies typically maintain a frontier, which is the collection of unexpanded nodes.
+They keep expanding the frontier, by choosing the next node to process (by adding all of its children nodes - i.e.
 those which are reachable via legal transitions - to the frontier for future evaluation) until a solution is found.
 The question is how to choose the next node to expand!
-Some graphs search strategies are said to be \textit{uninformed}, in the sense that they do not exploit any domain-specific knowledge or insight about what the graphs represent to choose the next node to expand.
-In that category fall for instance Breadth First Search (\textbf{BFS}) and Depth First Search (\textbf{DFS}), which as their name suggest expand nodes from the frontier based on, respectively, a \textbf{FIFO} and \textbf{LIFO} policy.
+Some graphs search strategies are said to be \textit{blind}, in the sense that they do not exploit any domain-specific knowledge or insight about what the graphs represent to choose the next node to expand.
+In that category fall Breadth First Search (\textbf{BFS}) and Depth First Search (\textbf{DFS}), which as their name suggest expand nodes from the frontier based on, respectively, a \textbf{FIFO} and \textbf{LIFO} policy (see pseudo-code \ref{alg:TheoryBFSDFS}).
 \textbf{BFS} and \textbf{DFS} will only work for modest problems where the number of transitions and states are reasonably small; they can however work in a wide variety of situations as they make no assumptions and require no knowledge.
@@
 \\
+\noindent \textit{Informed} strategies (see pseudo-code \ref{alg:TheoryAStar}) on the other hand exploit domain-specific knowledge.
+One popular algorithm is \textbf{A$^{*}$} (see \cite{DBLP:journals/jacm/DechterP85}), which first examines the node of smallest expected total cost (\textbf{f}), itself the sum of the cost-so-far from initial to current node (\textbf{g}) plus the expected cost-to-go from current node to goal (\textbf{h}).
+The expected cost-to-go \textbf{h} is called a \textit{heuristic}.
+The better the heuristic, the better \textbf{A$^{*}$} performs.
+An important property of heuristics is \textit{admissibility}: admissible heuristics never over-estimate the real cost-to-go, and can easily be shown to render \textbf{A$^{*}$} optimal.
 \teal
 \begin{algorithm}[H]
-\caption{Uninformed Graphs Search -- BFS \& DFS}\label{alg:TheoryBFSDFS}
+\caption{Blind Graphs Search -- BFS \& DFS}\label{alg:TheoryBFSDFS}
 \begin{algorithmic}
@@
 \black
-\noindent \textit{Informed} strategies, on the other hand, exploit domain-specific knowledge.
-One popular such strategy is \textbf{A$^{*}$} (see \cite{DBLP:journals/jacm/DechterP85}), which first examines the node of smallest expected total cost (\textbf{f}), itself the sum of the cost-so-far from initial to current node (\textbf{g}) plus the expected cost-to-go from current node to goal (\textbf{h}).
-The expected cost-to-go \textbf{h} is often called a \textit{heuristic}.
-The better the heuristic is, the better \textbf{A$^{*}$} usually performs.
-An important property of heuristics is \textit{admissibility}: admissible heuristics never over-estimate the real cost-to-go, and can easily be shown to render \textbf{A$^{*}$} optimal.
 \teal
@@
 
-As mentioned in the abstract, \textbf{RL} (see \cite{Sutton1998} for a brilliant exposition of the basic concepts and theory) has known a bit of false start in the 1950s and 1960s but recently been extremely successfully applied to a variety of problems (from Atari to board games and puzzles, robotics and else).
-\textbf{RL} is concerned with how intelligent agents can learn optimal decisions from observing/experiencing rewards while interacting in their environment.
-One of the fundamental concept in the field is that of value function $s \to \textbf{V}(s)$, which tells us the maximum expected reward - or equivalently and better suited to our puzzle solving task, the minimum cost - the agent can obtain from a given state $s$ (if it takes optimal decisions).
-In finite state and transitions spaces, the value function can be remarkably computed by a rather mechanical and magic-like procedure called value-iteration (kind of the equivalent of the Bellman equation from optimal control), where each state $s$'s value is iteratively refined by combining the value of $s$'s reachable children states.
+\textbf{RL} (see \cite{Sutton1998} for a brilliant exposition of the basic concepts and theory) has known a bit of false start in the 1950s and 1960s but recently been extremely successfully applied to a variety of problems (from Atari to board games and puzzles, robotics, \dots).
+\textbf{RL} is concerned with how intelligent agents can learn optimal decisions from experiencing rewards while interacting in their environment.
+One of the fundamental concept in the field is that of value function $s \to \textbf{V}(s)$, which tells us the maximum expected reward - or equivalently and better suited to our puzzle solving task, the minimum cost - the agent can obtain from a given state $s$ (assuming optimal decisions).
+In finite state and transitions spaces, the value function can be remarkably computed by a mechanical, rather magic-like procedure called value-iteration (equivalent to Optimal Control's Bellman equation, see pseudo-code below \ref{alg:RLTheory}), where each state $s$'s value is iteratively refined by combining the value of $s$'s reachable children states.
 
@@
 
-Quite similarly to \textbf{RL}, Deep Learning (\textbf{DL}) has not initially had the success the \textbf{AI} community was hoping it would.
+Similarly to \textbf{RL}, \textbf{DL} has not initially had the success the \textbf{AI} community was hoping it would.
 Marvin Minsky, who often took the blame for having \textit{killed} funding into the field, even regretted writing his 1969 book (\cite{minskypapert69}) in which he had proved that three-layer feed forward perceptrons were not quite the universal functions approximators some had hypothesized them to be.
 Since around 2006 (see \cite{GoodBengCour16} for more history of the recent burgeoning of \textbf{DL}), the field has known a rebirth, probably due to a combination of factors, from ever more powerful computers and larger storage, the availability of data sets large and small on the internet, the advances in many techniques and heuristics (backward propagation, normalisation, drop-outs, different optimisation schemes, etc \dots) and the huge amount of experimentation with different architectures (from very deep feed forward fully connected networks, to convolutional, recurrent and more exoctic networks).
-It is not an exageration to say that \textbf{DL} has rendered obsolete entire fields of research and bodies of knowledge, most notably in \textbf{CV}, robotics, computational biology and \textbf{NLP}.
+It is not an exageration to say that \textbf{DL} has revolutionized entire fields of research such as computer vision, robotics, computational biology and \textbf{NLP}.
 \textbf{DL} has often become the method of choice to learn or approximate highly dimensional functions or random processes.
@@
 So what is the link between all of these: A$^{*}$'s heuristics, \textbf{DL}, \textbf{RL}'s value iteration?
-\\
-\\
-Sometimes we have ways of computing good solutions/heuristics to our puzzles, but cannot computationally do this for all possible states (maybe the state space is simply too large!).
-One approach in that case is to compute the solution/heuristic for \textit{some manageable number of} states, and let a \textbf{DL} model learn how to extrapolate to the rest of the state space.
-This is a case of \textit{supervised} learning in some sense, since we need a teacher-solver to tell us good solutions or heuristics to extrapolate from.
-In very simplified pseudo code, the procedure looks like the following:
+Sometimes we have ways of computing good solutions/heuristics to our puzzles, but cannot computationally do this for all states (the state space might be too large!).
+One approach in that case is to compute solutions for \textit{some manageable number of} states, and let a \textbf{DL} model learn how to extrapolate to the whole state space.
+This is a case of \textit{supervised} learning, since we need a teacher-solver to tell us good solutions or heuristics to extrapolate from.
+Pseudo-code is provided below, see \ref{alg:TheoryDLDRL}:
 
@@
 \black
-
-For this project, I have implemented the above procedure of training a \textbf{DL} network from the solutions of other solvers/heuristics in a class called DeepLearner (see section \ref{sec:codelearners} for details).
-\\
-\\
-An alternative approach in order to compute the cost-to-go is via \textbf{RL} value-iteration as described in the previous section (\ref{sec:RLTheory}).
+\noindent An alternative approach to compute the cost-to-go is via \textbf{RL} value-iteration as described in the previous section (\ref{sec:RLTheory}).
 The state space is often so large however - and this certainly is the case for pretty much all \textbf{SP} and \textbf{RC}, except maybe the smallest dimensional \textbf{SP}s - that we cannot practically perform the value iteration procedure for all states.
 This is where \textbf{DL} comes again to the rescue to act as function approximator, combining with \textbf{RL} into what has become known as \textbf{DRL}.
-The subtelty is then that both the left-hand-side and right hand side in the value-iteration procedure are given by a \textbf{DL} network which we are constantly tweaking.
+The subtlety is that both the left-hand-side and right hand side in the value-iteration procedure are given by a \textbf{DL} network which we are constantly tweaking.
 For my implementation of \textbf{DRL}, I followed the same approach as in \cite{Mnih2013}, utilising two loops.
@@
 The outer loop is itself broken out of when deemed appropriate (some kind of convergence criteria).
-In simplified pseudo-code, this looks like:
+See pseudo-code \ref{alg:TheoryDRL} below:
 \\
 \\
-
 \teal
@@
 
-For this project, I have implemented the above algorithm in a class called DeepReinforcementLearner which I shall describe in details in section \ref{sec:codelearners}.
-
 %----------------------------------------------------------------------------------------
@@
 \end{equation}
-where $\mathcal{A}(s)$ are the transitions/actions available from state $s$. Conversely, we can recover $Q$ from $V$ since:
+where $\mathcal{A}(s)$ are the actions from state $s$. Conversely, we can recover $Q$ from $V$ since:
 \begin{equation} \label{eq:QI2}
@@
 \black
-
-\noindent For this project, I have implemented the above algorithm in a class called DeepQLearner which I shall describe in details in section \ref{sec:codelearners}.
 
@@
 
-\noindent I have implemented the exact \textbf{MCTS} procedure described in \cite{https://doi.org/10.48550/arxiv.1805.07470} and outlined above (\ref{alg:TheoryMCTS}) in a class called MonteCarloTreeSearch (see section \ref{sec:codesolvers} for details).
+\noindent Note that for this project, I have implemented all the above algorithms in Python (BFS, DFS, A$^{*}$, DeepLearner, DeepReinforcementLearner, DeepQLearner, MonteCarloTreeSearch), and more details can be found in section (\ref{sec:Implementation}).

### 2022-09-12 latex/Chapters/Puzzles.tex
update
trimming puzzles

@@
 
-\noindent It is rather easy to see why this is the case, let us discuss why: given a configuration c of the tiles, let us define the permutation p(c) of this configuration according to the following schema: we enumerate the tiles row by row (top to bottom), left to right for odd rows and right to left for the even rows, ignoring the empty compartment.
-For instance, the following 15-puzzle c:
+\noindent It is easy to see why this is the case, let me discuss why: given a configuration $c$ of the tiles, let us define the permutation $p(c)$ corresponding to enumerating the tiles row by row (top to bottom), left to right for odd rows and right to left for the even rows, ignoring the empty compartment.
+For the following left-hand-side15-puzzle $c_{l}$ we have $p(c_{l})$ = (1, 6, 2, 3, 4, 7, 10, 5, 9, 15, \red 14 \black, \blue 11, 8\black, 12, 13):
 
@@
 \end{fifteen}
-\\
-\\
-we have p(c) = (1, 6, 2, 3, 4, 7, 10, 5, 9, 15, \red 14 \black, \blue 11, 8\black, 12, 13).
-\\
-It is easy to see that the parity of p(c) cannot change by a legal move of the puzzle.
-Indeed, p(c) is clearly invariant by lateral move of a tile, so its parity is invariant too.
-A vertical move of a tile will displace a number in p(c) by an even number of positions right or left.
-For instance, moving tile 14 into the empty compartment below it results in a new configuration $c_{2}$:
-
+% 
 \begin{fifteen}
@@
 \\
-\\
-with $p(c_{2})$ = (1, 6, 2, 3, 4, 7, 10, 5, 9, 15, \blue11, 8\black, \red14\black, 12, 13), which is equivalent to moving 14 by 2 positions on the right.
+The parity $p(c)$ of a configuration $c$ cannot change via a legal move.
+Indeed, p(c) is clearly invariant by lateral move of a tile, so its parity is invariant too.
+A vertical move of a tile will displace a number in $p(c)$ by an even number of positions right or left.
+For instance, moving tile 14 above in $c_{l}$  into the empty compartment below it results in a new configuration $c_{r}$ with $p(c_{r})$ = (1, 6, 2, 3, 4, 7, 10, 5, 9, 15, \blue11, 8\black, \red14\black, 12, 13), which is equivalent to moving 14 by 2 positions on the right.
 This obviously cannot change the parity since exactly 2 pairs of numbers are now in a different order, that is (14, 11) and (14, 8) now appear in the respective opposite orders as (11, 14) and (8, 14).
-\\
-\\
 This is the crux of the proof of the well known necessary condition (even parity of p(c)) for a configuration c to be solvable (see part I of \cite{Johnson1879}).
-\\
-\\
-In the case of the coiled puzzle, we can clearly solve all configurations of even parity, since all the legal moves of the normal puzzle are allowed.
+In the case of the coiled puzzle, we can solve all configurations of even parity, since all the legal moves of the normal puzzle are allowed.
 In addition, we can for instance transition between the following 2 configurations, which clearly have respectively even and odd parities:
@@
 \Subsection{History -- from Erno Rubik to speed cubing competitions}
-The second puzzle I will focus on is the well known Rubik's Cube (\textbf{RC}), invented in 1974 by Erno Rubik, originally known under the name of \textit{Magic Cube} (see \cite{RubiksWiki}).
-Since it was commercialized in 1980, it has kown a worldwide success: countless people are playing with the famous original 3x3x3 cube, as well as with many variants of it, more or less difficult and contrived.
-Competitions nowadays bring together \textit{speedcubers} who have trained for years and memorized various algorithms to solve the Rubik's in astonishly effective and fast times (literally in seconds for the best ones).
-Interestingly, some of the principles used by speedcubers generalise to different dimensions.
+The second puzzle I will focus on is the Rubik's Cube (\textbf{RC}), invented in 1974 by Erno Rubik, originally known as \textit{Magic Cube} (see \cite{RubiksWiki}).
+Since it was commercialised in 1980, it has kown a worldwide success: countless people are playing with the original 3x3x3 cube, as well as with variants of it of all shapes and sizes.
+Competitions bring together \textit{speedcubers} who have trained for years and memorized algorithms to solve the Rubik's in astonishly fast times (seconds for the best ones).
+Some of the principles used by speedcubers generalise to different dimensions.
 As an example, it is quite impressive to watch \textit{Cubastic} solve a 15x15x15 cube in (only) two and a half hours!
 (see \cite{151515Rubiks}).
-The key insight to solving high dimensional \textbf{RC}s (short obviously of coming up with specific and sophisticated strategies) is the following: if we manage to group all the inside (i.e.
-not located on a corner or an edge of the cube) cubies by colour, and then all the non-corner edges on a given tranche of a face by colour, we are essentially reducing the cube to a 3x3x3, since no face rotation will ever undo the inside cubies and non-corner-edge cubies' grouping.
-For instance, the following picture shows faces of my own 4x4x4 and 3x3x3 cubes which I have arranged to be in identical configurations.
+The key insight to solving high dimensional \textbf{RC}s (short obviously of coming up with specific strategies) is the following: if we manage to group all the inside (i.e.
+not located on a corner or an edge) cubies by colour, and then all the non-corner edges on a given tranche of a face by colour, we are essentially reducing the cube to a 3x3x3, since no face rotation will ever undo these groupings.
+The picture below (\ref{fig:4442333}) shows my own 4x4x4 and 3x3x3 cubes arranged to be in identical configurations.
+\begin{figure}[H]
+  \noindent
+  \makebox[\textwidth]{\includegraphics[scale=0.25]{./Figures/RC333444}}
+  \caption[Reducing RC]{Reducing a 4x4x4 \textbf{RC} to a 3x3x3 -- home made picture}
+  \label{fig:4442333}
+\end{figure}
+\noindent If we can bring the 4x4x4 cube to such a configuration on all faces (which is not very hard), and if we know how to solve a 3x3x3, we are essentially done.
+(The additional degrees of freedom on the 4x4x4 complicate slightly the story since certain arrangements of colors which are possible on the 4x4x4 are not on the 3x3x3, e.g.
+red an and white on opposite faces).
+Overall, the difficuly of hand-solving cubes does not grow as much with dimensionality as one would think, and the time difference comes mostly from the difficulty of handling and manipulating larger cubes.
+
+\Subsection{Conventions: faces, moves and configurations}
+I follow the conventions and notations of the \textbf{RC} community and refer to the faces as F (front), B (back), U (up), D (down), L (left) and R (right).
+It is common in \textbf{RC} jargon to refer to moves  by the name of the face that is rotated clock-wise (when facing it), and by adding a \textit{'} for counter-clock-wise rotations.
+Sometimes numbers are used to indicate repeated moves, though I will count that as multiple moves, since obviously e.g.
+U2 = U U = U' U'.
+As another example: F B' F2 means rotating the front face clock-wise, the back face counter-clock-wise, followed by the front face twice.
+The convention I follow is called the \textit{quarter-turn-metric}, and obvioulsy the length of solutions is affected by which metric we use.
 \\
-\\
-It is quite interesting to note that the difficuly of hand-solving cubes does not grow nearly as much with increasing dimensionality (most of the time difference comes rather from the difficulty of handling and manipulating the cubes of high dimensions at speed).
+Since I am interested in experimenting with \textit{no-human-knowledge} methods such as \textbf{DRL} and \textbf{DQL}, I will consider cubes which are identical up to full cube rotation such as the \textit{flattened} cubes on figure \ref{fig:configurations} ) to be different.
+They will be represented by different tensors in my code, and this is up to the solvers to make sense of things.
 
 \begin{figure}[H]
-\centering
-\includegraphics[scale=0.3]{./Figures/RC333444}
-\caption[Reducing RC]{Reducing a 4x4x4 \textbf{RC} to a 3x3x3 -- home made picture}
-\end{figure}
-
-\noindent It is quite clear from that picture above that if we can bring the 4x4x4 cube to such a configuration on all faces (which is not super hard), and we know how to solve a 3x3x3, we are essentially done.
-(well, almost, since the additional degrees of freedom coming from the higher dimension of the 4x4x4 mean that certain arrangements of the colors which are possible on the 4x4x4 are not on the 3x3x3.
-For instance we can never have the center red and white on opposite faces of a 3x3x3, but we easily can on a 4x4x4!
-Fortunately that is nothing that a couple of easily learnable algorithms cannot fix).
-
-\Subsection{Conventions: faces, moves and configurations}
-In my code as well as in this write-up, I follow the conventions and notations of the \textbf{RC} literature and community as much as possible, I will refer to the faces as F (front), B (back), U (up), D (down), L (left) and R (right).
-The cubes' tiles are of 6 different colors, and without loss of generality (though the interested reader is referred to appendix \ref{sec:Kociemba} for some subtlety around that), we can consider that these colors are also called F, B, U, D, L and R, and consider the goal state as one where, up to rotations of the whole cube, the color's name match the faces (color F on face F, etc).
-The Rubik's has, as one would expect, an extremely large state space.
-I will compute the state space size for the 3x3x3 and 2x2x2 cubes in the next two subsections and discuss conditions under which  a cube is solvable.
-\\
-It is common in \textbf{RC} jargon to refer to moves  by the name of the face that is rotated clock-wise (when facing it), and by adding a prime \textit{'} for counter-clock-wise rotations.
-Sometimes people use a number after the face to indicate repeated similar moves, though I will count in my code that as 2 moves, since obviously e.g.
-U2 = U U = U' U'.
-As another example: F B' F2 means rotating the front face clock-wise, the back face counter-clock-wise, followed by the front face twice.
-Some people refer to the convention I have chosen as the \textit{quarter turn metric}, and obvioulsy the length of solutions is affected by which convention/metric we use.
-\\
-Finally, let me make a remark on what I consider in this project to be distinct configurations of the Rubik's cube.
-Since I am interested in experimenting with, among other methods, \textbf{DRL} and \textbf{DQL}, and that they are usually presented as methods which learn \textit{without-human-knowledge}, I thought it would be a bit akin to cheating to consider that two cubes equivalent by full space rotation are one and the same and represent them as such in my code.
-Ideally, we would rather let the algorithms learn (if they are able to) about symmetries.
-For instance, the two following cubes (which are clearly just a rotation away or equivalently a change-in-observer away from one another) are the same for us humans.
-I however represent them as different tensors in my code, and this is up to the solvers to make sense of things.
-
-\begin{figure}[H]
-\centering
-\includegraphics[scale=0.4]{./Figures/equivalentcubes}
+  \noindent
+  \makebox[\textwidth]{\includegraphics[scale=0.4]{./Figures/equivalentcubes}}
   \caption[Reducing RC]{Two equivalent configurations by full-cube-rotation}
@@
 
+\noindent The \textbf{RC} has, as one would expect, an extremely large state space.
+I will compute the state space size for the 3x3x3 and 2x2x2 cubes in the next two subsections and discuss conditions under which a cube is solvable.
+
 \Subsection{3x3x3 Rubik's -- search space, solvability \& God's number}
+\label{Theory:333RCSSS}
 

### 2022-09-12 latex/Chapters/Implementation.tex
update
trimming implementation

@@
 \\
-In addition, this module contains a \textbf{Puzzled} base class which most classes below inherit from.
-That allow e.g.
-heuristics, search algorithms, solvers and learners to know what puzzle and dimension they operate on, without having to reimplement these basic facts in each of them.
+A \textbf{Puzzled} base class which most classes below inherit from.
+It allows e.g.
+heuristics, search algorithms, solvers and learners to know what puzzle and dimension they operate on, without having to reimplement these facts in each of them.
 \end{itemize}
@@
 \begin{itemize}
-\item \textbf{Manhattan}: This heuristic is specific to the SP.
-It simply adds for each tile (other than the empty tile) the $L_{1}$ distance between their current position and their target position.
-We can quickly see that this heuristic is admissible.
-Indeeed, each move displaces one (and one only) tile by one position.
-In other puzzles, it is generally the case that a move will affect the position of many tiles (e.g.
-RC).
-Therefore, if tiles could somehow freely move on top of one-another (that is, we remove the constraint that there can be at most one tile per compartment, the number of moves necessary to solve the \textbf{SP} would exactly be the Manhattan distance.
+\item \textbf{Manhattan}: This heuristic is specific to the \textbf{SP}.
+It adds over non-empty tiles the $L_{1}$ distance between their current and their target position.
+This heuristic is admissible: indeeed, each move displaces one (and one only) tile by one position (notice that in other puzzles such as the \textbf{RC}, a move might affect multiple tiles).
+Therefore, if tiles could somehow freely move on top of one-another, the number of moves necessary to solve the \textbf{SP} would exactly be the Manhattan distance.
+Having an admissible heuristic will be useful to ascertain the optimality (or not) of other heuristics based on \textbf{DL}, \textbf{DRL} and \textbf{DQL}.
 \\
 \\
-The reason why it is interesting to have a known admissible heuristic is that we can obvioulsy compare other heuristics (DL, DRL, DQL, etc) in terms of optimality.
-\\
-\\
-I have also implemented an improvement to the Manhattan distance, which I shall call Manhattan++ in here (and in code logs and graphs) and can be activated by simply passing \textit{plus=True} to the Manhattan Heuristic (see example \ref{ASSS} as well as a thorougher performance comparison on the 2x5 \textbf{SP} in \ref{MHComp}).
-It is based on the concept of linear constraints that I read about in lecture notes from Washington University (\cite{SlidingPuzzleLectureNotes}).
-The idea is simply that when tiles are already placed on their target row or column, but not in the expected order, we will need to get the tiles out of the way of one another to get to the goal state, and this is not accounted for by the Manhattan distance.
-For instance, in the following 2x3 puzzle, the Manhattan distance does not account for the fact that, at best, tile \red 3 \black needs to get out of the way for tiles \blue 1 \black and \blue 2 \black to move, and then needs to get back in its row, adding a cost of 2 to the Manhattan distance.
+I have also implemented an improvement which I call Manhattan++.
+It can be activated by passing \textit{plus=True} to the Manhattan Heuristic (see example \ref{ASSS} as well as the 2x5 \textbf{SP} performance comparison in \ref{MHComp}).
+It is based on the concept of linear constraints (see lecture notes from Washington University \cite{SlidingPuzzleLectureNotes}): when tiles are already placed on their target row or column, but not in the expected order, we need to get them out of the way of one another to reach the goal, which the Manhattan distance does not account for.
+For instance, in the following 2x3 puzzle, tile \red 3 \black needs to get out of the way for tiles \blue 1 \black and \blue 2 \black to move left, and then needs to get back in its row, adding 2 to the total cost.
 
@@
 
-Two important things to notice are that linear constraints across rows and columns (which I might generically refer to as \textit{lines} in the following) can be added without breaking admissibility (hence giving more pessimistic, or accurate, cost estimates than simple Manhattan).
-This is because if a tile is involved in two linear constraints, it will need to get out of the way both horizontally and vertically.
-The second thing is that when several pairs in a line are not in order, we cannot simply add 2 for each distinct out-of-order pair, the right penalty to add is more subtle than that and needs to be computed recursively.
-For instance, let us now consider the following configuration:
+Linear constraints across rows and columns (which I generically refer to as \textit{lines}) can be added without breaking admissibility.
+This is because a tile involved in two linear constraints needs to get out of the way both horizontally and vertically.
+Also notice that when several pairs on a line are out-of-order, we cannot naively add 2 for each.
+The right penalty to add needs to be computed recursively.
+The following configuration for instance with all pairs (1, 2), (1, 3) and (2, 3) out-of-order only inccurs an extra penalty of 4, not 6.
 \begin{center}
@@
 \end{center}
-The correct penalty to add is not 6 (3 times 2 since all pairs (1, 2), (1, 3) and (2, 3) are out of order, but only 4.
-Indeed if, say, tile 3 got somehow out of the way at the back of the SP (imagine just another dimension there where we can move tiles) and tile 2 got out of the way by moving down to let tile 1 pass across, we could be done by simply adding 4 to the Manhattan distance (2 to move tile 3 out and back, and 2 to move tile 2 out and back).
-The correct way to compute the penalty cost for linear constraints is therefore to do it recursively, taking the minimum additional cost of moving either the left-most or the right-most tile of the line under consideration out of the way (that additional cost to move these left-most or right-most tile is 2 if not at their expected order in the line, 0 otherwise) plus the penalty of reordering the rest of the line.
+It would indeed be enough to get tile 3 out of the way above the puzzle, and tile 2 under the puzzle to let tile 1 pass across.
+My recursive implementation takes the minimum additional cost of moving either the left-most or the right-most tile of the line under consideration out of the way (that additional cost being 2 or 0, depending on whether or not the tile in question is currently in place) plus the penalty of reordering the rest of the line.
 \\
 \\
-Finally, as suggested by the reference lecture notes (which give very vague details about the above subtleties), I have precomputed all the penalties for all possible rows, columns and all possible tiles ordering they could have and saved the corresponding penalties in a database.
-For memory efficiency, I also only saved penalties which are non-zero.
-The very first time any call to Manhattan++ is made, for a given dimension nxm, the appropriate linear constraint penalties are computed and populated in a database.
-\\
-\\
-Notice that for an nxm \textbf{SP}, there are n rows, each of which can have $\frac{(n * m)!}{(n * m - m)!}$ different ordering of tiles and m columns which can each have $\frac{(n * m)!}{(n * m - n)!}$ different ordering of tiles.
-This means the pre-computations and data-base sizes for the Manhattan++ heuristic are actually manageable, as it grows much slower than the number of possible puzzles.
-The maximum number of penalties to compute for $nxm \leq 5x5$ are:
+Finally, as suggested in the rather vagues details of \cite{SlidingPuzzleLectureNotes}, I have precomputed and stored all the penalties for all possible rows, columns and tiles ordering.
+For memory efficiency, I only saved non-zero panalties.
+The very first time a call to Manhattan++ is made, for a given dimension nxm, the appropriate linear constraint penalties are computed and saved.
+For an nxm \textbf{SP}, each of the n rows can have $\frac{(n * m)!}{(n * m - m)!}$ different tiles orderings and each of the m columns $\frac{(n * m)!}{(n * m - n)!}$.
+As a result, my Manhattan++ pre-computations and data-bases are quite manageable, as they grow with n and m much slower than the state space.
+The number of penalties to compute for $nxm \leq 5x5$ are:
 
@@
 \end{center}
-Taking also into account that we only store non-zero penalties, we actually get the following (quite smaller) number of penalties in our data bases:
+Removing 0 penalties, we get the following (quite smaller) data base sizes:
 \begin{center}
@@
 That data is saved under a tree following the schema below:
-$$\$\{ "RUBIKSDATA"\}/data\_type/puzzle\_type/dimension/file\_name.pkl$$
+$$\$RUBIKSDATA/data\_type/puzzle\_type/dimension/file\_name.pkl$$

### 2022-09-12 latex/Chapters/SlidingPuzzle.tex
update
trimming SP

@@
 
-In this chapter, I show and discuss results on \textbf{SP} of various dimensions, going from \textit{small} dimensions (which I define as those I can reasonably \textit{fully} solve given my computing resources, the finite time I have to complete this project, and the choice of programming language), then I move to studying in much detail the case of the 3x3, and finally move on to slightly higher dimensions 3x4 and 4x4.
-Table \ref{tab:gridSP} below shows which solvers I have run on each of the dimensions attempted.
-The reader is referred to the implementation chapter \ref{sec:Implementation} for a detailed explanation of each of these solvers, and to the appendix section \ref{sec:Examples} for code examples on how to run them.
+I now discuss results on the \textbf{SP}, going from \textit{small} dimensions (which I can reasonably \textit{fully} solve given my computing resources, the finite time I have to complete this project, and the choice of programming language), then I move to studying in detail the 3x3 \textbf{SP}, and finally move to higher 3x4 and 4x4 dimensions.
+Table \ref{tab:gridSP} shows which solvers I have run on each dimension.
+The reader is referred to sections \ref{sec:Chapter1} and \ref{sec:Implementation} for details on each solver, and to the appendix section \ref{sec:Examples} for code examples.
 
@@
 
-\noindent I am neither making claims of depth (each solver could surely be tuned and optimized/adapted, possibly differently for each dimension) nor breadth (I have not run all of the solvers versus each dimension).
-Running these experiments takes a lot of computing time (even though I have very decent computing resources), so I had to be somewhat selective.
-I wanted however to run all of the solvers I have implemented one one particular dimension (the intermediary 3x3 seemed appropriate to do so) and try to answer as many of the following interesting questions as possible:
+\noindent I am not making claims of depth (each solver could be tuned and optimized much further) nor breadth (I have not run all solvers versus all dimensions).
+These experiments take a lot of computing time, so I had to be somewhat selective.
+I wanted however to run all of the solvers on at least one particular dimension (3x3 seemed appropriate) to try and answer, even if not in generality, as many of the following interesting questions:
 \begin{itemize}
@@
 \item How much, if any, improvement do we get with Manhattan++ over Manhattan?
-\item Is there much loss between \textbf{DL} and \textbf{DRL}, i.e. is losing the teacher used by \textbf{DL} a deal breaker?
-\item Is \textbf{DQL}, all things being equal, able to learn a better cost-to-go heuristic than \textbf{DRL}?
-\item How does \textbf{MCTS} perform? How important is it to tune the hyper-parameter $c$? How much improvement, if any, does the post trim of the \textbf{MCTS} tree via \textbf{BFS} bring?
-\item Are any of the \textbf{D*L} techniques able to perform on par with A$^{*}$ with Manhattan++ or perfect heuristic?
+\item Does \textbf{DL} perform much better than \textbf{DRL} due to supervision?
+\item Is \textbf{DQL}, all things being equal, able to learn a better value-heuristic than \textbf{DRL}?
+\item How does \textbf{MCTS} perform? How important is it to tune the hyper-parameter $c$? How much improvement does the final \textbf{BFS} tree-trimming bring?
+\item Are the \textbf{D*L} heuristics competitive versus Manhattan++ or the perfect heuristic?
 \end{itemize}
-All of these questions are answered (obviously not in generality, but in the context of the experiments I have been running) in this chapter.
 
@@
 
-As mentioned in chapter \ref{sec:Puzzles}, the state space cardinality for the \textbf{SP} grows very quickly with n and m.
+As mentioned in chapter \ref{sec:Puzzles}, the state space cardinality for the \textbf{SP} grows quickly with n and m.
 The only dimensions which have less than 239.5 millions states are shown in table \ref{tab:smallSP}.
-Note I am also only considering n $\leq$ m since dimension mxn can always be solved by symmetry from the solutions mxn:
 
@@
 \end{table}
-\noindent In this section, I will discuss \textit{full} results for these 5 dimensions.
-In order to fully solve them, one can simply use rubiks.scripts.learner, setting up the PerfectLearner with A* and manhattan heuristic, or instantiate directly a PerfectLearner as have seen in section \ref{PLSS}.
-I obtained the following God numbers (table \ref{tab:smallSPGN}) for these puzzles:
+\noindent Using my PerfectLearner with A* and Manhattan heuristic (see \ref{PLSS} for example of how to run this), I obtained the following God numbers (table \ref{tab:smallSPGN}) for these dimensions:
 \begin{table}[H]
@@
 \end{table}
-\noindent Among the above dimensions, 2x5 has the largest state space.
-My current result in table \ref{tab:smallSPGN} for that dimension is provisional, as I have only solved about 1.2 million puzzles out of the 1.8 million possible configurations, despite running a combined 600 hours (most of the time over 16 cores, except for a full week-end run on a 72-core c5.18xlarge Amazon EC2 instance).
-It is well possible that I haven't yet bumped into the hardest configuration though!
-\\
-\\
+\noindent Among those dimensions, 2x5 has the largest state space and my result for it is still provisional, as I have only solved about 1.2 out of 1.8 million puzzles at the time of writing, despite running a combined 600 hours (mostly over 16 cores, except for a full week-end run on a 72-core c5.18xlarge Amazon EC2 instance).
+It is well possible that I haven't yet bumped into the hardest 2x5 configuration!
 The perfectLearner also keeps track of the hardest puzzle it has encountered (i.e.
@@
 \label{MHComp}
-In this section, I verify empirically that, as expected, the overhead of adding penalty in Manhattan++ for the linear constraint (which have all been precomputed and stored in a database) is more than compensated for by the reduction in nodes expansion.
+Next, I verify empirically that, as expected, the overhead of adding penalty in Manhattan++ for the linear constraint is compensated for by the reduction in nodes expansion.
 I have run my solver script for 2x5 in performance test mode, for both Manhattan and Manhattan++, with 250 randomly shuffled puzzles with \textit{nb\_shuffles} from 0 to 60 by increment of 5, as well as with \textit{nb\_shuffles = $\inf$}.
-The resulting run time and nodes expansions are as follows:
+The resulting run time and nodes expansions are shown in table \ref{fig:25SPPerformanceManhattan}:
 
@@
 
-\noindent As can be seen, for low difficulty (up to \textit{nb\_shuffles = 20}), the node expansions are about the same in both cases, and the overhead of adding the linear constraints penalty increases the run time.
-However, for any non trivial case, Manhattan++ outperforms considerably (by a factor of about 2.5).
+\noindent For low difficulty (\textit{nb\_shuffles} $\leq$ 20), the node expansions are about the same in both cases, and the overhead of adding the linear constraints penalty increases the run time.
+However, for harder puzzles, Manhattan++ considerably outperforms, by a factor of 2.5+.
 Table \ref{tab:mppOutperformance} below summarizes the results for the 250 \textit{perfectly} shuffled configurations:
@@
   \makebox[\textwidth]{\includegraphics[scale=0.45]{./Figures/33SPnets}}
-  \caption[33SPnet]{Architecture used for the \textbf{DL} heuristics training on the 3x3 \textbf{SP}}
+  \caption[33SPnet]{Architecture I used for the \textbf{DL} heuristics training on the 3x3 \textbf{SP}}
   \label{fig:33SPnets}

### 2022-09-12 latex/Chapters/Rubiks.tex
trimming RC

@@
 
-In this second results section, I discuss my experiments on the Rubik's cube.
-I have run reasonably comprehensive experiments on the 2x2x2, throwing 6 of my solvers at the problem.
+I now discuss my experiments on the Rubik's cube.
+I have run reasonably comprehensive trial on the 2x2x2, throwing 6 of my solvers at the problem.
 As mentioned in \ref{sec:Puzzles}, the 2x2x2 \textbf{RC}, in spite of its deceivingly simple physical appearance, is actually not that much easier to solve (by hand) than the 3x3x3.
-For instance, it takes me usually just under 1 minute, wherras the 3x3x3 takes me only double that time, just under 2 minutes.
-I am obviously a beginner in Rubik's cube handsolving, but my personal ratio of time-to-solve and perceived difficulty is in line with that of experienced speedcubers who might take and average of, say, 2-3 seconds for the 2x2x2 and 10 seconds for the 3x3x3.
-\\
+For instance, it takes me usually just under one minute, versus two minutes for the 3x3x3.
+I am obviously no \textit{speedcuber}, but my time-to-solve and perceived difficulty ratios are in line with that of experienced players who might take 2-3 seconds for the 2x2x2 and 10 seconds for the 3x3x3.
 Computing wise, the 2x2x2 was of decent size for my purpose, with its 88 million configurations.
-Let me start with a table summarizing which solvers I have attempted to run on the 2 dimensions I have attempted to experiment with:
+Table \ref{tab:gridRC} summarizes which solvers I have attempted to run on the 2 dimensions:
 
+\begin{table}[H]
 \begin{center}
@@
 \end{tabular}
+\caption{\label{tab:gridRC} Solvers used vs \textbf{RC} dimension}
 \end{center}
+\end{table}
 
@@
 \Subsection{BFS}
-BFS is the only one for which I used a step of 1 on the difficuly and stopped at 5 as it was clear it would not go any further.
-Given that the run time and expanded nodes are clearly going to increase linearly, and given where 5 scrambling from goal took this solver already, there was no point going any further!
+For \textbf{BFS} I used a step of 1 on the difficuly and stopped at 5 as it was clear it would not go much further.
+Its run time and expanded nodes are increasing linearly with difficulty, and it was already taking an average close to a minute for 5 shuffles from goal.
 
 \Subsection{Kociemba, A$^{*}$[Kociemba], A$^{*}$[DL[A$^{*}$[Kociemba]]]}
-Next, let me discuss Kociemba (see \cite{HKociemba} and \cite{Kociemba}) and associated experiments.
-We obviously expect Kociemba to be by far the fastest solver since it is handcrafted with much specific knowledge about the Rubik's cube group theory and multi-stage solving which I discussed in \ref{Theory:RCMSS}.
-I found the 2x2x2 library I used to not be optimal however, despite claims to the contrary.
-I have not had the time to investigate why this was not the case, but empirically I found that for instance running A$^{*}$ with Kociemba's cost as a heuristic was resulting in higher quality (i.e.
-of lower cost) solutions, clearly showing the non-optimality of Kociemba to begin with.
-\\
-\noindent Similarly to the higher dimensions \textbf{SP} from the previous section, there was no hope for me to run a perfect solver on the 2x2x2 \textbf{RC}.
-For starter, I have no admissible heuristic to guarantee optimality!
-Even if I had found one such heuristic and managed to implement it, it is doubtful I could have solved all 88 millions configurations without considerable work and time to reduce the configurations via symmetry arguments.
-As a result, I used solutions from A$^{*}$[Kociemba] on (perfectly) randomly generated configurations to train my \textbf{DL} heuristic.
-Thinking of it, it is rather amusing since Kociemba is itself using a two-stage IDA$^{*}$ solver, so I end-up here with 3 levels of nested A$^{*}$ algorithms, namely A$^{*}$[DL[A$^{*}$[IDA$^{*}$[2-stage-subgroups]]]].
-\\
-Notice finally that, since I have no optimal solver for the \textbf{RC}, I have set Kociemba as my benchmark in the performance test for the purpose of computing the optimality score.
-This is why some (actually all) of the other solvers will show scores larger than 100\% (top-right graph of figure \ref{fig:222RCPerformance}).
+I obviously expected Kociemba (see \cite{HKociemba} and \cite{Kociemba}) to be by far the fastest solver since it is handcrafted with much specific knowledge about the \textbf{RC} group theory and multi-stage solving (see section \ref{Theory:RCMSS}).
+Contrary to its claims, I found the 2x2x2 library to not be optimal, running A$^{*}$ with Kociemba's cost as a heuristic indeed resulted in shorter solutions.
+I did however set Kociemba as my benchmark in the \textit{performance\_test} since I had no admissible heuristic to guarantee optimality otherwise.
+The optimality scores (top-right graph of figure \ref{fig:222RCPerformance}) therefore do reach more than 100\%, as all other solvers found shorter solutions than Kociemba.
+I used the solutions of A$^{*}$[Kociemba] on perfectly randomly generated configurations to train the \textbf{DL} heuristic.
+Since Kociemba is itself using a two-stage IDA$^{*}$ solver, I end-up here with 3 levels of nested A$^{*}$ algorithms: A$^{*}$[DL[A$^{*}$[IDA$^{*}$[2-stage-subgroups]]]]!
 
@@
 \begin{itemize}
-\item Update the target network every 1,000 epochs (\textit{update\_target\_network\_frequency=1000}) or when the MSE falls below 10 basis points of the maximum target cost-to-go (\textit{update\_target\_network\_threshold=1e-3})
-\item Run for a maximum of 11,000 epochs (\textit{nb\_epochs=11000}) or when the maximum target so far has not been increasing by more than 1\% in the last 5,500 epochs (\textit{max\_target\_uptick=1e-2} and \textit{max\_target\_not\_increasing\_epochs\_pct=0.5})
-\item At every update of the target network (\textit{training\_data\_every\_epoch=False}), generate 10,00 sequences of puzzles (\textit{nb\_sequences=10,000}), each comprised of a 15-scramble puzzle back to goal state, with all intermediary puzzles along the path (\textit{nb\_shuffles=15}).
-\item Use a fully connected network (\textit{network\_type=fully\_connected\_net}) with three hidden layers of size 600, 300, 100 (\textit{layers\_description=(600,300,100)}) and use torch.optim.RMSProp as optimiser (\textit{optimiser=rms\_prop}) with an exponential scheduler (\textit{scheduler=exponential\_scheduler}) with gamma 0.999 (\textit{gamma\_scheduler=0.9999}) and \textit{learning\_rate=0.001}.
-The networks architecture is shown in figure \ref{fig:222RCnets}.
+\item Update the target network every 1,000 epochs (\textit{update\_target\_network\_frequency}) or when $\frac{MSE}{max target}$ falls below 0.1\% (\textit{update\_target\_network\_threshold=1e-3})
+\item Stop at \textit{nb\_epochs}=11,000 or if the max target has not increased more than 1\% (\textit{max\_target\_uptick}) in the last 5,500 epochs (\textit{max\_target\_not\_increasing\_epochs\_pct}=0.5)
+\item At each target net update (\textit{training\_data\_every\_epoch}=False), generate \textit{nb\_sequences}=10,000 sequences of puzzles comprised of a 15-scramble puzzle along with its path to goal (\textit{nb\_shuffles}=15).
+\item Fully connected network (\textit{network\_type=fully\_connected\_net}) with three hidden layers (\textit{layers\_description}=(600,300,100)), an RMSProp optimiser (\textit{optimiser=rms\_prop}), \textit{scheduler=exponential\_scheduler} with \textit{gamma\_scheduler}=0.9999 and \textit{learning\_rate}=0.1\%.
+The networks architecture are shown in figure \ref{fig:222RCnets}.
 \end{itemize}
-Graphs \ref{fig:222RCDRL} and \ref{fig:222RCDQL} show both learners' following metrics tracked over the epochs:
+
+\begin{figure}[H]
+  \noindent
+  \makebox[\textwidth]{\includegraphics[scale=0.6]{./Figures/222RCnets}}
+  \caption[222RCnets]{2x2x2 \textbf{RC} architecture for \textbf{DRL} (left) \& \textbf{DQL} (right) heuristics training}
+  \label{fig:222RCnets}
+\end{figure}
+
+\noindent Graphs \ref{fig:222RCDRL} and \ref{fig:222RCDQL} show both learners' following metrics tracked over the epochs:
 \begin{itemize}
-\item learning rate, which decreases due to the scheduler, but gets reset at each network update.
-\item max target.
-One interesting dynamic during learning is that due to the way targets are constructed using value-iteration update, and starting from an all 0 network, the max target only increases by about 1 at best at every network update.
-That is, the first network only learns how to differentiate goal from non-goal, then the second one learns how to differentiate goals, cost 1 and cost 2+ \dots.
-This is why I have introduced the \textit{update\_target\_network\_threshold} discussed above.
-It makes sure we do not waste time initially training the left-hand-side network when training is easy.
-\textit{max\_target\_uptick} and \textit{max\_target\_not\_increasing\_epochs\_pct} on the other hand make sure that we stop training altogether when the max target does not grow anymore.
-This is because we do not know a-priori what the God number of a given puzzle/dimension is.
-Let us assume for the sake of the argument that it is 15: that means that beyond 15 updates of the target network we might not get much benefit to additional training and value-iterating.
-\item the loss (MSE for \textbf{DRL}, MSE + CrossEntropyLoss for \textbf{DQL})
-\item the loss divided by the max target.
-The exit criterion \textit{update\_target\_network\_threshold} is based on that quantity, since a loss of e.g.
-0.01 means little if the we do not compare it to the possible range that the cost-to-go can take.
+\item learning rate: decreasing due to the scheduler and reset at each network update.
+\item max target: an interesting dynamic during learning is that since targets are constructed using value-iteration update, starting from an all 0 network, the max target increases by at most about 1 at every network update, until we reach God's number.
+The first network only learns to differentiate goal from non-goal, the second one learns how to differentiate goals, cost 1 and cost 2+ \dots.
+This is why I introduced \textit{update\_target\_network\_threshold} discussed above, to makes sure we do not spend time initially training the left-hand-side network when it is easy.
+\textit{max\_target\_uptick} and \textit{max\_target\_not\_increasing\_epochs\_pct} make sure that we stop training when the max target does not grow anymore.
+\item loss: MSE for \textbf{DRL}, MSE + CrossEntropyLoss for \textbf{DQL}
+\item loss divided by max target: the exit criterion \textit{update\_target\_network\_threshold} is based on that quantity, since a loss of e.g.
+1\% means little if the we do not compare it to the possible range of the cost-to-go.
 \item The percentage of all the possible puzzles of that dimension \textit{seen} by the learner.
@@
 
-\begin{figure}
+\begin{figure}[H]
   \noindent
-  \makebox[\textwidth]{\includegraphics[scale=0.6]{./Figures/222RCnets}}
-  \caption[222RCnets]{2x2x2 \textbf{RC} architecture for \textbf{DRL} (left) \& \textbf{DQL} (right) heuristics training}
-  \label{fig:222RCnets}
-\end{figure}
-
-\begin{figure}
-  \noindent
-  \makebox[\textwidth]{\includegraphics[scale=0.5]{./Figures//222RCDRL}}
+  \makebox[\textwidth]{\includegraphics[scale=0.4]{./Figures//222RCDRL}}
   \caption[222RCDRL]{\textbf{DRL} of the 2x2x2 \textbf{RC}}
@@
 
-\begin{figure}
+\begin{figure}[H]
   \noindent
-  \makebox[\textwidth]{\includegraphics[scale=0.5]{./Figures//222RCDQL}}
+  \makebox[\textwidth]{\includegraphics[scale=0.4]{./Figures//222RCDQL}}
   \caption[222RCDQL]{\textbf{DQL} of the 2x2x2 \textbf{RC}}
@@
 
-We summarize as usual the 2x2x2 \textbf{RC} performance tests by our usual metrics \ref{fig:222RCPerformance}.
-Overall, the \textbf{DL} and \textbf{DRL} managed to solve all cubes (though \textbf{DL} took over an houd and \textbf{DRL} up to 6 hours for the slowest cube they encountered.
-Both methods outperformed in terms of optimality score Kociemba and got on par with A$^{*}[Kociemba]$.
+I summarize as usual the 2x2x2 \textbf{RC} performance tests with my usual metrics in figure \ref{fig:222RCPerformance}.
+Overall, the \textbf{DL} and \textbf{DRL} managed to solve all cubes (though \textbf{DL} took over an hour and \textbf{DRL} 6 hours for their slowest cube.
+Both methods outperformed Kociemba in terms of optimality-score and got on par with A$^{*}[Kociemba]$.
 Their max cost over all cubes encountered was 12 for up to 20 shuffles, and 13 for $\infty$ shuffling.
@@
 \\
-An very interesting remark is that if we check the learning graph \ref{fig:222RCDRL} of the \textbf{DRL}, we can see it only ever assigned a cost-to-go (value function) of 7.6 at most.
-Despite that fact, it was still able to solve cubes requiring 13 moves.
+Quite remarkably we can see (second graph of \ref{fig:222RCDRL}) that the \textbf{DRL} heuristic only ever assigned a cost-to-go of 7.6 at most.
+Despite that, it was still able to solve cubes requiring up to 13 moves (likely to have been among the hardest 2x2x2 cubes).
 How is that possible?
-Notice that the loss of 0.2 towards which it converged at the end of its training is misleading, since it is not a loss versus the actual true cost-to-go, but rather versus that of the target network.
-My theory is that while the value the \textbf{DRL} assigned to hard-to-solve cubes is off (by at least 4 or 5), it is likely the case that the network gets the relative value of neighbouring cubes about right (given that the value-iteration teaches the solver that, it is not super surprising either), and this is sufficient to guide A$^{*}$ extremely well.
-If both the absolute value of the cost-to-go and the relative (among neighbouring cubes) was totally off, we would indeed not expect the heuristic to make A$^{*}$ any better than, say, \textbf{BFS} and we would therefore not expect it to be able to solve \textit{costly} cubes in 10-13 moves.
+Notice that the loss of 0.2 towards which the network converged at the end of its training is misleading, since this is in-sample loss versus the target network.
+My theory is that while the value the \textbf{DRL} assigned to hard-to-solve cubes can be off (by at least 4 or 5), it is likely that the network gets the relative value of neighbouring cubes right (not that surprising given the value-iteration), and this is sufficient to guide A$^{*}$ extremely well.
+If both the absolute value of the cost-to-go and the relative (among neighbouring cubes) was totally off, we would indeed not expect the heuristic to make A$^{*}$ any better than, say, \textbf{BFS} and we would therefore not expect it to be able to solve \textit{costly} cubes requiring 10+ moves.
 \\
 \\
-Finally, we can see that \textbf{DQL}, despite showing seemingly/visually cleaner and more monotonic loss convergence than \textbf{DRL} (see \ref{fig:222RCDQL} versus \ref{fig:222RCDRL}), struggled a lot more to solve difficult cubes (above 10 shuffles).
-Clearly this shows that my earlier findings on the 3x3 \textbf{SP} are likely not universal, and very much depend on the parameters set during training, as well as on the problem at hand.
+Finally, we can see that \textbf{DQL}, despite showing visually cleaner and more monotonic loss convergence than \textbf{DRL} (see \ref{fig:222RCDQL} versus \ref{fig:222RCDRL}), struggled a lot more to solve difficult cubes (above 10 shuffles).
+This shows that my earlier findings on the 3x3 \textbf{SP} are likely not universal, and very much depend on the parameters set during training, as well as on the problem at hand.
 
@@
   \noindent
-  \makebox[\textwidth]{\includegraphics[scale=0.55]{./Figures/222RCPerformance}}
+  \makebox[\textwidth]{\includegraphics[scale=0.52]{./Figures/222RCPerformance}}
   \caption[222RCPerformance]{Solvers' performance comparison 2x2x2 \textbf{RC}}
@@
   \noindent
-  \makebox[\textwidth]{\includegraphics[scale=0.55]{./Figures/333RCPerformance}}
+  \makebox[\textwidth]{\includegraphics[scale=0.52]{./Figures/333RCPerformance}}
   \caption[333RCPerformance]{Solvers' performance comparison 3x3x3 \textbf{RC}}

### 2022-09-12 latex/Chapters/SlidingPuzzle.tex
trimming RC

@@
 Table \ref{tab:gridSP} shows which solvers I have run on each dimension.
-The reader is referred to sections \ref{sec:Chapter1} and \ref{sec:Implementation} for details on each solver, and to the appendix section \ref{sec:Examples} for code examples.
+The reader is referred to the theory (\ref{Chapter1}) and implementation (\ref{sec:Implementation}) sections for details on each solver, and to the appendix section \ref{sec:Examples} for code examples.
 

### 2022-09-12 latex/Chapters/Conclusion.tex
trimming conclusion

--- 2022-09-10 latex/Chapters/Conclusion.tex
@@
 
-\Section{Learning}
-Working on this project has been overall a lot of fun.
-Even setting aside implementation and computer-experimentation for a moment, I have learnt a great deal about puzzles in general and how to approach thinking about them in a more structured fashion.
-Specifically about the Rubik's cube, I have even spent the time to learn a few algorithms that are sufficient to solve them by hand.
-I am now able to solve the 2x2x2 under one minute, 3x3x3 usually around 2 minutes and the 4x4x4 in about 5 minutes.
-More specifically regarding the techniques discussed in this project, the principles that underly uninformed and informed search, A$^{*}$ and heuristics, multi-stage solvers, and of course the various deep learning methods I have played with (\textbf{DL}, \textbf{DRL} and \textbf{DQL}) all extend to many other puzzle and dimensions and are a great new tool-set to add to my knowledge and experience.
-
-\Section{Implementation}
-I am quite satistified to have had the time to implement and play with so many solvers and methods, spanning so many different \textbf{CS} and \textbf{AI} paradigms, from handcrafting my own \textbf{SP} \textit{naive} solver, playing with Kociemba, to implementing \textbf{BFS}, \textbf{DFS}, A$^{*}$, admissible heuristics such as Manhattan for the \textbf{SP} and improving on it via linear constraints penalty, and of course all the machinery necessary to fit, train and exploit some \textbf{DL}, \textbf{DRL} and \textbf{DQL} heuristics to guide search algorithms such as  A$^{*}$ and \textbf{MCTS}.
-Ideally I would have liked to implement things in C++ for speed and multithreading (especially for \textbf{MCTS}) but time was limited so I decided to prioritize breadth and ease of experimentation over depth and efficiency of implementation.
-The choice of Python definitely made it much easier to integrate with many of the ecosystem's libraries (matplotlib, pandas and pytorch in particular).
-
-\Section{Deep Reinforcement-Q Learning}
-It will not come as a surprise to anyone who has played with them that Deep Reinforcement Learning and Deep Q Learning models can be quite capricious to train: choosing the convergence mechanisms and criteria, the interplay of generation of new training data with the inner and outer loop of the \textbf{DRL} and \textbf{DQL} algorithms (see \ref{sec:TheoryDLDRL} and \ref{sec:TheoryDQL}) seemed more of an art than a sicence at times.
-Obviously a lot more tuning, as well as much longer training time and computing resources would have been necessary to make these techniques more competitive at higher dimensions, but overall I remain extremely impressed by how, all things considered - and given how daunting the project seemed at first - it took me little effort to train successful \textbf{DRL} and \textbf{DQL} models capable of solving puzzles with such large state and action spaces.
-
-\Section{Professional issues}
-Bug in Kociemba and software design\\
-AI safety 
-See \cite{ManyWorlds}
-See \cite{https://doi.org/10.1111/1758-5899.12718} and \cite{Bostrom2014}
+\textbf{Learning} -- Working on this project has been a lot of fun.
+Setting aside implementation and experimentation for a moment, I have learnt a great deal about puzzles in general and how to think about them in a structured fashion.
+I have also spent the time to learn a few \textbf{RC} algorithms allowing me to solve the 2x2x2 in under one minute the 3x3x3 in 2 minutes and the 4x4x4 in about 5 minutes.
+More specifically regarding the techniques discussed in this project, the principles that underly blind and informed search, A$^{*}$ and heuristics, multi-stage solvers, and of course the various \textbf{AI} methods I have played with (\textbf{DL}, \textbf{DRL} and \textbf{DQL}) all are a great new addition to my tool-set and knowledge.
+\\
+\\
+\textbf{Implementation} -- I am quite satistified to have implemented and played with so many solvers and methods, spanning different \textbf{CS} and \textbf{AI} paradigms, from handcrafting my own \textbf{SP} \textit{naive} solver, playing with Kociemba, to implementing \textbf{BFS}, \textbf{DFS}, A$^{*}$, admissible heuristics such as Manhattan with linear constraints penalty, and of course all the machinery necessary to fit, train and exploit some \textbf{DL}, \textbf{DRL} and \textbf{DQL} heuristics and even \textbf{MCTS}.
+Ideally I would have liked to implement things in C++ for speed and multithreading but time was limited so I decided to prioritize breadth and ease of experimentation over depth and efficiency of implementation.
+Python made it easier to integrate many of the ecosystem's libraries e.g.
+matplotlib, pandas and pytorch.
+\\
+\\
+\textbf{Deep Reinforcement-Q Learning} -- Anyone who has played with \textbf{DL}, \textbf{DRL} and \textbf{DQL} models (\ref{sec:TheoryDLDRL} and \ref{sec:TheoryDQL}) know they can be capricious to train: choosing the convergence criteria, the interplay of training data generation with the inner and outer loops felt more art than sicence.
+More tuning, longer training time and computing resources would have been necessary to make these techniques more competitive at higher dimensions, but I remain extremely impressed by how, all things considered - and given how daunting the project seemed at first - it took me little effort to train successful \textbf{DRL} and \textbf{DQL} models capable of solving large state and action spaces puzzles.
+\\
+\\
+\textbf{Professional issues} -- The BCS code of conduct (\cite{BCS}) tells us to `show what you know, learn what you don't'.
+Furthering my \textbf{ML} and \textbf{AI} knowledge with this MSc in AI probably has the `learn what you don't' part covered.
+As for the `show what you know', I have endeavoured to show honest and sometimes modest, results.
+I would have liked to go further with the \textbf{DxL} methods on the 3x3x3 \textbf{RC} but this was sadly not possible within the time imparted.
+As for issues of ethics and safety, outined in the ACM code of ethics (\cite{ACM}), anyone working in \textbf{AI} should in my opinion be aware of them.
+There is some divide in the community regarding the dangers of \textbf{AI}.
+I myself firmly stand in the Bostrom camp (see \cite{https://doi.org/10.1111/1758-5899.12718} and \cite{Bostrom2014}).
+His thought-provoking paper-clip-maximiser shows that the most benign looking \textbf{AI} agent, if sufficiently capable, could create disproportionate harm.
+We probably do not want either to see a Rubik's solver engulf all of the earth's and galaxy's resources to compute the nxnxn Rubik's God number for ever growing values of n :)

### 2022-09-12 latex/Abstract/Abstract.tex
trimming abstract and ack

@@
 Its framework is very appealing and intuitive.
-In particular, the mathematical beauty of Value iteration and Q iteration (\cite{WatkinsThesis}) for discrete state and action spaces, blindly iterating from \textit{any} initial value, is quite profound.
+In particular, the mathematical beauty of Value and Q iteration (\cite{WatkinsThesis}) for discrete state and action spaces, blindly iterating from \textit{any} initial value, is quite profound.
 Sadly, it had until recently proven hard to achieve practical success with these methods.
-Inspired however by the seminal success of Deep Mind's team in using Deep Reinforcement Learning (\textbf{DRL}) to play Atari games (\cite{Mnih2013})  and to master the game of Go (\cite{AlphaGo}), researchers have in recent years made a lot of progress towards designing algorithms capable of learning and solving, \textit{without human knowledge}, the Rubik's Cube (\textbf{RC}) - as well as similar single player puzzles - by using Deep Q-Learning (\textbf{DQL})  (\cite{DBLP:journals/corr/abs-1805-07470}) or search and (\textbf{DRL}) value iteration (\cite{https://doi.org/10.48550/arxiv.1805.07470}).
+Inspired however by the seminal work of Deep Mind's team in using Deep Reinforcement Learning (\textbf{DRL}) to play Atari games (\cite{Mnih2013})  and to master the game of Go (\cite{AlphaGo}), researchers have made a lot of progress towards designing algorithms capable of learning and solving, \textit{without human knowledge}, the Rubik's Cube (\textbf{RC}) by using Deep Q-Learning (\textbf{DQL})  (\cite{DBLP:journals/corr/abs-1805-07470}) or search and (\textbf{DRL}) value iteration (\cite{https://doi.org/10.48550/arxiv.1805.07470}).
+I will attempt to implement a variety of solvers combining $A^{*}$ and heuristics, some handcrafted, others trained on sequences of puzzles using Deep Learning (\textbf{DL}), \textbf{DRL} or \textbf{DQL}, to solve the 15-puzzle and the Rubik's cube.
 \\
 \\
-In this project, I will attempt to implement a variety of solvers combining $A^{*}$ search and heuristics, some of which will be handcrafted, others which I will train on randomly generated sequences of puzzles using \textbf{DL}, \textbf{DRL} or \textbf{DQL}, to solve the 15-puzzle (and variations of different dimensions), as well as the Rubik's cube.
+\textbf{Organisation of this thesis} -- In section 1, I discuss what I hope to get out of this project, both in terms of personal learning, as well as in terms of tangible results (solving some puzzles!).
+In section 2, I do a recap of the methods used in the project.
+Section 3 is dedicated to the mathematics of the sliding puzzle \textbf{SP} and the \textbf{RC}.
+I give then an overview of my implementation in section 4.
+Sections 6 and 7 present my results on respectively the \textbf{SP} and the \textbf{RC} and section 8 concludes.
+A couple of appendices complement the main text: the first one details how to use the code base, the second describes a limitation I bumped into while using the kociemba library.
 \\
 \\
-\textbf{Organisation of this thesis} -- In the first section, I describe what I hope to get out of this project, both in terms of personal learning, as well as in terms of tangible results (solving some puzzles!).
-In section 2, I will do a recap of the different methods used in the project.
-Section 3 is dedicated to discussing the mathematics of the sliding puzzle \textbf{SP} and the \textbf{RC}.
-I might make interesting observations in there and give some references for the keen reader.
-I give then an overview in section 4 of the code base I have developed (see my github page \cite{FB}) - for this project.
-Finally, sections 6 and 7 present my results on respectively the \textbf{SP} and the \textbf{RC} and section 8 concludes the thesis.
-\\
-\\
-A couple of appendix sections complement the main text.
-The first one details how to use the code base, from constructing puzzles, to learning and solving them using the various solvers I have developed.
-The second one describes a limitation I bumped into while using the kociemba library (a well-known open source 3x3x3 \textbf{RC} solver), and how I circumvented it.
-\newpage
-\thispagestyle{empty}
 \noindent \textbf{Acknowledgements} -- 

### 2022-09-12 latex/Acknowledgements/Acknowledgements.tex
update
trimming abstract and ack

@@
-I am grateful to my employer Bank of America for sponsoring my part time 2-year MSc in Artificial Intelligence at Royal Holloway.
-I entered in the study of AI with a heavy dose of skepticism and am delighted to have learnt so much and revised my judgment of this field.
-In particular I would like to thank my managers Mitrajit Dutta and Stephen Thompson for supporting me in pursuing this MSc.
+I am grateful to Bank of America for sponsoring my part time 2-year MSc in Artificial Intelligence at Royal Holloway.
+In particular I would like to thank my managers Mitrajit Dutta and Stephen Thompson for their support.
+The company of my colleague Saurabh Kumar, as he embarked on the same MSc journey and with whom I have had countless debates about the new cool stuff we learnt, has made the whole process much more enjoyable.
 \\
-\\
-I have been fortunate to have had the company of my colleague Saurabh Kumar, as he decided to follow me in this MSc journey.
-We have had countless opportunities to debate about all the new cool stuff we learnt during the MSc, and that has definitely made the whole process much more enjoyable.
-\\
-\\
-I would also like to thank Professor Chris Watkins, for agreeing to supervise my project.
-It was an honour to attend his very lively lectures on Deep Learning during the first year of the MSc, as well as to absorb some of his wisdom and enthusiasm at our meetings during my work on this project.
-I am also glad to have followed his advice at our last meeting: I stopped shaving for a month, which did free enough time for me to implement DQL (joint policy and value learning), as well as a Monte Carlo Tree Search algorithm, to nicely complement my assortment of sliding-puzzle and Rubik's cube solvers.
-\\
+I would also like to thank Professor Chris Watkins for supervising my project.
+It was an honour to attend his lectures on \textbf{DL} during the 2020-21 school year, and I tried to absorb some of his wisdom at our meetings during this project.
+I am glad to have followed his advice at our last meeting: I stopped shaving for a month, which did free enough time for me to implement \textbf{DQL} (joint policy and value learning), as well as an \textbf{MCTS} algorithm, to nicely complement my assortment of solvers.
 \\

### 2022-09-12 latex/Acknowledgements/Acknowledgements.tex
spell check

@@
-I am grateful to Bank of America for sponsoring my part time 2-year MSc in Artificial Intelligence at Royal Holloway.
-In particular I would like to thank my managers Mitrajit Dutta and Stephen Thompson for their support.
+I am grateful to Bank of America for sponsoring my part time 2-year MSc in Artificial Intelligence at Royal Holloway and would like to thank my managers Mitrajit Dutta and Stephen Thompson for their support.
 The company of my colleague Saurabh Kumar, as he embarked on the same MSc journey and with whom I have had countless debates about the new cool stuff we learnt, has made the whole process much more enjoyable.
@@
 \\
-Last but not least, my wife Thanh Nh\~a is my biggest supporter in everything I do, and her continued encouragement in my pursuit of this MSc has made it so much easier to juggle between a demanding banking job, running 50 to 70 miles a week in preparation for the Chicago marathon, and the coursework, lectures and exam revisions.
+Finally, my wife Thanh Nh\~a is my biggest supporter in everything I do, and her continued encouragement in my pursuit of this MSc has made it so much easier to juggle between a demanding banking job, running 50 to 70 miles a week in preparation for the Chicago marathon, and the coursework, lectures and exam revisions.

### 2022-09-12 latex/Chapters/Conclusion.tex
spell check

@@
 I have also spent the time to learn a few \textbf{RC} algorithms allowing me to solve the 2x2x2 in under one minute the 3x3x3 in 2 minutes and the 4x4x4 in about 5 minutes.
-More specifically regarding the techniques discussed in this project, the principles that underly blind and informed search, A$^{*}$ and heuristics, multi-stage solvers, and of course the various \textbf{AI} methods I have played with (\textbf{DL}, \textbf{DRL} and \textbf{DQL}) all are a great new addition to my tool-set and knowledge.
+More specifically regarding the techniques discussed in this project, the principles that underly blind and informed search, A$^{*}$ and heuristics, multi-stage solvers, and of course the various \textbf{AI} methods I have played with (\textbf{DL}, \textbf{DRL} and \textbf{DQL}) all are a great new addition to my toolset and knowledge.
 \\
 \\
-\textbf{Implementation} -- I am quite satistified to have implemented and played with so many solvers and methods, spanning different \textbf{CS} and \textbf{AI} paradigms, from handcrafting my own \textbf{SP} \textit{naive} solver, playing with Kociemba, to implementing \textbf{BFS}, \textbf{DFS}, A$^{*}$, admissible heuristics such as Manhattan with linear constraints penalty, and of course all the machinery necessary to fit, train and exploit some \textbf{DL}, \textbf{DRL} and \textbf{DQL} heuristics and even \textbf{MCTS}.
-Ideally I would have liked to implement things in C++ for speed and multithreading but time was limited so I decided to prioritize breadth and ease of experimentation over depth and efficiency of implementation.
+\textbf{Implementation} -- I am quite satisfied to have implemented and played with so many solvers and methods, spanning different \textbf{CS} and \textbf{AI} paradigms, from handcrafting my own \textbf{SP} \textit{naive} solver, playing with Kociemba, to implementing \textbf{BFS}, \textbf{DFS}, A$^{*}$, admissible heuristics such as Manhattan with linear constraints penalty, and of course all the machinery necessary to fit, train and exploit some \textbf{DL}, \textbf{DRL} and \textbf{DQL} heuristics and even \textbf{MCTS}.
+Ideally, I would have liked to implement things in C++ for speed and multithreading but time was limited so I decided to prioritize breadth and ease of experimentation over depth and efficiency of implementation.
 Python made it easier to integrate many of the ecosystem's libraries e.g.
@@
 \\
-\textbf{Deep Reinforcement-Q Learning} -- Anyone who has played with \textbf{DL}, \textbf{DRL} and \textbf{DQL} models (\ref{sec:TheoryDLDRL} and \ref{sec:TheoryDQL}) know they can be capricious to train: choosing the convergence criteria, the interplay of training data generation with the inner and outer loops felt more art than sicence.
+\textbf{Deep Reinforcement-Q Learning} -- Anyone who has played with \textbf{DL}, \textbf{DRL} and \textbf{DQL} models (\ref{sec:TheoryDLDRL} and \ref{sec:TheoryDQL}) know they can be capricious to train: choosing the convergence criteria, the interplay of training data generation with the inner and outer loops felt more art than science.
 More tuning, longer training time and computing resources would have been necessary to make these techniques more competitive at higher dimensions, but I remain extremely impressed by how, all things considered - and given how daunting the project seemed at first - it took me little effort to train successful \textbf{DRL} and \textbf{DQL} models capable of solving large state and action spaces puzzles.
@@
 I would have liked to go further with the \textbf{DxL} methods on the 3x3x3 \textbf{RC} but this was sadly not possible within the time imparted.
-As for issues of ethics and safety, outined in the ACM code of ethics (\cite{ACM}), anyone working in \textbf{AI} should in my opinion be aware of them.
+As for issues of ethics and safety, outlined in the ACM code of ethics (\cite{ACM}), anyone working in \textbf{AI} should in my opinion be aware of them.
 There is some divide in the community regarding the dangers of \textbf{AI}.
@@
 His thought-provoking paper-clip-maximiser shows that the most benign looking \textbf{AI} agent, if sufficiently capable, could create disproportionate harm.
-We probably do not want either to see a Rubik's solver engulf all of the earth's and galaxy's resources to compute the nxnxn Rubik's God number for ever growing values of n :)
+We probably do not want either to see a rogue Rubik's solver engulf all of the earth's and galaxy's resources to compute the nxnxn Rubik's God number for ever growing values of n!

### 2022-09-12 latex/Chapters/DRLSearch.tex
spell check

@@
 
-It is fruitful to think of single-player, deterministic, fully-observable puzzles such as the \textbf{SP} and \textbf{RC} as unit-cost graphs.
+It is fruitful to think of single-player, deterministic, fully observable puzzles such as the \textbf{SP} and \textbf{RC} as unit-cost graphs.
 We start from an initial \textit{node} (usually some scrambled configuration of the \textbf{SP} or \textbf{RC}), and via legal moves, we transition to new nodes, until hopefully we manage to get to the goal in the shortest number of moves possible.
@@
 \\
-Graphs search (or graphs traversal) is a rather large, sophisticated and mature branch of \textbf{CS} that studies strategies to find \textit{optimal} paths from initial to goal node in a graph.
-What is meant by \textit{optimal} can vary, but usually is concerned with one or several of minimizing the run-time of the strategies, their memory complexity/usage or the total cost/length of the solutions they come up with.
+Graphs search (or graphs traversal) is a rather large, sophisticated, and mature branch of \textbf{CS} that studies strategies to find \textit{optimal} paths from initial to goal node in a graph.
+What is meant by \textit{optimal} can vary, but usually is concerned with one or several of minimizing the run-time of the strategies, their memory complexity/usage, or the total cost/length of the solutions they come up with.
 In the general theory, different transitions can have different costs associated with them.
@@
 optimal).
-\textbf{DFS} is obvisouly not optimal as it could very well find a solution very deep in the graph while expanding the very first branch, not realising that a solution was one level away from the root on the next branch!
+\textbf{DFS} is obviously not optimal as it could very well find a solution very deep in the graph while expanding the very first branch, not realising that a solution was one level away from the root on the next branch!
 \\
@@
 The better the heuristic, the better \textbf{A$^{*}$} performs.
-An important property of heuristics is \textit{admissibility}: admissible heuristics never over-estimate the real cost-to-go, and can easily be shown to render \textbf{A$^{*}$} optimal.
+An important property of heuristics is \textit{admissibility}: admissible heuristics never over-estimate the real cost-to-go and can easily be shown to render \textbf{A$^{*}$} optimal.
 \teal
@@
 Marvin Minsky, who often took the blame for having \textit{killed} funding into the field, even regretted writing his 1969 book (\cite{minskypapert69}) in which he had proved that three-layer feed forward perceptrons were not quite the universal functions approximators some had hypothesized them to be.
-Since around 2006 (see \cite{GoodBengCour16} for more history of the recent burgeoning of \textbf{DL}), the field has known a rebirth, probably due to a combination of factors, from ever more powerful computers and larger storage, the availability of data sets large and small on the internet, the advances in many techniques and heuristics (backward propagation, normalisation, drop-outs, different optimisation schemes, etc \dots) and the huge amount of experimentation with different architectures (from very deep feed forward fully connected networks, to convolutional, recurrent and more exoctic networks).
-It is not an exageration to say that \textbf{DL} has revolutionized entire fields of research such as computer vision, robotics, computational biology and \textbf{NLP}.
+Since around 2006 (see \cite{GoodBengCour16} for more history of the recent burgeoning of \textbf{DL}), the field has known a rebirth, probably due to a combination of factors, from ever more powerful computers and larger storage, the availability of data sets large and small on the internet, the advances in many techniques and heuristics (backward propagation, normalisation, drop-outs, different optimisation schemes, etc \dots) and the huge amount of experimentation with different architectures (from very deep feed forward fully connected networks, to convolutional, recurrent and more exotic networks).
+It is not an exaggeration to say that \textbf{DL} has revolutionized entire fields of research such as computer vision, robotics, computational biology, and \textbf{NLP}.
 \textbf{DL} has often become the method of choice to learn or approximate highly dimensional functions or random processes.
-The beauty of it is that the relevant features are autodiscovered during the training of the network, via back-ward propagation and trial-and-error, rather than having to be postulated or handcrafted as is often the case with other \textbf{ML} techniques.
+The beauty of it is that the relevant features are auto discovered during the training of the network, via back-ward propagation and trial-and-error, rather than having to be postulated or handcrafted as is often the case with other \textbf{ML} techniques.
 
@@
 This is where \textbf{DL} comes again to the rescue to act as function approximator, combining with \textbf{RL} into what has become known as \textbf{DRL}.
-The subtlety is that both the left-hand-side and right hand side in the value-iteration procedure are given by a \textbf{DL} network which we are constantly tweaking.
+The subtlety is that both the left-hand-side and right-hand-side in the value-iteration procedure are given by a \textbf{DL} network which we are constantly tweaking.
 For my implementation of \textbf{DRL}, I followed the same approach as in \cite{Mnih2013}, utilising two loops.
 During iteration over the first (inner) loop, the right hand-side $V$ of the value-function update equation is computed using a fixed \textit{target-network}.
-The left hand side $V$ is using another \textit{current-network} which alone is modified by the \textbf{DL} learning (backward propagation and optimization), until convergence is deemed reached.
+The left-hand-side $V$ is using another \textit{current-network} which alone is modified by the \textbf{DL} learning (backward propagation and optimization), until convergence is deemed reached.
 When convergence in the inner loop is deemed obtained, we break out of it, and run the outer loop, which updates the \textit{target-network} via a copy of the \textit{current-network}.
@@
 \\
-Similarly to our discussion about \textbf{DRL}, in cases where the state space is large, it can be useful to approximate $Q$, or a somewhat related quantity (such as e.g.
+As in \textbf{DRL}, in cases where the state space is large, it can be useful to approximate $Q$, or a somewhat related quantity (such as e.g.
 a probability distribution over the actions that can be taken from $s$) by a neural network, and use a similar procedure to that of the previous section to update the right-hand-side and left-hand-side of the iterations.
@@
 \\
-I have for this project followed the exact same procedure described in \cite{https://doi.org/10.48550/arxiv.1805.07470} (and also used by alpha-go \cite{AlphaGo}): the output of the neural network that is learnt by \textbf{DQL}, which actually is surprisingly similar to the approach and pseudo code shown for \textbf{DRL} earlier.
+I have for this project followed the exact same procedure described in \cite{https://doi.org/10.48550/arxiv.1805.07470} (and also used by alpha-go \cite{AlphaGo}): the output of the neural network that is learnt by \textbf{DQL}, which is surprisingly similar to the approach and pseudo code shown for \textbf{DRL} earlier.
 The only 3 differences are that:
@@
 The paths followed from initial state to leaves are decided via a combination of the probability distribution over actions (allowing exploration) and value function (exploitation), according to a trade-off which can be tuned via a hyper-parameter $c$.
-To prevent following the same paths (which might lead the search far from the goal), the \textbf{MCTS} keeps track of actions already followed and downweight their probability so that subsequent paths explore different actions.
+To prevent following the same paths (which might lead the search far from the goal), the \textbf{MCTS} keeps track of actions already followed and down-weight their probability so that subsequent paths explore different actions.
 Let me first present in pseudo-code the paths generation, in particular the trade-off between actions-values (exploration-exploitation):

### 2022-09-12 latex/Chapters/Implementation.tex
trimming abstract and ack
spell check

@@
 
-The code I have developed for this project is all publicly available on my github page (\cite{FB}).
+The code I have developed for this project is all publicly available on my github page (see \cite{FB}).
 It can easily be installed using the setup file provided, which makes it easy to then use Python's customary import command to play with the code.
@@
 It adds over non-empty tiles the $L_{1}$ distance between their current and their target position.
-This heuristic is admissible: indeeed, each move displaces one (and one only) tile by one position (notice that in other puzzles such as the \textbf{RC}, a move might affect multiple tiles).
+This heuristic is admissible: indeed, each move displaces one (and one only) tile by one position (notice that in other puzzles such as the \textbf{RC}, a move might affect multiple tiles).
 Therefore, if tiles could somehow freely move on top of one-another, the number of moves necessary to solve the \textbf{SP} would exactly be the Manhattan distance.
@@
 The right penalty to add needs to be computed recursively.
-The following configuration for instance with all pairs (1, 2), (1, 3) and (2, 3) out-of-order only inccurs an extra penalty of 4, not 6.
+The following configuration for instance with all pairs (1, 2), (1, 3) and (2, 3) out-of-order only incurs an extra penalty of 4, not 6.
 \begin{center}
@@
 \\
-Finally, as suggested in the rather vagues details of \cite{SlidingPuzzleLectureNotes}, I have precomputed and stored all the penalties for all possible rows, columns and tiles ordering.
-For memory efficiency, I only saved non-zero panalties.
+Finally, as suggested in the rather vague details of \cite{SlidingPuzzleLectureNotes}, I have precomputed and stored all the penalties for all possible rows, columns and tiles ordering.
+For memory efficiency, I only saved non-zero penalties.
 The very first time a call to Manhattan++ is made, for a given dimension nxm, the appropriate linear constraint penalties are computed and saved.
@@
 \begin{itemize}
-\item \textbf{DeepLearning}: a Puzzled Loggable Factory that can instantiate some configurable deep networks, and provide the necessary glue with the rest of the code base so that puzzles be seemlessly passed to the networks and trained on.
+\item \textbf{DeepLearning}: a Puzzled Loggable Factory that can instantiate some configurable deep networks, and provide the necessary glue with the rest of the code base so that puzzles be seamlessly passed to the networks and trained on.
 \item \textbf{FullyConnected}: wrapper around a Pytorch fully connected network, with configurable hidden layers and size.
@@
 It provides some common code to learners (to save or purge what they learnt), kick off learning and plot results.
-Concrete derived implementation define what and how they learn, and what interesting they can display about this learning process.
+Concrete derived implementations define what and how they learn, and what interesting they can display about this learning process.
 The four implemented learners are:
@@
 Notice that for puzzles of too high dimension, where my computing resources will not allow to solve exhaustively all the configurations of a given dimension, this class can still be used to populate a data base of optimal costs, which can then be used by DeepLearner.
-If it is to be used this way, the PerfectLearner can be configured to use perfectly random configurations to learn from, rather than going through the configurations one by one in a well defined order.
+If it is to be used this way, the PerfectLearner can be configured to use perfectly random configurations to learn from, rather than going through the configurations one by one in a well-defined order.
 
-\item \textbf{DeepLearner} DeepLearner instanciates a DeepLarning (neural network), and trains it on training data by minimizing the mean square error between the cost-to-go from the target solver versus the neural network.
+\item \textbf{DeepLearner} DeepLearner instantiates a DeepLarning (neural network), and trains it on training data by minimizing the mean square error between the cost-to-go from the target solver versus the neural network.
 The main parameters driving this learner are:
@@
 \item \textit{nb\_epochs} and \textit{threshold} which control exit criteria for the network training.
-\item \textit{nb\_sequences}, \textit{nb\_shuffles\_min}, \textit{nb\_shuffles\_max}, \textit{training\_data\_freq} and \textit{training\_data\_every\_epoch} which control how often we regenerate new training data, how many sequences of puzzles it contais, how scrambled the puzzles should be.
+\item \textit{nb\_sequences}, \textit{nb\_shuffles\_min}, \textit{nb\_shuffles\_max}, \textit{training\_data\_freq} and \textit{training\_data\_every\_epoch} which control how often we regenerate new training data, how many sequences of puzzles it contains, how scrambled the puzzles should be.
 \item \textit{learning\_rate}, \textit{scheduler}, \textit{optimizer} and assorted parameters which control the optimisation at each backward propagation. 
@@
 \item \textit{\_\_construct\_target\_\_} which performs the the Q-V-iteration to update the targets, as described in pseudo code form in \ref{sec:TheoryDQL}.
-This function updates not only the value function of a node given its children nodes, but also updates the optimal action vector (0 for non optimal actions, 1 for the optimal action leading to the child node of highest value function).
+This function updates not only the value function of a node given its children nodes, but also updates the optimal action vector (0 for non-optimal actions, 1 for the optimal action leading to the child node of highest value function).
 \end{itemize}
@@
 This module implements the actual puzzles solvers.
-The base class Solver is a Factory, and in addition to being able to instantiating the following types of solvers, can run different solvers through a similal sequences of random puzzles (for various increasing degrees of difficulty (scrambling), and/or perfectly shuffled ones) and display a comparison of how they perform in a number of metrics (configurable from percentage of puzzles solved before time out, percentage of optimal puzzle (if optimality can be ascertained), mean-median-maximum run time, mean-median-maximum expanded nodes (for solvers where it makes sense), mean-median-max cost, and an optimality score which compares the total sum of cost over all puzzles versus a benchmark solver).
-Ideally the benchmark solver to compute the optimality score should provably optimal, but in the case of the \textbf{RC} that was not possible so I used A$^{*}$ with KociembaHeuristic as my benchmark.
+The base class Solver is a Factory, and in addition to being able to instantiate the following types of solvers, can run different solvers through a similar sequence of random puzzles (for various increasing degrees of difficulty (scrambling), and/or perfectly shuffled ones) and display a comparison of how they perform in a number of metrics (configurable from percentage of puzzles solved before time out, percentage of optimal puzzle (if optimality can be ascertained), mean-median-maximum run time, mean-median-maximum expanded nodes (for solvers where it makes sense), mean-median-max cost, and an optimality score which compares the total sum of cost over all puzzles versus a benchmark solver).
+Ideally the benchmark solver to compute the optimality score should provably optimal, but in the case of the \textbf{RC} that was not possible, so I used A$^{*}$ with KociembaHeuristic as my benchmark.
 For the \textbf{SP} I used the optimal A$^{*}$ with Manhattan++ heuristic.
@@
 This solver basically generates \textit{pseudo-random} paths from initial node to a new leaf at each iteration (see \ref{alg:TheoryMCTS}), until we obtain a goal leaf.Paths are, despite the name, completely deterministic: they indeed result from a trade-off between the probability of actions (penalised for repeated use) and the value function, but there are no random draws being performed.
-\\ As per McAleer \& Agostilenni's paper, I have implemented a final pruning via \textbf{BFS} of the tree resulting from the above \textit{Monte-Carlo} process.It is inded clear that the paths generated by this \textbf{MCTS} algorithm have little chance of being optimal since there is no correction mechanism to favour shortcuts over long loops.
+\\ As per McAleer \& Agostilenni's paper, I have implemented a final pruning via \textbf{BFS} of the tree resulting from the above \textit{Monte-Carlo} process.
+It is indeed clear that the paths generated by this \textbf{MCTS} algorithm have little chance of being optimal since there is no correction mechanism to favour shortcuts over long loops.
 I have however made this \textbf{BFS} trimming optional via a \textit{trim\_tree} parameter to study its effect on solutions' optimality and run time (see results section \ref{sec:MTCSHyperParamsEffect}).

### 2022-09-12 latex/Chapters/Objectives.tex
spell check

@@
 For the project component of the MSc, I thought it would be interesting (and fun) for me to apply some of the \textbf{DL}, \textbf{DRL} and search techniques from \textbf{AIPnT} to the sliding puzzle (\textbf{SP}) and of course the famous Rubik's cube (\textbf{RC}).
-I am in particular looking to solidify my understanding of \textbf{DRL} and \textbf{DQL} by  implementing and experimenting with concrete (though arguably of limited practical use) problems.
+I am looking to solidify my understanding of \textbf{DRL} and \textbf{DQL} by  implementing and experimenting with concrete (though arguably of limited practical use) problems.
 

### 2022-09-12 latex/Chapters/Puzzles.tex
spell check

@@
 Since then, several variations of the 15-puzzle have become popular, such as the 24-puzzle.
-A rather contrived but interesting one is the coiled 15-puzzle (\cite{Coiled15Puzzle}), where the bottom-right and the top-left compartments are adjascent; the additional move that this allows renders all $16!$ configurations solvable.
+A rather contrived but interesting one is the coiled 15-puzzle (\cite{Coiled15Puzzle}), where the bottom-right and the top-left compartments are adjacent; the additional move that this allows renders all $16!$ configurations solvable.
 
@@
 This obviously cannot change the parity since exactly 2 pairs of numbers are now in a different order, that is (14, 11) and (14, 8) now appear in the respective opposite orders as (11, 14) and (8, 14).
-This is the crux of the proof of the well known necessary condition (even parity of p(c)) for a configuration c to be solvable (see part I of \cite{Johnson1879}).
+This is the crux of the proof of the well-known necessary condition (even parity of p(c)) for a configuration c to be solvable (see part I of \cite{Johnson1879}).
 In the case of the coiled puzzle, we can solve all configurations of even parity, since all the legal moves of the normal puzzle are allowed.
@@
 In this thesis, as well as in the code base (\cite{FB}) I have written to do this project, we will consider the general case of a board with n columns and m rows, where $(n, m) \in {\mathbb{N}^{+}}^{2}$, forming $n * m$ compartments.
-$n * m - 1$ tiles, numbered 1 to $n * m - 1$ are placed in all the compartments but one (which is left empty), and we can slide a tile directly adjascent to the empty compartment into it.
+$n * m - 1$ tiles, numbered 1 to $n * m - 1$ are placed in all the compartments but one (which is left empty), and we can slide a tile directly adjacent to the empty compartment into it.
 Notice from a programming and mathematical analysis perspective, it is often easier to equivalently think of the empty compartment being moved into (or swapped with) an adjacent tile.
@@
 A favourite past time among computer scientists around the globe is therefore to search for more refined lower and upper bounds for $\mathcal{G}(n, m)$, for ever increasing values of n and m.
-For moderate n and m, we can actually solve optimally all possible configurations of the puzzle and compute exactly $\mathcal{G}(n, m)$  (using for instance $A^{*}$ and an admissible heuristic (recall \ref{GSH}, and we shall see modest examples of that in the results section later).
+For moderate n and m, we can solve optimally all possible configurations of the puzzle and compute exactly $\mathcal{G}(n, m)$  (using for instance $A^{*}$ and an admissible heuristic (recall \ref{GSH}, and we shall see modest examples of that in the results section later).
 For larger values of n and m (say 5 by 5), we do not know what the God number is.
@@
 The second puzzle I will focus on is the Rubik's Cube (\textbf{RC}), invented in 1974 by Erno Rubik, originally known as \textit{Magic Cube} (see \cite{RubiksWiki}).
-Since it was commercialised in 1980, it has kown a worldwide success: countless people are playing with the original 3x3x3 cube, as well as with variants of it of all shapes and sizes.
-Competitions bring together \textit{speedcubers} who have trained for years and memorized algorithms to solve the Rubik's in astonishly fast times (seconds for the best ones).
+Since it was commercialised in 1980, it has known a worldwide success: countless people are playing with the original 3x3x3 cube, as well as with variants of it of all shapes and sizes.
+Competitions bring together \textit{speedcubers} who have trained for years and memorized algorithms to solve the Rubik's in astonishingly fast times (seconds for the best ones).
 Some of the principles used by speedcubers generalise to different dimensions.
@@
   \makebox[\textwidth]{\includegraphics[scale=0.25]{./Figures/RC333444}}
-  \caption[Reducing RC]{Reducing a 4x4x4 \textbf{RC} to a 3x3x3 -- home made picture}
+  \caption[Reducing RC]{Reducing a 4x4x4 \textbf{RC} to a 3x3x3 -- home-made picture}
   \label{fig:4442333}
@@
 red an and white on opposite faces).
-Overall, the difficuly of hand-solving cubes does not grow as much with dimensionality as one would think, and the time difference comes mostly from the difficulty of handling and manipulating larger cubes.
+Overall, the difficulty of hand-solving cubes does not grow as much with dimensionality as one would think, and the time difference comes mostly from the difficulty of handling and manipulating larger cubes.
 
@@
 I follow the conventions and notations of the \textbf{RC} community and refer to the faces as F (front), B (back), U (up), D (down), L (left) and R (right).
-It is common in \textbf{RC} jargon to refer to moves  by the name of the face that is rotated clock-wise (when facing it), and by adding a \textit{'} for counter-clock-wise rotations.
+It is common in \textbf{RC} jargon to refer to moves  by the name of the face that is rotated clockwise (when facing it), and by adding a \textit{'} for counter-clockwise rotations.
 Sometimes numbers are used to indicate repeated moves, though I will count that as multiple moves, since obviously e.g.
 U2 = U U = U' U'.
-As another example: F B' F2 means rotating the front face clock-wise, the back face counter-clock-wise, followed by the front face twice.
-The convention I follow is called the \textit{quarter-turn-metric}, and obvioulsy the length of solutions is affected by which metric we use.
+As another example: F B' F2 means rotating the front face clockwise, the back face counter-clockwise, followed by the front face twice.
+The convention I follow is called the \textit{quarter-turn-metric}, and obviously the length of solutions is affected by which metric we use.
 \\
@@
 The second observation is that when facing a corner cubie, the order of its three color (e.g.
-enumerated clock-wise) is invariant.
+enumerated clockwise) is invariant.
 For instance, consider the corner cubie with colors (F, R, U).
-No scrambling of the cube will ever produce clock-wise order of e.g.
+No scrambling of the cube will ever produce clockwise order of e.g.
 (F, U, R).
@@
 Computing $N$ is a bit tricker.
-For the 3x3x3 Rubik's, each of the 8 corner cubies could be placed at one of the physical corner locations and oriented three different ways (rather than six, since as discussed earlier, the clock-wise orientation of a given corner cubie can never be changed).
+For the 3x3x3 Rubik's, each of the 8 corner cubies could be placed at one of the physical corner locations and oriented three different ways (rather than six, since as discussed earlier, the clockwise orientation of a given corner cubie can never be changed).
 We have therefore $8!
@@
 2^{12}$.
-There are however some well know parities (see \cite{Schoenert} for details) which are invariant under legal moves of the Rubik's cube and which reduce the number of arrangements that are physically attainable via face rotations.
+There are however some well-known parities (see \cite{Schoenert} for details) which are invariant under legal moves of the Rubik's cube and which reduce the number of arrangements that are physically attainable via face rotations.
 The \textit{corner orientation parity} (\textbf{COP}) for instance sums over the corner cubies the number of 120$^{o}$ rotations it would take to bring a given cubie to its standard orientation (i.e.
@@
 Notice that this is exactly how beginners and advanced speedcubers solve the Rubik's cube, except they usually have more than 2 stages as described above.
-$S_{1}$ might for instance be the set of cubes where both lower layers are solved, the cross on the top layer as well as all corners are in their respective places (but maybe not oriented as expected), and then they search within a well known and small set of algorithms how to get to the goal state.
+$S_{1}$ might for instance be the set of cubes where both lower layers are solved, the cross on the top layer as well as all corners are in their respective places (but maybe not oriented as expected), and then they search within a well-known and small set of algorithms how to get to the goal state.
 A previous stage might consist in $S_{2}$, the set of configurations where the two layers are solved, but the top layer is random, and then they will search for a sequence of moves to get into $S_{1}$, etc \dots
@@
 Pretty much all the papers about \textbf{RC} that I have surveyed chose a more compact representation for the cube than I have.
-They remove the center cubies of the 3x3x3 cube since, they argue, those never move via face-turns.
+They remove the centre cubies of the 3x3x3 cube since, they argue, those never move via face-turns.
 Some of them go much further and make use of various symmetry arguments to reduce the search space and reduce the \textbf{RC} to having one goal state (instead of 24).
@@
 This is still \textit{cheating} a bit in my opinion, especially if you are claiming to solve these puzzles \textit{without-human-knowledge}, as we can only perform these dimensionality reduction tricks if we have external world knowledge about the physical structure of the puzzles.
-In this project, I have chosen as pure an approach as I could, and not only did not reduce the dimensionality of the tensor representation of the puzzles, but also, specifically for the Rubik's cube, I did not make use of invariance by spacial rotation of the full cube.
+In this project, I have chosen as pure an approach as I could, and not only did not reduce the dimensionality of the tensor representation of the puzzles, but also, specifically for the Rubik's cube, I did not make use of invariance by spatial rotation of the full cube.

### 2022-09-12 latex/Chapters/Rubiks.tex
spell check

@@
 \Section{2x2x2}
-All the following experiments on the 2x2x2 \textbf{RC} have been run using my Solver's \textit{performance\_test} (see \ref{sec:codesolvers}): for each level of difficulty, the test present the solvers with the same 100 random configurations (for fairness and to increase the significance of the results).
+All the following experiments on the 2x2x2 \textbf{RC} have been run using my Solver's \textit{performance\_test} (see \ref{sec:codesolvers}): for each level of difficulty, the test presents the solvers with the same 100 random configurations (for fairness and to increase the significance of the results).
 The levels of difficulty (number of scrambles from goal) used here go from 2 to 20 in step of 2.
@@
 \Subsection{BFS}
-For \textbf{BFS} I used a step of 1 on the difficuly and stopped at 5 as it was clear it would not go much further.
+For \textbf{BFS} I used a step of 1 on the difficulty and stopped at 5 as it was clear it would not go much further.
 Its run time and expanded nodes are increasing linearly with difficulty, and it was already taking an average close to a minute for 5 shuffles from goal.
@@
 \item Fully connected network (\textit{network\_type=fully\_connected\_net}) with three hidden layers (\textit{layers\_description}=(600,300,100)), an RMSProp optimiser (\textit{optimiser=rms\_prop}), \textit{scheduler=exponential\_scheduler} with \textit{gamma\_scheduler}=0.9999 and \textit{learning\_rate}=0.1\%.
-The networks architecture are shown in figure \ref{fig:222RCnets}.
+The networks architectures are shown in figure \ref{fig:222RCnets}.
 \end{itemize}
@@
 From both graphs (\ref{fig:222RCDRL} \& \ref{fig:222RCDQL}) we can see that the learners saw less than 1\% of all puzzles.
-It is good to keep in mind however that \textit{seen} does not mean really much here, since the learning is \textit{unsupervised}.
-This is quite in contrast to the DeepLearner, whose traiing data consist of puzzles with their cost-to-go computed by a \textit{teacher} (optimal or other efficient solver).
+It is good to keep in mind however that \textit{seen} does not mean really much here since the learning is \textit{unsupervised}.
+This is quite in contrast to the DeepLearner, whose training data consist of puzzles with their cost-to-go computed by a \textit{teacher} (optimal or other efficient solver).
 \end{itemize}
@@
 A$^{*}[Kociemba]$ never took more than 30 moves on any of the cubes presented to it, which is not bad given the known God number of 26 in the quarter-turn-metric (as discussed earlier in \ref{Theory:333RCSSS})
-\item A$^{*}$[\textbf{DL}] managed to solve all puzzles scrambled less than 10 times, and found shorter worst case solutions than even A$^{*}[Kociemba]$ on which it was trained.
-Within the 2 hours time out, it failed to solve one of the 50 cubes for difficulty 10, and two of the 50 cubes of difficulty 12, so i did not try further than that.
+\item A$^{*}$[\textbf{DL}] managed to solve all puzzles scrambled less than 10 times, and found shorter worst-case solutions than even A$^{*}[Kociemba]$ on which it was trained.
+Within the 2 hours time-out, it failed to solve one of the 50 cubes for difficulty 10, and two of the 50 cubes of difficulty 12, so i did not try further than that.
 \item A$^{*}$[\textbf{DRL}] only managed to solve all puzzles of difficulty less than 6 but at least did find only optimal solutions (compared with \textbf{BFS}).

### 2022-09-12 latex/Chapters/SlidingPuzzle.tex
spell check

@@
 As mentioned in chapter \ref{sec:Puzzles}, the state space cardinality for the \textbf{SP} grows quickly with n and m.
-The only dimensions which have less than 239.5 millions states are shown in table \ref{tab:smallSP}.
+The only dimensions which have less than 239.5 million states are shown in table \ref{tab:smallSP}.
 
@@
 \end{table}
-\noindent Among those dimensions, 2x5 has the largest state space and my result for it is still provisional, as I have only solved about 1.2 out of 1.8 million puzzles at the time of writing, despite running a combined 600 hours (mostly over 16 cores, except for a full week-end run on a 72-core c5.18xlarge Amazon EC2 instance).
+\noindent Among those dimensions, 2x5 has the largest state space and my result for it is still provisional, as I have only solved about 1.2 out of 1.8 million puzzles at the time of writing, despite running a combined 600 hours (mostly over 16 cores, except for a full weekend run on a 72-core c5.18xlarge Amazon EC2 instance).
 It is well possible that I haven't yet bumped into the hardest 2x5 configuration!
@@
 \label{sec:S33PL}
-As discussed in the previous section section \ref{sec:SPLowDimension}, the 3x3 \textbf{SP} is one of the cases I have been able to solve perfectly, since it only has 181,440 possible configurations.
-Its God number is only 31, which definitely makes it manageable.
+As discussed in the previous section \ref{sec:SPLowDimension}, the 3x3 \textbf{SP} is one of the cases I have been able to solve perfectly, since it only has 181,440 possible configurations.
+Its God number is only 31, which makes it manageable.
 However, this is already an intermediary size, large enough that it is worth trying and comparing a few different methods, including \textbf{DL}, \textbf{DRL} and \textbf{DQL}.
@@
 In this subsection I explore the effect on \textbf{MCTS} of the trade-off hyper-parameter $c$ as well as of trimming the resulting tree via \textbf{BFS}.
-(see the threory section \ref{sec:TheoryMCTS} as well as the implementation section \ref{sec:codesolvers} for more details).
+(see the theory section \ref{sec:TheoryMCTS} as well as the implementation section \ref{sec:codesolvers} for more details).
 To that effect, I have run my Solver performance test with four different values of $c$ (0, 1, 10 and 100), and in each case, with and without \textbf{BFS} trimming.
 A$^{*}[Manhattan++]$ is also presented to have an optimal solver benchmark.
-For each level of difficulty, the test present the solvers with the same 150 random configurations (for fairness and to increase the significance of the results).
+For each level of difficulty, the test presents the solvers with the same 150 random configurations (for fairness and to increase the significance of the results).
 The levels of difficulty (number of scrambles from goal) used here go from 5 to 50 in step of 5.
@@
 \item As expected, \textbf{BFS} becomes quickly hopeless.
-\item The naive solver is obvioulsy by far the fastest but finds poor solutions (about 2.5 longer on average than optimal).
+\item The naive solver is obviously by far the fastest but finds poor solutions (about 2.5 longer on average than optimal).
 \item As discussed in the previous section, \textbf{MCTS} is slow, and even with much exploration and \textbf{BFS} tree trimming found solution about 25\% longer than optimal.
@@
 They ended up \textit{seeing} around 25\% of all the possible configurations during training (in an \textit{unsupervised} way though) so did show some reasonable amount of out-of-sample generalization.
-Their optimality score was an impressive 99\%+ at all levels of difficuly for \textbf{DRL} and 98\%+ for \textbf{DQL}.
+Their optimality score was an impressive 99\%+ at all levels of difficulty for \textbf{DRL} and 98\%+ for \textbf{DQL}.
 Even more interestingly, the \textbf{DQL} heuristic expanded almost 4 times less nodes than \textbf{DRL}, suggesting some benefit from the joint learning of the policy and value function.
@@
 In an implementation without duplicate pruning, \textbf{BFS} would time out no matter what the bound is set.
-Indeed, it would need to explore in the order of $3^{31}$ - roughly 617 trillions - nodes to reach the goal!
+Indeed, it would need to explore in the order of $3^{31}$, roughly 617 trillions nodes to reach the goal!
 Even with my implementation which does pruning, it would need to pretty much traverse the entire 3x3 \textbf{SP} tree.
 \\
-Rather interestingly, the \textbf{DRL} heuristic performs much better than the Manhattan heuristic (not super suprising), but also outperforms the perfect heuristic quite significantly both in terms of run time and of nodes expansion.
-Obviously there is no guarantee that the perfect heuristic will not be outperformed on some random configuration, and it does on this occasion.
-However, as we have seen in the previous subection \ref{ssec:33SPSC}, it is not the case on average.
+Rather interestingly, the \textbf{DRL} heuristic performs much better than the Manhattan heuristic (not super surprising), but also outperforms the perfect heuristic quite significantly both in terms of run time and of nodes expansion.
+Obviously, there is no guarantee that the perfect heuristic will not be outperformed on some random configuration, and it does on this occasion.
+However, as we have seen in the previous subsection \ref{ssec:33SPSC}, it is not the case on average.
 \\
@@
 Key take-aways are that all \textbf{DxL} outperformed A$^{*}$[Manhattan++] in terms of node expansion and got optimality scores of 99\%+ for all levels of difficulty, despite only seeing 0.009\% of all configurations for \textbf{DL}, and 0.27\% for \textbf{DRL} and \textbf{DQL}.
-Again this shows very impressive generalisation and performance from the\textbf{DxL} methods.
+Again this shows very impressive generalisation and performance from the \textbf{DxL} methods.
 
@@
 Again \textbf{DL} got trained on randomly generated configurations solved by A$^{*}$[Manhattan++].
-\textbf{DL} obtained an optimality scores of 97\%+ across all levels of difficulty and \textbf{DRL} managed 99\%+.
+\textbf{DL} obtained an optimality score of 97\%+ across all levels of difficulty and \textbf{DRL} managed 99\%+.
 \textbf{DL} and \textbf{DRL} saw of the order of respectively $10^{-7}$\% and $10^{-5}$\% of all configurations during training, so again showed impressive performance and generalization!

### 2022-09-13 latex/Abstract/Abstract.tex
last cleanup

@@
 \label{Abstract}
-\noindent \textbf{Motivation} -- Reinforcement Learning (\textbf{RL}), exposed with brilliant clarity in \cite{Sutton1998}, has until recently known less success than we might have hoped for.
+\noindent \textbf{Motivation} -- Reinforcement Learning (\textbf{RL}), exposed with brilliant clarity in \cite{Sutton1998}, had until recently known less success than we might have hoped for.
 Its framework is very appealing and intuitive.
 In particular, the mathematical beauty of Value and Q iteration (\cite{WatkinsThesis}) for discrete state and action spaces, blindly iterating from \textit{any} initial value, is quite profound.
-Sadly, it had until recently proven hard to achieve practical success with these methods.
+Sadly, it had proven hard to achieve practical success with these methods.
 Inspired however by the seminal work of Deep Mind's team in using Deep Reinforcement Learning (\textbf{DRL}) to play Atari games (\cite{Mnih2013})  and to master the game of Go (\cite{AlphaGo}), researchers have made a lot of progress towards designing algorithms capable of learning and solving, \textit{without human knowledge}, the Rubik's Cube (\textbf{RC}) by using Deep Q-Learning (\textbf{DQL})  (\cite{DBLP:journals/corr/abs-1805-07470}) or search and (\textbf{DRL}) value iteration (\cite{https://doi.org/10.48550/arxiv.1805.07470}).
@@
 In section 2, I do a recap of the methods used in the project.
-Section 3 is dedicated to the mathematics of the sliding puzzle \textbf{SP} and the \textbf{RC}.
+Section 3 is dedicated to the mathematics of the sliding puzzle (\textbf{SP}) and the \textbf{RC}.
 I give then an overview of my implementation in section 4.

### 2022-09-13 latex/Appendices/Examples.tex
last cleanup

@@
 \includegraphics[scale=0.41]{./Figures/exampleperfectlearner}
-%\decoRule
 \caption[Examples]{perfect learner example}
@@
 
-As can be seen in the code snippet, this example will generate 100 perfectly shuffled (n=5, m=2) SPs and solve them.
+\noindent As can be seen in the code snippet, this example will generate 100 perfectly shuffled 5x2 \textbf{SP}s and solve them.
 Once done, a summary of the training data is printed, indicating, for each optimal cost, how many sequences have been generated.
@@
 \includegraphics[scale=0.45]{./Figures/exampledeeplearnertraining}
-%\decoRule
 \caption[Examples]{deep learner training example}
@@
 
-Then the Deep Learner will get trained on this data for 999 epochs.
+\noindent Then the Deep Learner will get trained on this data for 999 epochs.
 In the above example, I have chosen to fetch, every 100 epochs, 50 random sequences of puzzles from the training data.
@@
 We can see that the (MSE) loss on the value function decreases rapidly, and jumps back up every time we change the training data (since it has not yet seen some of it).
-By the end of the training, the in-sample MSE loss has dropped to 1.5 and the Deep Learner has seen 0.14\% of the possible (n=5, m=2) puzzles.
+By the end of the training, the in-sample MSE loss has dropped to 1.5 and the Deep Learner has seen 0.14\% of the possible 5x2 puzzles.
 
-\newgeometry{top=0mm, bottom=0mm, left=0mm, right=0mm}
-\begin{landscape}
-\centering\vspace*{\fill}
 \begin{figure}[H]
-\centering
-\includegraphics[scale=0.45]{./Figures/exampledeeplearnerlearning}
-%\decoRule
+  \noindent
+  \makebox[\textwidth]{\includegraphics[scale=0.4]{./Figures/exampledeeplearnerlearning}}
   \caption[Examples]{deep learner learning data example}
@@
 \end{figure}
-\vfill
-\end{landscape}
-\restoregeometry
 
@@
 Unlike the Deep Learner, the Deep Reinforcement Learner learns unsupervised (in the sense that there is no need to pre-solve puzzles to tell if what the actual (target) costs are), since it generates its own target using a combination of a target network and the simple min rule described in \ref{sec:codelearners}.
-Let us see here how to run it on the (n=4, m=2) SP for instance.
+Let us see here how to run it on the 4x2 \textbf{SP} for instance.
 The following code snippet will run a Deep Reinforcement Learner for a maximum of 25,000 epochs, generating randomly 10 sequences of puzzles shuffled 50 times from goal state every time it updates the target network.
-That update will happen either after 500 epochs or when the (MSE) loss on the value function gets under one thousands of the max target value.
+That update will happen either after 500 epochs or when the (MSE) loss on the value function gets under one thousandth of the max target value.
 The learner will stop if it reaches 25,000 epochs or if the target network updates 40 times.
@@
 By that point, the DRL learner had seen 40\% of the possible puzzles, and the very last MSE loss (after update, so out-of-sample) was around 0.4, corresponding to 2\% of the max target cost.
-It is interesting to notice that since I have shuffled the sequences only 50 times, and since the (n=4, m=2) SP is quite constrained in terms of possible moves, the max target ever produced by the network was only around 25, whereas we know the God number for this dimension is 36 (see later section \ref{sec:SPLowDimension}).
+It is interesting to notice that since I have shuffled the sequences only 50 times, and since the 4x2 \textbf{SP} is quite constrained in terms of possible moves, the max target ever produced by the network was only around 25, whereas we know the God number for this dimension is 36 (see later section \ref{sec:SPLowDimension}).
 It is therefore likely that the resulting network would not produce very optimal solutions for puzzles whose cost is in the region [25, 36].
 
-\newgeometry{top=0mm, bottom=0mm, left=0mm, right=0mm}
-\begin{landscape}
-\centering\vspace*{\fill}
 \begin{figure}[H]
-\centering
-\includegraphics[scale=0.55]{./Figures/exampledeepreinforcementlearnerlearning}
-%\decoRule
+  \noindent
+  \makebox[\textwidth]{\includegraphics[scale=0.4]{./Figures/exampledeepreinforcementlearnerlearning}}
   \caption[Examples]{deep reinforcement learner learning data example}
@@
 \label{BFSSS}
-The following example shows how to use Breadth First Search to solve a (n=3, m=3) SP which we do not \textit{shuffle} too much.
+The following example shows how to use Breadth First Search to solve a 3x3 \textbf{SP} which we do not \textit{shuffle} too much.
 
 \afblue
-\paragraph{}{\textbf{python code -- breadth first search solver}}
+\paragraph{}{\textbf{python code -- \textbf{BFS} solver}}
 \begin{python}
@@
     m=3
-    nb_shuffles=10
+    nb_shuffles=5
     solver_type=Solver.bfs
@@
 
-As expected (since BFS is optimal), the solution found, printed by the snippet of code above, is optimal.
+As expected (since \textbf{BFS} is optimal), the solution found, printed by the snippet of code above, is optimal.
 The check\_optimal flag in the code snippet indicates that the solver should let us know if solution is optimal.
@@
 \begin{figure}[H]
-\centering
-\includegraphics[scale=0.25]{./Figures/examplebfssolver}
-%\decoRule
+  \noindent
+  \makebox[\textwidth]{\includegraphics[scale=0.32]{./Figures/examplebfssolver}}
   \caption[Examples]{breadth first search solver example}
@@
 \black
-Here we can see, in contrast with the previous example, that the solution obtained by DFS is not optimal.
+Here we can see, in contrast with the previous example, that the solution obtained by \textbf{DFS} is not optimal.
 The Solver indicates it and shows an optimal solution.
@@
 \begin{figure}[H]
-\centering
-\includegraphics[scale=0.2]{./Figures/exampledfssolver}
-%\decoRule
+  \noindent
+  \makebox[\textwidth]{\includegraphics[scale=0.22]{./Figures/exampledfssolver}}
   \caption[Examples]{depth first search solver example}
@@
 \item if n and m are both equal to 2, we just keep moving the empty tile clock-wise until the puzzle is solved.
-Notice that this is bound to work, since moving clock-wise or counter-clock-wise are the two ony possible moves, and one of them is just un-doing the other one, therefore the only possible sequence of move in a (n=2, m=2) puzzle is to either keep moving clock-wise or counter-clock-wise.
+Notice that this is bound to work, since moving clock-wise or counter-clock-wise are the two ony possible moves, and one of them is just un-doing the other one, therefore the only possible sequence of move in a 2x2 puzzle is to either keep moving clock-wise or counter-clock-wise.
 \item if n $\geq$ m, we solve the top row
@@
 \\
-As a concrete example, let us assume we started with the following (n=6, m=6) puzzle:
+As a concrete example, let us assume we started with the following 6x6 puzzle:
 \\
@@
 \\
-After one call to solve the top row and the left column, we are left with solving the (n=5,m=5) sub-puzzle in blue:
+After one call to solve the top row and the left column, we are left with solving the 5x5 sub-puzzle in blue:
 \\
@@
 \\
-and we are left with solving the bottom sub-puzzle (n=4,m=5):
+and we are left with solving the bottom 4x5 sub-puzzle:
 \\
@@
 \\
-Below is a simple code snippet to run the naive solver on a randomly generated (n=2, m=2) SP:
+Below is a simple code snippet to run the naive solver on a randomly generated 2x2 SP:
 \afblue
@@
 \begin{figure}[H]
-\centering
-\includegraphics[scale=0.39]{./Figures/examplenaivesolver}
-%\decoRule
+  \noindent
+  \makebox[\textwidth]{\includegraphics[scale=0.42]{./Figures/examplenaivesolver}}
   \caption[Examples]{naive solver example}
@@
 \begin{figure}[H]
-\centering
-\includegraphics[scale=0.55]{./Figures/examplekociemba}
-%\decoRule
+  \noindent
+  \makebox[\textwidth]{\includegraphics[scale=0.58]{./Figures/examplekociemba}}
   \caption[Examples]{Kociemba solver example}
@@
 \afblue
-\paragraph{}{\textbf{python code -- depth first search solver}}
+\paragraph{}{\textbf{python code -- A$^{*}$ solver}}
 \begin{python}
@@
 As can be seen from the output below, the Manhattan++ improves on the number of expanded nodes (as expected, since the heuristic is less optimistic while retaining its optimality property).
-See a more detailed analysis in the SP results section \ref{}
+See a more detailed analysis in the \textbf{SP} results section \ref{MHComp}
 
 \begin{figure}[H]
-\centering
-\includegraphics[scale=0.5]{./Figures/exampleastarmanhattansolver}
-%\decoRule
-\caption[Examples]{a* manhattan solver example}
+  \noindent
+  \makebox[\textwidth]{\includegraphics[scale=0.55]{./Figures/exampleastarmanhattansolver}}
+  \caption[Examples]{A$^{*}$[Manhattan] solver example}
   \label{fig:exampleastarmanhattansolver}
@@
 \begin{figure}[H]
-\centering
-\includegraphics[scale=0.5]{./Figures/exampleastarmanhattanplussolver}
-%\decoRule
-\caption[Examples]{a* manhattan++ solver example}
+  \noindent
+  \makebox[\textwidth]{\includegraphics[scale=0.55]{./Figures/exampleastarmanhattanplussolver}}
+  \caption[Examples]{A$^{*}$[Manhattan++] solver example}
   \label{fig:exampleastarmanhattanplussolver}
@@
 \afblue
-\paragraph{}{\textbf{python code -- depth first search solver}}
+\paragraph{}{\textbf{python code -- A$^{*}$[Perfect] solver}}
 \begin{python}
@@
 \begin{figure}[H]
-\centering
-\includegraphics[scale=0.5]{./Figures/exampleastarperfectsolver}
-%\decoRule
-\caption[Examples]{a* perfect solver example}
+  \noindent
+  \makebox[\textwidth]{\includegraphics[scale=0.55]{./Figures/exampleastarperfectsolver}}
+  \caption[Examples]{A$^{*}$[Perfect] solver example}
   \label{fig:exampleastarperfectsolver}
 \end{figure}
-
-\Subsubsection{Deep Learning Heuristic}
-\Subsubsection{Deep Reinforcement Learning Heuristic}
-\Subsubsection{Deep Q Learning Heuristic}

### 2022-09-13 latex/Appendices/Kociemba.tex
update
last cleanup

@@
 
-\chapter{Kociemba Bug} % Main chapter title
+\Appendix{Kociemba Bug} % Main chapter title
 
@@
 I personally like to follow the great Scott Meyers' advice (see item 18 of \cite{meyers2005effective}), which is to not merely make interfaces easy to use, but also difficult to misuse.
-In that instance, I would argue they have made it very easy to shoot oneself in the foot with their interface which assumes the cubes passed to the solver are in \textit{standard} form (my terminology), not only without making that explicit anywhere, but worse than that: sometimes non standard initial configuration raise an exception, sometimes not and returning a solution which is not really one!
+In that instance, I would argue they have made it very easy to shoot oneself in the foot with their interface which assumes the cubes passed to the solver are in \textit{standard} form (my terminology), not only without making that explicit anywhere, but worse than that: sometimes nonstandard initial configuration raise an exception, sometimes not and returning a solution which is not really one!
 So what is the problem exactly?
@@
 To start with, let us remember from chapter \ref{sec:Puzzles} that each \textbf{RC} configuration has 24 equivalent configurations under invariance by full rotation of the cube in space.
-Out of 24 equivalent configurations, and for an observer facing the cube, there is only one whose center front (F) cubie is red and center up (U) cubie is white.
+Out of 24 equivalent configurations, and for an observer facing the cube, there is only one whose centre front (F) cubie is red and centre up (U) cubie is white.
 \\
@@
 In the general case of a scrambled cube, it will also endeavour to find solutions to the closest among the 24 equivalent solved configurations.
-Let's look at the following concerete example:
+Let's look at the following concrete example:
 
@@
 In the 3x3x3 case, things are not quite that simple.
-There is no way to reproduce full cube rotation from faces rotations alone, since the center cubies will never move via the latter, but do via the former.
+There is no way to reproduce full cube rotation from faces rotations alone, since the centre cubies will never move via the latter but do via the former.
 If we run the equivalent of the above problem with the (3x3x3) kociemba library, we silently get to a false solution as is apparent from printing the resulting cube.
@@
 %\decoRule
-\caption[Kociemba Bug]{3x3x3 \textbf{RC} -- kociemba does not check center cubies and returns wrong solution}
+\caption[Kociemba Bug]{3x3x3 \textbf{RC} -- kociemba does not check centre cubies and returns wrong solution}
 \label{fig:KociembaBug333}

### 2022-09-13 latex/Chapters/DRLSearch.tex
last cleanup

@@
 \State{$puzzles\_generation\_criteria \gets$ \Call{FromParams}{**args}}
-\State{$n\_puzzles \gets \gets$ \Call{FromParams}{**args}}
+\State{$n\_puzzles \gets $ \Call{FromParams}{**args}}
 \State{$target\_network\_update\_params \gets$ \Call{FromParams}{**args}}
@@
 
-\While{epoch < max\_epochs \&\& other\_convergence\_criteria(network)}
+\While{epoch < max\_epochs \&\& other\_convergence\_criteria(net)}
         \State{$epoch \gets epoch + 1$}
@@
         \If{puzzles\_generation\_criteria($puzzles$)}
-            \State{$puzzles \gets generate\_random\_states(puzzle\_type)$}
+            \State{$puzzles \gets generate\_random\_states(puzzle\_type, n\_puzzles)$}
         \EndIf
@@
             \State{$V[puzzle] \gets \infty$}
-            \If \Call{IsGoal}{$puzzle$}:
+            \If{\Call{IsGoal}{$puzzle$}}:
                 \State{$V[puzzle] \gets 0$}
@@
 \\ \teal{\algorithmiccomment{perform value iteration}}\black
-		\State{$child_cost \gets \Call{Cost}{transition} + target\_net(child)$}
+		\State{$child\_cost \gets \Call{Cost}{transition} + target\_net(child)$}
                 	\State{$V[puzzle] \gets \Call{Min}{V[puzzle], child\_cost}$}
@@
 \\ \teal{\algorithmiccomment{iterate through network training running feed-forward and backward-prop}}\black
-	\State{$net \gets \Call{TrainNeuralNet}{V, loss\_function, net\_architecture}$}
+	\State{$net \gets \Call{TrainNeuralNet}{V, loss\_function, net}$}
 \\ \teal{\algorithmiccomment{update target network if criteria met}}\black
@@
 	\EndIf
+\EndWhile
 \Return{$net$}
-\EndWhile
 
@@
 	\EndFor
+\\ \teal{\algorithmiccomment{ ... same as \textbf{DRL} ...}}\black
+\State{$net \gets \Call{TrainNeuralNet}{Q\_and\_V, loss\_function, net}$}
 \EndWhile
@@
 \\
-Another possibility is to use an algorithm combining the actions distribution and the value function to inform its search.
-The Monte Carlo Tree Search (\textbf{MCTS}) is such an algorithm.
-It generates at each iteration (possibly in a distributed or multithreaded manner) paths to a \textit{leaf} on the frontier of unexpanded nodes, expands that node and checks if it is a goal state (and keeps going until the goal is found).
-The paths followed from initial state to leaves are decided via a combination of the probability distribution over actions (allowing exploration) and value function (exploitation), according to a trade-off which can be tuned via a hyper-parameter $c$.
-To prevent following the same paths (which might lead the search far from the goal), the \textbf{MCTS} keeps track of actions already followed and down-weight their probability so that subsequent paths explore different actions.
+Another possibility is to use an algorithm, such as the Monte Carlo Tree Search (\textbf{MCTS}), combining the actions distribution and the value function to inform its search.
+\textbf{MCTS} generates at each iteration (possibly in a distribute manner) paths to a \textit{leaf} on the frontier of unexpanded nodes, expands that node and checks if it is a goal state (and keeps going until it is).
+The paths followed from initial state to leaves are decided via a combination of the probability distribution over actions (exploration) and value function (exploitation), according to a trade-off which can be tuned via a hyper-parameter $c$.
+To prevent always following the same paths, the \textbf{MCTS} keeps track of actions already followed and down-weights their probability so that subsequent paths explore different actions.
 Let me first present in pseudo-code the paths generation, in particular the trade-off between actions-values (exploration-exploitation):
@@
 \State{\Call{InitialiseTimeOut}{$time\_out$}}
+\If{\Call{IsGoal}{$root$}}
 \State{$tree \gets \Call{Tree}{root}$}
-\If{\Call{IsGoal}{$root$}}
   \Return{$root$}
@@
 	\State\Call{CheckTimeOut}{void}
-	\State{$leaf \gets root$}
+	\State{$leaf \gets tree.root$}
 	\While{$leaf.expanded$}
@@
 
-\noindent Note that for this project, I have implemented all the above algorithms in Python (BFS, DFS, A$^{*}$, DeepLearner, DeepReinforcementLearner, DeepQLearner, MonteCarloTreeSearch), and more details can be found in section (\ref{sec:Implementation}).
+\noindent Note that for this project, I have implemented all the above algorithms in Python (\textbf{BFS}, \textbf{DFS}, A$^{*}$, DeepLearner, DeepReinforcementLearner, DeepQLearner, MonteCarloTreeSearch), and more details can be found in section \ref{sec:Implementation}.

### 2022-09-13 latex/Chapters/Implementation.tex
last cleanup

@@
 \end{figure}
-This modules contains graph search strategies. I have actually reused the code I implemented for one of the AIPnT assignments here. It contains the following classes:
+This modules contains graph search strategies. I have actually reused some of the code I implemented for an \textbf{AIPnT} assignment here. It contains the following classes:
 \begin{itemize}
@@
 The base class Solver is a Factory, and in addition to being able to instantiate the following types of solvers, can run different solvers through a similar sequence of random puzzles (for various increasing degrees of difficulty (scrambling), and/or perfectly shuffled ones) and display a comparison of how they perform in a number of metrics (configurable from percentage of puzzles solved before time out, percentage of optimal puzzle (if optimality can be ascertained), mean-median-maximum run time, mean-median-maximum expanded nodes (for solvers where it makes sense), mean-median-max cost, and an optimality score which compares the total sum of cost over all puzzles versus a benchmark solver).
-Ideally the benchmark solver to compute the optimality score should provably optimal, but in the case of the \textbf{RC} that was not possible, so I used A$^{*}$ with KociembaHeuristic as my benchmark.
+Ideally the benchmark solver to compute the optimality score should be provably optimal, but in the case of the \textbf{RC} that was not possible, so I used Kociemba as benchmark.
 For the \textbf{SP} I used the optimal A$^{*}$ with Manhattan++ heuristic.
@@
 \\
-This solver basically generates \textit{pseudo-random} paths from initial node to a new leaf at each iteration (see \ref{alg:TheoryMCTS}), until we obtain a goal leaf.Paths are, despite the name, completely deterministic: they indeed result from a trade-off between the probability of actions (penalised for repeated use) and the value function, but there are no random draws being performed.
+This solver basically generates \textit{pseudo-random} paths from initial node to a new leaf at each iteration (see \ref{alg:TheoryMCTS}), until we obtain a goal leaf.
+Paths are, despite the name, completely deterministic: they indeed result from a trade-off between the probability of actions (penalised for repeated use) and the value function, but there are no random draws being performed.
 \\ As per McAleer \& Agostilenni's paper, I have implemented a final pruning via \textbf{BFS} of the tree resulting from the above \textit{Monte-Carlo} process.

### 2022-09-13 latex/Chapters/SlidingPuzzle.tex
last cleanup

@@
 
-%\Subsection{Deep reinforcement learner}
-%\label{sec:S33DRL}
-
-\begin{comment}
-
-Next I have run the DeepReinforcementLearner with the following main parameters:
-\begin{itemize}
-\item Update the target network every 1,000 epochs (\textit{update\_target\_network\_frequency=1000}) or when the MSE falls below 10 basis points of the maximum target cost-to-go (\textit{update\_target\_network\_threshold=1e-3})
-\item Run for a maximum of 11,000 epochs (\textit{nb\_epochs=11000}) or when the maximum target so far has not been increasing by more than 1\% in the last 5,500 epochs (\textit{max\_target\_uptick=1e-2} and \textit{max\_target\_not\_increasing\_epochs\_pct=0.5})
-\item At every update of the target network (\textit{training\_data\_every\_epoch=False}), generate 100 sequences of puzzles (\textit{nb\_sequences=100}), each comprised of a 50-scramble puzzle back to goal state, with all intermediary puzzles along the path (\textit{nb\_shuffles=50}).
-\item Use a fully connected network (\textit{network\_type=fully\_connected\_net}) with three hidden layers of size 600, 300, 100 (\textit{layers\_description=(600,300,100)}) and use torch.optim.RMSProp as optimiser (\textit{optimiser=rms\_prop}) with an exponential scheduler (\textit{scheduler=exponential\_scheduler}) with gamma 0.999 (\textit{gamma\_scheduler=0.9999}).
-\end{itemize}
-Below I show some of the DeepReinforcementLearner's quantities tracked over the epochs (see \ref{fig:33SPDeepReinforcementLearning}). I have shown only the following quantities:
-\begin{itemize}
-\item learning rate, which decreases due to the scheduler, but gets reset at each network update.
-\item max target.
-One interesting dynamic during learning is that due to the way targets are constructed using value-iteration update, and starting from an all 0 network, the max target only increases by about 1 at best at every network update.
-That is, the first network only learns how to differentiate goal from non-goal, then the second one learns how to differentiate goals, cost 1 and cost 2+ in some sense.
-This is why I have introduced the first set of parameters discussed just above.
-The first one makes sure we do not waste time initially training the left-hand-side network when training is easy (as consist of just goal versus not goal).
-The second ones make sure that we stop training altogether when the max target does not grow anymore.
-This is because we do not know a-priori what the God number of a given puzzle/dimension is, but let's assume for the sake of the argument that it is 15.
-Then that means that after 15 updates of the target network we might not really get so much benefit by keeping training and value-iterating.
-\item the MSE loss
-\item the MSE loss divided by the max target.
-The exit criterion \textit{update\_target\_network\_threshold} is based on that quantity, since a loss of e.g.
-0.01 means little if the we do not compare it to the possible range that the cost-to-go can take.
-\item The percentage of all the possible puzzles of that dimension \textit{seen} by the DeepReinforcementLearner.
-This is an interesting quantity, as it tells us later on when testing out of sample and comparing with other solvers, how well the solvers are able to generalise.
-In this case (3x3) we can see that the DeepReinforcementLearner quickly has seen a very large percentage of all puzzles.
-That being said, it is good to keep in mind that \textit{seen} does not mean really much here, since the learning is unsupervised.
-This is quite in contrast to the DeepLearner, for which puzzles seen come along with the cost-to-go computed by a \textit{teacher}, usually an optimal or at least efficient solver.
-\end{itemize}
-
-\begin{figure}[H]
-\centering
-\includegraphics[align=c, scale=0.42]{./Figures/33SPDeepReinforcementLearning}
-\caption[33SPDeepReinforcementLearning]{\textbf{DRL} of the 3x3 \textbf{SP}}
-\label{fig:33SPDeepReinforcementLearning}
-\end{figure}
-
-Notice that the DeepLearner and DeepQLearner have very similar sets of parameters and track similar metrics during their learning for later display, but I shall not detail it here since there is not much extra insight to be had compared to the DeepReinforcementLearning example we just went through.
-
-\end{comment}
-
 \Subsection{MCTS DQL}

### 2022-09-13 rubiks/scripts/draft4.py

@@ -1,16 +1,14 @@
-# ###################################################################
-from math import inf
-# ###################################################################
-from rubiks . deeplearning . deeplearning import DeepLearning
-from rubiks . learners . learner import Learner
-from rubiks . learners . deeplearner import DeepLearner
-from rubiks . puzzle . puzzle import Puzzle
-from rubiks . puzzle . trainingdata import TrainingData
-# ###################################################################
+####################################################################
+from rubiks.solvers.solver import Solver
+from rubiks.puzzle.puzzle import Puzzle
+####################################################################
 if '__main__' == __name__:
-    puzzle_type = Puzzle . rubiks_cube
-    n=2
-    init_from_random_goal = False
-    goals = Puzzle.factory(**globals()).get_equivalent()
-    print(goals[0], goals[-1])
-# ##################################################################
+    puzzle_type = Puzzle.sliding_puzzle
+    n=3
+    m=3
+    nb_shuffles=5
+    solver_type=Solver.bfs
+    check_optimal=True
+    action_type=Solver.do_solve
+    print(Solver.factory(**globals()).action())
+####################################################################

### 2022-12-17 latex/Chapters/Puzzles.tex
typos

@@
 (The additional degrees of freedom on the 4x4x4 complicate slightly the story since certain arrangements of colors which are possible on the 4x4x4 are not on the 3x3x3, e.g.
-red an and white on opposite faces).
+red and white on opposite faces).
 Overall, the difficulty of hand-solving cubes does not grow as much with dimensionality as one would think, and the time difference comes mostly from the difficulty of handling and manipulating larger cubes.
@@
 \\
-Since I am interested in experimenting with \textit{no-human-knowledge} methods such as \textbf{DRL} and \textbf{DQL}, I will consider cubes which are identical up to full cube rotation such as the \textit{flattened} cubes on figure \ref{fig:configurations} ) to be different.
+Since I am interested in experimenting with \textit{no-human-knowledge} methods such as \textbf{DRL} and \textbf{DQL}, I will consider cubes which are identical up to full cube rotation such as the \textit{flattened} cubes on figure \ref{fig:configurations} to be different.
 They will be represented by different tensors in my code, and this is up to the solvers to make sense of things.

### 2022-12-17 latex/Chapters/Rubiks.tex
last cleanup
typos

@@
   \noindent
-  \makebox[\textwidth]{\includegraphics[scale=0.4]{./Figures//222RCDRL}}
+  \makebox[\textwidth]{\includegraphics[scale=0.4]{./Figures/222RCDRL}}
   \caption[222RCDRL]{\textbf{DRL} of the 2x2x2 \textbf{RC}}
@@
   \noindent
-  \makebox[\textwidth]{\includegraphics[scale=0.4]{./Figures//222RCDQL}}
+  \makebox[\textwidth]{\includegraphics[scale=0.4]{./Figures/222RCDQL}}
   \caption[222RCDQL]{\textbf{DQL} of the 2x2x2 \textbf{RC}}
@@
 Notice that the loss of 0.2 towards which the network converged at the end of its training is misleading, since this is in-sample loss versus the target network.
-My theory is that while the value the \textbf{DRL} assigned to hard-to-solve cubes can be off (by at least 4 or 5), it is likely that the network gets the relative value of neighbouring cubes right (not that surprising given the value-iteration), and this is sufficient to guide A$^{*}$ extremely well.
+My theory is that while the value the \textbf{DRL} network assigned to hard-to-solve cubes can be off (by at least 4 or 5), it is likely that it gets the relative value of neighbouring cubes right (not that surprising given value-iteration assigns values from that of neighbours!), and this is sufficient to guide A$^{*}$ extremely well.
 If both the absolute value of the cost-to-go and the relative (among neighbouring cubes) was totally off, we would indeed not expect the heuristic to make A$^{*}$ any better than, say, \textbf{BFS} and we would therefore not expect it to be able to solve \textit{costly} cubes requiring 10+ moves.

### 2022-12-17 latex/Chapters/SlidingPuzzle.tex
typos

@@
 \item How does \textbf{MCTS} perform? How important is it to tune the hyper-parameter $c$? How much improvement does the final \textbf{BFS} tree-trimming bring?
-\item Are the \textbf{D*L} heuristics competitive versus Manhattan++ or the perfect heuristic?
+\item Are the \textbf{DxL} heuristics competitive versus Manhattan++ or the perfect heuristic?
 \end{itemize}
@@
 \label{sec:S33DRLDQL}
-As promised earlier in table \ref{tab:gridSP}, I have run the Solver's comparison test with 10 different solvers (Naive, \textbf{MCTS}, \textbf{BFS}, A$^{*}$ with the following heuristics: Manhattan, Manhattan++, Perfect, \textbf{DRL}, \text{DQL} and \text{DL}), to get an idea of their respective success rate, strengths and weaknesses.
+As promised earlier in table \ref{tab:gridSP}, I have run the Solver's comparison test with 10 different solvers (Naive, \textbf{MCTS}, \textbf{BFS}, A$^{*}$ with the following heuristics: Manhattan, Manhattan++, Perfect, \textbf{DRL}, \textbf{DQL} and \textbf{DL}), to get an idea of their respective success rate, strengths and weaknesses.
 The \textbf{DRL} and \textbf{DQL} heuristics were trained with the exact same parameters and a fully connected network of 3 hidden layers with 600, 300 and 100 nodes.
@@
 That did improve things, but overall I saw little difference in performance from my simpler 3-hidden-layer fully connected network.
-\item \textbf{DRL} and \textbf{DQL} heuristics, though not as efficient in terms node-expansion as the perfect and the \textbf{DL} heuristics, were rather surprisingly doing better than Manhattan++.
+\item \textbf{DRL} and \textbf{DQL} heuristics, though not as efficient in terms of node-expansion as the perfect and the \textbf{DL} heuristics, were rather surprisingly doing better than Manhattan++.
 They ended up \textit{seeing} around 25\% of all the possible configurations during training (in an \textit{unsupervised} way though) so did show some reasonable amount of out-of-sample generalization.

### 2022-12-17 rubiks/scripts/rubiks22/perf.py

@@ -64,7 +64,7 @@ def performance(**kw_args):
 
 if '__main__' == __name__:
     logger = Loggable(name='Perf22RC')
-    action_type = Solver.do_performance_test
+    action_type = Solver.do_plot
     fig_size = (12, 11)
     if action_type == Solver.do_plot:
         exclude_solver_names = ['1000seq_20shf']
